### 2013-03-18 figurefactory/files/limno_bgape_cl.tex
First few pages of motivation

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (1478,0)(1478,-0.181818181818182)(1478,-0.363636363636364)(1478,-0.545454545454545)(1478,-0.727272727272727)(1478,-0.909090909090909)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454545)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454545)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454545)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478.00016587919,-9.09090909090909)(1478.00033175469,-9.27272727272727)(1478.00033175469,-9.45454545454546)(1478.00016587919,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454545)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454545)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454545)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.909090909090909)(0,-0.727272727272727)(0,-0.545454545454545)(0,-0.363636363636364)(0,-0.181818181818182)(0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1448.14141414141,-9.07482483753433)(1449.38005040762,-9.09090909090909)(1458.72266748613,-9.27272727272727)(1462.16461608434,-9.45454545454546)(1457.93739588908,-9.63636363636364)(1448.14141414141,-9.76880948045942)(1443.79101312806,-9.81818181818182)(1433.21212121212,-9.89042536657361)(1418.28282828283,-9.97220402070884)(1411.64380927072,-10)(1403.35353535354,-10.0268854789336)(1388.42424242424,-10.0624512420213)(1373.49494949495,-10.0858912269465)(1358.56565656566,-10.0987182864105)(1343.63636363636,-10.102962992112)(1328.70707070707,-10.1010832499639)(1313.77777777778,-10.0957594734832)(1298.84848484848,-10.0895855113828)(1283.91919191919,-10.0847193351369)(1268.9898989899,-10.0825928939416)(1254.06060606061,-10.083784643917)(1239.13131313131,-10.0881421355553)(1224.20202020202,-10.0951010924332)(1209.27272727273,-10.1039452569402)(1194.34343434343,-10.1139499357554)(1179.41414141414,-10.1244504589076)(1164.48484848485,-10.1348650752363)(1149.55555555556,-10.1446919253603)(1134.62626262626,-10.1534885135226)(1119.69696969697,-10.1608315698405)(1104.76767676768,-10.1662517457766)(1089.83838383838,-10.1691748423721)(1074.90909090909,-10.1689575399048)(1059.9797979798,-10.1649715724152)(1045.05050505051,-10.1566770588623)(1030.12121212121,-10.1436963503898)(1015.19191919192,-10.125889781726)(1000.26262626263,-10.1034124247913)(985.333333333333,-10.0767151822671)(970.40404040404,-10.0464552735906)(955.474747474747,-10.013307823769)(949.978893163733,-10)(940.545454545455,-9.97807181148041)(925.616161616162,-9.94087063608248)(910.686868686869,-9.90105170089348)(895.757575757576,-9.8577226257769)(883.509186627693,-9.81818181818182)(880.828282828283,-9.80934461405926)(865.89898989899,-9.75303239531643)(850.969696969697,-9.68820509581558)(840.658582570405,-9.63636363636364)(836.040404040404,-9.60972698978828)(821.111111111111,-9.51036769366592)(813.765122243986,-9.45454545454546)(806.181818181818,-9.37884030418043)(796.52670901727,-9.27272727272727)(791.252525252525,-9.18172850331714)(786.1120165418,-9.09090909090909)(781.350004048785,-8.90909090909091)(781.700292860412,-8.72727272727273)(787.055375547823,-8.54545454545455)(791.252525252525,-8.46953330263129)(797.389084244081,-8.36363636363636)(806.181818181818,-8.25527238039726)(812.491262287979,-8.18181818181818)(821.111111111111,-8.09778589767859)(831.615728238001,-8)(836.040404040404,-7.96248037720495)(850.969696969697,-7.83881685491404)(853.498612390076,-7.81818181818182)(865.89898989899,-7.72191737491956)(877.441030726608,-7.63636363636364)(880.828282828283,-7.61233969300808)(895.757575757576,-7.51269570109658)(905.474751568828,-7.45454545454546)(910.686868686869,-7.42514082302524)(925.616161616162,-7.35260757307406)(940.545454545455,-7.29483688348728)(948.20983898438,-7.27272727272727)(955.474747474747,-7.25281917827936)(970.40404040404,-7.22671168195057)(985.333333333333,-7.21599459418017)(1000.26262626263,-7.22086934225301)(1015.19191919192,-7.24145872752697)(1028.19889936542,-7.27272727272727)(1030.12121212121,-7.27753437771899)(1045.05050505051,-7.3274095902308)(1059.9797979798,-7.38713601418184)(1074.90909090909,-7.45220989134266)(1075.44825719334,-7.45454545454546)(1089.83838383838,-7.51534267826917)(1104.76767676768,-7.57160221512044)(1119.69696969697,-7.61795889862764)(1127.48665225104,-7.63636363636364)(1134.62626262626,-7.65266400945172)(1149.55555555556,-7.67569822949993)(1164.48484848485,-7.68847110284722)(1179.41414141414,-7.69265816226828)(1194.34343434343,-7.6901830729894)(1209.27272727273,-7.68316533122948)(1224.20202020202,-7.67397725334571)(1239.13131313131,-7.66538965087568)(1254.06060606061,-7.66076041767317)(1268.9898989899,-7.66414041509026)(1283.91919191919,-7.68014015763703)(1298.84848484848,-7.71349387267142)(1313.77777777778,-7.76842040428974)(1323.34244512792,-7.81818181818182)(1328.70707070707,-7.85007561262625)(1343.63636363636,-7.9609183391006)(1348.14561466569,-8)(1358.56565656566,-8.0977740197848)(1366.90866754043,-8.18181818181818)(1373.49494949495,-8.25076188175244)(1384.2176058536,-8.36363636363636)(1388.42424242424,-8.4089083491506)(1401.52980262361,-8.54545454545455)(1403.35353535354,-8.56527958618061)(1418.28282828283,-8.72039779999265)(1418.96751701876,-8.72727272727273)(1433.21212121212,-8.88456727515451)(1435.49658592742,-8.90909090909091)(1448.14141414141,-9.07482483753433)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (432.949494949495,-7.62537808110239)(447.878787878788,-7.49515945642039)(453.634273405795,-7.45454545454546)(462.808080808081,-7.40824862633082)(477.737373737374,-7.3473805919878)(492.666666666667,-7.3028531014644)(507.59595959596,-7.27314545133157)(507.971324441396,-7.27272727272727)(522.525252525253,-7.25757880472974)(537.454545454546,-7.25458343000404)(552.383838383838,-7.26514500602056)(556.76546517783,-7.27272727272727)(567.313131313131,-7.29243391259954)(582.242424242424,-7.34173271265313)(597.171717171717,-7.42061591326051)(601.732815751869,-7.45454545454546)(612.10101010101,-7.5550535180392)(618.323625701474,-7.63636363636364)(627.030303030303,-7.81386668638062)(627.208049689042,-7.81818181818182)(631.340965192909,-8)(632.004804095685,-8.18181818181818)(629.389256087845,-8.36363636363636)(627.030303030303,-8.43471171011093)(623.324643902992,-8.54545454545455)(613.424218243758,-8.72727272727273)(612.10101010101,-8.74457614261215)(598.308154116587,-8.90909090909091)(597.171717171717,-8.91954081438586)(582.242424242424,-9.03637501436889)(573.133372914227,-9.09090909090909)(567.313131313131,-9.11967949971697)(552.383838383838,-9.16910271704596)(537.454545454546,-9.1926915906799)(522.525252525253,-9.19017711734558)(507.59595959596,-9.16118788005291)(492.666666666667,-9.10554360347104)(489.876759750011,-9.09090909090909)(477.737373737374,-9.00950811684864)(465.842909877035,-8.90909090909091)(462.808080808081,-8.87505513198468)(450.744269474239,-8.72727272727273)(447.878787878788,-8.67986623616029)(439.49329629746,-8.54545454545455)(432.949494949495,-8.40359706858181)(430.814624854265,-8.36363636363636)(424.148612809654,-8.18181818181818)(420.777096197368,-8)(422.257083613328,-7.81818181818182)(431.799598096469,-7.63636363636364)(432.949494949495,-7.62537808110239)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (119.434343434343,-7.92037924231854)(122.79885412663,-7.81818181818182)(134.363636363636,-7.73778282591567)(149.292929292929,-7.69947037438622)(164.222222222222,-7.69597240685678)(179.151515151515,-7.7116912954455)(194.080808080808,-7.73959272188856)(209.010101010101,-7.77631163252306)(223.22348855859,-7.81818181818182)(223.939393939394,-7.82088268510329)(238.868686868687,-7.88589235730763)(253.79797979798,-7.96215394608578)(260.195186932805,-8)(268.727272727273,-8.0909207600127)(275.609446204187,-8.18181818181818)(278.558915117606,-8.36363636363636)(273.150031331953,-8.54545454545455)(268.727272727273,-8.60336562419625)(258.062723933781,-8.72727272727273)(253.79797979798,-8.75528927680203)(238.868686868687,-8.82877256167601)(223.939393939394,-8.87148838869149)(209.010101010101,-8.88240467359964)(194.080808080808,-8.86254345832509)(179.151515151515,-8.81420352665458)(164.222222222222,-8.7402069702458)(162.255837112136,-8.72727272727273)(149.292929292929,-8.61230482699614)(142.731707572687,-8.54545454545455)(134.363636363636,-8.42739270293778)(130.062418671086,-8.36363636363636)(121.748906826838,-8.18181818181818)(119.434343434343,-8.06796056232384)(117.963012369102,-8)(119.434343434343,-7.92037924231854)};

\end{axis}
\end{tikzpicture}%