### 2013-03-19 figurefactory/files/voned_1_2.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
% 
\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
label shift={-4pt},
xlabel shift={-6pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=2,
xtick={0, 0.5, 1, 1.5, 2},
ymin=-2, ymax=4,
tickwidth=0cm,
]

\addplot [fill=olive!20!lightgray,draw=none,forget plot] coordinates{ (0,3.00405948968096)(0.01,3.00426406394551)(0.02,3.00447860979409)(0.03,3.00470358994989)(0.04,3.00493948692744)(0.05,3.00518680375087)(0.06,3.00544606468753)(0.07,3.00571781599646)(0.08,3.00600262669061)(0.09,3.00630108931215)(0.1,3.00661382071944)(0.11,3.00694146288451)(0.12,3.00728468369953)(0.13,3.00764417779044)(0.14,3.00802066733599)(0.15,3.00841490288972)(0.16,3.00882766420274)(0.17,3.00925976104432)(0.18,3.00971203401727)(0.19,3.01018535536484)(0.2,3.0106806297652)(0.21,3.01119879510944)(0.22,3.01174082325852)(0.23,3.012307720774)(0.24,3.01290052961726)(0.25,3.01352032781092)(0.26,3.01416823005603)(0.27,3.01484538829772)(0.28,3.01555299223157)(0.29,3.0162922697421)(0.3,3.01706448726417)(0.31,3.01787095005733)(0.32,3.01871300238221)(0.33,3.01959202756734)(0.34,3.0205094479538)(0.35,3.02146672470404)(0.36,3.02246535746042)(0.37,3.02350688383777)(0.38,3.02459287873308)(0.39,3.02572495343459)(0.4,3.02690475451079)(0.41,3.0281339624591)(0.42,3.02941429009222)(0.43,3.03074748063894)(0.44,3.03213530553473)(0.45,3.03357956187592)(0.46,3.03508206950958)(0.47,3.03664466772985)(0.48,3.03826921154973)(0.49,3.0399575675154)(0.5,3.04171160902898)(0.51,3.04353321114326)(0.52,3.04542424479078)(0.53,3.04738657040727)(0.54,3.04942203090831)(0.55,3.0515324439758)(0.56,3.05371959360932)(0.57,3.05598522089594)(0.58,3.05833101394996)(0.59,3.06075859697291)(0.6,3.06326951838249)(0.61,3.06586523795771)(0.62,3.06854711294636)(0.63,3.07131638307979)(0.64,3.07417415443909)(0.65,3.07712138211621)(0.66,3.08015885161281)(0.67,3.08328715891981)(0.68,3.08650668922046)(0.69,3.08981759416052)(0.7,3.09321976762968)(0.71,3.09671281999994)(0.72,3.1002960507682)(0.73,3.10396841955286)(0.74,3.1077285153967)(0.75,3.11157452433218)(0.76,3.11550419516906)(0.77,3.11951480346934)(0.78,3.12360311367997)(0.79,3.12776533940035)(0.8,3.13199710176903)(0.81,3.13629338596207)(0.82,3.14064849580542)(0.83,3.14505600651351)(0.84,3.14950871557854)(0.85,3.1539985918473)(0.86,3.15851672283708)(0.87,3.16305326035778)(0.88,3.16759736452434)(0.89,3.172137146263)(0.9,3.17665960843504)(0.91,3.18115058572431)(0.92,3.18559468345955)(0.93,3.18997521556861)(0.94,3.19427414189104)(0.95,3.19847200510653)(0.96,3.20254786757065)(0.97,3.20647924838623)(0.98,3.21024206107832)(0.99,3.21381055228393)(1,3.21715724191413)(1.01,3.22025286529712)(1.02,3.223066317865)(1.03,3.22556460300701)(1.04,3.22771278377613)(1.05,3.22947393920636)(1.06,3.23080912607447)(1.07,3.23167734702335)(1.08,3.2320355260559)(1.09,3.23183849250862)(1.1,3.23103897472494)(1.11,3.22958760477095)(1.12,3.2274329356729)(1.13,3.22452147280872)(1.14,3.22079772125822)(1.15,3.21620425111303)(1.16,3.210681782972)(1.17,3.20416929610886)(1.18,3.19660416210425)(1.19,3.18792230709721)(1.2,3.17805840624695)(1.21,3.16694611452772)(1.22,3.15451833863763)(1.23,3.14070755562896)(1.24,3.12544618492303)(1.25,3.10866702174)(1.26,3.09030374177493)(1.27,3.07029148935924)(1.28,3.0485675646173)(1.29,3.02507222963895)(1.3,2.99974966000901)(1.31,2.9725490770283)(1.32,2.9434261089661)(1.33,2.91234444879772)(1.34,2.87927790445372)(1.35,2.84421298107019)(1.36,2.80715220205927)(1.37,2.76811848209943)(1.38,2.72716103618234)(1.39,2.68436358948199)(1.4,2.6398561218657)(1.41,2.59383217748261)(1.42,2.54657513623406)(1.43,2.49849917893294)(1.44,2.45021453042913)(1.45,2.40263220045413)(1.46,2.35712849040268)(1.47,2.31577983705733)(1.48,2.28160327978947)(1.49,2.25849531293989)(1.5,2.25023914277021)(1.51,2.25849531293989)(1.52,2.28160327978947)(1.53,2.31577983705733)(1.54,2.35712849040268)(1.55,2.40263220045413)(1.56,2.45021453042913)(1.57,2.49849917893294)(1.58,2.54657513623407)(1.59,2.59383217748261)(1.6,2.6398561218657)(1.61,2.68436358948199)(1.62,2.72716103618234)(1.63,2.76811848209943)(1.64,2.80715220205927)(1.65,2.84421298107019)(1.66,2.87927790445372)(1.67,2.91234444879772)(1.68,2.9434261089661)(1.69,2.97254907702831)(1.7,2.99974966000901)(1.71,3.02507222963895)(1.72,3.0485675646173)(1.73,3.07029148935924)(1.74,3.09030374177493)(1.75,3.10866702174)(1.76,3.12544618492303)(1.77,3.14070755562896)(1.78,3.15451833863763)(1.79,3.16694611452772)(1.8,3.17805840624695)(1.81,3.18792230709721)(1.82,3.19660416210425)(1.83,3.20416929610886)(1.84,3.21068178297201)(1.85,3.21620425111303)(1.86,3.22079772125822)(1.87,3.22452147280872)(1.88,3.2274329356729)(1.89,3.22958760477095)(1.9,3.23103897472494)(1.91,3.23183849250862)(1.92,3.2320355260559)(1.93,3.23167734702335)(1.94,3.23080912607447)(1.95,3.22947393920636)(1.96,3.22771278377613)(1.97,3.22556460300701)(1.98,3.223066317865)(1.99,3.22025286529712)(2,3.21715724191413)(2,-0.547855845922187)(1.99,-0.527263965504428)(1.98,-0.505788009766045)(1.97,-0.483396179083081)(1.96,-0.460056220235874)(1.95,-0.435735499999608)(1.94,-0.410401089005194)(1.93,-0.384019856679305)(1.92,-0.356558578115416)(1.91,-0.327984053773209)(1.9,-0.298263242952087)(1.89,-0.267363412036782)(1.88,-0.235252298570199)(1.87,-0.201898292272354)(1.86,-0.167270634196579)(1.85,-0.131339635298263)(1.84,-0.0940769157910251)(1.83,-0.0554556667863313)(1.82,-0.0154509358624957)(1.81,0.0259600616013633)(1.8,0.068797603246997)(1.79,0.113079090316272)(1.78,0.15881864516803)(1.77,0.206026672346912)(1.76,0.254709379052199)(1.75,0.304868249805843)(1.74,0.356499468669974)(1.73,0.409593280343953)(1.72,0.464133278627427)(1.71,0.52009560669452)(1.7,0.577448047820867)(1.69,0.636148976778726)(1.68,0.696146129733814)(1.67,0.757375132046748)(1.66,0.819757695559522)(1.65,0.883199354305343)(1.64,0.947586541150821)(1.63,1.0127827025836)(1.62,1.07862297889869)(1.61,1.1449066975866)(1.6,1.21138645995895)(1.59,1.27775180583925)(1.58,1.34360407631222)(1.57,1.40841676119969)(1.56,1.47147176728893)(1.55,1.53175640985815)(1.54,1.58780088375249)(1.53,1.63744566984478)(1.52,1.6776036197675)(1.51,1.7043239894427)(1.5,1.71378843730561)(1.49,1.7043239894427)(1.48,1.6776036197675)(1.47,1.63744566984478)(1.46,1.58780088375249)(1.45,1.53175640985815)(1.44,1.47147176728893)(1.43,1.40841676119969)(1.42,1.34360407631223)(1.41,1.27775180583925)(1.4,1.21138645995896)(1.39,1.1449066975866)(1.38,1.07862297889869)(1.37,1.0127827025836)(1.36,0.947586541150821)(1.35,0.883199354305343)(1.34,0.819757695559521)(1.33,0.757375132046748)(1.32,0.696146129733814)(1.31,0.636148976778728)(1.3,0.577448047820867)(1.29,0.520095606694521)(1.28,0.464133278627426)(1.27,0.409593280343953)(1.26,0.356499468669972)(1.25,0.304868249805843)(1.24,0.254709379052201)(1.23,0.206026672346912)(1.22,0.15881864516803)(1.21,0.113079090316273)(1.2,0.0687976032469968)(1.19,0.0259600616013622)(1.18,-0.0154509358624968)(1.17,-0.0554556667863313)(1.16,-0.094076915791024)(1.15,-0.131339635298263)(1.14,-0.167270634196579)(1.13,-0.201898292272354)(1.12,-0.235252298570199)(1.11,-0.267363412036783)(1.1,-0.298263242952087)(1.09,-0.327984053773209)(1.08,-0.356558578115416)(1.07,-0.384019856679306)(1.06,-0.410401089005195)(1.05,-0.435735499999608)(1.04,-0.460056220235874)(1.03,-0.483396179083081)(1.02,-0.505788009766045)(1.01,-0.527263965504428)(1,-0.547855845922187)(0.99,-0.567594932959715)(0.98,-0.58651193556049)(0.97,-0.604636942442146)(0.96,-0.62199938229867)(0.95,-0.638627990816008)(0.94,-0.654550783917867)(0.93,-0.669795036691896)(0.92,-0.684387267478763)(0.91,-0.698353226637922)(0.9,-0.711717889534112)(0.89,-0.724505453317808)(0.88,-0.736739337100937)(0.87,-0.748442185156271)(0.86,-0.759635872794845)(0.85,-0.77034151460069)(0.84,-0.780579474725984)(0.83,-0.790369378972483)(0.82,-0.799730128406764)(0.81,-0.808679914277439)(0.8,-0.817236234022034)(0.79,-0.825415908169748)(0.78,-0.833235097963765)(0.77,-0.840709323543267)(0.76,-0.847853482540755)(0.75,-0.854681868964777)(0.74,-0.861208192251741)(0.73,-0.867445596383093)(0.72,-0.873406678975932)(0.71,-0.879103510265999)(0.7,-0.884547651912082)(0.69,-0.889750175560155)(0.68,-0.894721681114113)(0.67,-0.899472314667793)(0.66,-0.904011786060114)(0.65,-0.908349386021669)(0.64,-0.912494002886998)(0.63,-0.916454138852106)(0.62,-0.92023792576156)(0.61,-0.923853140413804)(0.6,-0.927307219377159)(0.59,-0.930607273312352)(0.58,-0.933760100800408)(0.57,-0.936772201677387)(0.56,-0.939649789879698)(0.55,-0.9423988058057)(0.54,-0.945024928201005)(0.53,-0.9475335855763)(0.52,-0.9499299671677)(0.51,-0.952219033450653)(0.5,-0.954405526219157)(0.49,-0.956493978242716)(0.48,-0.958488722513901)(0.47,-0.960393901099713)(0.46,-0.962213473610178)(0.45,-0.963951225297682)(0.44,-0.965610774800598)(0.43,-0.967195581544668)(0.42,-0.9687089528155)(0.41,-0.970154050515327)(0.4,-0.971533897616946)(0.39,-0.972851384327491)(0.38,-0.974109273974362)(0.37,-0.975310208625306)(0.36,-0.976456714454296)(0.35,-0.977551206864464)(0.34,-0.978595995378974)(0.33,-0.97959328831033)(0.32,-0.980545197218213)(0.31,-0.981453741165549)(0.3,-0.982320850782134)(0.29,-0.983148372144719)(0.28,-0.983938070482126)(0.27,-0.984691633713544)(0.26,-0.985410675827823)(0.25,-0.986096740111221)(0.24,-0.986751302230704)(0.23,-0.987375773179584)(0.22,-0.987971502091947)(0.21,-0.988539778932026)(0.2,-0.989081837064361)(0.19,-0.989598855710323)(0.18,-0.990091962296299)(0.17,-0.990562234698562)(0.16,-0.991010703389634)(0.15,-0.991438353490681)(0.14,-0.991846126734268)(0.13,-0.992234923341592)(0.12,-0.992605603818107)(0.11,-0.992958990671239)(0.1,-0.993295870053748)(0.09,-0.993616993336077)(0.08,-0.993923078610886)(0.07,-0.994214812132822)(0.06,-0.994492849696389)(0.05,-0.994757817954686)(0.04,-0.995010315681626)(0.03,-0.995250914980116)(0.02,-0.995480162438577)(0.01,-0.995698580238063)(0,-0.995906667212118)};

\addplot [
color=olive!70!black,
solid,
line width=2.0pt,
forget plot
]
coordinates{
 (0,1.0040897452219)(0.01,1.00429675075101)(0.02,1.00451394068955)(0.03,1.00474179738852)(0.04,1.00498082483891)(0.05,1.00523154956363)(0.06,1.00549452154094)(0.07,1.00577031516001)(0.08,1.00605953020939)(0.09,1.00636279289909)(0.1,1.00668075691684)(0.11,1.0070141045194)(0.12,1.00736354765928)(0.13,1.00772982914765)(0.14,1.00811372385394)(0.15,1.00851603994254)(0.16,1.00893762014722)(0.17,1.00937934308348)(0.18,1.00984212459929)(0.19,1.01032691916444)(0.2,1.01083472129863)(0.21,1.01136656703844)(0.22,1.01192353544313)(0.23,1.01250675013919)(0.24,1.01311738090328)(0.25,1.01375664528338)(0.26,1.01442581025747)(0.27,1.01512619392907)(0.28,1.01585916725891)(0.29,1.01662615583157)(0.3,1.01742864165593)(0.31,1.01826816499799)(0.32,1.01914632624425)(0.33,1.02006478779388)(0.34,1.0210252759773)(0.35,1.02202958299879)(0.36,1.02307956890017)(0.37,1.02417716354244)(0.38,1.02532436860184)(0.39,1.0265232595763)(0.4,1.02777598779804)(0.41,1.02908478244737)(0.42,1.03045195256247)(0.43,1.03187988903927)(0.44,1.03337106661514)(0.45,1.03492804582929)(0.46,1.03655347495241)(0.47,1.03825009187723)(0.48,1.04002072596113)(0.49,1.04186829981092)(0.5,1.04379583099951)(0.51,1.04580643370286)(0.52,1.04790332024529)(0.53,1.05008980253964)(0.54,1.05236929340842)(0.55,1.05474530777048)(0.56,1.05722146367707)(0.57,1.0598014831797)(0.58,1.06248919301117)(0.59,1.06528852505978)(0.6,1.06820351661536)(0.61,1.07123831036455)(0.62,1.07439715411089)(0.63,1.07768440019436)(0.64,1.08110450458274)(0.65,1.08466202560602)(0.66,1.08836162230311)(0.67,1.0922080523484)(0.68,1.09620616952385)(0.69,1.10036092070029)(0.7,1.10467734228994)(0.71,1.10916055612972)(0.72,1.11381576475312)(0.73,1.11864824600634)(0.74,1.12366334696196)(0.75,1.12886647708148)(0.76,1.13426310057572)(0.77,1.13985872790991)(0.78,1.14565890639797)(0.79,1.15166920982834)(0.8,1.15789522706149)(0.81,1.16434254953709)(0.82,1.17101675762665)(0.83,1.17792340576555)(0.84,1.18506800629633)(0.85,1.19245601195347)(0.86,1.20009279691821)(0.87,1.20798363637049)(0.88,1.21613368446413)(0.89,1.22454795065031)(0.9,1.233231274274)(0.91,1.24218829736773)(0.92,1.2514234355676)(0.93,1.2609408470771)(0.94,1.27074439960581)(0.95,1.28083763521224)(0.96,1.29122373298292)(0.97,1.30190546948369)(0.98,1.31288517692387)(0.99,1.32416469897996)(1,1.33574534423277)(1.01,1.34762783718023)(1.02,1.35981226679868)(1.03,1.37229803263705)(1.04,1.3850837884424)(1.05,1.39816738333153)(1.06,1.41154580054166)(1.07,1.42521509381473)(1.08,1.43917032149395)(1.09,1.45340547843888)(1.1,1.46791342589672)(1.11,1.48268581950264)(1.12,1.49771303562191)(1.13,1.51298409629154)(1.14,1.52848659306892)(1.15,1.54420661015185)(1.16,1.56012864719645)(1.17,1.57623554233018)(1.18,1.59250839593491)(1.19,1.60892649586181)(1.2,1.6254672448364)(1.21,1.64210609091893)(1.22,1.65881646200356)(1.23,1.67556970547131)(1.24,1.69233503425594)(1.25,1.70907948074268)(1.26,1.72576786009534)(1.27,1.74236274480266)(1.28,1.75882445244819)(1.29,1.77511104894428)(1.3,1.79117836972934)(1.31,1.80698006171278)(1.32,1.82246764906407)(1.33,1.83759062628575)(1.34,1.85229658238609)(1.35,1.86653136037919)(1.36,1.88023925679178)(1.37,1.89336326634995)(1.38,1.90584537755949)(1.39,1.91762692548457)(1.4,1.92864900867555)(1.41,1.9388529779019)(1.42,1.94818100511719)(1.43,1.95657674192344)(1.44,1.96398607771993)(1.45,1.97035800872147)(1.46,1.97564563012208)(1.47,1.97980726486781)(1.48,1.98280774379692)(1.49,1.98461985331508)(1.5,1.98522596830673)(1.51,1.98461985331508)(1.52,1.98280774379692)(1.53,1.97980726486781)(1.54,1.97564563012208)(1.55,1.97035800872147)(1.56,1.96398607771993)(1.57,1.95657674192344)(1.58,1.94818100511719)(1.59,1.9388529779019)(1.6,1.92864900867555)(1.61,1.91762692548457)(1.62,1.90584537755949)(1.63,1.89336326634995)(1.64,1.88023925679178)(1.65,1.86653136037919)(1.66,1.85229658238609)(1.67,1.83759062628575)(1.68,1.82246764906407)(1.69,1.80698006171278)(1.7,1.79117836972934)(1.71,1.77511104894428)(1.72,1.75882445244819)(1.73,1.74236274480266)(1.74,1.72576786009534)(1.75,1.70907948074268)(1.76,1.69233503425594)(1.77,1.67556970547131)(1.78,1.65881646200356)(1.79,1.64210609091893)(1.8,1.6254672448364)(1.81,1.60892649586181)(1.82,1.59250839593491)(1.83,1.57623554233018)(1.84,1.56012864719645)(1.85,1.54420661015185)(1.86,1.52848659306892)(1.87,1.51298409629154)(1.88,1.49771303562191)(1.89,1.48268581950264)(1.9,1.46791342589672)(1.91,1.45340547843888)(1.92,1.43917032149395)(1.93,1.42521509381473)(1.94,1.41154580054166)(1.95,1.39816738333153)(1.96,1.3850837884424)(1.97,1.37229803263705)(1.98,1.35981226679868)(1.99,1.34762783718023)(2,1.33574534423277) 
};

\addplot [
color=darkgray,
solid,
only marks,
mark=*,
mark size=2.5pt,
forget plot
]
coordinates{
 (0,-2)(0.5,-2)(1,-2)(1.5,-2)(2,-2)
};

\addplot [
color=green!60!black,
densely dotted,
line width=1.0pt,
forget plot
]
coordinates{
 (1.5,-2)(1.5,2)
};

\addplot [
color=green!60!black,
only marks,
mark=*,
mark size=2.5pt,
line width=1.0pt,
forget plot
]
coordinates{
 (1.5,-2)
};

\addplot [
color=green!60!black,
only marks,
mark=x,
mark size=3.5pt,
line width=1.5pt,
forget plot
]
coordinates{
 (1.5,2)
};

\end{axis}
\end{tikzpicture}%