### 2013-03-19 figurefactory/files/voned_2_2.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
label shift={-4pt},
xlabel shift={-6pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=2,
xtick={0, 0.5, 1, 1.5, 2},
ymin=-2, ymax=4,
tickwidth=0cm,
]

\addplot [fill=olive!20!lightgray,draw=none,forget plot] coordinates{ (0,3.47503448812459)(0.01,3.48731049975084)(0.02,3.49954194559996)(0.03,3.51169580854478)(0.04,3.52373658736368)(0.05,3.53562620953933)(0.06,3.54732394772005)(0.07,3.5587863408048)(0.08,3.56996712072292)(0.09,3.58081714610148)(0.1,3.59128434414726)(0.11,3.60131366222062)(0.12,3.61084703074508)(0.13,3.61982333928357)(0.14,3.62817842782351)(0.15,3.63584509555128)(0.16,3.64275312967022)(0.17,3.64882935713003)(0.18,3.6539977225008)(0.19,3.658179395654)(0.2,3.66129291342205)(0.21,3.66325436002188)(0.22,3.66397759177629)(0.23,3.66337451259421)(0.24,3.66135540783692)(0.25,3.65782934568775)(0.26,3.65270465707885)(0.27,3.64588950778609)(0.28,3.63729257973882)(0.29,3.62682388328426)(0.3,3.61439572866668)(0.31,3.59992389419638)(0.32,3.58332904183065)(0.33,3.56453845026926)(0.34,3.54348816452895)(0.35,3.5201257047368)(0.36,3.49441354454039)(0.37,3.46633367614049)(0.38,3.43589375027455)(0.39,3.40313556025084)(0.4,3.36814710795555)(0.41,3.33108028520932)(0.42,3.29217756679695)(0.43,3.25181343758871)(0.44,3.21056010648069)(0.45,3.16929264219669)(0.46,3.12935359666131)(0.47,3.09278728032563)(0.48,3.06257850973023)(0.49,3.04258948800252)(0.5,3.03657032527763)(0.51,3.04616393918492)(0.52,3.06971970397476)(0.53,3.10348382437098)(0.54,3.14359463326134)(0.55,3.18707028212863)(0.56,3.23187043785894)(0.57,3.27665683261068)(0.58,3.3205586920134)(0.59,3.36300798850483)(0.6,3.40363425050598)(0.61,3.44219882010877)(0.62,3.47855340793708)(0.63,3.51261338586195)(0.64,3.5443400923476)(0.65,3.57372875283708)(0.66,3.6007999811883)(0.67,3.62559362391807)(0.68,3.64816417797094)(0.69,3.66857729356851)(0.7,3.68690704505127)(0.71,3.70323375925228)(0.72,3.71764225861017)(0.73,3.73022042001256)(0.74,3.74105797922785)(0.75,3.75024553016918)(0.76,3.7578736814865)(0.77,3.76403234220293)(0.78,3.76881011463808)(0.79,3.77229377756097)(0.8,3.7745678459584)(0.81,3.77571419637071)(0.82,3.77581174869288)(0.83,3.7749361968396)(0.84,3.77315978185159)(0.85,3.77055110196015)(0.86,3.76717495489046)(0.87,3.76309220831331)(0.88,3.75835969488208)(0.89,3.75303012873943)(0.9,3.74715204076432)(0.91,3.74076973016611)(0.92,3.73392323033013)(0.93,3.72664828708403)(0.94,3.71897634779377)(0.95,3.71093455991574)(0.96,3.7025457778319)(0.97,3.69382857698031)(0.98,3.68479727446696)(0.99,3.67546195550829)(1,3.66582850520942)(1.01,3.65589864533175)(1.02,3.64566997584782)(1.03,3.63513602122168)(1.04,3.62428628149047)(1.05,3.61310628836048)(1.06,3.60157766666724)(1.07,3.58967820168808)(1.08,3.57738191293649)(1.09,3.56465913521318)(1.1,3.55147660784019)(1.11,3.53779757316361)(1.12,3.52358188557994)(1.13,3.50878613252339)(1.14,3.49336376905008)(1.15,3.47726526787405)(1.16,3.46043828695531)(1.17,3.44282785701784)(1.18,3.42437659169612)(1.19,3.40502492338305)(1.2,3.38471136829835)(1.21,3.36337282483493)(1.22,3.34094490990433)(1.23,3.3173623388315)(1.24,3.29255935540499)(1.25,3.2664702200547)(1.26,3.23902976592618)(1.27,3.21017403502275)(1.28,3.17984100984943)(1.29,3.14797146049419)(1.3,3.11450993339101)(1.31,3.07940591698944)(1.32,3.04261523254619)(1.33,3.0041017173509)(1.34,2.96383929625123)(1.35,2.9218145807794)(1.36,2.8780302024732)(1.37,2.83250919319779)(1.38,2.78530089617875)(1.39,2.73648917180149)(1.4,2.68620413065109)(1.41,2.63463942142347)(1.42,2.58207846429752)(1.43,2.52893534728687)(1.44,2.47581993642005)(1.45,2.42364234178214)(1.46,2.37377683652712)(1.47,2.32829548695259)(1.48,2.29020653454021)(1.49,2.26339149122763)(1.5,2.25161608648422)(1.51,2.2565334667147)(1.52,2.2764929186673)(1.53,2.30772712055712)(1.54,2.34634874201905)(1.55,2.38934128375939)(1.56,2.43462340262617)(1.57,2.48081122987684)(1.58,2.52698503543485)(1.59,2.57252543016284)(1.6,2.61700895300433)(1.61,2.66014298752552)(1.62,2.70172488214706)(1.63,2.74161572983533)(1.64,2.77972309256712)(1.65,2.8159892812743)(1.66,2.85038316427975)(1.67,2.8828942721013)(1.68,2.91352843478383)(1.69,2.94230446819852)(1.7,2.96925159661914)(1.71,2.9944074050817)(1.72,3.01781618231458)(1.73,3.0395275584581)(1.74,3.05959537034055)(1.75,3.07807670617191)(1.76,3.09503109450287)(1.77,3.110519811274)(1.78,3.12460528508395)(1.79,3.13735058530242)(1.8,3.14881898090992)(1.81,3.15907356034207)(1.82,3.16817690440477)(1.83,3.17619080568347)(1.84,3.18317602891506)(1.85,3.1891921076087)(1.86,3.19429717285143)(1.87,3.19854781075901)(1.88,3.20199894546136)(1.89,3.2047037448685)(1.9,3.20671354676306)(1.91,3.20807780302101)(1.92,3.2088440399828)(1.93,3.20905783318921)(1.94,3.20876279486513)(1.95,3.20800057268415)(1.96,3.20681085848069)(1.97,3.20523140569645)(1.98,3.20329805445655)(1.99,3.20104476326931)(2,3.19850364643326)(2,-0.56626896144308)(1.99,-0.5462160000136)(1.98,-0.525283818332375)(1.97,-0.503439715088015)(1.96,-0.480650441382582)(1.95,-0.456882268460431)(1.94,-0.43210106548919)(1.93,-0.406272388206559)(1.92,-0.379361579291655)(1.91,-0.351333881367478)(1.9,-0.32215456359208)(1.89,-0.291789062851237)(1.88,-0.260203140625971)(1.87,-0.227363056675648)(1.86,-0.193235760754053)(1.85,-0.157789103664555)(1.84,-0.120992069065456)(1.83,-0.0828150275636714)(1.82,-0.0432300147913194)(1.81,-0.00221103535694378)(1.8,0.040265605185011)(1.79,0.0842209378749517)(1.78,0.129672934495549)(1.77,0.176636074845227)(1.76,0.225120872878144)(1.75,0.27513335638814)(1.74,0.326674493449472)(1.73,0.379739556789333)(1.72,0.434317414403876)(1.71,0.490389730666115)(1.7,0.547930056345932)(1.69,0.606902777508557)(1.68,0.667261880845712)(1.67,0.728949474529893)(1.66,0.791893975825184)(1.65,0.856007834013682)(1.64,0.921184590745083)(1.63,0.987294974628876)(1.62,1.05418155701083)(1.61,1.12165121668306)(1.6,1.18946419418054)(1.59,1.25731772274879)(1.58,1.3248208621786)(1.57,1.39145483785665)(1.56,1.45650935952315)(1.55,1.51897981458641)(1.54,1.57740530309109)(1.53,1.62963738594122)(1.52,1.67260476553454)(1.51,1.70238028131943)(1.5,1.71517476059299)(1.49,1.70933693899395)(1.48,1.68654270693122)(1.47,1.65060184815967)(1.46,1.60545792902056)(1.45,1.55419624455546)(1.44,1.49897825318704)(1.43,1.44127760590081)(1.42,1.3821117013744)(1.41,1.32220378072731)(1.4,1.26208579512996)(1.39,1.20216212070966)(1.38,1.14274924211746)(1.37,1.08410093677044)(1.36,1.02642466013697)(1.35,0.969892507044983)(1.34,0.914648762446204)(1.33,0.860815261205179)(1.32,0.808495309244627)(1.31,0.757776639115429)(1.3,0.70873370314298)(1.29,0.661429501990403)(1.28,0.615917080011931)(1.27,0.572240776083215)(1.26,0.530437290727531)(1.25,0.490536611882969)(1.24,0.452562829233078)(1.23,0.416534858558752)(1.22,0.38246709172952)(1.21,0.35036998387673)(1.2,0.320250586414766)(1.19,0.29211303252461)(1.18,0.265958980234497)(1.17,0.241788017153657)(1.16,0.219598030119439)(1.15,0.199385542423449)(1.14,0.181146020831719)(1.13,0.164874154266725)(1.12,0.150564105746791)(1.11,0.138209738960353)(1.1,0.127804820674168)(1.09,0.119343200024961)(1.08,0.112818965615813)(1.07,0.108226581225861)(1.06,0.10556100084091)(1.05,0.104817763620081)(1.04,0.105993069327776)(1.03,0.109083834679302)(1.02,0.114087730971477)(1.01,0.121003203295626)(1,0.129829471559083)(0.99,0.140566513472174)(0.98,0.153215029590611)(0.97,0.167776390437933)(0.96,0.184252565669204)(0.95,0.202646035175346)(0.94,0.222959681967449)(0.93,0.245196666621817)(0.92,0.269360283009454)(0.91,0.295453794977895)(0.9,0.323480253598301)(0.89,0.353442294536177)(0.88,0.385341915048931)(0.87,0.419180230056646)(0.86,0.454957206672098)(0.85,0.492671376509548)(0.84,0.532319525015719)(0.83,0.573896356975411)(0.82,0.617394137231257)(0.81,0.662802305512272)(0.8,0.710107064074811)(0.79,0.759290936602772)(0.78,0.810332296463266)(0.77,0.863204861930176)(0.76,0.917877155314581)(0.75,0.974311921997447)(0.74,1.0324655040292)(0.73,1.09228716107302)(0.72,1.15371832877268)(0.71,1.21669180074849)(0.7,1.28113081480325)(0.69,1.34694801569449)(0.68,1.41404425466938)(0.67,1.48230716777235)(0.66,1.55160944738328)(0.65,1.62180667910266)(0.64,1.69273455001136)(0.63,1.76420512943459)(0.62,1.8360017538758)(0.61,1.90787176901693)(0.6,1.97951591498485)(0.59,2.05057234780867)(0.58,2.12059192909028)(0.57,2.18899909122462)(0.56,2.25502875462593)(0.55,2.31762418490194)(0.54,2.37527572575478)(0.53,2.42579018557806)(0.52,2.46605587636579)(0.51,2.49210938695124)(0.5,2.5001289993864)(0.49,2.48843630260724)(0.48,2.45869035659747)(0.47,2.41469754570973)(0.46,2.36041015773334)(0.45,2.2989311730237)(0.44,2.23244606337767)(0.43,2.16245704556852)(0.42,2.09001339354071)(0.41,2.01587257779527)(0.4,1.94060234913175)(0.39,1.86464378940838)(0.38,1.78835042513833)(0.37,1.71201292093404)(0.36,1.63587504271835)(0.35,1.56014425747618)(0.34,1.48499897607437)(0.33,1.41059365269786)(0.32,1.33706248789253)(0.31,1.26452220350642)(0.3,1.19307418839348)(0.29,1.12280620886868)(0.28,1.05379381182811)(0.27,0.986101506117318)(0.26,0.919783780187773)(0.25,0.854885995903985)(0.24,0.79144518621219)(0.23,0.72949077616544)(0.22,0.669045241187887)(0.21,0.61012471259441)(0.2,0.552739537697143)(0.19,0.496894799954984)(0.18,0.442590803304707)(0.17,0.389823523882892)(0.16,0.338585031689705)(0.15,0.288863884278023)(0.14,0.240645494218025)(0.13,0.193912471848914)(0.12,0.14864494465775)(0.11,0.104820854500891)(0.1,0.0624162337921177)(0.09,0.0214054617129871)(0.08,-0.0182384985515351)(0.07,-0.0565438805909741)(0.06,-0.0935399126342655)(0.05,-0.129256631605251)(0.04,-0.163724712499596)(0.03,-0.196975312239681)(0.02,-0.229039927188963)(0.01,-0.259950263532067)(0,-0.289738119751751)};

\addplot [
color=olive!70!black,
solid,
line width=2.0pt,
forget plot
]
coordinates{
 (0,1.59454509563606)(0.01,1.61564433597658)(0.02,1.63728425737121)(0.03,1.65946424845868)(0.04,1.68218240555269)(0.05,1.70543542995183)(0.06,1.72921852101884)(0.07,1.75352526512571)(0.08,1.77834752060356)(0.09,1.8036752988852)(0.1,1.82949664208298)(0.11,1.85579749730685)(0.12,1.88256158809745)(0.13,1.90977028342819)(0.14,1.93740246481817)(0.15,1.96543439219707)(0.16,1.99383956927345)(0.17,2.02258860928123)(0.18,2.05164910211666)(0.19,2.08098548403023)(0.2,2.11055891120819)(0.21,2.14032713876594)(0.22,2.17024440688404)(0.23,2.20026133604814)(0.24,2.23032483360879)(0.25,2.26037801415861)(0.26,2.29036013653446)(0.27,2.32020656059454)(0.28,2.34984872729673)(0.29,2.37921416601959)(0.3,2.40822653352234)(0.31,2.43680568944179)(0.32,2.46486781377326)(0.33,2.49232557238627)(0.34,2.51908833728684)(0.35,2.5450624690635)(0.36,2.57015166974765)(0.37,2.59425741518815)(0.38,2.61727947699051)(0.39,2.63911654511093)(0.4,2.65966696333104)(0.41,2.67882959108098)(0.42,2.69650480643314)(0.43,2.71259566656815)(0.44,2.72700924362848)(0.45,2.73965815563429)(0.46,2.75046231405488)(0.47,2.7593509117187)(0.48,2.76626467702202)(0.49,2.7711584228752)(0.5,2.77400392152397)(0.51,2.77479307537906)(0.52,2.7735398209958)(0.53,2.7702782278499)(0.54,2.76505933023799)(0.55,2.75794825513593)(0.56,2.74902167585604)(0.57,2.73836556033998)(0.58,2.72607318562214)(0.59,2.71224339247685)(0.6,2.6969790565389)(0.61,2.6803857542764)(0.62,2.66257060411225)(0.63,2.64364126474949)(0.64,2.62370507436833)(0.65,2.60286831584075)(0.66,2.58123559446248)(0.67,2.55890931594234)(0.68,2.53598925352349)(0.69,2.51257219414939)(0.7,2.48875165453657)(0.71,2.46461765888356)(0.72,2.44025657073788)(0.73,2.41575097226585)(0.74,2.39117958483002)(0.75,2.36661722538023)(0.76,2.34213479371238)(0.77,2.31779928614789)(0.78,2.29367383164069)(0.79,2.26981774673115)(0.8,2.24628660614135)(0.81,2.22313232614613)(0.82,2.20040325816301)(0.83,2.1781442902837)(0.84,2.15639695472304)(0.85,2.13519953939011)(0.86,2.1145872019932)(0.87,2.09459208527738)(0.88,2.07524343216166)(0.89,2.05656769969487)(0.9,2.03858867088633)(0.91,2.02132756359039)(0.92,2.00480313573499)(0.93,1.98903178628391)(0.94,1.97402765141244)(0.95,1.95980269545653)(0.96,1.94636679626872)(0.97,1.93372782467942)(0.98,1.92189171782199)(0.99,1.91086254613373)(1,1.90064257389451)(1.01,1.89123231320991)(1.02,1.88263057138807)(1.03,1.8748344916985)(1.04,1.86783958753873)(1.05,1.86163977007047)(1.06,1.85622736942201)(1.07,1.85159314958828)(1.08,1.8477263171952)(1.09,1.84461452433064)(1.1,1.84224386568177)(1.11,1.84059887025782)(1.12,1.83966248801905)(1.13,1.83941607177821)(1.14,1.83983935478899)(1.15,1.84091042449025)(1.16,1.84260569293244)(1.17,1.84489986447787)(1.18,1.84776590143712)(1.19,1.85117498838264)(1.2,1.85509649596757)(1.21,1.85949794517392)(1.22,1.86434497302089)(1.23,1.86960130088174)(1.24,1.87522870668825)(1.25,1.88118700244557)(1.26,1.88743401863969)(1.27,1.89392559729533)(1.28,1.90061559563584)(1.29,1.90745590251041)(1.3,1.91439646998892)(1.31,1.92138536278378)(1.32,1.92836882844262)(1.33,1.93529139156831)(1.34,1.9420959756662)(1.35,1.94872405659499)(1.36,1.95511585201063)(1.37,1.96121055164501)(1.38,1.96694659375663)(1.39,1.97226199363248)(1.4,1.97709473061332)(1.41,1.98138320076256)(1.42,1.98506674300701)(1.43,1.98808624735089)(1.44,1.99038485460839)(1.45,1.99190875802053)(1.46,1.99260811812604)(1.47,1.99243810335014)(1.48,1.99136006996721)(1.49,1.98934289639183)(1.5,1.98636448816631)(1.51,1.98241343801314)(1.52,1.97749002123169)(1.53,1.97160519504836)(1.54,1.96477893335906)(1.55,1.95703871663312)(1.56,1.94841819265865)(1.57,1.93895599181062)(1.58,1.92869468193723)(1.59,1.91767984925921)(1.6,1.90595929287052)(1.61,1.89358232152374)(1.62,1.88059914238955)(1.63,1.86706033240106)(1.64,1.85301638364022)(1.65,1.83851731499822)(1.66,1.82361234305209)(1.67,1.80834960575026)(1.68,1.79277593309509)(1.69,1.77693665955524)(1.7,1.76087547343885)(1.71,1.74463429891398)(1.72,1.72825320677865)(1.73,1.71177035046275)(1.74,1.69522192409063)(1.75,1.67864213974918)(1.76,1.66206322139461)(1.77,1.64551541309303)(1.78,1.62902699952919)(1.79,1.61262433693457)(1.8,1.59633189278366)(1.81,1.58017229278628)(1.82,1.56416637386654)(1.83,1.54833324196633)(1.84,1.53269033364492)(1.85,1.51725348056665)(1.86,1.50203697607831)(1.87,1.48705364317591)(1.88,1.47231490324964)(1.89,1.45783084507571)(1.9,1.44361029359572)(1.91,1.42966087808879)(1.92,1.41598909939962)(1.93,1.40260039593744)(1.94,1.38949920820709)(1.95,1.37668904167482)(1.96,1.36417252780801)(1.97,1.35195148316067)(1.98,1.34002696640539)(1.99,1.3283993332378)(2,1.31706828910212) 
};

\addplot [
color=darkgray,
solid,
only marks,
mark=*,
mark size=2.5pt,
forget plot
]
coordinates{
 (0,-2)(0.5,-2)(1,-2)(1.5,-2)(2,-2)
};

\addplot [
color=green!60!black,
densely dotted,
line width=1.0pt,
forget plot
]
coordinates{
 (0.5,-2)(0.5,2.8)
};

\addplot [
color=green!60!black,
only marks,
mark=*,
mark size=2.5pt,
line width=1.0pt,
forget plot
]
coordinates{
 (0.5,-2)
};

\addplot [
color=green!60!black,
only marks,
mark=x,
mark size=3.5pt,
line width=1.5pt,
forget plot
]
coordinates{
 (0.5,2.8)
};

\end{axis}
\end{tikzpicture}%