### 2013-03-19 figurefactory/files/voned_3_2.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
label shift={-4pt},
xlabel shift={-6pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=2,
xtick={0, 0.5, 1, 1.5, 2},
ymin=-2, ymax=4,
tickwidth=0cm,
]

\addplot [fill=olive!20!lightgray,draw=none,forget plot] coordinates{ (0,3.58482942463052)(0.01,3.60004798891101)(0.02,3.61522857444221)(0.03,3.63033189586352)(0.04,3.64531565363331)(0.05,3.66013442184805)(0.06,3.67473953961915)(0.07,3.68907900710597)(0.08,3.70309738743443)(0.09,3.71673571587552)(0.1,3.72993141781888)(0.11,3.742618237255)(0.12,3.75472617767874)(0.13,3.76618145754966)(0.14,3.77690648269497)(0.15,3.78681983832409)(0.16,3.79583630364506)(0.17,3.80386689244192)(0.18,3.81081892339735)(0.19,3.81659612444073)(0.2,3.82109877598645)(0.21,3.82422389862386)(0.22,3.82586549166252)(0.23,3.82591482996835)(0.24,3.82426082780994)(0.25,3.82079048005803)(0.26,3.81538939316725)(0.27,3.80794242109654)(0.28,3.79833442495176)(0.29,3.78645118004131)(0.3,3.77218046079061)(0.31,3.7554133434235)(0.32,3.73604577980269)(0.33,3.71398051537493)(0.34,3.68912945302201)(0.35,3.66141660795913)(0.36,3.63078186508568)(0.37,3.59718585332289)(0.38,3.56061641480533)(0.39,3.52109740977021)(0.4,3.47870102745279)(0.41,3.43356548123227)(0.42,3.3859211324836)(0.43,3.33612996677294)(0.44,3.28474615977828)(0.45,3.23260879787809)(0.46,3.18097820535255)(0.47,3.13171088766885)(0.48,3.08739598582499)(0.49,3.05119669234731)(0.5,3.0259838611258)(0.51,3.01284772838793)(0.52,3.01029666091057)(0.53,3.01514563996159)(0.54,3.02402797128164)(0.55,3.0342360054454)(0.56,3.04384617507831)(0.57,3.0515590388948)(0.58,3.05651103425065)(0.59,3.05813122154211)(0.6,3.05604507798808)(0.61,3.05001223160369)(0.62,3.0398862161563)(0.63,3.02558807820813)(0.64,3.00708869268196)(0.65,2.9843966280902)(0.66,2.95754962228728)(0.67,2.92660846538683)(0.68,2.89165253089039)(0.69,2.85277646735342)(0.7,2.81008773105625)(0.71,2.76370474622489)(0.72,2.71375554757698)(0.73,2.66037680479363)(0.74,2.6037131586945)(0.75,2.54391681983164)(0.76,2.48114739535772)(0.77,2.41557192163831)(0.78,2.34736508976863)(0.79,2.27670966017863)(0.8,2.20379707201033)(0.81,2.12882826416686)(0.82,2.05201473942949)(0.83,1.97357992300915)(0.84,1.89376089562466)(0.85,1.81281062378027)(0.86,1.73100087447615)(0.87,1.64862610128955)(0.88,1.56600874517876)(0.89,1.48350664197121)(0.9,1.40152362731363)(0.91,1.32052507105654)(0.92,1.24106109190918)(0.93,1.16380175577295)(0.94,1.0895906611355)(0.95,1.01952516621653)(0.96,0.955069311764951)(0.97,0.898186752650931)(0.98,0.851416932475347)(0.99,0.817685171628646)(1,0.799565107889029)(1.01,0.798121861452096)(1.02,0.812289633856214)(1.03,0.8394941968923)(1.04,0.876809815423523)(1.05,0.921696894661266)(1.06,0.972191980008956)(1.07,1.026831670377)(1.08,1.08451977451554)(1.09,1.1444144761036)(1.1,1.2058481943895)(1.11,1.26827408639539)(1.12,1.33123093587662)(1.13,1.39432002549962)(1.14,1.45718968863576)(1.15,1.51952478969417)(1.16,1.58103940072838)(1.17,1.6414715831874)(1.18,1.70057958243273)(1.19,1.7581389911792)(1.2,1.81394059435028)(1.21,1.86778870745259)(1.22,1.91949988503488)(1.23,1.96890191826689)(1.24,2.01583306927621)(1.25,2.06014150971716)(1.26,2.10168494539283)(1.27,2.14033041980383)(1.28,2.17595429881624)(1.29,2.20844244746681)(1.3,2.23769061939598)(1.31,2.26360509077436)(1.32,2.28610358546564)(1.33,2.30511655881966)(1.34,2.32058893735021)(1.35,2.33248245603253)(1.36,2.34077880280757)(1.37,2.34548388554397)(1.38,2.34663370439796)(1.39,2.34430258329642)(1.4,2.33861495813319)(1.41,2.32976265446075)(1.42,2.31803080653477)(1.43,2.30383755357099)(1.44,2.28779567363942)(1.45,2.27080806509891)(1.46,2.25421017454742)(1.47,2.2399573025432)(1.48,2.23078349147602)(1.49,2.23007528043064)(1.5,2.24102962233239)(1.51,2.2651406710595)(1.52,2.30131039476205)(1.53,2.34665072790035)(1.54,2.3979733507103)(1.55,2.45265743944081)(1.56,2.50880945592376)(1.57,2.56512775906107)(1.58,2.62072860112149)(1.59,2.67501062618579)(1.6,2.72756287250157)(1.61,2.7781048370449)(1.62,2.82644754667784)(1.63,2.87246790701773)(1.64,2.91609141311241)(1.65,2.95728018449664)(1.66,2.99602445277282)(1.67,3.03233633720696)(1.68,3.06624517275588)(1.69,3.09779391742564)(1.7,3.12703632874307)(1.71,3.15403470183875)(1.72,3.17885802752753)(1.73,3.20158047176855)(1.74,3.22228010642895)(1.75,3.24103784054218)(1.76,3.25793651447589)(1.77,3.27306012864814)(1.78,3.28649318497019)(1.79,3.2983201239044)(1.8,3.30862484347432)(1.81,3.3174902891288)(1.82,3.32499810530132)(1.83,3.33122834099536)(1.84,3.3362592028899)(1.85,3.34016685038152)(1.86,3.34302522772289)(1.87,3.34490592902509)(1.88,3.34587809239502)(1.89,3.34600831990287)(1.9,3.34536062043468)(1.91,3.34399637279505)(1.92,3.34197430669431)(1.93,3.33935049949039)(1.94,3.33617838676422)(1.95,3.33250878499287)(1.96,3.32838992475032)(1.97,3.32386749301519)(1.98,3.31898468329879)(1.99,3.31378225242948)(2,3.30829858293919)(2,-0.444829374793357)(1.99,-0.421098651071882)(1.98,-0.396446336163294)(1.97,-0.370845345516832)(1.96,-0.344268698795504)(1.95,-0.316689632748587)(1.94,-0.288081726215209)(1.93,-0.25841903808521)(1.92,-0.227676259077674)(1.91,-0.195828878232804)(1.9,-0.162853365049562)(1.89,-0.128727368240436)(1.88,-0.0934299321168721)(1.87,-0.0569417316656629)(1.86,-0.0192453274299609)(1.85,0.0196745586284737)(1.84,0.0598307450330455)(1.83,0.101233269114178)(1.82,0.143889040279859)(1.81,0.187801466606901)(1.8,0.23297005291475)(1.79,0.279389968201148)(1.78,0.327051579935454)(1.77,0.375939952180139)(1.76,0.426034303786354)(1.75,0.477307421903277)(1.74,0.529725024634428)(1.73,0.583245064695554)(1.72,0.637816963125363)(1.71,0.693380758098698)(1.7,0.749866148131158)(1.69,0.807191400594332)(1.68,0.865262084174702)(1.67,0.923969565669119)(1.66,0.983189184093607)(1.65,1.04277797331801)(1.64,1.10257173990012)(1.63,1.16238120155058)(1.62,1.22198673039036)(1.61,1.28113098443977)(1.6,1.33950828377602)(1.59,1.39674888254904)(1.58,1.45239512476716)(1.57,1.50586459578307)(1.56,1.55639257115558)(1.55,1.60294274080495)(1.54,1.64407480352181)(1.53,1.67777397387581)(1.52,1.70131956298155)(1.51,1.71144924430573)(1.5,1.70523183993032)(1.49,1.68156264538277)(1.48,1.64196899135767)(1.47,1.58971783498313)(1.46,1.52829738913243)(1.45,1.46057039095885)(1.44,1.38864376767554)(1.43,1.31402394188051)(1.42,1.23780101275596)(1.41,1.1607884731007)(1.4,1.08361635276817)(1.39,1.00679085020874)(1.38,0.930732327842318)(1.37,0.855799800322401)(1.36,0.782307022733014)(1.35,0.710533304055332)(1.34,0.640730954228341)(1.33,0.573130539780742)(1.32,0.507944680783682)(1.31,0.445370852532745)(1.3,0.385593488936256)(1.29,0.328785580005791)(1.28,0.275109888906981)(1.27,0.224719870300608)(1.26,0.177760342439145)(1.25,0.134367945276811)(1.24,0.094671402268187)(1.23,0.0587915923031793)(1.22,0.0268414287544383)(1.21,-0.00107446635636632)(1.2,-0.0248603146781139)(1.19,-0.0444299934333956)(1.18,-0.0597081549437549)(1.17,-0.0706317224008323)(1.16,-0.0771518600220896)(1.15,-0.079236559279762)(1.14,-0.0768740486083972)(1.13,-0.0700773350515127)(1.12,-0.0588903442762789)(1.11,-0.0433963757642554)(1.1,-0.0237299907984804)(1.09,-9.40942984856141e-05)(1.08,0.0272150070885241)(1.07,0.0577702328946769)(1.06,0.0909576019586149)(1.05,0.125893055316186)(1.04,0.161306130482944)(1.03,0.195403125446262)(1.02,0.225786465151679)(1.01,0.249639542185894)(1,0.264458529378529)(0.99,0.26920285236244)(0.98,0.264913763770812)(0.97,0.254095681204893)(0.96,0.239565626824373)(0.95,0.223721326871451)(0.94,0.208356283085153)(0.93,0.194740318290634)(0.92,0.183756324482167)(0.91,0.176016500654449)(0.9,0.171945442125653)(0.89,0.171836179811568)(0.88,0.175887465025862)(0.87,0.184228740738409)(0.86,0.196937137231983)(0.85,0.214049274806336)(0.84,0.23556963487419)(0.83,0.261476617420922)(0.82,0.291727002053005)(0.81,0.326259279554266)(0.8,0.364996162981932)(0.79,0.407846486369675)(0.78,0.454706633488184)(0.77,0.505461595674604)(0.76,0.55998572834969)(0.75,0.618143255391288)(0.74,0.679788555740815)(0.73,0.744766255290415)(0.72,0.812911137667727)(0.71,0.884047878763875)(0.7,0.957990600596524)(0.69,1.0345422291118)(0.68,1.11349362620844)(0.67,1.19462244634792)(0.66,1.27769163916542)(0.65,1.36244747611301)(0.64,1.4486169126074)(0.63,1.53590399298656)(0.62,1.62398483960066)(0.61,1.71250049851602)(0.6,1.80104647262306)(0.59,1.88915704018206)(0.58,1.97628124047184)(0.57,2.06174542720433)(0.56,2.14469426911443)(0.55,2.22399833130533)(0.54,2.29811518586665)(0.53,2.36490617240151)(0.52,2.42148216079223)(0.51,2.46433509334006)(0.5,2.49018607872373)(0.49,2.49750526559355)(0.48,2.48740515404448)(0.47,2.46283413364432)(0.46,2.42707965816406)(0.45,2.38289409924225)(0.44,2.3323292750101)(0.43,2.27686680349494)(0.42,2.21758765612926)(0.41,2.15530373759552)(0.4,2.09064643872724)(0.39,2.02412355716508)(0.38,1.95615559851786)(0.37,1.88709914785574)(0.36,1.81726219187339)(0.35,1.7469143967805)(0.34,1.6762941843428)(0.33,1.60561374383708)(0.32,1.53506269122152)(0.31,1.46481082659219)(0.3,1.3950102801787)(0.29,1.32579723630126)(0.28,1.2572933605496)(0.27,1.18960701402354)(0.26,1.12283431137273)(0.25,1.05706006141912)(0.24,0.992358617120402)(0.23,0.928794653500352)(0.22,0.866423886627792)(0.21,0.805293742920604)(0.2,0.745443985426882)(0.19,0.686907301918829)(0.18,0.629709858375887)(0.17,0.573871820560741)(0.16,0.519407845788207)(0.15,0.466327546571051)(0.14,0.414635927542119)(0.13,0.364333796858902)(0.12,0.315418153166848)(0.11,0.267882549111694)(0.1,0.221717432334636)(0.09,0.176910464847661)(0.08,0.133446821662446)(0.07,0.091309469530374)(0.06,0.0504794266397142)(0.05,0.010936004106592)(0.04,-0.0273429699125187)(0.03,-0.0643809426684996)(0.02,-0.100202445019883)(0.01,-0.13483291459035)(0,-0.168298533102028)};

\addplot [
color=olive!70!black,
solid,
line width=2.0pt,
forget plot
]
coordinates{
  (0,1.70826544576424)(0.01,1.73260753716033)(0.02,1.75751306471116)(0.03,1.78297547659751)(0.04,1.8089863418604)(0.05,1.83553521297732)(0.06,1.86260948312943)(0.07,1.89019423831817)(0.08,1.91827210454844)(0.09,1.94682309036159)(0.1,1.97582442507676)(0.11,2.00525039318335)(0.12,2.0350721654228)(0.13,2.06525762720428)(0.14,2.09577120511855)(0.15,2.12657369244757)(0.16,2.15762207471663)(0.17,2.18886935650133)(0.18,2.22026439088662)(0.19,2.25175171317978)(0.2,2.28327138070667)(0.21,2.31475882077223)(0.22,2.34614468914516)(0.23,2.37735474173435)(0.24,2.40830972246517)(0.25,2.43892527073858)(0.26,2.46911185226999)(0.27,2.49877471756004)(0.28,2.52781389275068)(0.29,2.55612420817129)(0.3,2.58359537048465)(0.31,2.61011208500784)(0.32,2.6355542355121)(0.33,2.659797129606)(0.34,2.6827118186824)(0.35,2.70416550236982)(0.36,2.72402202847953)(0.37,2.74214250058932)(0.38,2.75838600666159)(0.39,2.77261048346765)(0.4,2.78467373309002)(0.41,2.7944346094139)(0.42,2.80175439430643)(0.43,2.80649838513394)(0.44,2.80853771739419)(0.45,2.80775144856017)(0.46,2.8040289317583)(0.47,2.79727251065658)(0.48,2.78740056993473)(0.49,2.77435097897043)(0.5,2.75808496992476)(0.51,2.738591410864)(0.52,2.7158894108514)(0.53,2.69002590618155)(0.54,2.66107157857415)(0.55,2.62911716837536)(0.56,2.59427022209637)(0.57,2.55665223304956)(0.58,2.51639613736124)(0.59,2.47364413086208)(0.6,2.42854577530557)(0.61,2.38125636505986)(0.62,2.33193552787848)(0.63,2.28074603559734)(0.64,2.22785280264468)(0.65,2.1734220521016)(0.66,2.11762063072635)(0.67,2.06061545586737)(0.68,2.00257307854941)(0.69,1.94365934823261)(0.7,1.88403916582639)(0.71,1.82387631249438)(0.72,1.76333334262236)(0.73,1.70257153004202)(0.74,1.64175085721766)(0.75,1.58103003761146)(0.76,1.5205665618537)(0.77,1.46051675865646)(0.78,1.40103586162841)(0.79,1.34227807327415)(0.8,1.28439661749613)(0.81,1.22754377186056)(0.82,1.17187087074125)(0.83,1.11752827021504)(0.84,1.06466526524942)(0.85,1.01342994929331)(0.86,0.963969005854064)(0.87,0.916427421013977)(0.88,0.87094810510231)(0.89,0.82767141089139)(0.9,0.786734534719644)(0.91,0.748270785855492)(0.92,0.712408708195676)(0.93,0.679271037031793)(0.94,0.648973472110324)(0.95,0.621623246543991)(0.96,0.597317469294662)(0.97,0.576141216927912)(0.98,0.55816534812308)(0.99,0.543444011995543)(1,0.532011818633779)(1.01,0.523880701818995)(1.02,0.519038049503947)(1.03,0.517448661169281)(1.04,0.519057972953233)(1.05,0.523794974988726)(1.06,0.531574790983785)(1.07,0.542300951635837)(1.08,0.555867390802034)(1.09,0.572160190902558)(1.1,0.591059101795511)(1.11,0.612438855315566)(1.12,0.63617029580017)(1.13,0.662121345224055)(1.14,0.690157820013684)(1.15,0.720144115207205)(1.16,0.751943770353143)(1.17,0.785419930393282)(1.18,0.820435713744487)(1.19,0.856854498872902)(1.2,0.894540139836085)(1.21,0.933357120548111)(1.22,0.973170656894659)(1.23,1.01384675528503)(1.24,1.0552522357722)(1.25,1.09725472749698)(1.26,1.13972264391599)(1.27,1.18252514505222)(1.28,1.22553209386161)(1.29,1.2686140137363)(1.3,1.31164205416612)(1.31,1.35448797165355)(1.32,1.39702413312466)(1.33,1.4391235493002)(1.34,1.48065994578927)(1.35,1.52150788004393)(1.36,1.56154291277029)(1.37,1.60064184293319)(1.38,1.63868301612014)(1.39,1.67554671675258)(1.4,1.71111565545068)(1.41,1.74527556378072)(1.42,1.77791590964536)(1.43,1.80893074772575)(1.44,1.83821972065748)(1.45,1.86568922802888)(1.46,1.89125378183992)(1.47,1.91483756876317)(1.48,1.93637624141684)(1.49,1.9558189629067)(1.5,1.97313073113136)(1.51,1.98829495768261)(1.52,2.0013149788718)(1.53,2.01221235088808)(1.54,2.02102407711605)(1.55,2.02780009012288)(1.56,2.03260101353967)(1.57,2.03549617742207)(1.58,2.03656186294433)(1.59,2.03587975436742)(1.6,2.0335355781388)(1.61,2.02961791074233)(1.62,2.0242171385341)(1.63,2.01742455428415)(1.64,2.00933157650626)(1.65,2.00002907890732)(1.66,1.98960681843321)(1.67,1.97815295143804)(1.68,1.96575362846529)(1.69,1.95249265900999)(1.7,1.93845123843711)(1.71,1.92370772996873)(1.72,1.90833749532644)(1.73,1.89241276823205)(1.74,1.87600256553169)(1.75,1.85917263122273)(1.76,1.84198540913112)(1.77,1.82450004041414)(1.78,1.80677238245282)(1.79,1.78885504605277)(1.8,1.77079744819454)(1.81,1.75264587786785)(1.82,1.73444357279059)(1.83,1.71623080505477)(1.84,1.69804497396147)(1.85,1.67992070450499)(1.86,1.66188995014647)(1.87,1.64398209867971)(1.88,1.62622408013907)(1.89,1.60864047583122)(1.9,1.59125362769256)(1.91,1.57408374728112)(1.92,1.55714902380832)(1.93,1.54046573070259)(1.94,1.52404833027451)(1.95,1.50790957612214)(1.96,1.49206061297741)(1.97,1.47651107374918)(1.98,1.46126917356775)(1.99,1.4463418006788)(2,1.43173460407292) 
};

\addplot [
color=darkgray,
solid,
only marks,
mark=*,
mark size=2.5pt,
forget plot
]
coordinates{
 (0,-2)(0.5,-2)(1,-2)(1.5,-2)(2,-2)
};

\addplot [
color=green!60!black,
densely dotted,
line width=1.0pt,
forget plot
]
coordinates{
 (1,-2)(1,0.5)
};

\addplot [
color=green!60!black,
only marks,
mark=*,
mark size=2.5pt,
line width=1.0pt,
forget plot
]
coordinates{
 (1,-2)
};

\addplot [
color=green!60!black,
only marks,
mark=x,
mark size=3.5pt,
line width=1.5pt,
forget plot
]
coordinates{
 (1,0.5)
};

\end{axis}
\end{tikzpicture}%