### 2013-03-19 figurefactory/files/limno_bgape_class_0.tex
LSE algorithm

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,1.0100953990001e-06)(358.30303030303,4.04031425463559e-06)(373.232323232323,7.07043210985092e-06)(388.161616161616,9.09045456864996e-06)(403.090909090909,9.09045456864996e-06)(418.020202020202,9.09045456864996e-06)(432.949494949495,9.09045456864996e-06)(447.878787878788,9.09045456864996e-06)(462.808080808081,9.09045456864996e-06)(477.737373737374,9.09045456864996e-06)(492.666666666667,9.09045456864996e-06)(507.59595959596,9.09045456864996e-06)(522.525252525253,9.09045456864996e-06)(537.454545454546,9.09045456864996e-06)(552.383838383838,9.09045456864996e-06)(567.313131313131,9.09045456864996e-06)(582.242424242424,9.09045456864996e-06)(597.171717171717,9.09045456864996e-06)(612.10101010101,9.09045456864996e-06)(627.030303030303,9.09045456864996e-06)(641.959595959596,9.09045456864996e-06)(656.888888888889,9.09045456864996e-06)(671.818181818182,9.09045456864996e-06)(686.747474747475,9.09045456864996e-06)(701.676767676768,9.09045456864996e-06)(716.606060606061,9.09045456864996e-06)(731.535353535354,9.09045456864996e-06)(746.464646464646,9.09045456864996e-06)(761.393939393939,9.09045456864996e-06)(776.323232323232,9.09045456864996e-06)(791.252525252525,9.09045456864996e-06)(806.181818181818,9.09045456864996e-06)(821.111111111111,9.09045456864996e-06)(836.040404040404,9.09045456864996e-06)(850.969696969697,9.09045456864996e-06)(865.89898989899,9.09045456864996e-06)(880.828282828283,9.09045456864996e-06)(895.757575757576,9.09045456864996e-06)(910.686868686869,9.09045456864996e-06)(925.616161616162,9.09045456864996e-06)(940.545454545455,9.09045456864996e-06)(955.474747474747,9.09045456864996e-06)(970.40404040404,9.09045456864996e-06)(985.333333333333,9.09045456864996e-06)(1000.26262626263,9.09045456864996e-06)(1015.19191919192,9.09045456864996e-06)(1030.12121212121,9.09045456864996e-06)(1045.05050505051,9.09045456864996e-06)(1059.9797979798,9.09045456864996e-06)(1074.90909090909,9.09045456864996e-06)(1089.83838383838,9.09045456864996e-06)(1104.76767676768,9.09045456864996e-06)(1119.69696969697,9.09045456864996e-06)(1134.62626262626,9.09045456864996e-06)(1149.55555555556,9.09045456864996e-06)(1164.48484848485,9.09045456864996e-06)(1179.41414141414,9.09045456864996e-06)(1194.34343434343,9.09045456864996e-06)(1209.27272727273,9.09045456864996e-06)(1224.20202020202,9.09045456864996e-06)(1239.13131313131,9.09045456864996e-06)(1254.06060606061,9.09045456864996e-06)(1268.9898989899,9.09045456864996e-06)(1283.91919191919,9.09045456864996e-06)(1298.84848484848,9.09045456864996e-06)(1313.77777777778,9.09045456864996e-06)(1328.70707070707,9.09045456864996e-06)(1343.63636363636,9.09045456864996e-06)(1358.56565656566,9.09045456864996e-06)(1373.49494949495,9.09045456864996e-06)(1388.42424242424,9.09045456864996e-06)(1403.35353535354,9.09045456864996e-06)(1418.28282828283,9.09045456864996e-06)(1433.21212121212,9.09045456864996e-06)(1448.14141414141,9.09045456864996e-06)(1463.07070707071,9.09045456864996e-06)(1478,9.09045456864996e-06)(1478.00074642733,0)(1478.00074642733,-0.181818181818183)(1478.00074642733,-0.363636363636363)(1478.00074642733,-0.545454545454547)(1478.00074642733,-0.727272727272727)(1478.00074642733,-0.90909090909091)(1478.00074642733,-1.09090909090909)(1478.00074642733,-1.27272727272727)(1478.00074642733,-1.45454545454545)(1478.00074642733,-1.63636363636364)(1478.00074642733,-1.81818181818182)(1478.00049762651,-2)(1478.0002488174,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478.0002488174,-6.54545454545455)(1478.00049762651,-6.72727272727273)(1478.00074642733,-6.90909090909091)(1478.00074642733,-7.09090909090909)(1478.00074642733,-7.27272727272727)(1478.00074642733,-7.45454545454546)(1478.00074642733,-7.63636363636364)(1478.00074642733,-7.81818181818182)(1478.00074642733,-8)(1478.00074642733,-8.18181818181818)(1478.00074642733,-8.36363636363636)(1478.00074642733,-8.54545454545455)(1478.00074642733,-8.72727272727273)(1478.00074642733,-8.90909090909091)(1478.00074642733,-9.09090909090909)(1478.00074642733,-9.27272727272727)(1478.00074642733,-9.45454545454546)(1478.00074642733,-9.63636363636364)(1478.00074642733,-9.81818181818182)(1478.00074642733,-10)(1478.00074642733,-10.1818181818182)(1478.00074642733,-10.3636363636364)(1478.00074642733,-10.5454545454545)(1478.00074642733,-10.7272727272727)(1478.00074642733,-10.9090909090909)(1478.00074642733,-11.0909090909091)(1478.00074642733,-11.2727272727273)(1478.00074642733,-11.4545454545455)(1478.00074642733,-11.6363636363636)(1478.00074642733,-11.8181818181818)(1478.00074642733,-12)(1478.00049762651,-12.1818181818182)(1478.0002488174,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478.0002488174,-17.2727272727273)(1478.00049762651,-17.4545454545455)(1478.00074642733,-17.6363636363636)(1478.00074642733,-17.8181818181818)(1478.00074642733,-18)(1478,-18.0000090904546)(1463.07070707071,-18.0000090904546)(1448.14141414141,-18.0000090904546)(1433.21212121212,-18.0000090904546)(1418.28282828283,-18.0000090904546)(1403.35353535354,-18.0000090904546)(1388.42424242424,-18.0000090904546)(1373.49494949495,-18.0000090904546)(1358.56565656566,-18.0000090904546)(1343.63636363636,-18.0000090904546)(1328.70707070707,-18.0000090904546)(1313.77777777778,-18.0000090904546)(1298.84848484848,-18.0000090904546)(1283.91919191919,-18.0000090904546)(1268.9898989899,-18.0000090904546)(1254.06060606061,-18.0000090904546)(1239.13131313131,-18.0000090904546)(1224.20202020202,-18.0000090904546)(1209.27272727273,-18.0000090904546)(1194.34343434343,-18.0000090904546)(1179.41414141414,-18.0000090904546)(1164.48484848485,-18.0000090904546)(1149.55555555556,-18.0000090904546)(1134.62626262626,-18.0000090904546)(1119.69696969697,-18.0000090904546)(1104.76767676768,-18.0000090904546)(1089.83838383838,-18.0000090904546)(1074.90909090909,-18.0000090904546)(1059.9797979798,-18.0000090904546)(1045.05050505051,-18.0000090904546)(1030.12121212121,-18.0000090904546)(1015.19191919192,-18.0000090904546)(1000.26262626263,-18.0000090904546)(985.333333333333,-18.0000090904546)(970.40404040404,-18.0000090904546)(955.474747474747,-18.0000090904546)(940.545454545455,-18.0000090904546)(925.616161616162,-18.0000090904546)(910.686868686869,-18.0000090904546)(895.757575757576,-18.0000090904546)(880.828282828283,-18.0000090904546)(865.89898989899,-18.0000090904546)(850.969696969697,-18.0000090904546)(836.040404040404,-18.0000090904546)(821.111111111111,-18.0000090904546)(806.181818181818,-18.0000090904546)(791.252525252525,-18.0000090904546)(776.323232323232,-18.0000090904546)(761.393939393939,-18.0000090904546)(746.464646464646,-18.0000090904546)(731.535353535354,-18.0000090904546)(716.606060606061,-18.0000090904546)(701.676767676768,-18.0000090904546)(686.747474747475,-18.0000090904546)(671.818181818182,-18.0000090904546)(656.888888888889,-18.0000090904546)(641.959595959596,-18.0000090904546)(627.030303030303,-18.0000090904546)(612.10101010101,-18.0000090904546)(597.171717171717,-18.0000090904546)(582.242424242424,-18.0000090904546)(567.313131313131,-18.0000090904546)(552.383838383838,-18.0000090904546)(537.454545454546,-18.0000090904546)(522.525252525253,-18.0000090904546)(507.59595959596,-18.0000090904546)(492.666666666667,-18.0000090904546)(477.737373737374,-18.0000090904546)(462.808080808081,-18.0000090904546)(447.878787878788,-18.0000090904546)(432.949494949495,-18.0000090904546)(418.020202020202,-18.000006060404)(403.090909090909,-18.0000030302525)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(-0.000248817401863321,-15.0909090909091)(-0.000497626510092015,-14.9090909090909)(-0.000746427325097138,-14.7272727272727)(-0.000746427325097138,-14.5454545454545)(-0.000746427325097138,-14.3636363636364)(-0.000746427325097138,-14.1818181818182)(-0.000746427325097138,-14)(-0.000746427325097138,-13.8181818181818)(-0.000746427325097138,-13.6363636363636)(-0.000746427325097138,-13.4545454545455)(-0.000746427325097138,-13.2727272727273)(-0.000746427325097138,-13.0909090909091)(-0.000746427325097138,-12.9090909090909)(-0.000746427325097138,-12.7272727272727)(-0.000746427325097138,-12.5454545454545)(-0.000746427325097138,-12.3636363636364)(-0.000746427325097138,-12.1818181818182)(-0.000746427325097138,-12)(-0.000746427325097138,-11.8181818181818)(-0.000746427325097138,-11.6363636363636)(-0.000746427325097138,-11.4545454545455)(-0.000746427325097138,-11.2727272727273)(-0.000746427325097138,-11.0909090909091)(-0.000746427325097138,-10.9090909090909)(-0.000746427325097138,-10.7272727272727)(-0.000746427325097138,-10.5454545454545)(-0.000746427325097138,-10.3636363636364)(-0.000746427325097138,-10.1818181818182)(-0.000746427325097138,-10)(-0.000746427325097138,-9.81818181818182)(-0.000497626510092015,-9.63636363636364)(-0.000248817401863321,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(-0.000248817401863321,-8)(-0.000497626510092015,-7.81818181818182)(-0.000746427325097138,-7.63636363636364)(-0.000746427325097138,-7.45454545454546)(-0.000746427325097138,-7.27272727272727)(-0.000497626510092015,-7.09090909090909)(-0.000248817401863321,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(-0.000248817401863321,-4.36363636363636)(-0.000497626510092015,-4.18181818181818)(-0.000746427325097138,-4)(-0.000746427325097138,-3.81818181818182)(-0.000746427325097138,-3.63636363636364)(-0.000746427325097138,-3.45454545454546)(-0.000746427325097138,-3.27272727272727)(-0.000746427325097138,-3.09090909090909)(-0.000746427325097138,-2.90909090909091)(-0.000746427325097138,-2.72727272727273)(-0.000497626510092015,-2.54545454545455)(-0.000248817401863321,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

%\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-2.45454545454545)(14.9292929292929,-2.45454545454545)(29.8585858585859,-2.45454545454545)(44.7878787878788,-2.45454545454545)(59.7171717171717,-2.45454545454545)(74.6464646464647,-2.45454545454545)(89.5757575757576,-2.39393939393939)(97.040404040404,-2.36363636363636)(104.505050505051,-2.33333333333333)(119.434343434343,-2.27272727272727)(134.363636363636,-2.27272727272727)(149.292929292929,-2.21212121212121)(156.757575757576,-2.18181818181818)(164.222222222222,-2.15151515151515)(179.151515151515,-2.09090909090909)(194.080808080808,-2.03030303030303)(201.545454545455,-2)(209.010101010101,-1.96969696969697)(223.939393939394,-1.84848484848485)(231.40404040404,-1.81818181818182)(238.868686868687,-1.78787878787879)(253.79797979798,-1.66666666666667)(257.530303030303,-1.63636363636364)(268.727272727273,-1.54545454545455)(279.924242424242,-1.45454545454545)(283.656565656566,-1.42424242424242)(298.585858585859,-1.3030303030303)(302.318181818182,-1.27272727272727)(313.515151515152,-1.18181818181818)(320.979797979798,-1.09090909090909)(328.444444444444,-0.954545454545455)(330.93265993266,-0.90909090909091)(340.885521885522,-0.727272727272727)(343.373737373737,-0.636363636363637)(345.861952861953,-0.545454545454547)(350.838383838384,-0.363636363636363)(355.814814814815,-0.181818181818183)(358.30303030303,-0.0909090909090917)(360.791245791246,0)(373.232323232323,2.52515432531587e-06)(388.161616161616,4.54522728432498e-06)(403.090909090909,4.54522728432498e-06)(418.020202020202,4.54522728432498e-06)(432.949494949495,4.54522728432498e-06)(447.878787878788,4.54522728432498e-06)(462.808080808081,4.54522728432498e-06)(477.737373737374,4.54522728432498e-06)(492.666666666667,4.54522728432498e-06)(507.59595959596,4.54522728432498e-06)(522.525252525253,4.54522728432498e-06)(537.454545454546,4.54522728432498e-06)(552.383838383838,4.54522728432498e-06)(567.313131313131,4.54522728432498e-06)(582.242424242424,4.54522728432498e-06)(597.171717171717,4.54522728432498e-06)(612.10101010101,4.54522728432498e-06)(627.030303030303,4.54522728432498e-06)(641.959595959596,4.54522728432498e-06)(656.888888888889,4.54522728432498e-06)(671.818181818182,4.54522728432498e-06)(686.747474747475,4.54522728432498e-06)(701.676767676768,4.54522728432498e-06)(716.606060606061,4.54522728432498e-06)(731.535353535354,4.54522728432498e-06)(746.464646464646,4.54522728432498e-06)(761.393939393939,4.54522728432498e-06)(776.323232323232,4.54522728432498e-06)(791.252525252525,4.54522728432498e-06)(806.181818181818,4.54522728432498e-06)(821.111111111111,4.54522728432498e-06)(836.040404040404,4.54522728432498e-06)(850.969696969697,4.54522728432498e-06)(865.89898989899,4.54522728432498e-06)(880.828282828283,4.54522728432498e-06)(895.757575757576,4.54522728432498e-06)(910.686868686869,4.54522728432498e-06)(925.616161616162,4.54522728432498e-06)(940.545454545455,4.54522728432498e-06)(955.474747474747,4.54522728432498e-06)(970.40404040404,4.54522728432498e-06)(985.333333333333,4.54522728432498e-06)(1000.26262626263,4.54522728432498e-06)(1015.19191919192,4.54522728432498e-06)(1030.12121212121,4.54522728432498e-06)(1045.05050505051,4.54522728432498e-06)(1059.9797979798,4.54522728432498e-06)(1074.90909090909,4.54522728432498e-06)(1089.83838383838,4.54522728432498e-06)(1104.76767676768,4.54522728432498e-06)(1119.69696969697,4.54522728432498e-06)(1134.62626262626,4.54522728432498e-06)(1149.55555555556,4.54522728432498e-06)(1164.48484848485,4.54522728432498e-06)(1179.41414141414,4.54522728432498e-06)(1194.34343434343,4.54522728432498e-06)(1209.27272727273,4.54522728432498e-06)(1224.20202020202,4.54522728432498e-06)(1239.13131313131,4.54522728432498e-06)(1254.06060606061,4.54522728432498e-06)(1268.9898989899,4.54522728432498e-06)(1283.91919191919,4.54522728432498e-06)(1298.84848484848,4.54522728432498e-06)(1313.77777777778,4.54522728432498e-06)(1328.70707070707,4.54522728432498e-06)(1343.63636363636,4.54522728432498e-06)(1358.56565656566,4.54522728432498e-06)(1373.49494949495,4.54522728432498e-06)(1388.42424242424,4.54522728432498e-06)(1403.35353535354,4.54522728432498e-06)(1418.28282828283,4.54522728432498e-06)(1433.21212121212,4.54522728432498e-06)(1448.14141414141,4.54522728432498e-06)(1463.07070707071,4.54522728432498e-06)(1478,4.54522728432498e-06)(1478.00037321366,0)(1478.00037321366,-0.181818181818183)(1478.00037321366,-0.363636363636363)(1478.00037321366,-0.545454545454547)(1478.00037321366,-0.727272727272727)(1478.00037321366,-0.90909090909091)(1478.00037321366,-1.09090909090909)(1478.00037321366,-1.27272727272727)(1478.00037321366,-1.45454545454545)(1478.00037321366,-1.63636363636364)(1478.00037321366,-1.81818181818182)(1478.00012440663,-2)(1478,-2.09090909090909)(1463.07070707071,-2.09090909090909)(1448.14141414141,-2.09090909090909)(1433.21212121212,-2.03030303030303)(1425.74747474747,-2)(1418.28282828283,-1.96969696969697)(1403.35353535354,-1.90909090909091)(1388.42424242424,-1.90909090909091)(1373.49494949495,-1.90909090909091)(1358.56565656566,-1.90909090909091)(1343.63636363636,-1.90909090909091)(1328.70707070707,-1.90909090909091)(1313.77777777778,-1.90909090909091)(1298.84848484848,-1.96969696969697)(1291.38383838384,-2)(1283.91919191919,-2.03030303030303)(1268.9898989899,-2.09090909090909)(1254.06060606061,-2.09090909090909)(1239.13131313131,-2.09090909090909)(1224.20202020202,-2.09090909090909)(1209.27272727273,-2.15151515151515)(1201.80808080808,-2.18181818181818)(1194.34343434343,-2.21212121212121)(1179.41414141414,-2.27272727272727)(1164.48484848485,-2.33333333333333)(1157.0202020202,-2.36363636363636)(1149.55555555556,-2.39393939393939)(1134.62626262626,-2.51515151515151)(1127.16161616162,-2.54545454545455)(1119.69696969697,-2.57575757575758)(1104.76767676768,-2.6969696969697)(1097.30303030303,-2.72727272727273)(1089.83838383838,-2.75757575757576)(1074.90909090909,-2.87878787878788)(1067.44444444444,-2.90909090909091)(1059.9797979798,-2.93939393939394)(1045.05050505051,-3.06060606060606)(1041.31818181818,-3.09090909090909)(1030.12121212121,-3.18181818181818)(1022.65656565657,-3.27272727272727)(1015.19191919192,-3.36363636363636)(1007.72727272727,-3.45454545454546)(1000.26262626263,-3.54545454545455)(992.79797979798,-3.63636363636364)(985.333333333333,-3.72727272727273)(977.868686868687,-3.81818181818182)(970.40404040404,-3.90909090909091)(962.939393939394,-4)(955.474747474747,-4.09090909090909)(948.010101010101,-4.18181818181818)(940.545454545455,-4.27272727272727)(933.080808080808,-4.36363636363636)(925.616161616162,-4.45454545454545)(914.419191919192,-4.54545454545455)(910.686868686869,-4.57575757575758)(895.757575757576,-4.6969696969697)(888.292929292929,-4.72727272727273)(880.828282828283,-4.75757575757576)(865.89898989899,-4.87878787878788)(858.434343434343,-4.90909090909091)(850.969696969697,-4.93939393939394)(836.040404040404,-5)(821.111111111111,-5)(806.181818181818,-5.06060606060606)(798.717171717172,-5.09090909090909)(791.252525252525,-5.12121212121212)(776.323232323232,-5.18181818181818)(761.393939393939,-5.18181818181818)(746.464646464646,-5.24242424242424)(739,-5.27272727272727)(731.535353535354,-5.3030303030303)(716.606060606061,-5.36363636363636)(701.676767676768,-5.36363636363636)(686.747474747475,-5.42424242424242)(679.282828282828,-5.45454545454546)(671.818181818182,-5.48484848484849)(656.888888888889,-5.54545454545455)(641.959595959596,-5.54545454545455)(627.030303030303,-5.54545454545455)(612.10101010101,-5.54545454545455)(597.171717171717,-5.54545454545455)(582.242424242424,-5.54545454545455)(567.313131313131,-5.54545454545455)(552.383838383838,-5.54545454545455)(537.454545454546,-5.54545454545455)(522.525252525253,-5.54545454545455)(507.59595959596,-5.54545454545455)(492.666666666667,-5.54545454545455)(477.737373737374,-5.54545454545455)(462.808080808081,-5.54545454545455)(447.878787878788,-5.54545454545455)(432.949494949495,-5.54545454545455)(418.020202020202,-5.54545454545455)(403.090909090909,-5.54545454545455)(388.161616161616,-5.54545454545455)(373.232323232323,-5.54545454545455)(358.30303030303,-5.60606060606061)(350.838383838384,-5.63636363636364)(343.373737373737,-5.66666666666667)(328.444444444444,-5.72727272727273)(313.515151515152,-5.78787878787879)(309.782828282828,-5.81818181818182)(298.585858585859,-5.95454545454546)(296.097643097643,-6)(291.121212121212,-6.18181818181818)(296.097643097643,-6.36363636363636)(298.585858585859,-6.40909090909091)(306.050505050505,-6.54545454545455)(313.515151515152,-6.63636363636364)(320.979797979798,-6.72727272727273)(328.444444444444,-6.81818181818182)(339.641414141414,-6.90909090909091)(343.373737373737,-6.93939393939394)(358.30303030303,-7.06060606060606)(365.767676767677,-7.09090909090909)(373.232323232323,-7.12121212121212)(388.161616161616,-7.18181818181818)(403.090909090909,-7.18181818181818)(418.020202020202,-7.24242424242424)(432.949494949495,-7.24242424242424)(447.878787878788,-7.24242424242424)(462.808080808081,-7.18181818181818)(477.737373737374,-7.18181818181818)(492.666666666667,-7.12121212121212)(500.131313131313,-7.09090909090909)(507.59595959596,-7.06060606060606)(522.525252525253,-7)(537.454545454546,-6.93939393939394)(544.919191919192,-6.90909090909091)(552.383838383838,-6.87878787878788)(567.313131313131,-6.81818181818182)(582.242424242424,-6.81818181818182)(597.171717171717,-6.81818181818182)(612.10101010101,-6.81818181818182)(627.030303030303,-6.81818181818182)(641.959595959596,-6.81818181818182)(656.888888888889,-6.81818181818182)(671.818181818182,-6.81818181818182)(686.747474747475,-6.87878787878788)(694.212121212121,-6.90909090909091)(701.676767676768,-6.93939393939394)(716.606060606061,-7)(731.535353535354,-7.06060606060606)(739,-7.09090909090909)(746.464646464646,-7.12121212121212)(761.393939393939,-7.18181818181818)(776.323232323232,-7.24242424242424)(783.787878787879,-7.27272727272727)(791.252525252525,-7.3030303030303)(806.181818181818,-7.36363636363636)(821.111111111111,-7.36363636363636)(836.040404040404,-7.36363636363636)(850.969696969697,-7.36363636363636)(865.89898989899,-7.36363636363636)(880.828282828283,-7.36363636363636)(895.757575757576,-7.3030303030303)(903.222222222222,-7.27272727272727)(910.686868686869,-7.24242424242424)(925.616161616162,-7.18181818181818)(940.545454545455,-7.18181818181818)(955.474747474747,-7.18181818181818)(970.40404040404,-7.18181818181818)(985.333333333333,-7.12121212121212)(992.79797979798,-7.09090909090909)(1000.26262626263,-7.06060606060606)(1015.19191919192,-7)(1030.12121212121,-7)(1045.05050505051,-7)(1059.9797979798,-7)(1074.90909090909,-7)(1089.83838383838,-7)(1104.76767676768,-7)(1119.69696969697,-7)(1134.62626262626,-7)(1149.55555555556,-7)(1164.48484848485,-7)(1179.41414141414,-7)(1194.34343434343,-7)(1209.27272727273,-7)(1224.20202020202,-7)(1239.13131313131,-7)(1254.06060606061,-7)(1268.9898989899,-7.06060606060606)(1276.45454545455,-7.09090909090909)(1283.91919191919,-7.12121212121212)(1298.84848484848,-7.18181818181818)(1313.77777777778,-7.18181818181818)(1328.70707070707,-7.18181818181818)(1343.63636363636,-7.18181818181818)(1358.56565656566,-7.12121212121212)(1366.0303030303,-7.09090909090909)(1373.49494949495,-7.06060606060606)(1388.42424242424,-6.93939393939394)(1395.88888888889,-6.90909090909091)(1403.35353535354,-6.87878787878788)(1418.28282828283,-6.75757575757576)(1425.74747474747,-6.72727272727273)(1433.21212121212,-6.6969696969697)(1448.14141414141,-6.63636363636364)(1463.07070707071,-6.63636363636364)(1478,-6.63636363636364)(1478.00012440663,-6.72727272727273)(1478.00037321366,-6.90909090909091)(1478.00037321366,-7.09090909090909)(1478.00037321366,-7.27272727272727)(1478.00037321366,-7.45454545454546)(1478.00037321366,-7.63636363636364)(1478.00037321366,-7.81818181818182)(1478.00037321366,-8)(1478.00037321366,-8.18181818181818)(1478.00037321366,-8.36363636363636)(1478.00037321366,-8.54545454545455)(1478.00037321366,-8.72727272727273)(1478.00037321366,-8.90909090909091)(1478.00037321366,-9.09090909090909)(1478.00037321366,-9.27272727272727)(1478.00037321366,-9.45454545454546)(1478.00037321366,-9.63636363636364)(1478.00037321366,-9.81818181818182)(1478.00037321366,-10)(1478.00037321366,-10.1818181818182)(1478.00037321366,-10.3636363636364)(1478.00037321366,-10.5454545454545)(1478.00037321366,-10.7272727272727)(1478.00037321366,-10.9090909090909)(1478.00037321366,-11.0909090909091)(1478.00037321366,-11.2727272727273)(1478.00037321366,-11.4545454545455)(1478.00037321366,-11.6363636363636)(1478.00037321366,-11.8181818181818)(1478.00037321366,-12)(1478.00012440663,-12.1818181818182)(1478,-12.2727272727273)(1463.07070707071,-12.2727272727273)(1448.14141414141,-12.2727272727273)(1433.21212121212,-12.2727272727273)(1418.28282828283,-12.2727272727273)(1403.35353535354,-12.2727272727273)(1388.42424242424,-12.2727272727273)(1373.49494949495,-12.2121212121212)(1366.0303030303,-12.1818181818182)(1358.56565656566,-12.1515151515152)(1343.63636363636,-12.0909090909091)(1328.70707070707,-12.030303030303)(1324.97474747475,-12)(1313.77777777778,-11.9090909090909)(1306.31313131313,-11.8181818181818)(1298.84848484848,-11.6818181818182)(1296.36026936027,-11.6363636363636)(1286.40740740741,-11.4545454545455)(1283.91919191919,-11.3636363636364)(1281.43097643098,-11.2727272727273)(1268.9898989899,-11.1212121212121)(1265.25757575758,-11.0909090909091)(1254.06060606061,-11)(1239.13131313131,-11)(1224.20202020202,-11)(1209.27272727273,-11)(1194.34343434343,-11.0606060606061)(1186.87878787879,-11.0909090909091)(1179.41414141414,-11.1212121212121)(1164.48484848485,-11.2424242424242)(1157.0202020202,-11.2727272727273)(1149.55555555556,-11.3030303030303)(1134.62626262626,-11.4242424242424)(1127.16161616162,-11.4545454545455)(1119.69696969697,-11.4848484848485)(1104.76767676768,-11.5454545454545)(1089.83838383838,-11.4848484848485)(1082.37373737374,-11.4545454545455)(1074.90909090909,-11.4242424242424)(1059.9797979798,-11.3030303030303)(1052.51515151515,-11.2727272727273)(1045.05050505051,-11.2424242424242)(1030.12121212121,-11.1212121212121)(1026.38888888889,-11.0909090909091)(1015.19191919192,-11)(1007.72727272727,-10.9090909090909)(1000.26262626263,-10.8181818181818)(992.79797979798,-10.7272727272727)(985.333333333333,-10.6363636363636)(974.136363636364,-10.5454545454545)(970.40404040404,-10.5151515151515)(955.474747474747,-10.3939393939394)(948.010101010101,-10.3636363636364)(940.545454545455,-10.3333333333333)(925.616161616162,-10.2727272727273)(910.686868686869,-10.2727272727273)(895.757575757576,-10.2727272727273)(880.828282828283,-10.2727272727273)(865.89898989899,-10.2121212121212)(858.434343434343,-10.1818181818182)(850.969696969697,-10.1515151515152)(836.040404040404,-10.0909090909091)(821.111111111111,-10.030303030303)(813.646464646465,-10)(806.181818181818,-9.96969696969697)(791.252525252525,-9.84848484848485)(783.787878787879,-9.81818181818182)(776.323232323232,-9.78787878787879)(761.393939393939,-9.66666666666667)(753.929292929293,-9.63636363636364)(746.464646464646,-9.60606060606061)(731.535353535354,-9.48484848484848)(724.070707070707,-9.45454545454546)(716.606060606061,-9.42424242424243)(701.676767676768,-9.36363636363636)(686.747474747475,-9.36363636363636)(671.818181818182,-9.36363636363636)(656.888888888889,-9.36363636363636)(641.959595959596,-9.36363636363636)(627.030303030303,-9.36363636363636)(612.10101010101,-9.36363636363636)(597.171717171717,-9.36363636363636)(582.242424242424,-9.36363636363636)(567.313131313131,-9.36363636363636)(552.383838383838,-9.42424242424243)(544.919191919192,-9.45454545454546)(537.454545454546,-9.48484848484848)(522.525252525253,-9.60606060606061)(515.060606060606,-9.63636363636364)(507.59595959596,-9.66666666666667)(492.666666666667,-9.78787878787879)(485.20202020202,-9.81818181818182)(477.737373737374,-9.84848484848485)(462.808080808081,-9.96969696969697)(455.343434343434,-10)(447.878787878788,-10.030303030303)(432.949494949495,-10.0909090909091)(418.020202020202,-10.0909090909091)(403.090909090909,-10.0909090909091)(388.161616161616,-10.1515151515152)(380.69696969697,-10.1818181818182)(373.232323232323,-10.2121212121212)(358.30303030303,-10.2727272727273)(343.373737373737,-10.2727272727273)(328.444444444444,-10.2727272727273)(313.515151515152,-10.2727272727273)(298.585858585859,-10.2727272727273)(283.656565656566,-10.2727272727273)(268.727272727273,-10.2727272727273)(253.79797979798,-10.2727272727273)(238.868686868687,-10.3333333333333)(231.40404040404,-10.3636363636364)(223.939393939394,-10.3939393939394)(209.010101010101,-10.4545454545455)(194.080808080808,-10.4545454545455)(179.151515151515,-10.5151515151515)(171.686868686869,-10.5454545454545)(164.222222222222,-10.5757575757576)(149.292929292929,-10.6969696969697)(145.560606060606,-10.7272727272727)(134.363636363636,-10.8181818181818)(126.89898989899,-10.9090909090909)(119.434343434343,-11)(111.969696969697,-11.0909090909091)(104.505050505051,-11.1818181818182)(97.040404040404,-11.2727272727273)(89.5757575757576,-11.4090909090909)(87.0875420875421,-11.4545454545455)(82.1111111111111,-11.6363636363636)(82.1111111111111,-11.8181818181818)(82.1111111111111,-12)(82.1111111111111,-12.1818181818182)(82.1111111111111,-12.3636363636364)(87.0875420875421,-12.5454545454545)(89.5757575757576,-12.6363636363636)(92.0639730639731,-12.7272727272727)(102.016835016835,-12.9090909090909)(104.505050505051,-13)(106.993265993266,-13.0909090909091)(116.946127946128,-13.2727272727273)(119.434343434343,-13.3181818181818)(126.89898989899,-13.4545454545455)(134.363636363636,-13.5454545454545)(145.560606060606,-13.6363636363636)(149.292929292929,-13.6666666666667)(164.222222222222,-13.7878787878788)(171.686868686869,-13.8181818181818)(179.151515151515,-13.8484848484849)(194.080808080808,-13.969696969697)(201.545454545455,-14)(209.010101010101,-14.030303030303)(223.939393939394,-14.0909090909091)(238.868686868687,-14.0909090909091)(253.79797979798,-14.0909090909091)(268.727272727273,-14.0909090909091)(283.656565656566,-14.0909090909091)(298.585858585859,-14.0909090909091)(313.515151515152,-14.0909090909091)(328.444444444444,-14.0909090909091)(343.373737373737,-14.0909090909091)(358.30303030303,-14.0909090909091)(373.232323232323,-14.0909090909091)(388.161616161616,-14.0909090909091)(403.090909090909,-14.030303030303)(410.555555555556,-14)(418.020202020202,-13.969696969697)(432.949494949495,-13.9090909090909)(447.878787878788,-13.9090909090909)(462.808080808081,-13.8484848484849)(470.272727272727,-13.8181818181818)(477.737373737374,-13.7878787878788)(492.666666666667,-13.7272727272727)(507.59595959596,-13.7272727272727)(522.525252525253,-13.6666666666667)(529.989898989899,-13.6363636363636)(537.454545454546,-13.6060606060606)(552.383838383838,-13.5454545454545)(567.313131313131,-13.5454545454545)(582.242424242424,-13.4848484848485)(589.707070707071,-13.4545454545455)(597.171717171717,-13.4242424242424)(612.10101010101,-13.3636363636364)(627.030303030303,-13.3030303030303)(634.494949494949,-13.2727272727273)(641.959595959596,-13.2424242424242)(656.888888888889,-13.1818181818182)(671.818181818182,-13.1212121212121)(679.282828282828,-13.0909090909091)(686.747474747475,-13.0606060606061)(701.676767676768,-13)(716.606060606061,-12.9393939393939)(724.070707070707,-12.9090909090909)(731.535353535354,-12.8787878787879)(746.464646464646,-12.8181818181818)(761.393939393939,-12.7575757575758)(768.858585858586,-12.7272727272727)(776.323232323232,-12.6969696969697)(791.252525252525,-12.6363636363636)(806.181818181818,-12.6363636363636)(821.111111111111,-12.5757575757576)(828.575757575758,-12.5454545454545)(836.040404040404,-12.5151515151515)(850.969696969697,-12.4545454545455)(865.89898989899,-12.4545454545455)(880.828282828283,-12.4545454545455)(895.757575757576,-12.4545454545455)(910.686868686869,-12.3939393939394)(918.151515151515,-12.3636363636364)(925.616161616162,-12.3333333333333)(940.545454545455,-12.2727272727273)(955.474747474747,-12.2727272727273)(970.40404040404,-12.2727272727273)(985.333333333333,-12.2727272727273)(1000.26262626263,-12.2121212121212)(1007.72727272727,-12.1818181818182)(1015.19191919192,-12.1515151515152)(1030.12121212121,-12.0909090909091)(1045.05050505051,-12.0909090909091)(1059.9797979798,-12.030303030303)(1067.44444444444,-12)(1074.90909090909,-11.969696969697)(1089.83838383838,-11.9090909090909)(1104.76767676768,-11.9090909090909)(1119.69696969697,-11.969696969697)(1123.42929292929,-12)(1134.62626262626,-12.1363636363636)(1137.11447811448,-12.1818181818182)(1142.09090909091,-12.3636363636364)(1142.09090909091,-12.5454545454545)(1137.11447811448,-12.7272727272727)(1134.62626262626,-12.8181818181818)(1132.13804713805,-12.9090909090909)(1122.18518518519,-13.0909090909091)(1119.69696969697,-13.1818181818182)(1117.20875420875,-13.2727272727273)(1107.25589225589,-13.4545454545455)(1104.76767676768,-13.5454545454545)(1102.27946127946,-13.6363636363636)(1097.30303030303,-13.8181818181818)(1097.30303030303,-14)(1092.3265993266,-14.1818181818182)(1089.83838383838,-14.2727272727273)(1087.35016835017,-14.3636363636364)(1087.35016835017,-14.5454545454545)(1089.83838383838,-14.6363636363636)(1092.3265993266,-14.7272727272727)(1097.30303030303,-14.9090909090909)(1097.30303030303,-15.0909090909091)(1102.27946127946,-15.2727272727273)(1104.76767676768,-15.3181818181818)(1112.23232323232,-15.4545454545455)(1119.69696969697,-15.5909090909091)(1122.18518518519,-15.6363636363636)(1132.13804713805,-15.8181818181818)(1134.62626262626,-15.8636363636364)(1142.09090909091,-16)(1149.55555555556,-16.0909090909091)(1160.75252525253,-16.1818181818182)(1164.48484848485,-16.2121212121212)(1179.41414141414,-16.3333333333333)(1183.14646464646,-16.3636363636364)(1194.34343434343,-16.4545454545455)(1205.5404040404,-16.5454545454545)(1209.27272727273,-16.5757575757576)(1224.20202020202,-16.6969696969697)(1231.66666666667,-16.7272727272727)(1239.13131313131,-16.7575757575758)(1254.06060606061,-16.8181818181818)(1268.9898989899,-16.8787878787879)(1276.45454545455,-16.9090909090909)(1283.91919191919,-16.9393939393939)(1298.84848484848,-17)(1313.77777777778,-17.0606060606061)(1321.24242424242,-17.0909090909091)(1328.70707070707,-17.1212121212121)(1343.63636363636,-17.1818181818182)(1358.56565656566,-17.1818181818182)(1373.49494949495,-17.1818181818182)(1388.42424242424,-17.1818181818182)(1403.35353535354,-17.1818181818182)(1418.28282828283,-17.2424242424242)(1425.74747474747,-17.2727272727273)(1433.21212121212,-17.3030303030303)(1448.14141414141,-17.3636363636364)(1463.07070707071,-17.3636363636364)(1478,-17.3636363636364)(1478.00012440663,-17.4545454545455)(1478.00037321366,-17.6363636363636)(1478.00037321366,-17.8181818181818)(1478.00037321366,-18)(1478,-18.0000045452273)(1463.07070707071,-18.0000045452273)(1448.14141414141,-18.0000045452273)(1433.21212121212,-18.0000045452273)(1418.28282828283,-18.0000045452273)(1403.35353535354,-18.0000045452273)(1388.42424242424,-18.0000045452273)(1373.49494949495,-18.0000045452273)(1358.56565656566,-18.0000045452273)(1343.63636363636,-18.0000045452273)(1328.70707070707,-18.0000045452273)(1313.77777777778,-18.0000045452273)(1298.84848484848,-18.0000045452273)(1283.91919191919,-18.0000045452273)(1268.9898989899,-18.0000045452273)(1254.06060606061,-18.0000045452273)(1239.13131313131,-18.0000045452273)(1224.20202020202,-18.0000045452273)(1209.27272727273,-18.0000045452273)(1194.34343434343,-18.0000045452273)(1179.41414141414,-18.0000045452273)(1164.48484848485,-18.0000045452273)(1149.55555555556,-18.0000045452273)(1134.62626262626,-18.0000045452273)(1119.69696969697,-18.0000045452273)(1104.76767676768,-18.0000045452273)(1089.83838383838,-18.0000045452273)(1074.90909090909,-18.0000045452273)(1059.9797979798,-18.0000045452273)(1045.05050505051,-18.0000045452273)(1030.12121212121,-18.0000045452273)(1015.19191919192,-18.0000045452273)(1000.26262626263,-18.0000045452273)(985.333333333333,-18.0000045452273)(970.40404040404,-18.0000045452273)(955.474747474747,-18.0000045452273)(940.545454545455,-18.0000045452273)(925.616161616162,-18.0000045452273)(910.686868686869,-18.0000045452273)(895.757575757576,-18.0000045452273)(880.828282828283,-18.0000045452273)(865.89898989899,-18.0000045452273)(850.969696969697,-18.0000045452273)(836.040404040404,-18.0000045452273)(821.111111111111,-18.0000045452273)(806.181818181818,-18.0000045452273)(791.252525252525,-18.0000045452273)(776.323232323232,-18.0000045452273)(761.393939393939,-18.0000045452273)(746.464646464646,-18.0000045452273)(731.535353535354,-18.0000045452273)(716.606060606061,-18.0000045452273)(701.676767676768,-18.0000045452273)(686.747474747475,-18.0000045452273)(671.818181818182,-18.0000045452273)(656.888888888889,-18.0000045452273)(641.959595959596,-18.0000045452273)(627.030303030303,-18.0000045452273)(612.10101010101,-18.0000045452273)(597.171717171717,-18.0000045452273)(582.242424242424,-18.0000045452273)(567.313131313131,-18.0000045452273)(552.383838383838,-18.0000045452273)(537.454545454546,-18.0000045452273)(522.525252525253,-18.0000045452273)(507.59595959596,-18.0000045452273)(492.666666666667,-18.0000045452273)(477.737373737374,-18.0000045452273)(462.808080808081,-18.0000045452273)(447.878787878788,-18.0000045452273)(432.949494949495,-18.0000045452273)(418.020202020202,-18.000001515101)(410.555555555556,-18)(410.555555555556,-17.8181818181818)(410.555555555556,-17.6363636363636)(405.579124579125,-17.4545454545455)(403.090909090909,-17.3636363636364)(400.602693602694,-17.2727272727273)(390.649831649832,-17.0909090909091)(388.161616161616,-17)(385.673400673401,-16.9090909090909)(375.720538720539,-16.7272727272727)(373.232323232323,-16.6818181818182)(365.767676767677,-16.5454545454545)(358.30303030303,-16.4545454545455)(350.838383838384,-16.3636363636364)(343.373737373737,-16.2727272727273)(335.909090909091,-16.1818181818182)(328.444444444444,-16.0909090909091)(320.979797979798,-16)(313.515151515152,-15.9090909090909)(306.050505050505,-15.8181818181818)(298.585858585859,-15.7272727272727)(291.121212121212,-15.6363636363636)(283.656565656566,-15.5454545454545)(272.459595959596,-15.4545454545455)(268.727272727273,-15.4242424242424)(253.79797979798,-15.3030303030303)(246.333333333333,-15.2727272727273)(238.868686868687,-15.2424242424242)(223.939393939394,-15.1212121212121)(216.474747474747,-15.0909090909091)(209.010101010101,-15.0606060606061)(194.080808080808,-15)(179.151515151515,-15)(164.222222222222,-15)(149.292929292929,-15)(134.363636363636,-15)(119.434343434343,-15)(104.505050505051,-15)(89.5757575757576,-15)(74.6464646464647,-15)(59.7171717171717,-15)(44.7878787878788,-15)(29.8585858585859,-15)(14.9292929292929,-15)(0,-15)(-0.000124406627521346,-14.9090909090909)(-0.000373213662548569,-14.7272727272727)(-0.000373213662548569,-14.5454545454545)(-0.000373213662548569,-14.3636363636364)(-0.000373213662548569,-14.1818181818182)(-0.000373213662548569,-14)(-0.000373213662548569,-13.8181818181818)(-0.000373213662548569,-13.6363636363636)(-0.000373213662548569,-13.4545454545455)(-0.000373213662548569,-13.2727272727273)(-0.000373213662548569,-13.0909090909091)(-0.000373213662548569,-12.9090909090909)(-0.000373213662548569,-12.7272727272727)(-0.000373213662548569,-12.5454545454545)(-0.000373213662548569,-12.3636363636364)(-0.000373213662548569,-12.1818181818182)(-0.000373213662548569,-12)(-0.000373213662548569,-11.8181818181818)(-0.000373213662548569,-11.6363636363636)(-0.000373213662548569,-11.4545454545455)(-0.000373213662548569,-11.2727272727273)(-0.000373213662548569,-11.0909090909091)(-0.000373213662548569,-10.9090909090909)(-0.000373213662548569,-10.7272727272727)(-0.000373213662548569,-10.5454545454545)(-0.000373213662548569,-10.3636363636364)(-0.000373213662548569,-10.1818181818182)(-0.000373213662548569,-10)(-0.000373213662548569,-9.81818181818182)(-0.000124406627521346,-9.63636363636364)(0,-9.54545454545455)(14.9292929292929,-9.54545454545455)(29.8585858585859,-9.54545454545455)(44.7878787878788,-9.48484848484848)(48.520202020202,-9.45454545454546)(59.7171717171717,-9.36363636363636)(67.1818181818182,-9.27272727272727)(74.6464646464647,-9.13636363636364)(77.1346801346801,-9.09090909090909)(87.0875420875421,-8.90909090909091)(87.0875420875421,-8.72727272727273)(87.0875420875421,-8.54545454545455)(74.6464646464647,-8.39393939393939)(72.1582491582492,-8.36363636363636)(59.7171717171717,-8.21212121212121)(52.2525252525253,-8.18181818181818)(44.7878787878788,-8.15151515151515)(29.8585858585859,-8.03030303030303)(22.3939393939394,-8)(14.9292929292929,-7.96969696969697)(0,-7.90909090909091)(-0.000124406627521346,-7.81818181818182)(-0.000373213662548569,-7.63636363636364)(-0.000373213662548569,-7.45454545454546)(-0.000373213662548569,-7.27272727272727)(-0.000124406627521346,-7.09090909090909)(0,-7)(14.9292929292929,-6.93939393939394)(22.3939393939394,-6.90909090909091)(29.8585858585859,-6.87878787878788)(44.7878787878788,-6.81818181818182)(59.7171717171717,-6.75757575757576)(67.1818181818182,-6.72727272727273)(74.6464646464647,-6.6969696969697)(89.5757575757576,-6.63636363636364)(104.505050505051,-6.57575757575758)(111.969696969697,-6.54545454545455)(119.434343434343,-6.51515151515151)(134.363636363636,-6.45454545454545)(149.292929292929,-6.39393939393939)(156.757575757576,-6.36363636363636)(164.222222222222,-6.33333333333333)(179.151515151515,-6.21212121212121)(186.616161616162,-6.18181818181818)(194.080808080808,-6.15151515151515)(206.521885521886,-6)(209.010101010101,-5.95454545454546)(216.474747474747,-5.81818181818182)(216.474747474747,-5.63636363636364)(211.498316498317,-5.45454545454546)(209.010101010101,-5.40909090909091)(201.545454545455,-5.27272727272727)(194.080808080808,-5.18181818181818)(186.616161616162,-5.09090909090909)(179.151515151515,-5)(167.954545454545,-4.90909090909091)(164.222222222222,-4.87878787878788)(149.292929292929,-4.75757575757576)(141.828282828283,-4.72727272727273)(134.363636363636,-4.6969696969697)(119.434343434343,-4.57575757575758)(111.969696969697,-4.54545454545455)(104.505050505051,-4.51515151515151)(89.5757575757576,-4.45454545454545)(74.6464646464647,-4.39393939393939)(67.1818181818182,-4.36363636363636)(59.7171717171717,-4.33333333333333)(44.7878787878788,-4.27272727272727)(29.8585858585859,-4.27272727272727)(14.9292929292929,-4.27272727272727)(0,-4.27272727272727)(-0.000124406627521346,-4.18181818181818)(-0.000373213662548569,-4)(-0.000373213662548569,-3.81818181818182)(-0.000373213662548569,-3.63636363636364)(-0.000373213662548569,-3.45454545454546)(-0.000373213662548569,-3.27272727272727)(-0.000373213662548569,-3.09090909090909)(-0.000373213662548569,-2.90909090909091)(-0.000373213662548569,-2.72727272727273)(-0.000124406627521346,-2.54545454545455)(0,-2.45454545454545)};

%\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1298.84848484848,-8.68181818181818)(1301.3367003367,-8.72727272727273)(1306.31313131313,-8.90909090909091)(1301.3367003367,-9.09090909090909)(1298.84848484848,-9.13636363636364)(1291.38383838384,-9.27272727272727)(1283.91919191919,-9.36363636363636)(1272.72222222222,-9.45454545454546)(1268.9898989899,-9.48484848484848)(1254.06060606061,-9.54545454545455)(1239.13131313131,-9.60606060606061)(1231.66666666667,-9.63636363636364)(1224.20202020202,-9.66666666666667)(1209.27272727273,-9.72727272727273)(1194.34343434343,-9.72727272727273)(1179.41414141414,-9.72727272727273)(1164.48484848485,-9.72727272727273)(1149.55555555556,-9.72727272727273)(1134.62626262626,-9.66666666666667)(1127.16161616162,-9.63636363636364)(1119.69696969697,-9.60606060606061)(1104.76767676768,-9.54545454545455)(1089.83838383838,-9.54545454545455)(1074.90909090909,-9.48484848484848)(1067.44444444444,-9.45454545454546)(1059.9797979798,-9.42424242424243)(1045.05050505051,-9.3030303030303)(1037.58585858586,-9.27272727272727)(1030.12121212121,-9.24242424242424)(1017.68013468013,-9.09090909090909)(1015.19191919192,-9.06060606060606)(1002.75084175084,-8.90909090909091)(1002.75084175084,-8.72727272727273)(1011.4595959596,-8.54545454545455)(1015.19191919192,-8.51515151515152)(1030.12121212121,-8.39393939393939)(1037.58585858586,-8.36363636363636)(1045.05050505051,-8.33333333333333)(1059.9797979798,-8.27272727272727)(1074.90909090909,-8.21212121212121)(1082.37373737374,-8.18181818181818)(1089.83838383838,-8.15151515151515)(1104.76767676768,-8.09090909090909)(1119.69696969697,-8.09090909090909)(1134.62626262626,-8.09090909090909)(1149.55555555556,-8.09090909090909)(1164.48484848485,-8.09090909090909)(1179.41414141414,-8.09090909090909)(1194.34343434343,-8.09090909090909)(1209.27272727273,-8.15151515151515)(1216.73737373737,-8.18181818181818)(1224.20202020202,-8.21212121212121)(1239.13131313131,-8.27272727272727)(1254.06060606061,-8.33333333333333)(1261.52525252525,-8.36363636363636)(1268.9898989899,-8.39393939393939)(1283.91919191919,-8.51515151515152)(1287.65151515152,-8.54545454545455)(1298.84848484848,-8.68181818181818)};

\end{axis}
\end{tikzpicture}%