### 2013-03-19 figurefactory/files/limno_bgape_class_160.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478.0002488174,-8.72727272727273)(1478.00049762651,-8.90909090909091)(1478.00074642733,-9.09090909090909)(1478.00066349464,-9.27272727272727)(1478.00058056104,-9.45454545454546)(1478.00058056104,-9.63636363636364)(1478.00066349464,-9.81818181818182)(1478.00049762651,-10)(1478.0002488174,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1478,-8.81818181818182)(1478.00012440663,-8.90909090909091)(1478.00037321366,-9.09090909090909)(1478.00029027891,-9.27272727272727)(1478.00020734323,-9.45454545454546)(1478.00020734323,-9.63636363636364)(1478.00029027891,-9.81818181818182)(1478.00012440663,-10)(1478,-10.0909090909091)(1463.07070707071,-10.0909090909091)(1448.14141414141,-10.1590909090909)(1440.67676767677,-10.1818181818182)(1433.21212121212,-10.2121212121212)(1418.28282828283,-10.2272727272727)(1403.35353535354,-10.2272727272727)(1388.42424242424,-10.2727272727273)(1373.49494949495,-10.3454545454545)(1358.56565656566,-10.3454545454545)(1343.63636363636,-10.3409090909091)(1328.70707070707,-10.2727272727273)(1313.77777777778,-10.2727272727273)(1298.84848484848,-10.2727272727273)(1283.91919191919,-10.2954545454545)(1268.9898989899,-10.2727272727273)(1254.06060606061,-10.2954545454545)(1239.13131313131,-10.2727272727273)(1224.20202020202,-10.2954545454545)(1209.27272727273,-10.2727272727273)(1194.34343434343,-10.2727272727273)(1179.41414141414,-10.2727272727273)(1164.48484848485,-10.2121212121212)(1149.55555555556,-10.2121212121212)(1134.62626262626,-10.2121212121212)(1119.69696969697,-10.2121212121212)(1104.76767676768,-10.25)(1089.83838383838,-10.2727272727273)(1074.90909090909,-10.3484848484848)(1067.44444444444,-10.3636363636364)(1059.9797979798,-10.3818181818182)(1045.05050505051,-10.4090909090909)(1033.85353535354,-10.3636363636364)(1030.12121212121,-10.3484848484848)(1015.19191919192,-10.2727272727273)(1000.26262626263,-10.2045454545455)(985.333333333333,-10.2121212121212)(970.40404040404,-10.2727272727273)(955.474747474748,-10.25)(944.277777777778,-10.1818181818182)(940.545454545455,-10.1666666666667)(925.616161616162,-10.0909090909091)(910.686868686869,-10.0227272727273)(895.757575757576,-10.030303030303)(888.292929292929,-10)(880.828282828283,-9.98181818181818)(865.89898989899,-9.94545454545455)(852.835858585859,-9.81818181818182)(850.969696969697,-9.8)(836.040404040404,-9.78787878787879)(826.709595959596,-9.63636363636364)(821.111111111111,-9.54545454545455)(806.181818181818,-9.5)(798.717171717172,-9.45454545454546)(791.252525252525,-9.42424242424243)(776.323232323232,-9.31818181818182)(773.835016835017,-9.27272727272727)(772.590909090909,-9.09090909090909)(765.126262626263,-8.90909090909091)(761.393939393939,-8.81818181818182)(757.661616161616,-8.72727272727273)(758.905723905724,-8.54545454545455)(761.393939393939,-8.45454545454545)(765.126262626263,-8.36363636363636)(776.323232323232,-8.22727272727273)(778.811447811448,-8.18181818181818)(791.252525252525,-8.06818181818182)(798.717171717172,-8)(806.181818181818,-7.93181818181818)(821.111111111111,-7.83636363636364)(822.60404040404,-7.81818181818182)(833.552188552189,-7.63636363636364)(836.040404040404,-7.61363636363636)(850.969696969697,-7.52272727272727)(858.434343434343,-7.45454545454546)(865.89898989899,-7.4)(880.828282828283,-7.36363636363636)(893.26936026936,-7.27272727272727)(895.757575757576,-7.25)(910.686868686869,-7.24242424242424)(925.616161616162,-7.15909090909091)(940.545454545455,-7.15909090909091)(955.474747474747,-7.10606060606061)(970.40404040404,-7.13636363636364)(985.333333333333,-7.10606060606061)(1000.26262626263,-7.10606060606061)(1003.99494949495,-7.09090909090909)(1015.19191919192,-7.03636363636364)(1030.12121212121,-7.06818181818182)(1037.58585858586,-7.09090909090909)(1045.05050505051,-7.10606060606061)(1059.9797979798,-7.13636363636364)(1074.90909090909,-7.20454545454546)(1089.83838383838,-7.25)(1092.3265993266,-7.27272727272727)(1104.76767676768,-7.42424242424243)(1112.23232323232,-7.45454545454546)(1119.69696969697,-7.46969696969697)(1134.62626262626,-7.56060606060606)(1149.55555555556,-7.58181818181818)(1164.48484848485,-7.61363636363636)(1171.94949494949,-7.63636363636364)(1179.41414141414,-7.65151515151515)(1194.34343434343,-7.68181818181818)(1209.27272727273,-7.68181818181818)(1224.20202020202,-7.65151515151515)(1231.66666666667,-7.63636363636364)(1239.13131313131,-7.61818181818182)(1254.06060606061,-7.61818181818182)(1261.52525252525,-7.63636363636364)(1268.9898989899,-7.66666666666667)(1283.91919191919,-7.72727272727273)(1298.84848484848,-7.70454545454546)(1313.77777777778,-7.72727272727273)(1328.70707070707,-7.72727272727273)(1339.90404040404,-7.81818181818182)(1343.63636363636,-7.84848484848485)(1358.56565656566,-7.84848484848485)(1371.00673400673,-8)(1373.49494949495,-8.09090909090909)(1380.9595959596,-8.18181818181818)(1388.42424242424,-8.27272727272727)(1395.88888888889,-8.36363636363636)(1403.35353535354,-8.39393939393939)(1418.28282828283,-8.43181818181818)(1433.21212121212,-8.47727272727273)(1440.67676767677,-8.54545454545455)(1448.14141414141,-8.61363636363636)(1460.58249158249,-8.72727272727273)(1463.07070707071,-8.77272727272727)(1478,-8.81818181818182)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1433.21212121212,-9.43181818181818)(1434.70505050505,-9.45454545454546)(1435.07828282828,-9.63636363636364)(1433.21212121212,-9.72727272727273)(1418.28282828283,-9.72727272727273)(1403.35353535354,-9.70454545454546)(1388.42424242424,-9.78787878787879)(1385.93602693603,-9.81818181818182)(1375.36111111111,-10)(1373.49494949495,-10.0181818181818)(1358.56565656566,-10.0454545454545)(1343.63636363636,-10.0151515151515)(1339.90404040404,-10)(1328.70707070707,-9.90909090909091)(1313.77777777778,-9.90909090909091)(1298.84848484848,-9.90909090909091)(1283.91919191919,-9.96969696969697)(1268.9898989899,-9.90909090909091)(1254.06060606061,-9.96969696969697)(1239.13131313131,-9.90909090909091)(1224.20202020202,-9.96969696969697)(1209.27272727273,-9.90909090909091)(1194.34343434343,-9.90909090909091)(1179.41414141414,-9.90909090909091)(1164.48484848485,-9.88636363636364)(1149.55555555556,-9.88636363636364)(1134.62626262626,-9.88636363636364)(1119.69696969697,-9.88636363636364)(1104.76767676768,-9.93181818181818)(1089.83838383838,-9.97727272727273)(1087.35016835017,-10)(1074.90909090909,-10.0757575757576)(1059.9797979798,-10.1060606060606)(1045.05050505051,-10.1363636363636)(1030.12121212121,-10.0757575757576)(1017.68013468013,-10)(1015.19191919192,-9.97727272727273)(1000.26262626263,-9.90909090909091)(985.333333333333,-9.88636363636364)(970.40404040404,-9.90909090909091)(955.474747474748,-9.93181818181818)(940.545454545455,-9.87272727272727)(929.348484848485,-9.81818181818182)(925.616161616162,-9.79545454545455)(910.686868686869,-9.72727272727273)(895.757575757576,-9.70454545454546)(884.560606060606,-9.63636363636364)(880.828282828283,-9.61363636363636)(865.89898989899,-9.61363636363636)(850.969696969697,-9.47272727272727)(848.481481481482,-9.45454545454546)(836.040404040404,-9.3030303030303)(834.174242424242,-9.27272727272727)(821.111111111111,-9.11363636363636)(819.244949494949,-9.09090909090909)(808.04797979798,-8.90909090909091)(807.674747474747,-8.72727272727273)(806.181818181818,-8.63636363636364)(803.693602693603,-8.54545454545455)(806.181818181818,-8.45454545454545)(808.04797979798,-8.36363636363636)(821.111111111111,-8.20454545454546)(823.599326599327,-8.18181818181818)(834.547474747475,-8)(836.040404040404,-7.98181818181818)(850.969696969697,-7.90909090909091)(856.568181818182,-7.81818181818182)(865.89898989899,-7.72727272727273)(880.828282828283,-7.70454545454546)(888.292929292929,-7.63636363636364)(895.757575757576,-7.58181818181818)(910.686868686869,-7.61363636363636)(925.616161616162,-7.48484848484849)(940.545454545455,-7.48484848484849)(943.03367003367,-7.45454545454546)(955.474747474747,-7.37878787878788)(970.40404040404,-7.40909090909091)(985.333333333333,-7.37878787878788)(1000.26262626263,-7.37878787878788)(1015.19191919192,-7.32727272727273)(1030.12121212121,-7.36363636363636)(1045.05050505051,-7.4)(1059.9797979798,-7.40909090909091)(1065.57828282828,-7.45454545454546)(1074.90909090909,-7.56818181818182)(1089.83838383838,-7.61818181818182)(1092.3265993266,-7.63636363636364)(1104.76767676768,-7.75)(1119.69696969697,-7.79545454545455)(1127.16161616162,-7.81818181818182)(1134.62626262626,-7.83636363636364)(1149.55555555556,-7.87272727272727)(1164.48484848485,-7.90909090909091)(1179.41414141414,-7.94545454545455)(1194.34343434343,-7.95454545454546)(1209.27272727273,-7.95454545454546)(1224.20202020202,-7.92424242424242)(1239.13131313131,-7.92424242424242)(1254.06060606061,-7.95454545454546)(1265.25757575758,-8)(1268.9898989899,-8.04545454545455)(1283.91919191919,-8.09090909090909)(1298.84848484848,-8.09090909090909)(1310.04545454545,-8.18181818181818)(1313.77777777778,-8.21212121212121)(1328.70707070707,-8.27272727272727)(1339.90404040404,-8.36363636363636)(1343.63636363636,-8.39393939393939)(1356.07744107744,-8.54545454545455)(1358.56565656566,-8.56818181818182)(1373.49494949495,-8.67272727272727)(1379.09343434343,-8.72727272727273)(1388.42424242424,-8.84090909090909)(1399.62121212121,-8.90909090909091)(1403.35353535354,-8.95454545454546)(1418.28282828283,-9.04545454545455)(1420.14898989899,-9.09090909090909)(1422.76161616162,-9.27272727272727)(1433.21212121212,-9.43181818181818)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (59.7171717171717,-7.77272727272727)(67.1818181818182,-7.63636363636364)(74.6464646464647,-7.54545454545455)(85.8434343434344,-7.45454545454546)(89.5757575757576,-7.42424242424243)(104.505050505051,-7.36363636363636)(119.434343434343,-7.3030303030303)(126.89898989899,-7.27272727272727)(134.363636363636,-7.24242424242424)(149.292929292929,-7.18181818181818)(164.222222222222,-7.24242424242424)(171.686868686869,-7.27272727272727)(179.151515151515,-7.31818181818182)(194.080808080808,-7.40909090909091)(209.010101010101,-7.36363636363636)(223.939393939394,-7.36363636363636)(238.868686868687,-7.36363636363636)(253.79797979798,-7.42424242424243)(261.262626262626,-7.45454545454546)(268.727272727273,-7.48484848484849)(283.656565656566,-7.54545454545455)(298.585858585859,-7.54545454545455)(313.515151515152,-7.60606060606061)(317.247474747475,-7.63636363636364)(328.444444444444,-7.77272727272727)(335.909090909091,-7.81818181818182)(343.373737373737,-7.84848484848485)(358.30303030303,-7.86363636363636)(360.791245791246,-7.81818181818182)(373.232323232323,-7.66666666666667)(388.161616161616,-7.66666666666667)(403.090909090909,-7.66666666666667)(406.823232323232,-7.63636363636364)(414.287878787879,-7.45454545454546)(418.020202020202,-7.43181818181818)(432.949494949495,-7.36363636363636)(447.878787878788,-7.32727272727273)(462.808080808081,-7.29545454545455)(470.272727272727,-7.27272727272727)(477.737373737374,-7.22727272727273)(492.666666666667,-7.18181818181818)(507.59595959596,-7.18181818181818)(522.525252525253,-7.24242424242424)(537.454545454546,-7.25)(552.383838383838,-7.21818181818182)(567.313131313131,-7.20454545454546)(582.242424242424,-7.15909090909091)(597.171717171717,-7.22727272727273)(612.10101010101,-7.18181818181818)(623.29797979798,-7.27272727272727)(627.030303030303,-7.29545454545455)(641.959595959596,-7.3030303030303)(656.888888888889,-7.42424242424243)(660.621212121212,-7.45454545454546)(671.818181818182,-7.54545454545455)(679.282828282828,-7.63636363636364)(686.747474747475,-7.77272727272727)(689.23569023569,-7.81818181818182)(694.212121212121,-8)(694.212121212121,-8.18181818181818)(689.23569023569,-8.36363636363636)(686.747474747475,-8.40909090909091)(679.282828282828,-8.54545454545455)(671.818181818182,-8.63636363636364)(660.621212121212,-8.72727272727273)(656.888888888889,-8.75757575757576)(641.959595959596,-8.81818181818182)(627.030303030303,-8.81818181818182)(612.10101010101,-8.86363636363636)(604.636363636364,-8.90909090909091)(597.171717171717,-8.93939393939394)(582.242424242424,-9.02272727272727)(567.313131313131,-9.07272727272727)(563.580808080808,-9.09090909090909)(552.383838383838,-9.15909090909091)(537.454545454546,-9.18181818181818)(522.525252525253,-9.20454545454546)(507.59595959596,-9.21818181818182)(492.666666666667,-9.25757575757576)(477.737373737374,-9.25454545454546)(462.808080808081,-9.20454545454546)(450.367003367003,-9.09090909090909)(447.878787878788,-9.06060606060606)(432.949494949495,-8.93939393939394)(425.484848484849,-8.90909090909091)(418.020202020202,-8.88636363636364)(403.090909090909,-8.81818181818182)(388.161616161616,-8.77272727272727)(373.232323232323,-8.77272727272727)(358.30303030303,-8.81818181818182)(343.373737373737,-8.81818181818182)(328.444444444444,-8.81818181818182)(313.515151515152,-8.84090909090909)(298.585858585859,-8.88636363636364)(291.121212121212,-8.90909090909091)(283.656565656566,-8.93939393939394)(268.727272727273,-9.06060606060606)(264.99494949495,-9.09090909090909)(253.79797979798,-9.18181818181818)(242.60101010101,-9.27272727272727)(238.868686868687,-9.3030303030303)(223.939393939394,-9.42424242424243)(216.474747474747,-9.45454545454546)(209.010101010101,-9.48484848484848)(194.080808080808,-9.54545454545455)(179.151515151515,-9.54545454545455)(164.222222222222,-9.54545454545455)(149.292929292929,-9.54545454545455)(138.09595959596,-9.45454545454546)(134.363636363636,-9.40909090909091)(126.89898989899,-9.27272727272727)(121.922558922559,-9.09090909090909)(119.434343434343,-9)(116.946127946128,-8.90909090909091)(111.969696969697,-8.72727272727273)(106.993265993266,-8.54545454545455)(104.505050505051,-8.45454545454545)(100.772727272727,-8.36363636363636)(89.5757575757576,-8.27272727272727)(78.3787878787879,-8.18181818181818)(74.6464646464647,-8.15151515151515)(62.2053872053872,-8)(59.7171717171717,-7.90909090909091)(57.2289562289562,-7.81818181818182)(59.7171717171717,-7.77272727272727)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (447.878787878788,-7.78787878787879)(462.808080808081,-7.75)(477.737373737374,-7.70454545454546)(488.934343434343,-7.63636363636364)(492.666666666667,-7.60606060606061)(507.59595959596,-7.54545454545455)(522.525252525253,-7.56818181818182)(537.454545454546,-7.54545454545455)(552.383838383838,-7.50909090909091)(567.313131313131,-7.52272727272727)(582.242424242424,-7.54545454545455)(589.707070707071,-7.63636363636364)(597.171717171717,-7.70454545454546)(609.612794612795,-7.81818181818182)(612.10101010101,-7.90909090909091)(614.589225589226,-8)(614.589225589226,-8.18181818181818)(612.10101010101,-8.21212121212121)(599.659932659933,-8.36363636363636)(597.171717171717,-8.39393939393939)(584.73063973064,-8.54545454545455)(582.242424242424,-8.57575757575757)(569.801346801347,-8.72727272727273)(567.313131313131,-8.74545454545454)(552.383838383838,-8.79545454545455)(537.454545454546,-8.81818181818182)(522.525252525253,-8.87878787878788)(507.59595959596,-8.87878787878788)(500.131313131313,-8.90909090909091)(492.666666666667,-8.93181818181818)(488.934343434343,-8.90909090909091)(480.225589225589,-8.72727272727273)(477.737373737374,-8.63636363636364)(474.005050505051,-8.54545454545455)(462.808080808081,-8.45454545454545)(455.343434343434,-8.36363636363636)(447.878787878788,-8.22727272727273)(445.390572390572,-8.18181818181818)(446.012626262626,-8)(445.390572390572,-7.81818181818182)(447.878787878788,-7.78787878787879)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (477.737373737374,-14.6818181818182)(488.934343434343,-14.5454545454545)(492.666666666667,-14.5151515151515)(507.59595959596,-14.3939393939394)(515.060606060606,-14.3636363636364)(522.525252525253,-14.3333333333333)(537.454545454546,-14.2727272727273)(552.383838383838,-14.2727272727273)(567.313131313131,-14.2727272727273)(582.242424242424,-14.2727272727273)(597.171717171717,-14.2727272727273)(612.10101010101,-14.2727272727273)(627.030303030303,-14.2727272727273)(641.959595959596,-14.2727272727273)(656.888888888889,-14.2727272727273)(671.818181818182,-14.2727272727273)(686.747474747475,-14.3181818181818)(690.479797979798,-14.3636363636364)(690.479797979798,-14.5454545454545)(686.747474747475,-14.5757575757576)(671.818181818182,-14.6969696969697)(664.353535353535,-14.7272727272727)(656.888888888889,-14.7575757575758)(641.959595959596,-14.8787878787879)(634.494949494949,-14.9090909090909)(627.030303030303,-14.9393939393939)(612.10101010101,-15.0606060606061)(604.636363636364,-15.0909090909091)(597.171717171717,-15.1212121212121)(582.242424242424,-15.1818181818182)(567.313131313131,-15.2424242424242)(559.848484848485,-15.2727272727273)(552.383838383838,-15.3030303030303)(537.454545454546,-15.3636363636364)(522.525252525253,-15.3636363636364)(507.59595959596,-15.3636363636364)(492.666666666667,-15.3030303030303)(488.934343434343,-15.2727272727273)(477.737373737374,-15.1363636363636)(475.249158249158,-15.0909090909091)(470.272727272727,-14.9090909090909)(475.249158249158,-14.7272727272727)(477.737373737374,-14.6818181818182)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (164.222222222222,-8.31818181818182)(179.151515151515,-8.21212121212121)(194.080808080808,-8.21212121212121)(206.521885521886,-8.36363636363636)(209.010101010101,-8.45454545454545)(211.498316498317,-8.54545454545455)(209.010101010101,-8.63636363636364)(194.080808080808,-8.68181818181818)(179.151515151515,-8.65909090909091)(166.710437710438,-8.54545454545455)(164.222222222222,-8.45454545454545)(161.734006734007,-8.36363636363636)(164.222222222222,-8.31818181818182)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (1358.56565656566,-4.54545454545455)(0,-18)(0,0)(1478,-14.7272727272727)(0,-8.72727272727273)(477.737373737374,-9.09090909090909)(432.949494949495,-6.90909090909091)(358.30303030303,-11.2727272727273)(806.181818181818,-8.72727272727273)(895.757575757576,-10.5454545454545)(836.040404040404,-7.09090909090909)(1164.48484848485,-8.90909090909091)(1433.21212121212,-9.27272727272727)(1313.77777777778,-7.63636363636364)(1239.13131313131,-10.5454545454545)(1089.83838383838,-7.45454545454546)(0,-5.81818181818182)(656.888888888889,-9.63636363636364)(298.585858585859,-8.54545454545455)(567.313131313131,-8.18181818181818)(1015.19191919192,-10.1818181818182)(1478,-8.36363636363636)(910.686868686869,0)(880.828282828283,-18)(1478,-10.5454545454545)(343.373737373737,-8)(850.969696969697,-9.45454545454546)(701.676767676768,-8)(343.373737373737,-9.09090909090909)(1000.26262626263,-7.63636363636364)(1358.56565656566,-10)(179.151515151515,-7.63636363636364)(1209.27272727273,-7.63636363636364)(1373.49494949495,-8.36363636363636)(477.737373737374,-7.81818181818182)(0,-13.2727272727273)(865.89898989899,-7.63636363636364)(1134.62626262626,-10.3636363636364)(1478,0)(701.676767676768,-8.72727272727273)(1433.21212121212,-9.63636363636364)(552.383838383838,-7.81818181818182)(567.313131313131,-7.45454545454546)(865.89898989899,-9.81818181818182)(507.59595959596,-8.90909090909091)(343.373737373737,-7.81818181818182)(791.252525252525,-7.81818181818182)(1030.12121212121,-10.1818181818182)(1104.76767676768,-7.45454545454546)(1343.63636363636,-8.18181818181818)(1358.56565656566,-10.1818181818182)(925.616161616162,-7.45454545454546)(1448.14141414141,-9.27272727272727)(418.020202020202,-8.90909090909091)(612.10101010101,-7.63636363636364)(776.323232323232,-8.90909090909091)(597.171717171717,-8.54545454545455)(910.686868686869,-10)(1194.34343434343,-7.63636363636364)(582.242424242424,-7.45454545454546)(432.949494949495,-7.63636363636364)(806.181818181818,-8)(1268.9898989899,-10.3636363636364)(1149.55555555556,-10.1818181818182)(1418.28282828283,-8.72727272727273)(1463.07070707071,-9.45454545454546)(821.111111111111,-9.45454545454546)(1254.06060606061,-7.81818181818182)(940.545454545455,-7.45454545454546)(1104.76767676768,-10.1818181818182)(388.161616161616,-8.54545454545455)(821.111111111111,-8.18181818181818)(552.383838383838,-8.72727272727273)(462.808080808081,-7.45454545454546)(1328.70707070707,-8)(1403.35353535354,-10)(925.616161616162,-10)(1074.90909090909,-7.45454545454546)(641.959595959596,-8.18181818181818)(1388.42424242424,-8.36363636363636)(373.232323232323,-8.18181818181818)(791.252525252525,-8.90909090909091)(582.242424242424,-8.72727272727273)(1179.41414141414,-7.63636363636364)(1059.9797979798,-10.1818181818182)(567.313131313131,-8.90909090909091)(1448.14141414141,-9.09090909090909)(880.828282828283,-7.63636363636364)(641.959595959596,-7.81818181818182)(477.737373737374,-8.90909090909091)(1254.06060606061,-10)(1239.13131313131,-10)(552.383838383838,-7.45454545454546)(850.969696969697,-9.63636363636364)(1149.55555555556,-7.63636363636364)(1059.9797979798,-7.45454545454546)(985.333333333333,-7.27272727272727)(597.171717171717,-7.45454545454546)(552.383838383838,-7.27272727272727)(403.090909090909,-7.81818181818182)(432.949494949495,-3.27272727272727)(194.080808080808,-8.36363636363636)(194.080808080808,-8.54545454545455)(1045.05050505051,-10.1818181818182)(1313.77777777778,-8)(1373.49494949495,-10)(1448.14141414141,-9.63636363636364)(1000.26262626263,-10.1818181818182)(582.242424242424,-8.90909090909091)(134.363636363636,-8)(44.7878787878788,-7.63636363636364)(179.151515151515,-8.18181818181818)(164.222222222222,-8.36363636363636)(223.939393939394,-8)(179.151515151515,-8.54545454545455)(179.151515151515,-8.72727272727273)(149.292929292929,-7.81818181818182)(119.434343434343,-8.18181818181818)(253.79797979798,-8)(209.010101010101,-7.81818181818182)(253.79797979798,-8.54545454545455)(418.020202020202,-8.54545454545455)(791.252525252525,-8.54545454545455)(850.969696969697,-7.81818181818182)(612.10101010101,-7.45454545454546)(806.181818181818,-9.27272727272727)(1239.13131313131,-7.81818181818182)(1478,-18)(940.545454545455,-10)(612.10101010101,-8.72727272727273)(1268.9898989899,-10)(641.959595959596,-8.36363636363636)(1433.21212121212,-9.81818181818182)(1463.07070707071,-9.63636363636364)(418.020202020202,-7.81818181818182)(1403.35353535354,-8.54545454545455)(1433.21212121212,-8.90909090909091)(164.222222222222,-8.54545454545455)(836.040404040404,-9.63636363636364)(447.878787878788,-7.63636363636364)(537.454545454546,-7.27272727272727)(268.727272727273,-8.36363636363636)(194.080808080808,-7.81818181818182)(268.727272727273,-8.54545454545455)(1134.62626262626,-7.63636363636364)(910.686868686869,-7.45454545454546)(970.40404040404,-7.27272727272727)(1000.26262626263,-7.27272727272727)(1015.19191919192,-7.27272727272727)(1015.19191919192,-7.09090909090909)(1358.56565656566,-8.18181818181818)(388.161616161616,-8.36363636363636)(492.666666666667,-8.90909090909091)(194.080808080808,-8.72727272727273)(194.080808080808,-8.90909090909091)(164.222222222222,-8.72727272727273)(179.151515151515,-8.90909090909091)(209.010101010101,-8.90909090909091)(164.222222222222,-8.90909090909091)(223.939393939394,-8.90909090909091) 
};
\end{axis}
\end{tikzpicture}%