### 2013-03-19 figurefactory/files/limno_bgape_class_340.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478.00016587919,-9.09090909090909)(1478.00016587919,-9.27272727272727)(1478.00016587919,-9.45454545454546)(1478,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1448.14141414141,-8.93181818181818)(1456.85016835017,-9.09090909090909)(1456.85016835017,-9.27272727272727)(1461.8265993266,-9.45454545454546)(1456.85016835017,-9.63636363636364)(1453.7398989899,-9.81818181818182)(1448.14141414141,-9.88636363636364)(1433.21212121212,-9.92424242424243)(1418.28282828283,-9.98484848484848)(1415.79461279461,-10)(1403.35353535354,-10.1136363636364)(1392.15656565657,-10.1818181818182)(1388.42424242424,-10.2045454545455)(1373.49494949495,-10.3181818181818)(1358.56565656566,-10.2727272727273)(1343.63636363636,-10.25)(1336.17171717172,-10.1818181818182)(1328.70707070707,-10.1363636363636)(1321.24242424242,-10.1818181818182)(1313.77777777778,-10.25)(1298.84848484848,-10.2727272727273)(1283.91919191919,-10.2878787878788)(1270.85606060606,-10.1818181818182)(1268.9898989899,-10.1666666666667)(1254.06060606061,-10.1666666666667)(1239.13131313131,-10.1060606060606)(1224.20202020202,-10.1666666666667)(1209.27272727273,-10.1666666666667)(1205.5404040404,-10.1818181818182)(1194.34343434343,-10.2272727272727)(1183.14646464646,-10.1818181818182)(1179.41414141414,-10.1666666666667)(1164.48484848485,-10.1666666666667)(1149.55555555556,-10.1363636363636)(1142.09090909091,-10.1818181818182)(1134.62626262626,-10.25)(1119.69696969697,-10.25)(1104.76767676768,-10.2727272727273)(1089.83838383838,-10.2272727272727)(1074.90909090909,-10.2878787878788)(1059.9797979798,-10.3484848484848)(1056.24747474747,-10.3636363636364)(1045.05050505051,-10.4090909090909)(1033.85353535354,-10.3636363636364)(1030.12121212121,-10.3484848484848)(1015.19191919192,-10.2878787878788)(1000.26262626263,-10.2)(985.333333333333,-10.2)(970.40404040404,-10.2)(955.474747474747,-10.2272727272727)(944.277777777778,-10.1818181818182)(940.545454545455,-10.1666666666667)(925.616161616162,-10.1060606060606)(910.686868686869,-10.0454545454545)(895.757575757576,-10.0757575757576)(880.828282828283,-10.0227272727273)(877.09595959596,-10)(865.89898989899,-9.95454545454546)(852.462626262626,-9.81818181818182)(850.969696969697,-9.8030303030303)(836.040404040404,-9.70454545454546)(824.843434343434,-9.63636363636364)(821.111111111111,-9.61363636363636)(808.04797979798,-9.45454545454546)(806.181818181818,-9.43939393939394)(791.252525252525,-9.37878787878788)(778.189393939394,-9.27272727272727)(778.189393939394,-9.09090909090909)(776.323232323232,-9.04545454545455)(770.724747474747,-8.90909090909091)(776.323232323232,-8.84090909090909)(782.543771043771,-8.72727272727273)(776.323232323232,-8.57575757575757)(774.457070707071,-8.54545454545455)(776.323232323232,-8.5)(780.055555555556,-8.36363636363636)(781.921717171717,-8.18181818181818)(791.252525252525,-8.09090909090909)(806.181818181818,-8.01515151515152)(808.04797979798,-8)(821.111111111111,-7.89393939393939)(830.441919191919,-7.81818181818182)(836.040404040404,-7.75)(850.969696969697,-7.65151515151515)(852.462626262626,-7.63636363636364)(863.410774410775,-7.45454545454546)(865.89898989899,-7.43181818181818)(880.828282828283,-7.37878787878788)(895.757575757576,-7.31818181818182)(910.686868686869,-7.34848484848485)(925.616161616162,-7.28787878787879)(929.348484848485,-7.27272727272727)(940.545454545455,-7.20454545454546)(955.474747474747,-7.10606060606061)(970.40404040404,-7.10606060606061)(985.333333333333,-7.10606060606061)(1000.26262626263,-7.10606060606061)(1015.19191919192,-7.10606060606061)(1030.12121212121,-7.16666666666667)(1045.05050505051,-7.22727272727273)(1059.9797979798,-7.22727272727273)(1071.17676767677,-7.27272727272727)(1074.90909090909,-7.28787878787879)(1089.83838383838,-7.28787878787879)(1104.76767676768,-7.34848484848485)(1119.69696969697,-7.38636363636364)(1125.29545454545,-7.45454545454546)(1134.62626262626,-7.53030303030303)(1149.55555555556,-7.59090909090909)(1164.48484848485,-7.61818181818182)(1179.41414141414,-7.61818181818182)(1194.34343434343,-7.61818181818182)(1201.80808080808,-7.63636363636364)(1209.27272727273,-7.65151515151515)(1224.20202020202,-7.65151515151515)(1239.13131313131,-7.65151515151515)(1242.86363636364,-7.63636363636364)(1254.06060606061,-7.59090909090909)(1268.9898989899,-7.59090909090909)(1283.91919191919,-7.59090909090909)(1295.11616161616,-7.63636363636364)(1298.84848484848,-7.65151515151515)(1313.77777777778,-7.75)(1328.70707070707,-7.79545454545455)(1343.63636363636,-7.79545454545455)(1358.56565656566,-7.79545454545455)(1361.05387205387,-7.81818181818182)(1366.0303030303,-8)(1372.25084175084,-8.18181818181818)(1373.49494949495,-8.20454545454546)(1388.42424242424,-8.29545454545455)(1399.62121212121,-8.36363636363636)(1403.35353535354,-8.37878787878788)(1418.28282828283,-8.47727272727273)(1423.88131313131,-8.54545454545455)(1433.21212121212,-8.65909090909091)(1438.81060606061,-8.72727272727273)(1446.89730639731,-8.90909090909091)(1448.14141414141,-8.93181818181818)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1433.21212121212,-9.40909090909091)(1435.07828282828,-9.45454545454546)(1433.21212121212,-9.5)(1427.61363636364,-9.63636363636364)(1418.28282828283,-9.71212121212121)(1403.35353535354,-9.75)(1397.75505050505,-9.81818181818182)(1388.42424242424,-9.90909090909091)(1379.09343434343,-10)(1373.49494949495,-10.0454545454545)(1362.29797979798,-10)(1358.56565656566,-9.97727272727273)(1343.63636363636,-9.93181818181818)(1328.70707070707,-9.86363636363636)(1313.77777777778,-9.93181818181818)(1298.84848484848,-9.97727272727273)(1291.38383838384,-10)(1283.91919191919,-10.0151515151515)(1282.05303030303,-10)(1268.9898989899,-9.89393939393939)(1254.06060606061,-9.89393939393939)(1239.13131313131,-9.83333333333333)(1224.20202020202,-9.89393939393939)(1209.27272727273,-9.89393939393939)(1194.34343434343,-9.95454545454546)(1179.41414141414,-9.89393939393939)(1164.48484848485,-9.89393939393939)(1149.55555555556,-9.86363636363636)(1134.62626262626,-9.93181818181818)(1119.69696969697,-9.93181818181818)(1104.76767676768,-9.97727272727273)(1089.83838383838,-9.95454545454546)(1078.64141414141,-10)(1074.90909090909,-10.0151515151515)(1059.9797979798,-10.0757575757576)(1045.05050505051,-10.1363636363636)(1030.12121212121,-10.0757575757576)(1015.19191919192,-10.0151515151515)(1012.7037037037,-10)(1000.26262626263,-9.92424242424243)(985.333333333333,-9.92424242424243)(970.40404040404,-9.92424242424243)(955.474747474747,-9.95454545454546)(940.545454545455,-9.89393939393939)(925.616161616162,-9.83333333333333)(921.883838383838,-9.81818181818182)(910.686868686869,-9.77272727272727)(895.757575757576,-9.8)(880.828282828283,-9.72727272727273)(865.89898989899,-9.65454545454546)(864.032828282828,-9.63636363636364)(850.969696969697,-9.53030303030303)(841.638888888889,-9.45454545454546)(836.040404040404,-9.38636363636364)(821.111111111111,-9.28787878787879)(819.244949494949,-9.27272727272727)(811.780303030303,-9.09090909090909)(806.181818181818,-9.02272727272727)(796.85101010101,-8.90909090909091)(804.93771043771,-8.72727272727273)(798.717171717172,-8.54545454545455)(804.688888888889,-8.36363636363636)(806.181818181818,-8.34090909090909)(819.244949494949,-8.18181818181818)(821.111111111111,-8.16666666666667)(836.040404040404,-8.06818181818182)(847.237373737374,-8)(850.969696969697,-7.97727272727273)(864.032828282828,-7.81818181818182)(865.89898989899,-7.8)(880.828282828283,-7.72727272727273)(890.159090909091,-7.63636363636364)(895.757575757576,-7.59090909090909)(910.686868686869,-7.62121212121212)(925.616161616162,-7.56060606060606)(940.545454545455,-7.52272727272727)(946.143939393939,-7.45454545454546)(955.474747474747,-7.37878787878788)(970.40404040404,-7.37878787878788)(985.333333333333,-7.37878787878788)(1000.26262626263,-7.37878787878788)(1015.19191919192,-7.37878787878788)(1030.12121212121,-7.43939393939394)(1033.85353535354,-7.45454545454546)(1045.05050505051,-7.5)(1059.9797979798,-7.5)(1074.90909090909,-7.56060606060606)(1089.83838383838,-7.56060606060606)(1104.76767676768,-7.62121212121212)(1108.5,-7.63636363636364)(1119.69696969697,-7.70454545454546)(1134.62626262626,-7.8030303030303)(1138.35858585859,-7.81818181818182)(1149.55555555556,-7.86363636363636)(1164.48484848485,-7.89393939393939)(1179.41414141414,-7.89393939393939)(1194.34343434343,-7.89393939393939)(1209.27272727273,-7.92424242424242)(1224.20202020202,-7.92424242424242)(1239.13131313131,-7.92424242424242)(1254.06060606061,-7.86363636363636)(1268.9898989899,-7.86363636363636)(1283.91919191919,-7.86363636363636)(1298.84848484848,-7.92424242424242)(1308.17929292929,-8)(1313.77777777778,-8.06818181818182)(1328.70707070707,-8.09090909090909)(1343.63636363636,-8.16363636363636)(1345.50252525253,-8.18181818181818)(1352.96717171717,-8.36363636363636)(1358.56565656566,-8.43181818181818)(1373.49494949495,-8.53030303030303)(1377.22727272727,-8.54545454545455)(1388.42424242424,-8.61363636363636)(1403.35353535354,-8.70454545454545)(1405.2196969697,-8.72727272727273)(1418.28282828283,-8.83333333333333)(1424.50336700337,-8.90909090909091)(1429.4797979798,-9.09090909090909)(1429.4797979798,-9.27272727272727)(1433.21212121212,-9.40909090909091)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (403.090909090909,-7.77272727272727)(414.287878787879,-7.63636363636364)(418.020202020202,-7.5)(419.886363636364,-7.45454545454546)(432.949494949495,-7.29545454545455)(447.878787878788,-7.28787878787879)(462.808080808081,-7.28787878787879)(466.540404040404,-7.27272727272727)(477.737373737374,-7.20454545454546)(492.666666666667,-7.10606060606061)(507.59595959596,-7.10606060606061)(522.525252525253,-7.16666666666667)(537.454545454546,-7.22727272727273)(552.383838383838,-7.16666666666667)(567.313131313131,-7.20454545454546)(582.242424242424,-7.20454545454546)(587.840909090909,-7.27272727272727)(597.171717171717,-7.34848484848485)(612.10101010101,-7.38636363636364)(617.699494949495,-7.45454545454546)(627.030303030303,-7.56818181818182)(632.628787878788,-7.63636363636364)(641.959595959596,-7.75)(647.558080808081,-7.81818181818182)(656.888888888889,-7.93181818181818)(662.487373737374,-8)(662.487373737374,-8.18181818181818)(660.621212121212,-8.36363636363636)(656.888888888889,-8.38636363636364)(641.959595959596,-8.5)(634.494949494949,-8.54545454545455)(627.030303030303,-8.61363636363636)(617.699494949495,-8.72727272727273)(612.10101010101,-8.86363636363636)(608.368686868687,-8.90909090909091)(597.171717171717,-8.97727272727273)(582.242424242424,-9.07575757575758)(567.313131313131,-9.07575757575758)(563.580808080808,-9.09090909090909)(552.383838383838,-9.13636363636364)(537.454545454546,-9.1969696969697)(522.525252525253,-9.25757575757576)(507.59595959596,-9.25757575757576)(492.666666666667,-9.25757575757576)(477.737373737374,-9.1969696969697)(462.808080808081,-9.15909090909091)(457.209595959596,-9.09090909090909)(449.122895622896,-8.90909090909091)(447.878787878788,-8.89393939393939)(432.949494949495,-8.83333333333333)(419.886363636364,-8.72727272727273)(418.020202020202,-8.68181818181818)(412.421717171717,-8.54545454545455)(409.311447811448,-8.36363636363636)(403.090909090909,-8.25)(391.893939393939,-8.18181818181818)(391.893939393939,-8)(391.893939393939,-7.81818181818182)(403.090909090909,-7.77272727272727)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (432.949494949495,-7.77272727272727)(438.54797979798,-7.63636363636364)(447.878787878788,-7.56060606060606)(462.808080808081,-7.56060606060606)(477.737373737374,-7.52272727272727)(483.335858585859,-7.45454545454546)(492.666666666667,-7.37878787878788)(507.59595959596,-7.37878787878788)(522.525252525253,-7.43939393939394)(526.257575757576,-7.45454545454546)(537.454545454546,-7.5)(548.651515151515,-7.45454545454546)(552.383838383838,-7.43939393939394)(556.116161616162,-7.45454545454546)(567.313131313131,-7.52272727272727)(582.242424242424,-7.52272727272727)(597.171717171717,-7.62121212121212)(600.90404040404,-7.63636363636364)(612.10101010101,-7.70454545454546)(621.431818181818,-7.81818181818182)(627.030303030303,-7.88636363636364)(636.361111111111,-8)(636.361111111111,-8.18181818181818)(627.030303030303,-8.29545454545455)(615.833333333333,-8.36363636363636)(612.10101010101,-8.38636363636364)(599.037878787879,-8.54545454545455)(597.171717171717,-8.56818181818182)(584.108585858586,-8.72727272727273)(582.242424242424,-8.75)(567.313131313131,-8.8030303030303)(552.383838383838,-8.86363636363636)(541.186868686869,-8.90909090909091)(537.454545454546,-8.92424242424242)(522.525252525253,-8.98484848484848)(507.59595959596,-8.98484848484848)(492.666666666667,-8.98484848484848)(483.335858585859,-8.90909090909091)(477.737373737374,-8.84090909090909)(462.808080808081,-8.75)(460.941919191919,-8.72727272727273)(447.878787878788,-8.56818181818182)(444.146464646465,-8.54545454545455)(432.949494949495,-8.40909090909091)(431.705387205387,-8.36363636363636)(431.083333333333,-8.18181818181818)(431.083333333333,-8)(431.083333333333,-7.81818181818182)(432.949494949495,-7.77272727272727)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (89.5757575757576,-8.13636363636364)(104.505050505051,-8.13636363636364)(119.434343434343,-8.02272727272727)(134.363636363636,-8.06818181818182)(139.962121212121,-8)(149.292929292929,-7.84848484848485)(164.222222222222,-7.83636363636364)(167.954545454545,-7.81818181818182)(179.151515151515,-7.77272727272727)(194.080808080808,-7.77272727272727)(209.010101010101,-7.77272727272727)(220.207070707071,-7.81818181818182)(223.939393939394,-7.83333333333333)(238.868686868687,-7.84090909090909)(253.79797979798,-7.93181818181818)(268.727272727273,-7.93181818181818)(279.924242424242,-8)(283.656565656566,-8.02272727272727)(296.719696969697,-8.18181818181818)(298.585858585859,-8.22727272727273)(304.184343434343,-8.36363636363636)(298.585858585859,-8.5)(296.719696969697,-8.54545454545455)(286.144781144781,-8.72727272727273)(283.656565656566,-8.75)(268.727272727273,-8.77272727272727)(257.530303030303,-8.90909090909091)(253.79797979798,-8.95454545454546)(238.868686868687,-8.97727272727273)(229.537878787879,-9.09090909090909)(223.939393939394,-9.15909090909091)(209.010101010101,-9.15909090909091)(194.080808080808,-9.15909090909091)(188.482323232323,-9.09090909090909)(179.151515151515,-8.97727272727273)(167.954545454545,-8.90909090909091)(164.222222222222,-8.89090909090909)(149.292929292929,-8.89090909090909)(134.363636363636,-8.81818181818182)(121.922558922559,-8.72727272727273)(119.434343434343,-8.70454545454545)(106.371212121212,-8.54545454545455)(104.505050505051,-8.5)(93.3080808080808,-8.36363636363636)(89.5757575757576,-8.27272727272727)(87.709595959596,-8.18181818181818)(89.5757575757576,-8.13636363636364)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (149.292929292929,-8.27272727272727)(153.025252525253,-8.18181818181818)(164.222222222222,-8.13636363636364)(179.151515151515,-8.04545454545455)(194.080808080808,-8.04545454545455)(209.010101010101,-8.04545454545455)(223.939393939394,-8.10606060606061)(238.868686868687,-8.16666666666667)(242.60101010101,-8.18181818181818)(253.79797979798,-8.31818181818182)(257.530303030303,-8.36363636363636)(253.79797979798,-8.40909090909091)(242.60101010101,-8.54545454545455)(238.868686868687,-8.56818181818182)(225.805555555556,-8.72727272727273)(223.939393939394,-8.75)(209.010101010101,-8.84090909090909)(194.080808080808,-8.75)(190.348484848485,-8.72727272727273)(179.151515151515,-8.65909090909091)(164.222222222222,-8.59090909090909)(149.292929292929,-8.59090909090909)(143.694444444444,-8.54545454545455)(147.426767676768,-8.36363636363636)(149.292929292929,-8.27272727272727)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (1358.56565656566,-4.54545454545455)(0,-18)(0,0)(1478,-14.7272727272727)(0,-8.72727272727273)(477.737373737374,-9.09090909090909)(432.949494949495,-6.90909090909091)(358.30303030303,-11.2727272727273)(806.181818181818,-8.72727272727273)(895.757575757576,-10.5454545454545)(836.040404040404,-7.09090909090909)(1164.48484848485,-8.90909090909091)(1433.21212121212,-9.27272727272727)(1313.77777777778,-7.63636363636364)(1239.13131313131,-10.5454545454545)(1089.83838383838,-7.45454545454546)(0,-5.81818181818182)(656.888888888889,-9.63636363636364)(298.585858585859,-8.54545454545455)(567.313131313131,-8.18181818181818)(1015.19191919192,-10.1818181818182)(1478,-8.36363636363636)(910.686868686869,0)(880.828282828283,-18)(1478,-10.5454545454545)(343.373737373737,-8)(850.969696969697,-9.45454545454546)(701.676767676768,-8)(343.373737373737,-9.09090909090909)(1000.26262626263,-7.63636363636364)(1358.56565656566,-10)(179.151515151515,-7.63636363636364)(1209.27272727273,-7.63636363636364)(1373.49494949495,-8.36363636363636)(477.737373737374,-7.81818181818182)(0,-13.2727272727273)(865.89898989899,-7.63636363636364)(1134.62626262626,-10.3636363636364)(1478,0)(701.676767676768,-8.72727272727273)(1433.21212121212,-9.63636363636364)(552.383838383838,-7.81818181818182)(567.313131313131,-7.45454545454546)(865.89898989899,-9.81818181818182)(507.59595959596,-8.90909090909091)(343.373737373737,-7.81818181818182)(791.252525252525,-7.81818181818182)(1030.12121212121,-10.1818181818182)(1104.76767676768,-7.45454545454546)(1343.63636363636,-8.18181818181818)(1358.56565656566,-10.1818181818182)(925.616161616162,-7.45454545454546)(1448.14141414141,-9.27272727272727)(418.020202020202,-8.90909090909091)(612.10101010101,-7.63636363636364)(776.323232323232,-8.90909090909091)(597.171717171717,-8.54545454545455)(910.686868686869,-10)(1194.34343434343,-7.63636363636364)(582.242424242424,-7.45454545454546)(432.949494949495,-7.63636363636364)(806.181818181818,-8)(1268.9898989899,-10.3636363636364)(1149.55555555556,-10.1818181818182)(1418.28282828283,-8.72727272727273)(1463.07070707071,-9.45454545454546)(821.111111111111,-9.45454545454546)(1254.06060606061,-7.81818181818182)(940.545454545455,-7.45454545454546)(1104.76767676768,-10.1818181818182)(388.161616161616,-8.54545454545455)(821.111111111111,-8.18181818181818)(552.383838383838,-8.72727272727273)(462.808080808081,-7.45454545454546)(1328.70707070707,-8)(1403.35353535354,-10)(925.616161616162,-10)(1074.90909090909,-7.45454545454546)(641.959595959596,-8.18181818181818)(1388.42424242424,-8.36363636363636)(373.232323232323,-8.18181818181818)(791.252525252525,-8.90909090909091)(582.242424242424,-8.72727272727273)(1179.41414141414,-7.63636363636364)(1059.9797979798,-10.1818181818182)(567.313131313131,-8.90909090909091)(1448.14141414141,-9.09090909090909)(880.828282828283,-7.63636363636364)(641.959595959596,-7.81818181818182)(477.737373737374,-8.90909090909091)(1254.06060606061,-10)(1239.13131313131,-10)(552.383838383838,-7.45454545454546)(850.969696969697,-9.63636363636364)(1149.55555555556,-7.63636363636364)(1059.9797979798,-7.45454545454546)(985.333333333333,-7.27272727272727)(597.171717171717,-7.45454545454546)(552.383838383838,-7.27272727272727)(403.090909090909,-7.81818181818182)(432.949494949495,-3.27272727272727)(194.080808080808,-8.36363636363636)(194.080808080808,-8.54545454545455)(1045.05050505051,-10.1818181818182)(1313.77777777778,-8)(1373.49494949495,-10)(1448.14141414141,-9.63636363636364)(1000.26262626263,-10.1818181818182)(582.242424242424,-8.90909090909091)(134.363636363636,-8)(44.7878787878788,-7.63636363636364)(179.151515151515,-8.18181818181818)(164.222222222222,-8.36363636363636)(223.939393939394,-8)(179.151515151515,-8.54545454545455)(179.151515151515,-8.72727272727273)(149.292929292929,-7.81818181818182)(119.434343434343,-8.18181818181818)(253.79797979798,-8)(209.010101010101,-7.81818181818182)(253.79797979798,-8.54545454545455)(418.020202020202,-8.54545454545455)(791.252525252525,-8.54545454545455)(850.969696969697,-7.81818181818182)(612.10101010101,-7.45454545454546)(806.181818181818,-9.27272727272727)(1239.13131313131,-7.81818181818182)(1478,-18)(940.545454545455,-10)(612.10101010101,-8.72727272727273)(1268.9898989899,-10)(641.959595959596,-8.36363636363636)(1433.21212121212,-9.81818181818182)(1463.07070707071,-9.63636363636364)(418.020202020202,-7.81818181818182)(1403.35353535354,-8.54545454545455)(1433.21212121212,-8.90909090909091)(164.222222222222,-8.54545454545455)(836.040404040404,-9.63636363636364)(447.878787878788,-7.63636363636364)(537.454545454546,-7.27272727272727)(268.727272727273,-8.36363636363636)(194.080808080808,-7.81818181818182)(268.727272727273,-8.54545454545455)(1134.62626262626,-7.63636363636364)(910.686868686869,-7.45454545454546)(970.40404040404,-7.27272727272727)(1000.26262626263,-7.27272727272727)(1015.19191919192,-7.27272727272727)(1015.19191919192,-7.09090909090909)(1358.56565656566,-8.18181818181818)(388.161616161616,-8.36363636363636)(492.666666666667,-8.90909090909091)(194.080808080808,-8.72727272727273)(194.080808080808,-8.90909090909091)(164.222222222222,-8.72727272727273)(179.151515151515,-8.90909090909091)(209.010101010101,-8.90909090909091)(164.222222222222,-8.90909090909091)(223.939393939394,-8.90909090909091)(119.434343434343,-8.36363636363636)(791.252525252525,-8.72727272727273)(627.030303030303,-7.63636363636364)(1224.20202020202,-7.81818181818182)(791.252525252525,-9.09090909090909)(1089.83838383838,-10.1818181818182)(806.181818181818,-8.18181818181818)(641.959595959596,-8)(821.111111111111,-8)(940.545454545455,-7.27272727272727)(1030.12121212121,-7.27272727272727)(403.090909090909,-8.36363636363636)(1418.28282828283,-10)(283.656565656566,-8.18181818181818)(268.727272727273,-8)(477.737373737374,-7.45454545454546)(1224.20202020202,-10)(1209.27272727273,-10)(1283.91919191919,-10)(134.363636363636,-8.54545454545455)(134.363636363636,-8.72727272727273)(791.252525252525,-8.36363636363636)(447.878787878788,-8.72727272727273)(955.474747474747,-7.27272727272727)(522.525252525253,-7.27272727272727)(268.727272727273,-8.72727272727273)(776.323232323232,-8.72727272727273)(627.030303030303,-8.54545454545455)(1209.27272727273,-7.81818181818182)(1448.14141414141,-9.81818181818182)(880.828282828283,-9.81818181818182)(253.79797979798,-8.72727272727273)(895.757575757576,-7.45454545454546)(149.292929292929,-8)(462.808080808081,-8.90909090909091)(298.585858585859,-8.36363636363636)(403.090909090909,-8)(388.161616161616,-8.18181818181818)(1074.90909090909,-10.1818181818182)(432.949494949495,-8.72727272727273)(1268.9898989899,-7.81818181818182)(119.434343434343,-8.54545454545455)(1283.91919191919,-7.81818181818182)(149.292929292929,-8.90909090909091)(298.585858585859,-8.18181818181818)(1194.34343434343,-10)(1388.42424242424,-10)(1343.63636363636,-10.1818181818182)(283.656565656566,-8.54545454545455)(1463.07070707071,-9.27272727272727)(836.040404040404,-7.81818181818182)(238.868686868687,-8)(552.383838383838,-8.90909090909091)(522.525252525253,-9.09090909090909)(507.59595959596,-9.09090909090909)(537.454545454546,-9.09090909090909)(492.666666666667,-9.09090909090909)(1478,-9.45454545454546)(641.959595959596,-8.54545454545455)(1343.63636363636,-8)(1045.05050505051,-7.27272727272727)(1119.69696969697,-7.63636363636364)(418.020202020202,-7.63636363636364)(119.434343434343,-8)(836.040404040404,-8)(134.363636363636,-8.18181818181818)(164.222222222222,-8)(806.181818181818,-9.45454545454546)(791.252525252525,-9.27272727272727)(850.969696969697,-9.81818181818182)(283.656565656566,-8.36363636363636)(388.161616161616,-8)(1373.49494949495,-8.18181818181818)(134.363636363636,-8.36363636363636)(671.818181818182,-14.3636363636364)(507.59595959596,-7.27272727272727)(567.313131313131,-7.27272727272727)(582.242424242424,-7.27272727272727)(447.878787878788,-7.45454545454546)(492.666666666667,-7.27272727272727)(1298.84848484848,-7.81818181818182)(597.171717171717,-8.90909090909091)(552.383838383838,-9.09090909090909)(627.030303030303,-8.72727272727273)(627.030303030303,-7.81818181818182)(955.474747474747,-10)(1119.69696969697,-10.1818181818182)(1313.77777777778,-7.81818181818182)(149.292929292929,-8.72727272727273)(1164.48484848485,-7.63636363636364)(1448.14141414141,-9.45454545454546)(1433.21212121212,-9.09090909090909)(1433.21212121212,-8.72727272727273)(1448.14141414141,-8.90909090909091)(403.090909090909,-8.54545454545455)(403.090909090909,-8.18181818181818)(477.737373737374,-7.27272727272727)(238.868686868687,-8.72727272727273)(268.727272727273,-8.18181818181818)(223.939393939394,-7.81818181818182)(656.888888888889,-8.18181818181818)(656.888888888889,-8)(776.323232323232,-8.54545454545455)(283.656565656566,-8)(1328.70707070707,-10.1818181818182)(432.949494949495,-8.54545454545455)(1418.28282828283,-8.54545454545455)(656.888888888889,-8.36363636363636)(1134.62626262626,-10.1818181818182)(418.020202020202,-8.36363636363636)(1179.41414141414,-10)(432.949494949495,-7.45454545454546)(179.151515151515,-8)(179.151515151515,-7.81818181818182)(1373.49494949495,-10.1818181818182)(597.171717171717,-7.27272727272727)(985.333333333333,-10.1818181818182)(238.868686868687,-8.90909090909091)(253.79797979798,-8.90909090909091)(821.111111111111,-9.63636363636364)(403.090909090909,-7.63636363636364)(418.020202020202,-8.18181818181818)(418.020202020202,-8)(641.959595959596,-7.63636363636364)(104.505050505051,-8.18181818181818)(104.505050505051,-8.36363636363636)(104.505050505051,-8)(89.5757575757576,-8.18181818181818)(1358.56565656566,-8)(209.010101010101,-8)(164.222222222222,-7.81818181818182)(1298.84848484848,-10)(1313.77777777778,-10.1818181818182)(970.40404040404,-10)(895.757575757576,-10)(1463.07070707071,-9.09090909090909)(925.616161616162,-7.27272727272727)(1403.35353535354,-8.36363636363636)(1388.42424242424,-8.18181818181818)(1328.70707070707,-7.81818181818182)(612.10101010101,-8.90909090909091)(447.878787878788,-8.90909090909091)(1059.9797979798,-7.27272727272727)(1418.28282828283,-9.81818181818182)(1194.34343434343,-7.81818181818182)(432.949494949495,-8.36363636363636)(627.030303030303,-7.45454545454546)(283.656565656566,-8.72727272727273)(119.434343434343,-8.72727272727273)(597.171717171717,-8.72727272727273)(238.868686868687,-7.81818181818182)(194.080808080808,-8)(776.323232323232,-9.09090909090909)(865.89898989899,-7.81818181818182)(761.393939393939,-8.72727272727273)(1298.84848484848,-10.1818181818182)(1164.48484848485,-10)(313.515151515152,-8.36363636363636)(253.79797979798,-8.18181818181818)(223.939393939394,-8.72727272727273)(462.808080808081,-8.72727272727273)(627.030303030303,-8.36363636363636)(656.888888888889,-7.81818181818182)(627.030303030303,-8)(1149.55555555556,-10)(1283.91919191919,-10.1818181818182)(209.010101010101,-9.09090909090909)(194.080808080808,-9.09090909090909)(223.939393939394,-9.09090909090909)(447.878787878788,-8.54545454545455)(134.363636363636,-7.81818181818182)(462.808080808081,-7.27272727272727)(1433.21212121212,-10)(432.949494949495,-7.81818181818182)(970.40404040404,-10.1818181818182)(1104.76767676768,-7.63636363636364)(253.79797979798,-8.36363636363636)(641.959595959596,-8.72727272727273)(462.808080808081,-9.09090909090909)(1373.49494949495,-8) 
};
\end{axis}
\end{tikzpicture}%