### 2013-03-19 figurefactory/files/limno_bgape_class_354.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478,-9.09090909090909)(1478,-9.27272727272727)(1478,-9.45454545454546)(1478,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1448.14141414141,-9.04545454545455)(1450.62962962963,-9.09090909090909)(1455.60606060606,-9.27272727272727)(1450.62962962963,-9.45454545454546)(1448.14141414141,-9.5)(1440.67676767677,-9.63636363636364)(1433.21212121212,-9.72727272727273)(1422.01515151515,-9.81818181818182)(1418.28282828283,-9.84848484848485)(1403.35353535354,-9.96969696969697)(1395.88888888889,-10)(1388.42424242424,-10.030303030303)(1373.49494949495,-10.0909090909091)(1358.56565656566,-10.0909090909091)(1343.63636363636,-10.0909090909091)(1328.70707070707,-10.0909090909091)(1313.77777777778,-10.0909090909091)(1298.84848484848,-10.0909090909091)(1283.91919191919,-10.0909090909091)(1268.9898989899,-10.0909090909091)(1254.06060606061,-10.0909090909091)(1239.13131313131,-10.0909090909091)(1224.20202020202,-10.0909090909091)(1209.27272727273,-10.0909090909091)(1194.34343434343,-10.0909090909091)(1179.41414141414,-10.0909090909091)(1164.48484848485,-10.0909090909091)(1149.55555555556,-10.0909090909091)(1134.62626262626,-10.0909090909091)(1119.69696969697,-10.0909090909091)(1104.76767676768,-10.0909090909091)(1089.83838383838,-10.0909090909091)(1074.90909090909,-10.0909090909091)(1059.9797979798,-10.0909090909091)(1045.05050505051,-10.0909090909091)(1030.12121212121,-10.0909090909091)(1015.19191919192,-10.0909090909091)(1000.26262626263,-10.0909090909091)(985.333333333333,-10.0909090909091)(970.40404040404,-10.0909090909091)(955.474747474747,-10.0909090909091)(940.545454545455,-10.030303030303)(933.080808080808,-10)(925.616161616162,-9.96969696969697)(910.686868686869,-9.90909090909091)(895.757575757576,-9.90909090909091)(880.828282828283,-9.84848484848485)(873.363636363636,-9.81818181818182)(865.89898989899,-9.78787878787879)(850.969696969697,-9.72727272727273)(836.040404040404,-9.66666666666667)(832.308080808081,-9.63636363636364)(821.111111111111,-9.54545454545455)(813.646464646465,-9.45454545454546)(806.181818181818,-9.36363636363636)(798.717171717172,-9.27272727272727)(791.252525252525,-9.13636363636364)(788.76430976431,-9.09090909090909)(783.787878787879,-8.90909090909091)(783.787878787879,-8.72727272727273)(788.76430976431,-8.54545454545455)(791.252525252525,-8.5)(798.717171717172,-8.36363636363636)(806.181818181818,-8.27272727272727)(813.646464646465,-8.18181818181818)(821.111111111111,-8.09090909090909)(832.308080808081,-8)(836.040404040404,-7.96969696969697)(850.969696969697,-7.84848484848485)(854.70202020202,-7.81818181818182)(865.89898989899,-7.72727272727273)(877.09595959596,-7.63636363636364)(880.828282828283,-7.60606060606061)(895.757575757576,-7.48484848484849)(903.222222222222,-7.45454545454546)(910.686868686869,-7.42424242424243)(925.616161616162,-7.36363636363636)(940.545454545455,-7.36363636363636)(955.474747474747,-7.3030303030303)(962.939393939394,-7.27272727272727)(970.40404040404,-7.24242424242424)(985.333333333333,-7.18181818181818)(1000.26262626263,-7.18181818181818)(1015.19191919192,-7.24242424242424)(1022.65656565657,-7.27272727272727)(1030.12121212121,-7.3030303030303)(1045.05050505051,-7.36363636363636)(1059.9797979798,-7.36363636363636)(1074.90909090909,-7.42424242424243)(1082.37373737374,-7.45454545454546)(1089.83838383838,-7.48484848484849)(1104.76767676768,-7.54545454545455)(1119.69696969697,-7.54545454545455)(1134.62626262626,-7.60606060606061)(1142.09090909091,-7.63636363636364)(1149.55555555556,-7.66666666666667)(1164.48484848485,-7.72727272727273)(1179.41414141414,-7.72727272727273)(1194.34343434343,-7.72727272727273)(1209.27272727273,-7.72727272727273)(1224.20202020202,-7.72727272727273)(1239.13131313131,-7.72727272727273)(1254.06060606061,-7.72727272727273)(1268.9898989899,-7.72727272727273)(1283.91919191919,-7.78787878787879)(1291.38383838384,-7.81818181818182)(1298.84848484848,-7.84848484848485)(1313.77777777778,-7.90909090909091)(1328.70707070707,-7.90909090909091)(1343.63636363636,-7.96969696969697)(1347.36868686869,-8)(1358.56565656566,-8.09090909090909)(1369.76262626263,-8.18181818181818)(1373.49494949495,-8.21212121212121)(1388.42424242424,-8.33333333333333)(1392.15656565657,-8.36363636363636)(1403.35353535354,-8.45454545454545)(1410.81818181818,-8.54545454545455)(1418.28282828283,-8.63636363636364)(1425.74747474747,-8.72727272727273)(1433.21212121212,-8.81818181818182)(1440.67676767677,-8.90909090909091)(1448.14141414141,-9.04545454545455)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (418.020202020202,-7.77272727272727)(429.217171717172,-7.63636363636364)(432.949494949495,-7.60606060606061)(447.878787878788,-7.48484848484849)(455.343434343434,-7.45454545454546)(462.808080808081,-7.42424242424243)(477.737373737374,-7.36363636363636)(492.666666666667,-7.3030303030303)(500.131313131313,-7.27272727272727)(507.59595959596,-7.24242424242424)(522.525252525253,-7.18181818181818)(537.454545454546,-7.18181818181818)(552.383838383838,-7.24242424242424)(559.848484848485,-7.27272727272727)(567.313131313131,-7.3030303030303)(582.242424242424,-7.36363636363636)(597.171717171717,-7.42424242424243)(600.90404040404,-7.45454545454546)(612.10101010101,-7.54545454545455)(619.565656565657,-7.63636363636364)(627.030303030303,-7.77272727272727)(629.518518518518,-7.81818181818182)(639.47138047138,-8)(639.47138047138,-8.18181818181818)(639.47138047138,-8.36363636363636)(629.518518518518,-8.54545454545455)(627.030303030303,-8.59090909090909)(615.833333333333,-8.72727272727273)(612.10101010101,-8.75757575757576)(597.171717171717,-8.87878787878788)(589.707070707071,-8.90909090909091)(582.242424242424,-8.93939393939394)(567.313131313131,-9)(552.383838383838,-9)(537.454545454546,-9)(522.525252525253,-9)(507.59595959596,-9)(492.666666666667,-9)(477.737373737374,-8.93939393939394)(470.272727272727,-8.90909090909091)(462.808080808081,-8.87878787878788)(447.878787878788,-8.75757575757576)(444.146464646465,-8.72727272727273)(432.949494949495,-8.63636363636364)(425.484848484849,-8.54545454545455)(418.020202020202,-8.40909090909091)(415.531986531986,-8.36363636363636)(410.555555555556,-8.18181818181818)(410.555555555556,-8)(415.531986531986,-7.81818181818182)(418.020202020202,-7.77272727272727)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (134.363636363636,-8.13636363636364)(145.560606060606,-8)(149.292929292929,-7.96969696969697)(164.222222222222,-7.90909090909091)(179.151515151515,-7.90909090909091)(194.080808080808,-7.90909090909091)(209.010101010101,-7.90909090909091)(223.939393939394,-7.96969696969697)(231.40404040404,-8)(238.868686868687,-8.03030303030303)(253.79797979798,-8.09090909090909)(264.99494949495,-8.18181818181818)(268.727272727273,-8.22727272727273)(276.191919191919,-8.36363636363636)(271.215488215488,-8.54545454545455)(268.727272727273,-8.59090909090909)(257.530303030303,-8.72727272727273)(253.79797979798,-8.75757575757576)(238.868686868687,-8.81818181818182)(223.939393939394,-8.81818181818182)(209.010101010101,-8.81818181818182)(194.080808080808,-8.81818181818182)(179.151515151515,-8.81818181818182)(164.222222222222,-8.75757575757576)(156.757575757576,-8.72727272727273)(149.292929292929,-8.6969696969697)(136.851851851852,-8.54545454545455)(134.363636363636,-8.5)(126.89898989899,-8.36363636363636)(131.875420875421,-8.18181818181818)(134.363636363636,-8.13636363636364)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (1358.56565656566,-4.54545454545455)(0,-18)(0,0)(1478,-14.7272727272727)(0,-8.72727272727273)(477.737373737374,-9.09090909090909)(432.949494949495,-6.90909090909091)(358.30303030303,-11.2727272727273)(806.181818181818,-8.72727272727273)(895.757575757576,-10.5454545454545)(836.040404040404,-7.09090909090909)(1164.48484848485,-8.90909090909091)(1433.21212121212,-9.27272727272727)(1313.77777777778,-7.63636363636364)(1239.13131313131,-10.5454545454545)(1089.83838383838,-7.45454545454546)(0,-5.81818181818182)(656.888888888889,-9.63636363636364)(298.585858585859,-8.54545454545455)(567.313131313131,-8.18181818181818)(1015.19191919192,-10.1818181818182)(1478,-8.36363636363636)(910.686868686869,0)(880.828282828283,-18)(1478,-10.5454545454545)(343.373737373737,-8)(850.969696969697,-9.45454545454546)(701.676767676768,-8)(343.373737373737,-9.09090909090909)(1000.26262626263,-7.63636363636364)(1358.56565656566,-10)(179.151515151515,-7.63636363636364)(1209.27272727273,-7.63636363636364)(1373.49494949495,-8.36363636363636)(477.737373737374,-7.81818181818182)(0,-13.2727272727273)(865.89898989899,-7.63636363636364)(1134.62626262626,-10.3636363636364)(1478,0)(701.676767676768,-8.72727272727273)(1433.21212121212,-9.63636363636364)(552.383838383838,-7.81818181818182)(567.313131313131,-7.45454545454546)(865.89898989899,-9.81818181818182)(507.59595959596,-8.90909090909091)(343.373737373737,-7.81818181818182)(791.252525252525,-7.81818181818182)(1030.12121212121,-10.1818181818182)(1104.76767676768,-7.45454545454546)(1343.63636363636,-8.18181818181818)(1358.56565656566,-10.1818181818182)(925.616161616162,-7.45454545454546)(1448.14141414141,-9.27272727272727)(418.020202020202,-8.90909090909091)(612.10101010101,-7.63636363636364)(776.323232323232,-8.90909090909091)(597.171717171717,-8.54545454545455)(910.686868686869,-10)(1194.34343434343,-7.63636363636364)(582.242424242424,-7.45454545454546)(432.949494949495,-7.63636363636364)(806.181818181818,-8)(1268.9898989899,-10.3636363636364)(1149.55555555556,-10.1818181818182)(1418.28282828283,-8.72727272727273)(1463.07070707071,-9.45454545454546)(821.111111111111,-9.45454545454546)(1254.06060606061,-7.81818181818182)(940.545454545455,-7.45454545454546)(1104.76767676768,-10.1818181818182)(388.161616161616,-8.54545454545455)(821.111111111111,-8.18181818181818)(552.383838383838,-8.72727272727273)(462.808080808081,-7.45454545454546)(1328.70707070707,-8)(1403.35353535354,-10)(925.616161616162,-10)(1074.90909090909,-7.45454545454546)(641.959595959596,-8.18181818181818)(1388.42424242424,-8.36363636363636)(373.232323232323,-8.18181818181818)(791.252525252525,-8.90909090909091)(582.242424242424,-8.72727272727273)(1179.41414141414,-7.63636363636364)(1059.9797979798,-10.1818181818182)(567.313131313131,-8.90909090909091)(1448.14141414141,-9.09090909090909)(880.828282828283,-7.63636363636364)(641.959595959596,-7.81818181818182)(477.737373737374,-8.90909090909091)(1254.06060606061,-10)(1239.13131313131,-10)(552.383838383838,-7.45454545454546)(850.969696969697,-9.63636363636364)(1149.55555555556,-7.63636363636364)(1059.9797979798,-7.45454545454546)(985.333333333333,-7.27272727272727)(597.171717171717,-7.45454545454546)(552.383838383838,-7.27272727272727)(403.090909090909,-7.81818181818182)(432.949494949495,-3.27272727272727)(194.080808080808,-8.36363636363636)(194.080808080808,-8.54545454545455)(1045.05050505051,-10.1818181818182)(1313.77777777778,-8)(1373.49494949495,-10)(1448.14141414141,-9.63636363636364)(1000.26262626263,-10.1818181818182)(582.242424242424,-8.90909090909091)(134.363636363636,-8)(44.7878787878788,-7.63636363636364)(179.151515151515,-8.18181818181818)(164.222222222222,-8.36363636363636)(223.939393939394,-8)(179.151515151515,-8.54545454545455)(179.151515151515,-8.72727272727273)(149.292929292929,-7.81818181818182)(119.434343434343,-8.18181818181818)(253.79797979798,-8)(209.010101010101,-7.81818181818182)(253.79797979798,-8.54545454545455)(418.020202020202,-8.54545454545455)(791.252525252525,-8.54545454545455)(850.969696969697,-7.81818181818182)(612.10101010101,-7.45454545454546)(806.181818181818,-9.27272727272727)(1239.13131313131,-7.81818181818182)(1478,-18)(940.545454545455,-10)(612.10101010101,-8.72727272727273)(1268.9898989899,-10)(641.959595959596,-8.36363636363636)(1433.21212121212,-9.81818181818182)(1463.07070707071,-9.63636363636364)(418.020202020202,-7.81818181818182)(1403.35353535354,-8.54545454545455)(1433.21212121212,-8.90909090909091)(164.222222222222,-8.54545454545455)(836.040404040404,-9.63636363636364)(447.878787878788,-7.63636363636364)(537.454545454546,-7.27272727272727)(268.727272727273,-8.36363636363636)(194.080808080808,-7.81818181818182)(268.727272727273,-8.54545454545455)(1134.62626262626,-7.63636363636364)(910.686868686869,-7.45454545454546)(970.40404040404,-7.27272727272727)(1000.26262626263,-7.27272727272727)(1015.19191919192,-7.27272727272727)(1015.19191919192,-7.09090909090909)(1358.56565656566,-8.18181818181818)(388.161616161616,-8.36363636363636)(492.666666666667,-8.90909090909091)(194.080808080808,-8.72727272727273)(194.080808080808,-8.90909090909091)(164.222222222222,-8.72727272727273)(179.151515151515,-8.90909090909091)(209.010101010101,-8.90909090909091)(164.222222222222,-8.90909090909091)(223.939393939394,-8.90909090909091)(119.434343434343,-8.36363636363636)(791.252525252525,-8.72727272727273)(627.030303030303,-7.63636363636364)(1224.20202020202,-7.81818181818182)(791.252525252525,-9.09090909090909)(1089.83838383838,-10.1818181818182)(806.181818181818,-8.18181818181818)(641.959595959596,-8)(821.111111111111,-8)(940.545454545455,-7.27272727272727)(1030.12121212121,-7.27272727272727)(403.090909090909,-8.36363636363636)(1418.28282828283,-10)(283.656565656566,-8.18181818181818)(268.727272727273,-8)(477.737373737374,-7.45454545454546)(1224.20202020202,-10)(1209.27272727273,-10)(1283.91919191919,-10)(134.363636363636,-8.54545454545455)(134.363636363636,-8.72727272727273)(791.252525252525,-8.36363636363636)(447.878787878788,-8.72727272727273)(955.474747474747,-7.27272727272727)(522.525252525253,-7.27272727272727)(268.727272727273,-8.72727272727273)(776.323232323232,-8.72727272727273)(627.030303030303,-8.54545454545455)(1209.27272727273,-7.81818181818182)(1448.14141414141,-9.81818181818182)(880.828282828283,-9.81818181818182)(253.79797979798,-8.72727272727273)(895.757575757576,-7.45454545454546)(149.292929292929,-8)(462.808080808081,-8.90909090909091)(298.585858585859,-8.36363636363636)(403.090909090909,-8)(388.161616161616,-8.18181818181818)(1074.90909090909,-10.1818181818182)(432.949494949495,-8.72727272727273)(1268.9898989899,-7.81818181818182)(119.434343434343,-8.54545454545455)(1283.91919191919,-7.81818181818182)(149.292929292929,-8.90909090909091)(298.585858585859,-8.18181818181818)(1194.34343434343,-10)(1388.42424242424,-10)(1343.63636363636,-10.1818181818182)(283.656565656566,-8.54545454545455)(1463.07070707071,-9.27272727272727)(836.040404040404,-7.81818181818182)(238.868686868687,-8)(552.383838383838,-8.90909090909091)(522.525252525253,-9.09090909090909)(507.59595959596,-9.09090909090909)(537.454545454546,-9.09090909090909)(492.666666666667,-9.09090909090909)(1478,-9.45454545454546)(641.959595959596,-8.54545454545455)(1343.63636363636,-8)(1045.05050505051,-7.27272727272727)(1119.69696969697,-7.63636363636364)(418.020202020202,-7.63636363636364)(119.434343434343,-8)(836.040404040404,-8)(134.363636363636,-8.18181818181818)(164.222222222222,-8)(806.181818181818,-9.45454545454546)(791.252525252525,-9.27272727272727)(850.969696969697,-9.81818181818182)(283.656565656566,-8.36363636363636)(388.161616161616,-8)(1373.49494949495,-8.18181818181818)(134.363636363636,-8.36363636363636)(671.818181818182,-14.3636363636364)(507.59595959596,-7.27272727272727)(567.313131313131,-7.27272727272727)(582.242424242424,-7.27272727272727)(447.878787878788,-7.45454545454546)(492.666666666667,-7.27272727272727)(1298.84848484848,-7.81818181818182)(597.171717171717,-8.90909090909091)(552.383838383838,-9.09090909090909)(627.030303030303,-8.72727272727273)(627.030303030303,-7.81818181818182)(955.474747474747,-10)(1119.69696969697,-10.1818181818182)(1313.77777777778,-7.81818181818182)(149.292929292929,-8.72727272727273)(1164.48484848485,-7.63636363636364)(1448.14141414141,-9.45454545454546)(1433.21212121212,-9.09090909090909)(1433.21212121212,-8.72727272727273)(1448.14141414141,-8.90909090909091)(403.090909090909,-8.54545454545455)(403.090909090909,-8.18181818181818)(477.737373737374,-7.27272727272727)(238.868686868687,-8.72727272727273)(268.727272727273,-8.18181818181818)(223.939393939394,-7.81818181818182)(656.888888888889,-8.18181818181818)(656.888888888889,-8)(776.323232323232,-8.54545454545455)(283.656565656566,-8)(1328.70707070707,-10.1818181818182)(432.949494949495,-8.54545454545455)(1418.28282828283,-8.54545454545455)(656.888888888889,-8.36363636363636)(1134.62626262626,-10.1818181818182)(418.020202020202,-8.36363636363636)(1179.41414141414,-10)(432.949494949495,-7.45454545454546)(179.151515151515,-8)(179.151515151515,-7.81818181818182)(1373.49494949495,-10.1818181818182)(597.171717171717,-7.27272727272727)(985.333333333333,-10.1818181818182)(238.868686868687,-8.90909090909091)(253.79797979798,-8.90909090909091)(821.111111111111,-9.63636363636364)(403.090909090909,-7.63636363636364)(418.020202020202,-8.18181818181818)(418.020202020202,-8)(641.959595959596,-7.63636363636364)(104.505050505051,-8.18181818181818)(104.505050505051,-8.36363636363636)(104.505050505051,-8)(89.5757575757576,-8.18181818181818)(1358.56565656566,-8)(209.010101010101,-8)(164.222222222222,-7.81818181818182)(1298.84848484848,-10)(1313.77777777778,-10.1818181818182)(970.40404040404,-10)(895.757575757576,-10)(1463.07070707071,-9.09090909090909)(925.616161616162,-7.27272727272727)(1403.35353535354,-8.36363636363636)(1388.42424242424,-8.18181818181818)(1328.70707070707,-7.81818181818182)(612.10101010101,-8.90909090909091)(447.878787878788,-8.90909090909091)(1059.9797979798,-7.27272727272727)(1418.28282828283,-9.81818181818182)(1194.34343434343,-7.81818181818182)(432.949494949495,-8.36363636363636)(627.030303030303,-7.45454545454546)(283.656565656566,-8.72727272727273)(119.434343434343,-8.72727272727273)(597.171717171717,-8.72727272727273)(238.868686868687,-7.81818181818182)(194.080808080808,-8)(776.323232323232,-9.09090909090909)(865.89898989899,-7.81818181818182)(761.393939393939,-8.72727272727273)(1298.84848484848,-10.1818181818182)(1164.48484848485,-10)(313.515151515152,-8.36363636363636)(253.79797979798,-8.18181818181818)(223.939393939394,-8.72727272727273)(462.808080808081,-8.72727272727273)(627.030303030303,-8.36363636363636)(656.888888888889,-7.81818181818182)(627.030303030303,-8)(1149.55555555556,-10)(1283.91919191919,-10.1818181818182)(209.010101010101,-9.09090909090909)(194.080808080808,-9.09090909090909)(223.939393939394,-9.09090909090909)(447.878787878788,-8.54545454545455)(134.363636363636,-7.81818181818182)(462.808080808081,-7.27272727272727)(1433.21212121212,-10)(432.949494949495,-7.81818181818182)(970.40404040404,-10.1818181818182)(1104.76767676768,-7.63636363636364)(253.79797979798,-8.36363636363636)(641.959595959596,-8.72727272727273)(462.808080808081,-9.09090909090909)(1373.49494949495,-8)(1134.62626262626,-10)(1343.63636363636,-10)(1119.69696969697,-10)(149.292929292929,-8.54545454545455)(149.292929292929,-8.18181818181818)(268.727272727273,-8.90909090909091)(1179.41414141414,-7.81818181818182)(89.5757575757576,-8)(985.333333333333,-10)(656.888888888889,-8.54545454545455)(1388.42424242424,-10.1818181818182)(776.323232323232,-8.36363636363636)(880.828282828283,-7.45454545454546)(1343.63636363636,-7.81818181818182) 
};
\end{axis}
\end{tikzpicture}%