### 2013-03-19 figurefactory/files/limno_bgape_class_40.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478.0002488174,-8.54545454545455)(1478.00049762651,-8.72727272727273)(1478.00074642733,-8.90909090909091)(1478.00074642733,-9.09090909090909)(1478.00074642733,-9.27272727272727)(1478.00074642733,-9.45454545454546)(1478.00074642733,-9.63636363636364)(1478.00074642733,-9.81818181818182)(1478.00058056104,-10)(1478.00033175469,-10.1818181818182)(1478.00008294006,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478.0002488174,-17.4545454545455)(1478.00049762651,-17.6363636363636)(1478.00074642733,-17.8181818181818)(1478.00074642733,-18)(1478,-18.0000090904546)(1463.07070707071,-18.0000090904546)(1448.14141414141,-18.0000090904546)(1433.21212121212,-18.0000090904546)(1418.28282828283,-18.0000090904546)(1403.35353535354,-18.000008080449)(1388.42424242424,-18.0000070704321)(1373.49494949495,-18.0000040403143)(1358.56565656566,-18.0000020201796)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(-0.000248817401863321,-8)(-0.000497626510092015,-7.81818181818182)(-0.000746427325097138,-7.63636363636364)(-0.000746427325097138,-7.45454545454546)(-0.000746427325097138,-7.27272727272727)(-0.000497626510092015,-7.09090909090909)(-0.000248817401863321,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(-0.000248817401863321,-3.81818181818182)(-0.000497626510092015,-3.63636363636364)(-0.000746427325097138,-3.45454545454546)(-0.000746427325097138,-3.27272727272727)(-0.000746427325097138,-3.09090909090909)(-0.000497626510092015,-2.90909090909091)(-0.000248817401863321,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-7)(14.9292929292929,-7)(29.8585858585859,-7)(44.7878787878788,-7)(59.7171717171717,-7)(74.6464646464647,-7)(89.5757575757576,-7)(104.505050505051,-6.93939393939394)(111.969696969697,-6.90909090909091)(119.434343434343,-6.87878787878788)(134.363636363636,-6.81818181818182)(149.292929292929,-6.75757575757576)(156.757575757576,-6.72727272727273)(164.222222222222,-6.6969696969697)(179.151515151515,-6.63636363636364)(194.080808080808,-6.57575757575758)(201.545454545455,-6.54545454545455)(209.010101010101,-6.51515151515151)(223.939393939394,-6.45454545454545)(238.868686868687,-6.45454545454545)(253.79797979798,-6.51515151515151)(261.262626262626,-6.54545454545455)(268.727272727273,-6.57575757575758)(283.656565656566,-6.6969696969697)(287.388888888889,-6.72727272727273)(298.585858585859,-6.81818181818182)(309.782828282828,-6.90909090909091)(313.515151515152,-6.93939393939394)(328.444444444444,-7.06060606060606)(335.909090909091,-7.09090909090909)(343.373737373737,-7.12121212121212)(358.30303030303,-7.18181818181818)(373.232323232323,-7.24242424242424)(380.69696969697,-7.27272727272727)(388.161616161616,-7.3030303030303)(403.090909090909,-7.36363636363636)(418.020202020202,-7.36363636363636)(432.949494949495,-7.3030303030303)(440.414141414141,-7.27272727272727)(447.878787878788,-7.24242424242424)(462.808080808081,-7.18181818181818)(477.737373737374,-7.18181818181818)(492.666666666667,-7.18181818181818)(507.59595959596,-7.18181818181818)(522.525252525253,-7.18181818181818)(537.454545454546,-7.18181818181818)(552.383838383838,-7.18181818181818)(567.313131313131,-7.18181818181818)(582.242424242424,-7.18181818181818)(597.171717171717,-7.18181818181818)(612.10101010101,-7.18181818181818)(627.030303030303,-7.18181818181818)(641.959595959596,-7.18181818181818)(656.888888888889,-7.24242424242424)(664.353535353535,-7.27272727272727)(671.818181818182,-7.3030303030303)(686.747474747475,-7.36363636363636)(701.676767676768,-7.36363636363636)(716.606060606061,-7.36363636363636)(731.535353535354,-7.36363636363636)(746.464646464646,-7.42424242424243)(753.929292929293,-7.45454545454546)(761.393939393939,-7.48484848484849)(776.323232323232,-7.54545454545455)(791.252525252525,-7.48484848484849)(798.717171717172,-7.45454545454546)(806.181818181818,-7.42424242424243)(821.111111111111,-7.36363636363636)(836.040404040404,-7.36363636363636)(850.969696969697,-7.34090909090909)(865.89898989899,-7.34090909090909)(880.828282828283,-7.34090909090909)(895.757575757576,-7.3030303030303)(903.222222222222,-7.27272727272727)(910.686868686869,-7.24242424242424)(925.616161616162,-7.18181818181818)(940.545454545455,-7.18181818181818)(955.474747474747,-7.18181818181818)(970.40404040404,-7.18181818181818)(985.333333333333,-7.18181818181818)(1000.26262626263,-7.12121212121212)(1007.72727272727,-7.09090909090909)(1015.19191919192,-7.06060606060606)(1030.12121212121,-7)(1045.05050505051,-7)(1059.9797979798,-7)(1074.90909090909,-7)(1089.83838383838,-7)(1104.76767676768,-7)(1119.69696969697,-7)(1134.62626262626,-7)(1149.55555555556,-7.06060606060606)(1157.0202020202,-7.09090909090909)(1164.48484848485,-7.12121212121212)(1179.41414141414,-7.18181818181818)(1194.34343434343,-7.18181818181818)(1209.27272727273,-7.24242424242424)(1216.73737373737,-7.27272727272727)(1224.20202020202,-7.31818181818182)(1239.13131313131,-7.36363636363636)(1254.06060606061,-7.36363636363636)(1268.9898989899,-7.42424242424243)(1276.45454545455,-7.45454545454546)(1283.91919191919,-7.48484848484849)(1298.84848484848,-7.60606060606061)(1313.77777777778,-7.60606060606061)(1321.24242424242,-7.63636363636364)(1328.70707070707,-7.66666666666667)(1343.63636363636,-7.66666666666667)(1358.56565656566,-7.78787878787879)(1366.0303030303,-7.81818181818182)(1373.49494949495,-7.84848484848485)(1388.42424242424,-7.96969696969697)(1392.15656565657,-8)(1403.35353535354,-8.09090909090909)(1414.5505050505,-8.18181818181818)(1418.28282828283,-8.21212121212121)(1433.21212121212,-8.33333333333333)(1436.94444444444,-8.36363636363636)(1448.14141414141,-8.45454545454545)(1459.33838383838,-8.54545454545455)(1463.07070707071,-8.57575757575757)(1478,-8.63636363636364)(1478.00012440663,-8.72727272727273)(1478.00037321366,-8.90909090909091)(1478.00037321366,-9.09090909090909)(1478.00037321366,-9.27272727272727)(1478.00037321366,-9.45454545454546)(1478.00037321366,-9.63636363636364)(1478.00037321366,-9.81818181818182)(1478.00020734323,-10)(1478,-10.1515151515152)(1470.53535353535,-10.1818181818182)(1463.07070707071,-10.2121212121212)(1448.14141414141,-10.3333333333333)(1440.67676767677,-10.3636363636364)(1433.21212121212,-10.3939393939394)(1418.28282828283,-10.4545454545455)(1403.35353535354,-10.5151515151515)(1395.88888888889,-10.5454545454545)(1388.42424242424,-10.5757575757576)(1373.49494949495,-10.6969696969697)(1366.0303030303,-10.7272727272727)(1358.56565656566,-10.7575757575758)(1343.63636363636,-10.8181818181818)(1328.70707070707,-10.8181818181818)(1313.77777777778,-10.8181818181818)(1298.84848484848,-10.8181818181818)(1283.91919191919,-10.8181818181818)(1268.9898989899,-10.7575757575758)(1265.25757575758,-10.7272727272727)(1254.06060606061,-10.6363636363636)(1239.13131313131,-10.5757575757576)(1224.20202020202,-10.5757575757576)(1209.27272727273,-10.6363636363636)(1194.34343434343,-10.6363636363636)(1179.41414141414,-10.6363636363636)(1164.48484848485,-10.6363636363636)(1149.55555555556,-10.5909090909091)(1134.62626262626,-10.5909090909091)(1119.69696969697,-10.5909090909091)(1104.76767676768,-10.6363636363636)(1089.83838383838,-10.6363636363636)(1074.90909090909,-10.6363636363636)(1059.9797979798,-10.5757575757576)(1052.51515151515,-10.5454545454545)(1045.05050505051,-10.5151515151515)(1030.12121212121,-10.4090909090909)(1015.19191919192,-10.4090909090909)(1000.26262626263,-10.4090909090909)(985.333333333333,-10.4545454545455)(970.40404040404,-10.4545454545455)(955.474747474747,-10.3939393939394)(948.010101010101,-10.3636363636364)(940.545454545455,-10.3333333333333)(925.616161616162,-10.2727272727273)(910.686868686869,-10.2727272727273)(895.757575757576,-10.2121212121212)(888.292929292929,-10.1818181818182)(880.828282828283,-10.1515151515152)(865.89898989899,-10.0909090909091)(850.969696969697,-10.0909090909091)(836.040404040404,-10.030303030303)(828.575757575758,-10)(821.111111111111,-9.96969696969697)(806.181818181818,-9.84848484848485)(798.717171717172,-9.81818181818182)(791.252525252525,-9.78787878787879)(776.323232323232,-9.66666666666667)(768.858585858586,-9.63636363636364)(761.393939393939,-9.60606060606061)(746.464646464646,-9.54545454545455)(731.535353535354,-9.48484848484848)(724.070707070707,-9.45454545454546)(716.606060606061,-9.42424242424243)(701.676767676768,-9.36363636363636)(686.747474747475,-9.36363636363636)(671.818181818182,-9.36363636363636)(656.888888888889,-9.3030303030303)(641.959595959596,-9.3030303030303)(627.030303030303,-9.3030303030303)(612.10101010101,-9.36363636363636)(597.171717171717,-9.36363636363636)(582.242424242424,-9.36363636363636)(567.313131313131,-9.36363636363636)(552.383838383838,-9.36363636363636)(537.454545454546,-9.36363636363636)(522.525252525253,-9.42424242424243)(515.060606060606,-9.45454545454546)(507.59595959596,-9.48484848484848)(492.666666666667,-9.54545454545455)(477.737373737374,-9.54545454545455)(462.808080808081,-9.54545454545455)(447.878787878788,-9.54545454545455)(432.949494949495,-9.54545454545455)(418.020202020202,-9.48484848484848)(410.555555555556,-9.45454545454546)(403.090909090909,-9.42424242424243)(388.161616161616,-9.36363636363636)(373.232323232323,-9.36363636363636)(358.30303030303,-9.31818181818182)(343.373737373737,-9.31818181818182)(328.444444444444,-9.31818181818182)(313.515151515152,-9.36363636363636)(298.585858585859,-9.36363636363636)(283.656565656566,-9.36363636363636)(268.727272727273,-9.42424242424243)(261.262626262626,-9.45454545454546)(253.79797979798,-9.48484848484848)(238.868686868687,-9.54545454545455)(223.939393939394,-9.60606060606061)(216.474747474747,-9.63636363636364)(209.010101010101,-9.66666666666667)(194.080808080808,-9.72727272727273)(179.151515151515,-9.72727272727273)(164.222222222222,-9.72727272727273)(149.292929292929,-9.72727272727273)(134.363636363636,-9.66666666666667)(130.631313131313,-9.63636363636364)(119.434343434343,-9.5)(116.946127946128,-9.45454545454546)(111.969696969697,-9.27272727272727)(111.969696969697,-9.09090909090909)(111.969696969697,-8.90909090909091)(111.969696969697,-8.72727272727273)(106.993265993266,-8.54545454545455)(104.505050505051,-8.5)(97.040404040404,-8.36363636363636)(89.5757575757576,-8.27272727272727)(78.3787878787879,-8.18181818181818)(74.6464646464647,-8.15151515151515)(59.7171717171717,-8.09090909090909)(44.7878787878788,-8.03030303030303)(37.3232323232323,-8)(29.8585858585859,-7.96969696969697)(14.9292929292929,-7.90909090909091)(0,-7.90909090909091)(-0.000124406627521346,-7.81818181818182)(-0.000373213662548569,-7.63636363636364)(-0.000373213662548569,-7.45454545454546)(-0.000373213662548569,-7.27272727272727)(-0.000124406627521346,-7.09090909090909)(0,-7)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-2.81818181818182)(14.9292929292929,-2.81818181818182)(29.8585858585859,-2.81818181818182)(44.7878787878788,-2.81818181818182)(59.7171717171717,-2.81818181818182)(74.6464646464647,-2.81818181818182)(89.5757575757576,-2.75757575757576)(97.040404040404,-2.72727272727273)(104.505050505051,-2.6969696969697)(119.434343434343,-2.63636363636364)(134.363636363636,-2.63636363636364)(149.292929292929,-2.63636363636364)(164.222222222222,-2.57575757575758)(171.686868686869,-2.54545454545455)(179.151515151515,-2.51515151515151)(194.080808080808,-2.45454545454545)(209.010101010101,-2.39393939393939)(216.474747474747,-2.36363636363636)(223.939393939394,-2.33333333333333)(238.868686868687,-2.27272727272727)(253.79797979798,-2.21212121212121)(261.262626262626,-2.18181818181818)(268.727272727273,-2.15151515151515)(283.656565656566,-2.09090909090909)(298.585858585859,-2.03030303030303)(306.050505050505,-2)(313.515151515152,-1.96969696969697)(328.444444444444,-1.90909090909091)(343.373737373737,-1.84848484848485)(350.838383838384,-1.81818181818182)(358.30303030303,-1.78787878787879)(373.232323232323,-1.72727272727273)(388.161616161616,-1.72727272727273)(403.090909090909,-1.72727272727273)(418.020202020202,-1.72727272727273)(432.949494949495,-1.72727272727273)(447.878787878788,-1.72727272727273)(462.808080808081,-1.72727272727273)(477.737373737374,-1.78787878787879)(485.20202020202,-1.81818181818182)(492.666666666667,-1.84848484848485)(507.59595959596,-1.90909090909091)(522.525252525253,-1.96969696969697)(526.257575757576,-2)(537.454545454546,-2.09090909090909)(548.651515151515,-2.18181818181818)(552.383838383838,-2.21212121212121)(567.313131313131,-2.33333333333333)(571.045454545455,-2.36363636363636)(582.242424242424,-2.45454545454545)(589.707070707071,-2.54545454545455)(597.171717171717,-2.63636363636364)(604.636363636364,-2.72727272727273)(612.10101010101,-2.81818181818182)(619.565656565657,-2.90909090909091)(627.030303030303,-3.04545454545455)(629.518518518518,-3.09090909090909)(634.494949494949,-3.27272727272727)(629.518518518518,-3.45454545454546)(627.030303030303,-3.5)(615.833333333333,-3.63636363636364)(612.10101010101,-3.66666666666667)(597.171717171717,-3.78787878787879)(589.707070707071,-3.81818181818182)(582.242424242424,-3.84848484848485)(567.313131313131,-3.90909090909091)(552.383838383838,-3.96969696969697)(544.919191919192,-4)(537.454545454546,-4.03030303030303)(522.525252525253,-4.09090909090909)(507.59595959596,-4.15151515151515)(500.131313131313,-4.18181818181818)(492.666666666667,-4.21212121212121)(477.737373737374,-4.27272727272727)(462.808080808081,-4.27272727272727)(447.878787878788,-4.33333333333333)(440.414141414141,-4.36363636363636)(432.949494949495,-4.39393939393939)(418.020202020202,-4.45454545454545)(403.090909090909,-4.45454545454545)(388.161616161616,-4.51515151515151)(380.69696969697,-4.54545454545455)(373.232323232323,-4.57575757575758)(358.30303030303,-4.63636363636364)(343.373737373737,-4.63636363636364)(328.444444444444,-4.63636363636364)(313.515151515152,-4.6969696969697)(306.050505050505,-4.72727272727273)(298.585858585859,-4.75757575757576)(283.656565656566,-4.75757575757576)(276.191919191919,-4.72727272727273)(268.727272727273,-4.6969696969697)(253.79797979798,-4.63636363636364)(238.868686868687,-4.63636363636364)(223.939393939394,-4.57575757575758)(216.474747474747,-4.54545454545455)(209.010101010101,-4.51515151515151)(194.080808080808,-4.39393939393939)(186.616161616162,-4.36363636363636)(179.151515151515,-4.33333333333333)(164.222222222222,-4.21212121212121)(156.757575757576,-4.18181818181818)(149.292929292929,-4.15151515151515)(134.363636363636,-4.09090909090909)(119.434343434343,-4.03030303030303)(111.969696969697,-4)(104.505050505051,-3.96969696969697)(89.5757575757576,-3.90909090909091)(74.6464646464647,-3.84848484848485)(67.1818181818182,-3.81818181818182)(59.7171717171717,-3.78787878787879)(44.7878787878788,-3.72727272727273)(29.8585858585859,-3.72727272727273)(14.9292929292929,-3.72727272727273)(0,-3.72727272727273)(-0.000124406627521346,-3.63636363636364)(-0.000373213662548569,-3.45454545454546)(-0.000373213662548569,-3.27272727272727)(-0.000373213662548569,-3.09090909090909)(-0.000124406627521346,-2.90909090909091)(0,-2.81818181818182)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1343.63636363636,-9.18181818181818)(1346.12457912458,-9.27272727272727)(1343.63636363636,-9.36363636363636)(1341.14814814815,-9.45454545454546)(1328.70707070707,-9.60606060606061)(1321.24242424242,-9.63636363636364)(1313.77777777778,-9.66666666666667)(1298.84848484848,-9.72727272727273)(1283.91919191919,-9.72727272727273)(1268.9898989899,-9.72727272727273)(1254.06060606061,-9.78787878787879)(1246.59595959596,-9.81818181818182)(1239.13131313131,-9.84848484848485)(1224.20202020202,-9.90909090909091)(1209.27272727273,-9.90909090909091)(1194.34343434343,-9.90909090909091)(1179.41414141414,-9.90909090909091)(1164.48484848485,-9.90909090909091)(1149.55555555556,-9.90909090909091)(1134.62626262626,-9.90909090909091)(1119.69696969697,-9.90909090909091)(1104.76767676768,-9.90909090909091)(1089.83838383838,-9.90909090909091)(1074.90909090909,-9.90909090909091)(1059.9797979798,-9.90909090909091)(1045.05050505051,-9.90909090909091)(1030.12121212121,-9.88636363636364)(1015.19191919192,-9.84090909090909)(1007.72727272727,-9.81818181818182)(1000.26262626263,-9.78787878787879)(985.333333333333,-9.72727272727273)(970.40404040404,-9.72727272727273)(955.474747474747,-9.72727272727273)(940.545454545455,-9.72727272727273)(925.616161616162,-9.66666666666667)(918.151515151515,-9.63636363636364)(910.686868686869,-9.60606060606061)(895.757575757576,-9.54545454545455)(880.828282828283,-9.48484848484848)(865.89898989899,-9.48484848484848)(862.166666666667,-9.45454545454546)(850.969696969697,-9.36363636363636)(836.040404040404,-9.3030303030303)(833.552188552189,-9.27272727272727)(821.111111111111,-9.12121212121212)(817.378787878788,-9.09090909090909)(806.181818181818,-8.95454545454546)(803.693602693603,-8.90909090909091)(798.717171717172,-8.72727272727273)(798.717171717172,-8.54545454545455)(806.181818181818,-8.40909090909091)(808.670033670034,-8.36363636363636)(821.111111111111,-8.21212121212121)(824.843434343434,-8.18181818181818)(836.040404040404,-8.09090909090909)(847.237373737374,-8)(850.969696969697,-7.95454545454546)(865.89898989899,-7.86363636363636)(873.363636363636,-7.81818181818182)(880.828282828283,-7.78787878787879)(895.757575757576,-7.78787878787879)(910.686868686869,-7.72727272727273)(925.616161616162,-7.72727272727273)(940.545454545455,-7.72727272727273)(955.474747474747,-7.72727272727273)(970.40404040404,-7.72727272727273)(985.333333333333,-7.66666666666667)(1000.26262626263,-7.66666666666667)(1015.19191919192,-7.66666666666667)(1030.12121212121,-7.72727272727273)(1045.05050505051,-7.72727272727273)(1059.9797979798,-7.72727272727273)(1074.90909090909,-7.75)(1089.83838383838,-7.75)(1104.76767676768,-7.75)(1119.69696969697,-7.78787878787879)(1127.16161616162,-7.81818181818182)(1134.62626262626,-7.84848484848485)(1149.55555555556,-7.90909090909091)(1164.48484848485,-7.90909090909091)(1179.41414141414,-7.96969696969697)(1186.87878787879,-8)(1194.34343434343,-8.03030303030303)(1209.27272727273,-8.09090909090909)(1224.20202020202,-8.09090909090909)(1239.13131313131,-8.15151515151515)(1246.59595959596,-8.18181818181818)(1254.06060606061,-8.21212121212121)(1268.9898989899,-8.33333333333333)(1272.72222222222,-8.36363636363636)(1283.91919191919,-8.45454545454545)(1295.11616161616,-8.54545454545455)(1298.84848484848,-8.57575757575757)(1313.77777777778,-8.6969696969697)(1317.5101010101,-8.72727272727273)(1328.70707070707,-8.86363636363636)(1331.19528619529,-8.90909090909091)(1341.14814814815,-9.09090909090909)(1343.63636363636,-9.18181818181818)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (462.808080808081,-14.6818181818182)(470.272727272727,-14.5454545454545)(477.737373737374,-14.4545454545455)(488.934343434343,-14.3636363636364)(492.666666666667,-14.3333333333333)(507.59595959596,-14.2727272727273)(522.525252525253,-14.2727272727273)(537.454545454546,-14.2727272727273)(552.383838383838,-14.2727272727273)(567.313131313131,-14.2727272727273)(582.242424242424,-14.2727272727273)(597.171717171717,-14.2727272727273)(612.10101010101,-14.2727272727273)(627.030303030303,-14.2727272727273)(641.959595959596,-14.2727272727273)(656.888888888889,-14.2727272727273)(671.818181818182,-14.2727272727273)(686.747474747475,-14.3181818181818)(690.479797979798,-14.3636363636364)(690.479797979798,-14.5454545454545)(686.747474747475,-14.5757575757576)(671.818181818182,-14.6969696969697)(664.353535353535,-14.7272727272727)(656.888888888889,-14.7575757575758)(641.959595959596,-14.8787878787879)(634.494949494949,-14.9090909090909)(627.030303030303,-14.9393939393939)(612.10101010101,-15.0606060606061)(604.636363636364,-15.0909090909091)(597.171717171717,-15.1212121212121)(582.242424242424,-15.1818181818182)(567.313131313131,-15.2424242424242)(559.848484848485,-15.2727272727273)(552.383838383838,-15.3030303030303)(537.454545454546,-15.3636363636364)(522.525252525253,-15.3636363636364)(507.59595959596,-15.3636363636364)(492.666666666667,-15.3636363636364)(477.737373737374,-15.3030303030303)(474.005050505051,-15.2727272727273)(462.808080808081,-15.1363636363636)(460.319865319865,-15.0909090909091)(455.343434343434,-14.9090909090909)(460.319865319865,-14.7272727272727)(462.808080808081,-14.6818181818182)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1478,-17.5454545454545)(1478.00012440663,-17.6363636363636)(1478.00037321366,-17.8181818181818)(1478.00037321366,-18)(1478,-18.0000045452273)(1463.07070707071,-18.0000045452273)(1448.14141414141,-18.0000045452273)(1433.21212121212,-18.0000045452273)(1418.28282828283,-18.0000045452273)(1403.35353535354,-18.0000035351964)(1388.42424242424,-18.0000025251543)(1375.98316498317,-18)(1388.42424242424,-17.8484848484848)(1395.88888888889,-17.8181818181818)(1403.35353535354,-17.7878787878788)(1418.28282828283,-17.7272727272727)(1433.21212121212,-17.7272727272727)(1448.14141414141,-17.6666666666667)(1455.60606060606,-17.6363636363636)(1463.07070707071,-17.6060606060606)(1478,-17.5454545454545)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (1358.56565656566,-4.54545454545455)(0,-18)(0,0)(1478,-14.7272727272727)(0,-8.72727272727273)(477.737373737374,-9.09090909090909)(432.949494949495,-6.90909090909091)(358.30303030303,-11.2727272727273)(806.181818181818,-8.72727272727273)(895.757575757576,-10.5454545454545)(836.040404040404,-7.09090909090909)(1164.48484848485,-8.90909090909091)(1433.21212121212,-9.27272727272727)(1313.77777777778,-7.63636363636364)(1239.13131313131,-10.5454545454545)(1089.83838383838,-7.45454545454546)(0,-5.81818181818182)(656.888888888889,-9.63636363636364)(298.585858585859,-8.54545454545455)(567.313131313131,-8.18181818181818)(1015.19191919192,-10.1818181818182)(1478,-8.36363636363636)(910.686868686869,0)(880.828282828283,-18)(1478,-10.5454545454545)(343.373737373737,-8)(850.969696969697,-9.45454545454546)(701.676767676768,-8)(343.373737373737,-9.09090909090909)(1000.26262626263,-7.63636363636364)(1358.56565656566,-10)(179.151515151515,-7.63636363636364)(1209.27272727273,-7.63636363636364)(1373.49494949495,-8.36363636363636)(477.737373737374,-7.81818181818182)(0,-13.2727272727273)(865.89898989899,-7.63636363636364)(1134.62626262626,-10.3636363636364)(1478,0)(701.676767676768,-8.72727272727273) 
};
\end{axis}
\end{tikzpicture}%