### 2013-03-20 figurefactory/files/ev_bgape_eps.tex
LSE theory and experiments

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{mycolor1}{rgb}{1,0.04,0}
\definecolor{mycolor2}{rgb}{1,0.08,0}
\definecolor{mycolor3}{rgb}{1,0.12,0}
\definecolor{mycolor4}{rgb}{1,0.16,0}
\definecolor{mycolor5}{rgb}{1,0.2,0}
\definecolor{mycolor6}{rgb}{1,0.24,0}
\definecolor{mycolor7}{rgb}{1,0.28,0}
\definecolor{mycolor8}{rgb}{1,0.32,0}
\definecolor{mycolor9}{rgb}{1,0.36,0}
\definecolor{mycolor10}{rgb}{1,0.4,0}
\definecolor{mycolor11}{rgb}{1,0.44,0}
\definecolor{mycolor12}{rgb}{1,0.48,0}
\definecolor{mycolor13}{rgb}{1,0.52,0}
\definecolor{mycolor14}{rgb}{1,0.56,0}
\definecolor{mycolor15}{rgb}{1,0.6,0}
\definecolor{mycolor16}{rgb}{1,0.68,0}
\definecolor{mycolor17}{rgb}{1,0.72,0}
\definecolor{mycolor18}{rgb}{1,0.76,0}
\definecolor{mycolor19}{rgb}{1,0.84,0}
\definecolor{mycolor20}{rgb}{1,0.88,0}
\definecolor{mycolor21}{rgb}{1,0.96,0}
\definecolor{mycolor22}{rgb}{1,1,0}
\definecolor{mycolor23}{rgb}{0.92,1,0.08}
\definecolor{mycolor24}{rgb}{0.84,1,0.16}
\definecolor{mycolor25}{rgb}{0.76,1,0.24}
\definecolor{mycolor26}{rgb}{0.68,1,0.32}
\definecolor{mycolor27}{rgb}{0.6,1,0.4}
\definecolor{mycolor28}{rgb}{0.52,1,0.48}
\definecolor{mycolor29}{rgb}{0.44,1,0.56}
\definecolor{mycolor30}{rgb}{0.36,1,0.64}
\definecolor{mycolor31}{rgb}{0.24,1,0.76}
\definecolor{mycolor32}{rgb}{0.16,1,0.84}
\definecolor{mycolor33}{rgb}{0.04,1,0.96}
\definecolor{mycolor34}{rgb}{0,0.92,1}
\definecolor{mycolor35}{rgb}{0,0.8,1}
\definecolor{mycolor36}{rgb}{0,0.68,1}
\definecolor{mycolor37}{rgb}{0,0.56,1}
\definecolor{mycolor38}{rgb}{0,0.44,1}
\definecolor{mycolor39}{rgb}{0,0.28,1}
\definecolor{mycolor40}{rgb}{0,0.12,1}
\definecolor{mycolor41}{rgb}{0,0.52,1}
\definecolor{mycolor42}{rgb}{0,0.48,1}
\definecolor{mycolor43}{rgb}{0,0.4,1}
\definecolor{mycolor44}{rgb}{0,0.36,1}
\definecolor{mycolor45}{rgb}{0,0.32,1}
\definecolor{mycolor46}{rgb}{0,0.24,1}
\definecolor{mycolor47}{rgb}{0,0.2,1}
\definecolor{mycolor48}{rgb}{0,0.16,1}
\definecolor{mycolor49}{rgb}{0,0.08,1}
\definecolor{mycolor50}{rgb}{0,0.04,1}
\definecolor{mycolor51}{rgb}{1,0.64,0}
\definecolor{mycolor52}{rgb}{0.96,1,0.04}
\definecolor{mycolor53}{rgb}{0.88,1,0.12}
\definecolor{mycolor54}{rgb}{0.8,1,0.2}
\definecolor{mycolor55}{rgb}{0.72,1,0.28}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-6pt},
ylabel shift={-4pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=400,
xlabel={Samples},
ymin=0.5, ymax=1,
ylabel={$F_1$-score},
name=plot1,
axis lines*=left,
colormap={mymap}{[1pt] rgb(0pt)=(0.5,0,0); rgb(8pt)=(1,0,0); rgb(24pt)=(1,1,0); rgb(40pt)=(0,1,1); rgb(56pt)=(0,0,1); rgb(63pt)=(0,0,0.5625)},
colorbar right,
colorbar style={
  tickwidth=0.03cm,
  at={(1.06, 0)},
  anchor=south east,
  yticklabel pos=right,
  yticklabel shift={-2pt}
},
tickwidth=0.1cm,
colorbar/width=0.15cm,
point meta min=0,
point meta max=2
]

\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (294,0.930862437633642) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (352,0.974216027874564) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (308,0.960111966410077) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (258,0.885196374622356) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.942997888810697) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (350,0.965034965034965) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (281,0.945454545454545) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (360,0.974107767669699) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (298,0.951219512195122) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (305,0.972515856236786) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (230,0.867078825347759) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (226,0.878971255673222) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (316,0.953439888811675) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (214,0.876712328767123) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (276,0.956951305575159) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.947220267417312) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.952843273231623) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (285,0.961884961884962) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (189,0.879818594104308) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (261,0.945404284727021) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (233,0.951269732326699) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.896860986547085) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (251,0.963838664812239) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.926898509581263) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (213,0.952247191011236) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (159,0.877488514548239) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.936319104268719) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (190,0.939958592132505) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (175,0.913105413105413) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (128,0.86697247706422) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (135,0.856278366111951) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (143,0.917018284106892) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (108,0.862147753236862) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (102,0.842105263157895) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.922865013774105) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (92,0.908821349147517) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (98,0.864216054013503) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (99,0.885409252669039) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (82,0.849438202247191) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (85,0.930428671820098) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (70,0.879219804951238) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (79,0.874243443174176) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.883755588673621) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (55,0.815052041633307) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.85912560721721) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (52,0.851851851851852) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (55,0.823275862068965) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.803951367781155) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.834905660377358) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.847107438016529) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.789259560618389) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.832298136645963) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.858564321250888) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.859408795962509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.813404417364813) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.771747211895911) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.78893280632411) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.769230769230769) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.763840224246671) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.776377952755905) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.760746147607461) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.760655737704918) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.768333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.784873949579832) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.73195020746888) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.728129205921938) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.745874587458746) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.766265060240964) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (355,0.966433566433566) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (372,0.970159611380985) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (364,0.962095106822881) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (354,0.967877094972067) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (330,0.964754664823773) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (327,0.968728283530229) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (371,0.966461327857632) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (331,0.964509394572025) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (237,0.881559220389805) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (353,0.959666203059805) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (301,0.944793850454228) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (325,0.959778085991678) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (325,0.948986722571628) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (280,0.945086705202312) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (221,0.874536005939124) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (226,0.928011404133999) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (266,0.946938775510204) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (200,0.933914306463326) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (166,0.919540229885057) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (183,0.907246376811594) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (187,0.877643504531722) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (215,0.945908460471567) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (230,0.948251748251748) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (226,0.953179594689029) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (213,0.938186813186813) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (184,0.943181818181818) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (147,0.907246376811594) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (176,0.936140350877193) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (165,0.929190751445086) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (160,0.929235167977126) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (163,0.950102529049897) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (168,0.930679478380233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (133,0.917915774446823) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (114,0.881638846737481) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (119,0.930826057756884) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (80,0.818110850897736) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (86,0.876371616678859) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (80,0.867415730337079) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (94,0.93183415319747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.905606813342796) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.899470899470899) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.90146750524109) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.830135039090263) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (59,0.85348506401138) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.829268292682927) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.820705176294073) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.845814977973568) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (50,0.836473247927656) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.833333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.835481425322214) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.813008130081301) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.845986984815618) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.852805280528053) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.810566037735849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.802684563758389) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.81413612565445) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.817219477769936) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.814555633310007) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.789515488482923) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.79409594095941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.769356153219234) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.79288256227758) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.776957163958641) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.724842767295597) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.789598108747045) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.763212079615649) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.819464033850493) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.54728370221328) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (387,0.95859649122807) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (348,0.961702127659574) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (275,0.876233864844343) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.932559825960841) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (358,0.965034965034965) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.946191474493361) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (263,0.92463768115942) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.956829440905874) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (314,0.954766875434934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (323,0.953389830508474) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (263,0.926300578034682) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (329,0.96056338028169) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.959056210964608) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.960339943342776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (308,0.957182320441989) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (285,0.954321855235418) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.953191489361702) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (204,0.918681318681319) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (258,0.940925266903915) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.876795162509448) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.947001394700139) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (203,0.924652523774689) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (187,0.917888563049853) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (213,0.952112676056338) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (230,0.948936170212766) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (207,0.938628158844765) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (194,0.944674965421853) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (148,0.921511627906977) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (166,0.92600422832981) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.879939209726444) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (182,0.944979367262723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (149,0.905109489051095) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (129,0.913868613138686) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (132,0.937588652482269) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.90050107372942) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (97,0.841221374045801) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (93,0.853374709076804) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (88,0.843822843822844) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (89,0.921052631578947) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.885838150289017) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (74,0.884241971620612) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (79,0.903724525650035) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.863768115942029) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (66,0.875280059746079) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.877167630057803) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (60,0.844137931034483) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (56,0.888730385164051) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.869433962264151) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.818722139673105) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.875524475524475) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.840245775729647) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.825590251332825) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.813559322033898) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.79126213592233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.809160305343511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.802896218825422) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.805687203791469) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.7808) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.786610878661088) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.764350453172205) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.78475935828877) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.785046728971962) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.788083953960731) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.811755361397935) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.696673189823875) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.806646525679758) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.788888888888889) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.785829307568438) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (374,0.962649753347428) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (338,0.97194950911641) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (345,0.963380281690141) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (362,0.956645344705046) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (365,0.968197879858657) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (340,0.962754743499649) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (343,0.959212376933896) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (334,0.959497206703911) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.933621933621934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (296,0.926007326007326) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (266,0.942628903413217) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.93371757925072) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (313,0.960839160839161) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (244,0.938040345821326) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.961379310344827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (255,0.934296028880866) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (232,0.935740072202166) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.951983298538622) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (204,0.880061115355233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.941761363636364) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (218,0.897744360902256) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (275,0.964260686755431) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.903846153846154) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.938804895608351) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (233,0.956036287508723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (172,0.923076923076923) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (210,0.961779013203613) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (181,0.909479686386315) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (188,0.949224259520451) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (169,0.950071326676177) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (171,0.96078431372549) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (161,0.945736434108527) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (148,0.936376210235131) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (146,0.935817805383023) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (116,0.925631768953068) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (105,0.914487632508834) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (94,0.90961262553802) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (97,0.905003625815808) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (90,0.895774647887324) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (90,0.900808229243203) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.836573074154068) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.835666912306559) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.853675945753034) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.884798909338787) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.823346303501945) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (59,0.831426392067124) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.826291079812207) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.839524517087667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (48,0.826706676669167) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (41,0.843465045592705) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.813450760608487) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.804705882352941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.808176100628931) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.793774319066148) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.810879190385832) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.82641252552757) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.805194805194805) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.813559322033898) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.761835396941005) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.810727969348659) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.809930178432894) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.793080505655356) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.822238478419898) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.805890227576974) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.789350039154268) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.785581106277191) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.759136212624585) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.766476810414971) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (343,0.964234620886981) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.954220314735336) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (311,0.936635105608157) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (306,0.967329545454545) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (307,0.94201861130995) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (336,0.94592645998558) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (323,0.960507757404795) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (262,0.938271604938271) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (251,0.917325664989216) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (314,0.954022988505747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.94608195542775) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.946280991735537) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.941007194244604) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (290,0.961918194640338) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.950819672131147) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (291,0.962078651685393) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (231,0.933235509904622) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (231,0.929824561403509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (209,0.918639053254438) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.951902368987796) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (233,0.953439888811675) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (262,0.957142857142857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (236,0.948699929725931) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (249,0.957507082152974) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.886706948640483) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (168,0.920704845814978) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (160,0.924198250728863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (180,0.930635838150289) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (163,0.930664760543245) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (143,0.92557111274871) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (162,0.932871972318339) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (138,0.920634920634921) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (139,0.943342776203966) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (134,0.928208846990573) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (100,0.864446165762975) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (101,0.841221374045801) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (101,0.869629629629629) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (89,0.893862815884476) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (84,0.901387874360847) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (81,0.874051593323217) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.913528591352859) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (77,0.897079276773296) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (60,0.880239520958084) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (57,0.87832973362131) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.797720797720798) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.88) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.814132104454685) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.796516231195566) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.890921885995777) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.872135994087213) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.832535885167464) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.827862289831865) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.880056777856636) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.81037037037037) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.81259842519685) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.778089887640449) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.791666666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.809523809523809) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.84012539184953) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.811614730878187) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.827169811320755) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.815450643776824) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.843490304709141) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.733676975945017) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.817365269461078) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.76030534351145) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.503267973856209) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.701214574898785) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (350,0.964936886395512) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (269,0.885022692889561) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (360,0.967559943582511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (248,0.885196374622356) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (202,0.900221729490022) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (333,0.96575821104123) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (352,0.963687150837989) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (302,0.963585434173669) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (323,0.959717314487632) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (246,0.951167728237792) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (217,0.929248723559446) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (268,0.934971098265896) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (256,0.940841054882395) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (331,0.97071129707113) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (269,0.972554539057002) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (284,0.960334029227557) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (280,0.962237762237762) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (274,0.955602536997886) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (220,0.925952552120776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (220,0.932374100719424) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (217,0.946695095948827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (238,0.965901183020181) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (163,0.884498480243161) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (218,0.948699929725931) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (207,0.946022727272727) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (198,0.951236749116608) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (194,0.942512420156139) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (154,0.934876989869754) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (154,0.944323933477946) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (190,0.956583629893238) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (168,0.952447552447552) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (117,0.893961708394698) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (93,0.894930198383541) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (123,0.915523465703971) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (102,0.867017280240421) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (103,0.888272033310201) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (92,0.869899923017706) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (96,0.913597733711048) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (79,0.840513983371126) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (66,0.824561403508772) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (74,0.884502923976608) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (72,0.895144264602393) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.885072655217966) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.876534296028881) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.888106966924701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.847250509164969) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.801872074882995) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (48,0.851421188630491) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.802873104549082) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.834258524980174) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.814927646610815) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.861493836113125) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.818466353677621) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.833222591362126) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.804861580013504) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.828685258964143) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.773712737127371) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.767326732673267) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.809005083514887) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.774603174603175) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.7859375) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.804012345679012) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.802666666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.749022673964034) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.724046140195208) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.689313517338995) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.725352112676056) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.414866032843561) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (279,0.928930366116296) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (356,0.954577218728162) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (362,0.955369595536959) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.955801104972376) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (385,0.956224350205198) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (309,0.938833570412518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (300,0.944405348346235) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (274,0.926164874551971) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (329,0.945224719101123) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.931929824561403) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (355,0.956461644782308) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (257,0.924177396280401) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (335,0.955431754874652) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (266,0.935988620199146) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.949480968858131) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (262,0.957550452331246) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (281,0.948824343015214) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (298,0.96551724137931) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (249,0.943213296398892) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (234,0.941836019621584) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (256,0.948311509303928) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (254,0.942657342657343) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (192,0.868035190615836) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (228,0.943632567849687) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.923516797712652) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (212,0.923076923076923) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (161,0.922190201729106) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (181,0.922859164897381) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.947960618846695) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (151,0.915645277577505) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (148,0.865470852017937) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (139,0.928923293455313) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (143,0.92) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (144,0.934886908841672) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (105,0.888888888888889) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (98,0.889374090247453) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (108,0.930327868852459) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (90,0.908037653874004) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (84,0.893430656934307) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (69,0.835463258785942) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (71,0.887060583395662) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.816358024691358) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (65,0.858433734939759) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (60,0.831360946745562) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (58,0.821939586645469) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.836193447737909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.84692417739628) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (45,0.802091112770724) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.794952681388013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.78975131876413) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.795509222133119) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.838056680161943) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.802996254681648) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.825816485225505) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.836532097948378) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.781542898341745) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.811111111111111) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.806842480399144) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.813507290867229) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.817610062893082) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.768) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.802278275020342) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.782546494992847) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.810483870967742) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.764627659574468) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.772486772486773) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.757249378624689) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.081180811808118) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (348,0.959326788218794) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (359,0.961864406779661) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (378,0.968242766407904) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (364,0.959660297239915) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (304,0.951977401129943) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (350,0.960227272727273) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (314,0.948884089272858) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (268,0.93644996347699) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.955903271692745) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (246,0.932748538011696) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (230,0.928571428571428) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (290,0.959553695955369) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.923976608187134) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (316,0.963752665245202) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.96113074204947) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (249,0.941092973740241) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.878533231474408) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (237,0.934105720492397) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (237,0.936727272727273) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (187,0.878347360367253) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (265,0.9593837535014) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.952313167259786) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (218,0.941595441595441) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (206,0.934936350777935) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (164,0.877566539923954) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.949681077250177) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (174,0.92836676217765) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (204,0.946695095948827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (142,0.859714928732183) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (134,0.922302158273381) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (173,0.940677966101695) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (156,0.933428775948461) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (152,0.934191702432046) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (88,0.83015873015873) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (112,0.896151053013798) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (92,0.881670533642691) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (112,0.935805991440799) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (78,0.892512972572276) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (90,0.915679442508711) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (77,0.897168405365127) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (74,0.881458966565349) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (66,0.883211678832117) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (65,0.877840909090909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.898509581263307) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (57,0.861731843575419) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (61,0.871259568545581) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.869318181818182) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.847430830039526) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.814756671899529) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.814814814814815) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.800656275635767) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.795819935691318) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.813688212927757) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.845605700712589) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.831967213114754) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.821917808219178) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.808444096950743) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.7936) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.789712556732224) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.778435239973701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.793650793650794) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.79874213836478) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.819007686932215) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.785333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.782292298362644) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.717832957110609) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.805946791862285) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.70697012802276) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (324,0.965322009907997) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (344,0.954967834167262) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (306,0.952116585704372) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (380,0.961702127659574) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (294,0.959603118355776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (254,0.928312816799421) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (324,0.96218487394958) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (309,0.947444204463643) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.96011396011396) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (275,0.951359084406295) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (293,0.950749464668094) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (298,0.955523672883788) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (291,0.959603118355776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (289,0.951236749116608) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (296,0.950634696755994) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (203,0.870109546165884) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (251,0.941672522839072) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (198,0.878274268104776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (226,0.920883820384889) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (246,0.947519769949676) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (170,0.910029498525074) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (228,0.952924393723252) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (223,0.951841359773371) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (249,0.956583629893238) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (158,0.87024087024087) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (202,0.954738330975955) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (148,0.860182370820669) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (184,0.915032679738562) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.860046911649726) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (160,0.926007326007326) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (131,0.861154446177847) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (138,0.861989205859676) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (133,0.92138133725202) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (117,0.922855082912761) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (111,0.903597122302158) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (92,0.902616279069767) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (109,0.928923293455313) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (93,0.892392049348869) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (98,0.935323383084577) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.899042004421518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (83,0.923829489867226) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (65,0.886750555144337) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (70,0.880116959064327) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (66,0.867292225201072) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (60,0.865609924190214) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.815642458100559) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (48,0.799365582870737) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (48,0.824742268041237) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.819512195121951) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.819018404907975) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.82648401826484) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.783387622149837) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.795962509012257) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.773246329526917) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.819334389857369) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.738095238095238) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.78010878010878) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.771565495207668) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.754601226993865) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.773480662983425) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.723842195540309) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.788663967611336) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.773584905660377) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.727117194183062) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.770359501100513) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.755212355212355) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.718446601941747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.673060156931125) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (325,0.95591322603219) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (302,0.952654232424677) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (309,0.958512160228898) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.955287437899219) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.954415954415954) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (332,0.968421052631579) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (327,0.972593113141251) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.97191011235955) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (337,0.957887223411849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (256,0.941944847605225) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (280,0.958185683912119) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.951029098651526) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (232,0.877030162412993) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (222,0.945165945165945) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (198,0.880860876249039) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (257,0.947444204463643) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (204,0.871439568899153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (303,0.969187675070028) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (204,0.937226277372263) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.961294862772695) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (191,0.940493468795356) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (209,0.935272727272727) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (164,0.909090909090909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (227,0.950508474576271) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (185,0.932851985559567) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (164,0.875576036866359) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.951305575158786) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (201,0.951497860199715) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (140,0.881996974281392) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (146,0.950980392156863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (179,0.958827634333566) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (122,0.932153392330383) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (115,0.865900383141762) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (138,0.929264909847434) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (107,0.896408839779005) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (107,0.909336941813261) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.869961977186312) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (80,0.860700389105058) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (84,0.86969696969697) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.80521597392013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.89900426742532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor14,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.892957746478873) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (78,0.902571041948579) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor16,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (59,0.85949177877429) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.88967468175389) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.877611940298507) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.860973888496824) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.829975825946817) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.885174418604651) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor22,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.811945117029863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor23,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.838709677419355) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor24,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.835409836065574) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor25,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.80250347705146) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor26,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.804043545878694) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.805673758865248) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.775038520801233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor29,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.799139167862267) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor30,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.7546875) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor31,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.783258594917788) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor32,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.812089356110381) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor33,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.790844514601421) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor34,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.798069187449718) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor35,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.750402576489533) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor36,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.766584766584767) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor37,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.70631970260223) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.741100323624595) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.820405310971349) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.694850115295926) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.659279778393352) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.750405186385737) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.701298701298701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.794076163610719) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.656502242152466) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.79048349961627) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.804532577903683) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.789808917197452) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.758827948910593) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.772635814889336) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.732796032238066) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.807417974322396) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.758152173913043) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.779639175257732) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.795505617977528) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.764754779717373) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.765809589993051) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.618556701030928) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.8140625) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.779050736497545) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.399612027158099) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.734463276836158) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.760072158749248) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.816816816816817) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.645510835913313) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.804580152671756) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (2,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.76523151909017) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.782808902532617) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.7) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.795768917819365) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.781575037147102) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.785104732350659) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.593448940269749) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.783402489626556) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.749190938511327) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.73890142964635) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.0355329949238578) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.746498599439776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.768066070199587) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.233009708737864) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.771929824561403) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.689100219458669) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.76663993584603) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.262008733624454) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.740207833733013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.794871794871795) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.207142857142857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.797404470079308) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.763665594855305) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.783333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.779527559055118) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.227657572906867) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.761666666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.800963081861958) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.236509758897819) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.782138024357239) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.782973621103117) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.770653514180025) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (25,0.790960451977401) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.78499106611078) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.75583596214511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.790289952798382) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.774193548387097) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.72264631043257) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.76103714085494) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.763432237369687) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.782475019215988) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.730407523510972) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.773790951638065) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.761982128350934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.783053323593864) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.788710907704043) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.689165186500888) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.772525849335303) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.76969696969697) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.758314855875831) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.724884080370943) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.651079136690647) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.764423076923077) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.774355751099937) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.795969773299748) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.708588957055215) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.762096774193548) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.762198888202594) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.766025641025641) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.740412979351032) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.762996941896024) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.637873754152824) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.568627450980392) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.801925722145804) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.778235779060181) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.456404736275565) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.790863668807994) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.774007220216606) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.778894472361809) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.706779661016949) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.758990053557766) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.746347941567065) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.726739926739927) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.758139534883721) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.702113156100886) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.738664468260511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.793674698795181) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.759656652360515) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.210161662817552) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (5,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.790123456790123) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.781542898341745) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.516765285996055) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.671197960917587) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.748224151539068) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.701505757307352) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.744479495268139) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.806451612903226) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.776200135226504) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.782215523737754) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.744279946164199) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.775055679287305) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.800936768149883) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.797822706065319) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.785714285714286) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.822794691647151) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.791354945968412) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.775572519083969) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.699846860643185) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.734864300626305) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.668453976764969) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.798732171156894) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.713804713804714) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.780162842339008) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.750773993808049) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.780373831775701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.73195020746888) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.795144157814871) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.747813411078717) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.743071161048689) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.77602523659306) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.771641791044776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.700087950747581) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.804628999319265) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.748172757475083) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.786301369863013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.774522292993631) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.777609682299546) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.789559543230016) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.788310762651461) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.744645799011532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.811819595645412) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.777847702957835) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.673862760215883) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.765498652291105) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.611913357400722) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.757692307692308) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.797101449275362) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.804511278195489) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.780285035629454) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.0408163265306122) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.760104302477184) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.819672131147541) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.741312741312741) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.680456490727532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.654418197725284) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.746172441579371) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.783279742765273) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.703216374269006) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.72249822569198) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.597320724980299) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.673196794300979) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.760330578512396) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.71701244813278) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.706355591311344) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.759316770186335) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.778294573643411) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.723646723646724) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.691455696202532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.778248587570621) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.790697674418605) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.783634933123525) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.225058004640371) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.755555555555556) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.746196957566053) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.75984555984556) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.775732788002726) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.8) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.799467021985343) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.735372340425532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.687105500450857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.756134969325153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.743388134381701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.70028818443804) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.800623052959501) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.783171521035599) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.234866828087167) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.742351046698873) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.792971734148205) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (2,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.776906628652887) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.78129713423831) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.768768768768769) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.817864619678995) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.783004552352048) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.784167289021658) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.792805755395683) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.792421746293245) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.747051114023591) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.816434724983433) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.00262467191601047) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.779761904761905) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.258064516129032) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.807692307692308) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.808629088378566) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.800528401585205) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.757062146892655) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.703894195444526) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.694262890341322) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.802387267904509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.751219512195122) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.766905330151153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.664383561643836) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.748019017432647) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.717391304347826) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.754448398576512) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.35361216730038) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.795902285263987) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.73663101604278) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.793313069908815) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.75) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.755656108597285) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.776595744680851) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.78726483357453) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.736220472440945) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.76499647141849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.768051118210863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.666666666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.76602086438152) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.695652173913043) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.626262626262626) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.785813630041725) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.708404802744425) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.629294755877034) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.802575107296137) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.789799072642967) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.619961612284069) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.713906111603189) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.783699059561129) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.779179810725552) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.670999187652315) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.428321678321678) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0.56007226738934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.633943427620632) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.75085910652921) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.765349032800673) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.779179810725552) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (2,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.784682080924855) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.774860779634049) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.780373831775701) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.244705882352941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.785763175906913) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.837570621468926) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.7601246105919) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.784067085953878) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.762811127379209) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.776432606941081) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.753911806543385) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.744379683597002) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.748858447488584) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.710108604845447) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.723051409618574) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.775181305398872) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.800929512006197) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.726307808946171) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.743791641429437) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.712005551700208) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.763969974979149) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.793650793650794) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.733113673805601) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0.702362204724409) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.655021834061135) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.72210953346856) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0.686119873817035) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.770338372930165) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.788990825688073) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.799222797927461) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.732888146911519) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.768601798855274) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.80835734870317) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.750147666863556) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.796502384737679) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.780766096169519) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.275334608030593) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.767203513909224) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.754047802621434) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.748137108792846) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.0344827586206896) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.755020080321285) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.76253081347576) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.80253766851705) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.771293375394322) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.544268406337372) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.743790849673202) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.239651416122004) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.757695343330702) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.654993514915694) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.768979591836735) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.782407407407407) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.241918665276329) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.754613807245386) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.78587528174305) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.736842105263158) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.741827885256838) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.758620689655172) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.776173285198556) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.791192103264996) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.749185667752443) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.787037037037037) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.71850699844479) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.778635778635778) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.745198463508322) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.774877650897227) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.786499215070643) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.700348432055749) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.753462603878116) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.792626728110599) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.741176470588235) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.765957446808511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.658008658008658) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.7765625) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.795050270688322) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.555133079847909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.674740484429066) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.685673556664291) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.782824112303881) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.538314176245211) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.752365930599369) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.707638279192274) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.701612903225806) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.708487084870849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.76875) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.618410700236035) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.154440154440154) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.803559206023272) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.768802228412256) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.707070707070707) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.705693664795509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.792251169004676) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.747268754552076) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.776588151320485) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.766265060240964) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.163265306122449) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.579512195121951) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.742808798646362) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.715702479338843) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.73938679245283) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.80327868852459) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.769945778466305) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.787923416789396) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.600749063670412) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.748376623376623) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.622540250447227) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.707350901525659) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.778501628664495) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.767616191904048) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.783748361730013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.785150078988941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.778123057799875) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.699236641221374) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.796795338674435) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.764617691154423) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.745791245791246) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.815028901734104) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.784883720930232) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.784711388455538) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.754426481909161) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.755795981452859) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.812742812742813) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.673497267759563) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.761437908496732) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.461095100864553) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.774668630338733) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.759148936170213) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.774395003903201) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.783608914450036) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.757939308398024) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.687631027253669) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.718964204112719) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.722778891115564) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.793243243243243) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.750716332378223) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.76459741856177) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.739656269891788) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.746395250212044) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.767284991568297) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.798223350253807) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (24,0.770580296896086) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.788355625491739) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.729684908789386) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.709849157054126) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.803459119496855) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.753898305084746) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.785658612626656) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.55378858746492) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.792452830188679) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.689908256880734) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.771604938271605) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.807888970051132) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.24896265560166) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.803041825095057) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.8125) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.724003127443315) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.768353528153956) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.754257907542579) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.751152073732719) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.711920529801324) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.773333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.765990639625585) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.247058823529412) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.743548387096774) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.746177370030581) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.650392327811683) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.693706293706294) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.610915492957747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.808646917534027) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.8218966846569) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.734099616858237) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.769911504424779) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.753184713375796) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.722488038277512) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.765990639625585) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.799199466310874) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.761389961389961) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.747336377473364) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.784810126582278) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.796570537802026) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.791758646063282) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.253521126760563) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.705432287681714) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.790584415584416) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.564727954971857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.608775137111517) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.736022646850672) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.787456445993031) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.798114689709348) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.780337941628264) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.792018419033001) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.11070110701107) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (7,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.772200772200772) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.171683389074693) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (5,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.749797570850202) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.736688548504741) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.780790085205267) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.796758946657664) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.731966590736522) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.745393634840871) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.800959232613909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.754349338900487) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.731297709923664) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.682512733446519) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.761020881670533) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.782870022539444) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.670618120237087) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.569811320754717) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.654594232059021) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.796403339755941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.771084337349397) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (9,0.0164609053497942) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.729751403368083) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.72) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.753558052434457) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.713656387665198) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.772655007949125) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.732336956521739) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.781666666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.762043795620438) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.738284066330209) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.716917922948074) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.692116182572614) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.774531835205992) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.746987951807229) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.716862745098039) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.652766639935846) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.749521988527725) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (23,0.789259560618389) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.770949720670391) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.708383961117861) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.792511700468019) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.697916666666667) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.695876288659794) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.727420667209113) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.747044917257683) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.762225969645868) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.77491961414791) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.752040175768989) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.767019667170953) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.790697674418605) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.75642965204236) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.788519637462236) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.774821544451655) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.240252897787144) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.696110744891233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.759247242050616) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.772264631043257) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.65300727032386) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.790629575402635) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.259593679458239) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.755277560594214) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.765543071161049) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.762805358550039) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.746590093323762) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.822342901474011) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.711356466876972) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.793478260869565) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.634275618374558) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.650709219858156) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.773680404916847) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.73394495412844) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.515410958904109) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.754978354978355) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.773448773448773) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.768020969855832) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.786707882534776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.722662440570523) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.766638584667228) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (20,0.72477693891558) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.734987990392314) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.786254295532646) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.641095890410959) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.806318681318681) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.779794790844515) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (10,0.727149627623561) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.799061767005473) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.811428571428571) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor41,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.758718861209964) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.831444759206799) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.786479802143446) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor42,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.756756756756757) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.760224538893344) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (22,0.799682034976153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor38,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.705882352941176) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.764468371467025) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor43,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (21,0.756159728122345) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.755134281200632) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.758169934640523) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor44,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.74591381872214) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.755852842809364) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor45,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (13,0.785657370517928) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.786567164179104) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor39,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.611218568665377) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (12,0.752857142857143) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (18,0.756602426837973) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor46,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.697961704756022) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.802884615384615) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor47,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (15,0.765306122448979) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (17,0.749058025621703) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor48,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (19,0.79053583855254) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.798771121351766) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (16,0.800316957210777) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor40,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (8,0.643799472295514) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (14,0.763369963369963) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor49,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (11,0.738001573564123) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor50,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (302,0.935897435897436) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (262,0.930594900849858) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (373,0.958133150308854) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (383,0.970954356846473) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (309,0.916850625459897) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (283,0.943396226415094) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (323,0.95676429567643) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (254,0.932107496463932) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.959112959112959) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (308,0.961379310344827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (248,0.938571428571429) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.963215258855586) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (303,0.969529085872576) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (295,0.954102920723227) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (207,0.884498480243161) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.92867332382311) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (221,0.906020558002937) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (243,0.952249134948097) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.949035812672176) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (236,0.954102920723227) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (183,0.924608819345661) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (198,0.946191474493361) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (206,0.939118264520644) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (163,0.92264017033357) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (171,0.922970159611381) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.887397464578673) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (145,0.943262411347518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (144,0.926058865757358) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (136,0.919212491513917) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (109,0.902015288394718) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (102,0.857566765578635) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (113,0.949080622347949) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.831913245546088) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (79,0.894183601962158) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.891061452513967) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.84299674267101) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.841710427606902) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.873280231716148) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.874743326488706) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (56,0.830838323353293) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (55,0.822840409956076) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.857571214392803) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.821165438713999) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.843235504652827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.838805970149254) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.804444444444444) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.798737174427782) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.807782101167315) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.747307373653687) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (329,0.950980392156863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (349,0.960553633217993) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (338,0.949057920446615) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (262,0.934782608695652) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (279,0.935028248587571) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (347,0.958333333333333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.970588235294118) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (275,0.945280437756498) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (214,0.8763197586727) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (236,0.935553946415641) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (292,0.954577218728162) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (329,0.971468336812804) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (255,0.946991404011461) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (281,0.959889349930844) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (274,0.940308517773306) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (188,0.919793966151582) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (228,0.918918918918919) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (237,0.94468085106383) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.945195729537366) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (209,0.948033707865168) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (241,0.954929577464789) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (174,0.922068463219228) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (172,0.924198250728863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (180,0.925435540069686) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (175,0.954293628808864) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (159,0.929824561403509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (136,0.870552611657835) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.87175572519084) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (153,0.92494639027877) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (112,0.82992125984252) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.915351506456241) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (133,0.922764227642276) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (96,0.917547568710359) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (91,0.894508670520231) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (78,0.854802680565897) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (80,0.914862914862915) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.863204096561814) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.893995552260934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (58,0.850691915513474) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (60,0.888888888888889) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (56,0.849315068493151) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (50,0.825019794140934) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (48,0.848845867460908) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.77502001601281) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.830788804071247) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.784988272087568) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.819858156028369) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.822373393801965) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.832572298325723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.823693379790941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (360,0.958654519971969) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (355,0.953405017921147) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.962025316455696) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (289,0.934497816593886) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (269,0.86910197869102) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (317,0.948611111111111) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.953910614525139) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.875384615384615) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (292,0.945583038869258) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (300,0.950175438596491) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (278,0.948336871903751) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (258,0.932489451476793) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.957805907172996) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (300,0.963493199713672) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (318,0.964210526315789) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (238,0.927641099855282) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.956036287508723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (222,0.921511627906977) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.925764192139738) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (257,0.953323903818953) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (221,0.939265536723164) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (215,0.927234927234927) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (176,0.914705882352941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (203,0.947368421052632) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (158,0.920245398773006) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (148,0.911917098445596) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (169,0.922749822820695) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (151,0.90028901734104) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (165,0.930914166085136) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (136,0.914893617021276) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.912398921832884) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (101,0.922413793103448) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (101,0.89900426742532) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (100,0.926359256710254) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (86,0.888571428571428) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (73,0.86046511627907) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (65,0.822981366459627) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (78,0.875714285714286) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (72,0.884917798427448) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (61,0.909732016925247) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (57,0.86377245508982) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (54,0.893124574540504) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.795811518324607) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.866176470588235) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.825278810408922) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.839319470699433) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.814390842191333) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.791124713083397) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.85878962536023) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.811361981063365) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (361,0.961990324809951) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (400,0.974180041870202) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (384,0.973121984838043) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (356,0.973756906077348) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.975103734439834) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (338,0.962025316455696) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (369,0.973463687150838) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (234,0.890566037735849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (272,0.964459930313589) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (340,0.969739619985925) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (299,0.964607911172797) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (297,0.970954356846473) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (307,0.970034843205575) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (291,0.966850828729282) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (273,0.969571230982019) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (286,0.97464016449623) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (224,0.943078004216444) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (264,0.967474048442907) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (234,0.961618981158409) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (237,0.957627118644068) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (205,0.967559943582511) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (212,0.934579439252336) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (184,0.953103448275862) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (188,0.936886395511921) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (168,0.952908587257618) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (126,0.870769230769231) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (120,0.879154078549849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (147,0.952836637047163) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (147,0.93408929836995) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (130,0.930808448652586) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (119,0.912778904665314) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (99,0.91399416909621) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (116,0.922330097087379) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.871064467766117) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (86,0.901387874360847) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (85,0.88626907073509) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (72,0.890792291220557) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.868827160493827) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (50,0.844514601420679) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.843226788432268) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (55,0.858630952380952) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (57,0.869446343130553) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.837430610626487) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (45,0.853087295954578) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.821917808219178) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.802241793434748) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.827933765298776) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.842654735272185) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.816577129700691) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (29,0.814296814296814) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (336,0.957431960921144) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (344,0.953310104529617) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (317,0.956645344705046) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (375,0.963066202090592) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (340,0.960283687943262) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (244,0.882175226586103) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (353,0.966480446927374) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (322,0.959774170783345) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (328,0.958017894012388) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (292,0.958066808813077) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (304,0.964838255977496) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (297,0.950608446671439) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (228,0.878012048192771) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (255,0.940406976744186) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.964985994397759) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (264,0.963117606123869) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (223,0.93915533285612) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (183,0.925872093023256) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (266,0.949856733524355) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.951856946354883) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (219,0.927075812274368) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.946206896551724) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (179,0.928312816799421) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (192,0.944793850454228) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.955096222380613) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.926724137931034) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.929618768328446) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (161,0.935574229691877) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (155,0.952249134948097) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.912152269399707) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (119,0.910791366906475) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (90,0.836335160532498) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (97,0.900293255131965) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (88,0.896746817538897) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (84,0.905011219147345) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (77,0.850152905198777) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.887943262411347) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (67,0.882058613295211) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (73,0.902404526166902) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (65,0.871794871794872) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.820109976433621) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (45,0.853582554517134) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.836781609195402) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.839452843772498) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (41,0.819314641744548) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.808208366219416) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.824727272727273) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.791967871485944) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.81650700073692) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.792297111416781) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (380,0.967428967428967) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (354,0.970446735395189) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (295,0.965948575399583) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (267,0.92790863668808) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (367,0.964754664823773) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (313,0.948682385575589) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.881608339538347) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (349,0.965469613259669) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (329,0.955570745044429) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (313,0.965181058495822) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (294,0.962131837307153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (304,0.95807560137457) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (223,0.880418535127055) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (228,0.881688018085908) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.95807560137457) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (222,0.931623931623932) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (264,0.942934782608695) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (208,0.926690391459075) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (205,0.879761015683346) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (239,0.935135135135135) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (179,0.882572924457741) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (201,0.926375982844889) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (160,0.869179600886918) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (165,0.875748502994012) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (161,0.92972972972973) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (176,0.946116165150455) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.947079037800687) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (152,0.918497519489723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (163,0.951523545706371) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (111,0.88138030194105) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (124,0.917142857142857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (112,0.902266288951841) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (96,0.864253393665158) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (86,0.919945725915875) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (83,0.88569374550683) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (88,0.918518518518518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.856512141280353) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (73,0.896232650363516) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (63,0.885359116022099) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.842911877394636) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (52,0.853658536585366) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.870481927710843) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (44,0.875636363636364) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (45,0.84558327714093) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.856948228882834) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.843326885880077) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.811102544333076) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (34,0.832151300236407) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.776772247360482) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.862800565770863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (288,0.893129770992366) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (316,0.942266571632217) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (380,0.973163841807909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (368,0.976256983240223) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (359,0.968332160450387) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (240,0.891603053435114) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (322,0.962290502793296) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (335,0.966386554621849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (340,0.968287526427061) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (342,0.977591036414566) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (273,0.967247386759582) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (293,0.971187631763879) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (303,0.977746870653685) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (255,0.9593837535014) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (250,0.936344969199179) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (227,0.93768115942029) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (290,0.968011126564673) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (189,0.893262679788039) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (264,0.960111966410077) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (191,0.93103448275862) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (232,0.956276445698166) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (132,0.856477889837083) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (177,0.927745664739884) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (205,0.958479943701618) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (175,0.93768115942029) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (134,0.882664647993944) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (127,0.925872093023256) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (151,0.913169319826339) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (122,0.906724511930586) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (142,0.93173821252639) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (109,0.858461538461538) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (121,0.909708078750849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (85,0.884241971620612) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (89,0.900785153461813) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.896451846488052) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (84,0.894623655913978) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.853435114503817) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.832558139534884) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (62,0.835886214442013) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.866914498141264) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (50,0.852281515854602) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.858654572940287) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.829411764705882) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.862637362637362) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.804453723034099) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.809220985691574) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.811361981063365) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.888059701492537) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.826830937713894) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.823620823620824) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (363,0.963380281690141) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (343,0.967651195499297) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (363,0.963276836158192) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (352,0.967109867039888) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (232,0.87044220325834) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (231,0.867132867132867) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (349,0.962702322308233) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (344,0.96629213483146) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (339,0.97752808988764) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (285,0.940836940836941) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (288,0.959887403237157) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (219,0.868421052631579) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (247,0.934687953555878) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (291,0.946554149085794) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (272,0.938775510204082) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.950749464668094) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (245,0.927641099855282) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (195,0.92882818116463) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (242,0.945351312987935) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (189,0.873174481168332) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (169,0.86910197869102) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (178,0.917510853835022) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (203,0.936139332365747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (185,0.942737430167598) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (176,0.919308357348703) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (179,0.941342756183746) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (128,0.910384068278805) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (156,0.942172073342736) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (144,0.943632567849687) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (115,0.921052631578947) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (119,0.87832973362131) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (98,0.871910112359551) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (101,0.911485003657644) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (75,0.838862559241706) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (87,0.924285714285714) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (73,0.884588804422944) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (69,0.877037037037037) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (68,0.883227176220807) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (61,0.85439763488544) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (58,0.87598944591029) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.819466248037676) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (51,0.836003051106026) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (46,0.856711915535445) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.786011656952539) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (40,0.808340016038492) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (32,0.838810641627543) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.87454677302393) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.784345047923323) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (31,0.805212620027435) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (27,0.752553024351925) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (395,0.962251201098147) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (281,0.934296028880866) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (381,0.963673749143249) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (331,0.949128919860627) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (316,0.937238493723849) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (271,0.937544867193108) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (252,0.878877400295421) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (287,0.937845303867403) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.96986301369863) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (312,0.960055096418733) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (301,0.9648033126294) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (302,0.950105411103303) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (323,0.964137931034483) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (272,0.9637234770705) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (294,0.954607977991747) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (260,0.939353099730458) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (259,0.962042788129744) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (225,0.929336188436831) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (269,0.953232462173315) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (211,0.938833570412518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (230,0.943189596167009) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.884241971620612) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (171,0.919540229885057) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (169,0.908296943231441) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (186,0.934871099050203) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (183,0.95407813570939) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.883582089552239) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (151,0.941828254847645) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (130,0.91027496382055) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (152,0.907518296739854) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (117,0.909698996655518) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (118,0.886111111111111) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (99,0.913454545454545) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (106,0.891628812459442) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (79,0.875798438608942) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (82,0.887142857142857) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.877840909090909) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (66,0.888230940044412) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (64,0.885793871866295) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.808868501529052) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.825) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (49,0.860411899313501) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.836913285600636) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (45,0.874285714285714) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.791862284820031) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (39,0.827208756841282) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (37,0.793242156074014) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.859479553903346) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (35,0.799363057324841) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (30,0.809266409266409) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (281,0.930935251798561) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (321,0.952646239554317) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!68!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (387,0.962131837307153) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (248,0.877458396369138) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (372,0.963117606123869) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (325,0.957805907172996) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!72!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (312,0.951135581555402) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (316,0.955182072829132) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (357,0.96) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!76!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (226,0.873369148119724) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (308,0.96455872133426) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (313,0.9721059972106) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!80!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (317,0.963636363636364) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (279,0.95049504950495) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!84!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (207,0.876438986953185) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (278,0.963380281690141) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!88!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (223,0.904129793510324) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (256,0.96140350877193) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!92!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (235,0.945659844742413) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (239,0.954732510288066) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red!96!black,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (177,0.874905231235785) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (179,0.857798165137615) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=red,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (220,0.959212376933896) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor1,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (171,0.925287356321839) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor2,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (192,0.956036287508723) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (152,0.909752547307132) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor3,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (175,0.95575221238938) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor4,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (144,0.929394812680115) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor5,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (136,0.922651933701657) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor6,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (137,0.943342776203966) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor7,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (110,0.861132660977502) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor8,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (111,0.91114245416079) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor9,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (107,0.90144727773949) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor10,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (100,0.899786780383795) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor11,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (91,0.858571428571429) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor12,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (76,0.839694656488549) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor13,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (71,0.821321321321321) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor15,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (74,0.898813677599442) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor51,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (58,0.857988165680473) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor17,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (61,0.852897473997028) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor18,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (53,0.884968242766408) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor19,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.816064257028112) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor20,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (43,0.857798165137615) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor21,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (47,0.833212996389892) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor52,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (38,0.824508320726172) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor53,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (42,0.835666912306559) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor54,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (33,0.841622035195103) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor55,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (36,0.827823691460055) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor27,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (28,0.793388429752066) 
};
\addplot [
color=blue,
mark size=1.3pt,
only marks,
mark=*,
mark options={solid,fill=mycolor28,draw=black,line width=0.1pt},
forget plot
]
coordinates{
 (26,0.806826997672614) 
};
\end{axis}
\end{tikzpicture}%