### 2013-03-20 figurefactory/files/limno_bgape_str_class_400.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (1478,0)(1478,-0.181818181818182)(1478,-0.363636363636364)(1478,-0.545454545454545)(1478,-0.727272727272727)(1478,-0.909090909090909)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454545)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454545)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454545)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478,-9.09090909090909)(1478,-9.27272727272727)(1478,-9.45454545454546)(1478,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454545)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454545)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454545)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.909090909090909)(0,-0.727272727272727)(0,-0.545454545454545)(0,-0.363636363636364)(0,-0.181818181818182)(0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (574.777777777778,-7.27272727272727)(582.242424242424,-7.3030303030303)(597.171717171717,-7.42424242424242)(600.90404040404,-7.45454545454545)(612.10101010101,-7.54545454545455)(619.565656565657,-7.63636363636364)(627.030303030303,-7.77272727272727)(629.518518518518,-7.81818181818182)(634.494949494949,-8)(634.494949494949,-8.18181818181818)(629.518518518518,-8.36363636363636)(627.030303030303,-8.40909090909091)(619.565656565657,-8.54545454545455)(612.10101010101,-8.63636363636364)(604.636363636364,-8.72727272727273)(597.171717171717,-8.81818181818182)(585.974747474747,-8.90909090909091)(582.242424242424,-8.93939393939394)(567.313131313131,-9.06060606060606)(559.848484848485,-9.09090909090909)(552.383838383838,-9.12121212121212)(537.454545454546,-9.18181818181818)(522.525252525253,-9.18181818181818)(507.59595959596,-9.18181818181818)(492.666666666667,-9.12121212121212)(485.20202020202,-9.09090909090909)(477.737373737374,-9.06060606060606)(462.808080808081,-8.93939393939394)(459.075757575758,-8.90909090909091)(447.878787878788,-8.77272727272727)(445.390572390572,-8.72727272727273)(435.43771043771,-8.54545454545455)(432.949494949495,-8.45454545454545)(430.461279461279,-8.36363636363636)(420.508417508418,-8.18181818181818)(418.020202020202,-8.09090909090909)(415.531986531986,-8)(415.531986531986,-7.81818181818182)(418.020202020202,-7.77272727272727)(425.484848484849,-7.63636363636364)(432.949494949495,-7.54545454545455)(444.146464646465,-7.45454545454545)(447.878787878788,-7.42424242424242)(462.808080808081,-7.36363636363636)(477.737373737374,-7.3030303030303)(485.20202020202,-7.27272727272727)(492.666666666667,-7.24242424242424)(507.59595959596,-7.18181818181818)(522.525252525253,-7.18181818181818)(537.454545454546,-7.18181818181818)(552.383838383838,-7.18181818181818)(567.313131313131,-7.24242424242424)(574.777777777778,-7.27272727272727)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (231.40404040404,-7.81818181818182)(238.868686868687,-7.84848484848485)(253.79797979798,-7.90909090909091)(268.727272727273,-7.96969696969697)(272.459595959596,-8)(283.656565656566,-8.13636363636364)(286.144781144781,-8.18181818181818)(286.144781144781,-8.36363636363636)(283.656565656566,-8.40909090909091)(276.191919191919,-8.54545454545455)(268.727272727273,-8.63636363636364)(257.530303030303,-8.72727272727273)(253.79797979798,-8.75757575757576)(238.868686868687,-8.81818181818182)(223.939393939394,-8.81818181818182)(209.010101010101,-8.81818181818182)(194.080808080808,-8.81818181818182)(179.151515151515,-8.75757575757576)(171.686868686869,-8.72727272727273)(164.222222222222,-8.6969696969697)(149.292929292929,-8.57575757575757)(145.560606060606,-8.54545454545455)(134.363636363636,-8.45454545454545)(126.89898989899,-8.36363636363636)(119.434343434343,-8.22727272727273)(116.946127946128,-8.18181818181818)(119.434343434343,-8.09090909090909)(121.922558922559,-8)(134.363636363636,-7.84848484848485)(141.828282828283,-7.81818181818182)(149.292929292929,-7.78787878787879)(164.222222222222,-7.72727272727273)(179.151515151515,-7.72727272727273)(194.080808080808,-7.72727272727273)(209.010101010101,-7.72727272727273)(223.939393939394,-7.78787878787879)(231.40404040404,-7.81818181818182)};

\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (821.111111111111,-13.6363636363636)(1224.20202020202,-3.27272727272727)(0,-18)(0,-8)(0,-6)(253.79797979798,-8.90909090909091)(104.505050505051,-10.1818181818182)(388.161616161616,-7.63636363636364)(492.666666666667,-8.90909090909091)(209.010101010101,-7.63636363636364)(656.888888888889,-7.81818181818182)(388.161616161616,-9.27272727272727)(507.59595959596,-8)(507.59595959596,-7.27272727272727)(567.313131313131,-8.72727272727273)(612.10101010101,-8.90909090909091)(238.868686868687,-8.54545454545455)(552.383838383838,-7.27272727272727)(418.020202020202,-7.27272727272727)(209.010101010101,-8.36363636363636)(89.5757575757576,-8.54545454545455)(134.363636363636,-8)(119.434343434343,-8.72727272727273)(328.444444444444,-7.63636363636364)(119.434343434343,-7.81818181818182)(89.5757575757576,-7.63636363636364)(627.030303030303,-8.18181818181818)(477.737373737374,-9.09090909090909)(537.454545454546,-7.27272727272727)(223.939393939394,-8.90909090909091)(462.808080808081,-9.09090909090909)(74.6464646464647,-7.63636363636364)(194.080808080808,-7.63636363636364)(597.171717171717,-7.45454545454545)(612.10101010101,-7.27272727272727)(612.10101010101,-8.54545454545455)(194.080808080808,-8.90909090909091)(343.373737373737,-8.90909090909091)(477.737373737374,-7.27272727272727)(597.171717171717,-7.27272727272727)(373.232323232323,-7.63636363636364)(89.5757575757576,-8.36363636363636)(567.313131313131,-8.90909090909091)(522.525252525253,-9.09090909090909)(567.313131313131,-9.09090909090909)(328.444444444444,-8.90909090909091)(627.030303030303,-7.81818181818182)(388.161616161616,-8.90909090909091)(313.515151515152,-7.81818181818182)(223.939393939394,-7.63636363636364)(119.434343434343,-7.63636363636364)(582.242424242424,-9.09090909090909)(627.030303030303,-8.36363636363636)(462.808080808081,-7.27272727272727)(403.090909090909,-8.90909090909091)(358.30303030303,-8.72727272727273)(149.292929292929,-8.72727272727273)(388.161616161616,-8.72727272727273)(328.444444444444,-7.81818181818182)(238.868686868687,-7.63636363636364)(358.30303030303,-8.54545454545455)(612.10101010101,-8.72727272727273)(582.242424242424,-7.27272727272727)(373.232323232323,-8.54545454545455)(373.232323232323,-8.36363636363636)(627.030303030303,-8)(388.161616161616,-8.36363636363636)(432.949494949495,-8.54545454545455)(432.949494949495,-7.45454545454545)(597.171717171717,-8.90909090909091)(313.515151515152,-8.36363636363636)(104.505050505051,-7.63636363636364)(74.6464646464647,-8)(134.363636363636,-8.54545454545455)(238.868686868687,-8.72727272727273)(522.525252525253,-7.09090909090909)(298.585858585859,-7.81818181818182)(403.090909090909,-7.63636363636364)(447.878787878788,-8.72727272727273)(492.666666666667,-9.09090909090909)(552.383838383838,-9.09090909090909)(209.010101010101,-8.72727272727273)(104.505050505051,-8.18181818181818)(179.151515151515,-8.72727272727273)(627.030303030303,-8.54545454545455)(388.161616161616,-8.18181818181818)(298.585858585859,-8.54545454545455)(283.656565656566,-7.81818181818182)(298.585858585859,-8.36363636363636)(164.222222222222,-7.45454545454545)(388.161616161616,-7.81818181818182)(388.161616161616,-8)(418.020202020202,-8.36363636363636)(537.454545454546,-9.27272727272727)(552.383838383838,-9.27272727272727)(432.949494949495,-8.36363636363636)(89.5757575757576,-8)(119.434343434343,-8.36363636363636)(104.505050505051,-8.36363636363636)(298.585858585859,-8.18181818181818)(567.313131313131,-7.27272727272727)(403.090909090909,-8)(179.151515151515,-7.45454545454545)(507.59595959596,-9.09090909090909)(627.030303030303,-7.63636363636364)(612.10101010101,-7.45454545454545)(418.020202020202,-7.63636363636364)(89.5757575757576,-8.18181818181818)(268.727272727273,-7.81818181818182)(403.090909090909,-7.81818181818182)(164.222222222222,-8.72727272727273)(253.79797979798,-8.72727272727273)(283.656565656566,-8.54545454545455)(447.878787878788,-7.27272727272727)(298.585858585859,-8)(134.363636363636,-7.63636363636364)(104.505050505051,-7.81818181818182)(492.666666666667,-7.27272727272727)(209.010101010101,-8.90909090909091)(418.020202020202,-8.18181818181818)(119.434343434343,-8.54545454545455)(641.959595959596,-8)(641.959595959596,-8.18181818181818)(477.737373737374,-8.90909090909091)(641.959595959596,-7.81818181818182)(89.5757575757576,-7.81818181818182)(268.727272727273,-8.72727272727273)(283.656565656566,-8)(462.808080808081,-8.90909090909091)(447.878787878788,-7.45454545454545)(418.020202020202,-7.81818181818182)(432.949494949495,-7.63636363636364)(462.808080808081,-7.45454545454545)(522.525252525253,-7.27272727272727)(418.020202020202,-8)(432.949494949495,-8.18181818181818)(432.949494949495,-7.81818181818182)(253.79797979798,-7.63636363636364)(149.292929292929,-7.63636363636364)(179.151515151515,-7.63636363636364)(164.222222222222,-7.63636363636364)(253.79797979798,-7.81818181818182)(283.656565656566,-8.36363636363636)(447.878787878788,-8.54545454545455)(522.525252525253,-9.27272727272727)(641.959595959596,-8.36363636363636)(238.868686868687,-7.81818181818182)(432.949494949495,-8)(104.505050505051,-8)(238.868686868687,-8.90909090909091)(283.656565656566,-8.18181818181818)(582.242424242424,-8.90909090909091)(447.878787878788,-7.63636363636364)(268.727272727273,-8.54545454545455)(552.383838383838,-7.09090909090909)(537.454545454546,-7.09090909090909)(462.808080808081,-8.72727272727273)(447.878787878788,-8.90909090909091)(627.030303030303,-7.45454545454545)(537.454545454546,-9.09090909090909)(134.363636363636,-7.81818181818182)(612.10101010101,-7.63636363636364)(313.515151515152,-8.18181818181818)(627.030303030303,-8.72727272727273)(477.737373737374,-7.45454545454545)(194.080808080808,-8.72727272727273)(223.939393939394,-8.72727272727273)(313.515151515152,-8)(268.727272727273,-8)(597.171717171717,-8.72727272727273)(223.939393939394,-7.81818181818182)(567.313131313131,-7.09090909090909)(507.59595959596,-9.27272727272727)(447.878787878788,-8.36363636363636)(432.949494949495,-8.72727272727273)(149.292929292929,-8.54545454545455)(567.313131313131,-9.27272727272727)(283.656565656566,-8.72727272727273)(582.242424242424,-7.45454545454545)(492.666666666667,-9.27272727272727)(403.090909090909,-8.18181818181818)(119.434343434343,-8)(268.727272727273,-7.63636363636364)(74.6464646464647,-8.18181818181818)(597.171717171717,-9.09090909090909)(268.727272727273,-8.36363636363636)(134.363636363636,-8.72727272727273)(104.505050505051,-8.54545454545455)(179.151515151515,-8.90909090909091)(447.878787878788,-7.81818181818182)(507.59595959596,-7.09090909090909)(612.10101010101,-8.36363636363636)(209.010101010101,-7.81818181818182)(328.444444444444,-8.18181818181818)(268.727272727273,-8.18181818181818)(418.020202020202,-8.54545454545455)(74.6464646464647,-7.81818181818182)(612.10101010101,-7.81818181818182)(641.959595959596,-7.63636363636364)(328.444444444444,-8)(462.808080808081,-8.54545454545455)(403.090909090909,-8.36363636363636)(253.79797979798,-8.54545454545455)(149.292929292929,-7.81818181818182)(194.080808080808,-7.81818181818182)(253.79797979798,-8)(164.222222222222,-7.81818181818182)(179.151515151515,-7.81818181818182)(119.434343434343,-8.18181818181818)(134.363636363636,-8.36363636363636)(418.020202020202,-7.45454545454545)(492.666666666667,-7.09090909090909)(164.222222222222,-8.90909090909091)(164.222222222222,-8.54545454545455)(373.232323232323,-7.81818181818182)(313.515151515152,-8.54545454545455)(373.232323232323,-8)(268.727272727273,-8.90909090909091)(298.585858585859,-8.72727272727273)(238.868686868687,-8)(59.7171717171717,-8)(134.363636363636,-8.18181818181818)(328.444444444444,-8.36363636363636)(477.737373737374,-9.27272727272727)(432.949494949495,-7.27272727272727)(403.090909090909,-7.45454545454545)(477.737373737374,-7.09090909090909)(462.808080808081,-7.09090909090909)(447.878787878788,-7.09090909090909)(283.656565656566,-7.63636363636364)(432.949494949495,-8.90909090909091)(447.878787878788,-8.18181818181818)(388.161616161616,-7.45454545454545)(612.10101010101,-8.18181818181818)(343.373737373737,-8)(358.30303030303,-7.81818181818182)(343.373737373737,-7.81818181818182)(582.242424242424,-7.09090909090909)(597.171717171717,-7.63636363636364)(567.313131313131,-7.45454545454545)(253.79797979798,-8.18181818181818)(477.737373737374,-8.72727272727273)(149.292929292929,-8)(149.292929292929,-8.36363636363636)(223.939393939394,-8)(253.79797979798,-8.36363636363636)(283.656565656566,-8.90909090909091)(358.30303030303,-8)(641.959595959596,-8.54545454545455)(179.151515151515,-8.54545454545455)(343.373737373737,-8.18181818181818)(447.878787878788,-8)(447.878787878788,-9.09090909090909)(149.292929292929,-7.45454545454545)(403.090909090909,-7.27272727272727)(358.30303030303,-7.63636363636364)(373.232323232323,-7.45454545454545)(612.10101010101,-8)(656.888888888889,-8)(164.222222222222,-8)(298.585858585859,-7.63636363636364)(74.6464646464647,-8.36363636363636)(492.666666666667,-7.45454545454545)(552.383838383838,-8.90909090909091)(507.59595959596,-8.90909090909091)(582.242424242424,-8.72727272727273)(418.020202020202,-8.72727272727273)(209.010101010101,-8)(373.232323232323,-8.18181818181818)(149.292929292929,-8.18181818181818)(179.151515151515,-8)(194.080808080808,-8)(343.373737373737,-7.63636363636364)(313.515151515152,-7.63636363636364)(656.888888888889,-8.18181818181818)(134.363636363636,-7.45454545454545)(119.434343434343,-7.45454545454545)(388.161616161616,-7.27272727272727)(358.30303030303,-7.45454545454545)(223.939393939394,-8.54545454545455)(358.30303030303,-8.18181818181818)(403.090909090909,-8.54545454545455)(432.949494949495,-7.09090909090909)(462.808080808081,-7.63636363636364)(238.868686868687,-8.18181818181818)(194.080808080808,-8.54545454545455)(597.171717171717,-8.54545454545455)(164.222222222222,-8.36363636363636)(164.222222222222,-8.18181818181818)(209.010101010101,-8.54545454545455)(238.868686868687,-8.36363636363636)(194.080808080808,-7.45454545454545)(462.808080808081,-8.36363636363636)(223.939393939394,-9.09090909090909)(343.373737373737,-7.45454545454545)(522.525252525253,-9.45454545454546)(223.939393939394,-8.18181818181818)(462.808080808081,-9.27272727272727)(343.373737373737,-8.36363636363636)(59.7171717171717,-7.81818181818182)(328.444444444444,-8.54545454545455)(179.151515151515,-8.18181818181818)(179.151515151515,-8.36363636363636)(522.525252525253,-8.90909090909091)(537.454545454546,-8.90909090909091)(462.808080808081,-8.18181818181818)(641.959595959596,-7.45454545454545)(627.030303030303,-7.27272727272727)(656.888888888889,-7.63636363636364)(597.171717171717,-7.09090909090909)(612.10101010101,-7.09090909090909)(641.959595959596,-7.27272727272727)(507.59595959596,-7.45454545454545)(552.383838383838,-7.45454545454545)(462.808080808081,-7.81818181818182)(522.525252525253,-7.45454545454545)(537.454545454546,-7.45454545454545)(477.737373737374,-7.63636363636364)(418.020202020202,-7.09090909090909)(582.242424242424,-9.27272727272727)(627.030303030303,-8.90909090909091)(492.666666666667,-8.72727272727273)(358.30303030303,-8.36363636363636)(537.454545454546,-9.45454545454546)(313.515151515152,-8.72727272727273)(209.010101010101,-8.18181818181818)(223.939393939394,-8.36363636363636)(194.080808080808,-8.18181818181818)(477.737373737374,-8.54545454545455)(149.292929292929,-8.90909090909091)(209.010101010101,-7.45454545454545)(656.888888888889,-7.45454545454545)(507.59595959596,-9.45454545454546)(552.383838383838,-9.45454545454546)(104.505050505051,-7.45454545454545)(89.5757575757576,-7.45454545454545)(223.939393939394,-7.45454545454545)(328.444444444444,-7.45454545454545)(238.868686868687,-7.45454545454545)(194.080808080808,-8.36363636363636)(492.666666666667,-9.45454545454546)(537.454545454546,-6.90909090909091)(522.525252525253,-6.90909090909091)(552.383838383838,-6.90909090909091)(507.59595959596,-6.90909090909091)(567.313131313131,-6.90909090909091)(567.313131313131,-4.72727272727273)(492.666666666667,-6.90909090909091)(462.808080808081,-8)(388.161616161616,-8.54545454545455)(343.373737373737,-8.54545454545455)(403.090909090909,-8.72727272727273)(104.505050505051,-8.72727272727273)(59.7171717171717,-7.63636363636364)(656.888888888889,-8.36363636363636)(582.242424242424,-7.63636363636364)(597.171717171717,-7.81818181818182)(238.868686868687,-9.09090909090909)(209.010101010101,-9.09090909090909)(253.79797979798,-9.09090909090909)(194.080808080808,-9.09090909090909)(298.585858585859,-8.90909090909091)(268.727272727273,-9.09090909090909)(418.020202020202,-8.90909090909091)(597.171717171717,-8.36363636363636)(582.242424242424,-6.90909090909091)(627.030303030303,-7.09090909090909)(328.444444444444,-8.72727272727273)(612.10101010101,-9.09090909090909)(373.232323232323,-7.27272727272727)(432.949494949495,-9.09090909090909)(492.666666666667,-7.63636363636364)(477.737373737374,-6.90909090909091)(313.515151515152,-8.90909090909091)(283.656565656566,-9.09090909090909)(179.151515151515,-9.09090909090909)(582.242424242424,-8.54545454545455)(597.171717171717,-8)(313.515151515152,-7.45454545454545)(134.363636363636,-8.90909090909091)(447.878787878788,-9.27272727272727)(343.373737373737,-8.72727272727273)(477.737373737374,-9.45454545454546)(477.737373737374,-7.81818181818182)(641.959595959596,-8.72727272727273)(298.585858585859,-7.45454545454545)(253.79797979798,-7.45454545454545)(283.656565656566,-7.45454545454545)(59.7171717171717,-8.18181818181818)(567.313131313131,-7.63636363636364)(597.171717171717,-8.18181818181818)(477.737373737374,-8.36363636363636)(462.808080808081,-6.90909090909091)(507.59595959596,-7.63636363636364)(507.59595959596,-8.72727272727273)(597.171717171717,-6.90909090909091)(298.585858585859,-9.09090909090909)(552.383838383838,-8.72727272727273)(418.020202020202,-9.09090909090909)(373.232323232323,-8.72727272727273) 
};

\end{axis}
\end{tikzpicture}%