### 2013-03-20 figurefactory/files/limno_bgape_var_class_400.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (1478,0)(1478,-0.181818181818182)(1478,-0.363636363636364)(1478,-0.545454545454545)(1478,-0.727272727272727)(1478,-0.909090909090909)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454545)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454545)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454545)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478.00016587919,-9.09090909090909)(1478.00033175469,-9.27272727272727)(1478.00033175469,-9.45454545454546)(1478.00016587919,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454545)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454545)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454545)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.909090909090909)(0,-0.727272727272727)(0,-0.545454545454545)(0,-0.363636363636364)(0,-0.181818181818182)(0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1216.73737373737,-7.63636363636364)(1224.20202020202,-7.66666666666667)(1239.13131313131,-7.72727272727273)(1254.06060606061,-7.72727272727273)(1268.9898989899,-7.72727272727273)(1283.91919191919,-7.78787878787879)(1291.38383838384,-7.81818181818182)(1298.84848484848,-7.84848484848485)(1313.77777777778,-7.90909090909091)(1328.70707070707,-7.96969696969697)(1336.17171717172,-8)(1343.63636363636,-8.03030303030303)(1358.56565656566,-8.15151515151515)(1362.29797979798,-8.18181818181818)(1373.49494949495,-8.27272727272727)(1384.69191919192,-8.36363636363636)(1388.42424242424,-8.39393939393939)(1403.35353535354,-8.51515151515152)(1407.08585858586,-8.54545454545455)(1418.28282828283,-8.63636363636364)(1425.74747474747,-8.72727272727273)(1433.21212121212,-8.81818181818182)(1440.67676767677,-8.90909090909091)(1448.14141414141,-9)(1455.60606060606,-9.09090909090909)(1463.07070707071,-9.22727272727273)(1465.55892255892,-9.27272727272727)(1465.55892255892,-9.45454545454546)(1463.07070707071,-9.5)(1455.60606060606,-9.63636363636364)(1448.14141414141,-9.72727272727273)(1436.94444444444,-9.81818181818182)(1433.21212121212,-9.84848484848485)(1418.28282828283,-9.96969696969697)(1410.81818181818,-10)(1403.35353535354,-10.030303030303)(1388.42424242424,-10.0909090909091)(1373.49494949495,-10.0909090909091)(1358.56565656566,-10.0909090909091)(1343.63636363636,-10.0909090909091)(1328.70707070707,-10.0909090909091)(1313.77777777778,-10.0909090909091)(1298.84848484848,-10.0909090909091)(1283.91919191919,-10.0909090909091)(1268.9898989899,-10.0909090909091)(1254.06060606061,-10.0909090909091)(1239.13131313131,-10.0909090909091)(1224.20202020202,-10.0909090909091)(1209.27272727273,-10.0909090909091)(1194.34343434343,-10.0909090909091)(1179.41414141414,-10.0909090909091)(1164.48484848485,-10.0909090909091)(1149.55555555556,-10.0909090909091)(1134.62626262626,-10.0909090909091)(1119.69696969697,-10.0909090909091)(1104.76767676768,-10.0909090909091)(1089.83838383838,-10.0909090909091)(1074.90909090909,-10.0909090909091)(1059.9797979798,-10.0909090909091)(1045.05050505051,-10.0909090909091)(1030.12121212121,-10.0909090909091)(1015.19191919192,-10.0909090909091)(1000.26262626263,-10.0909090909091)(985.333333333333,-10.0909090909091)(970.40404040404,-10.030303030303)(962.939393939394,-10)(955.474747474747,-9.96969696969697)(940.545454545455,-9.90909090909091)(925.616161616162,-9.90909090909091)(910.686868686869,-9.84848484848485)(903.222222222222,-9.81818181818182)(895.757575757576,-9.78787878787879)(880.828282828283,-9.72727272727273)(865.89898989899,-9.66666666666667)(858.434343434343,-9.63636363636364)(850.969696969697,-9.60606060606061)(836.040404040404,-9.48484848484848)(832.308080808081,-9.45454545454546)(821.111111111111,-9.36363636363636)(813.646464646465,-9.27272727272727)(806.181818181818,-9.18181818181818)(798.717171717172,-9.09090909090909)(791.252525252525,-9)(783.787878787879,-8.90909090909091)(776.323232323232,-8.77272727272727)(773.835016835017,-8.72727272727273)(773.835016835017,-8.54545454545455)(776.323232323232,-8.45454545454545)(778.811447811448,-8.36363636363636)(791.252525252525,-8.21212121212121)(793.740740740741,-8.18181818181818)(806.181818181818,-8.03030303030303)(813.646464646465,-8)(821.111111111111,-7.96969696969697)(836.040404040404,-7.84848484848485)(843.505050505051,-7.81818181818182)(850.969696969697,-7.78787878787879)(865.89898989899,-7.72727272727273)(880.828282828283,-7.66666666666667)(888.292929292929,-7.63636363636364)(895.757575757576,-7.60606060606061)(910.686868686869,-7.54545454545455)(925.616161616162,-7.54545454545455)(940.545454545455,-7.54545454545455)(955.474747474747,-7.54545454545455)(970.40404040404,-7.54545454545455)(985.333333333333,-7.54545454545455)(1000.26262626263,-7.54545454545455)(1015.19191919192,-7.54545454545455)(1030.12121212121,-7.54545454545455)(1045.05050505051,-7.54545454545455)(1059.9797979798,-7.54545454545455)(1074.90909090909,-7.54545454545455)(1089.83838383838,-7.54545454545455)(1104.76767676768,-7.54545454545455)(1119.69696969697,-7.54545454545455)(1134.62626262626,-7.54545454545455)(1149.55555555556,-7.54545454545455)(1164.48484848485,-7.54545454545455)(1179.41414141414,-7.54545454545455)(1194.34343434343,-7.54545454545455)(1209.27272727273,-7.60606060606061)(1216.73737373737,-7.63636363636364)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (559.848484848485,-7.45454545454545)(567.313131313131,-7.48484848484848)(582.242424242424,-7.60606060606061)(585.974747474747,-7.63636363636364)(597.171717171717,-7.77272727272727)(599.659932659933,-7.81818181818182)(609.612794612795,-8)(609.612794612795,-8.18181818181818)(604.636363636364,-8.36363636363636)(597.171717171717,-8.45454545454545)(585.974747474747,-8.54545454545455)(582.242424242424,-8.57575757575757)(567.313131313131,-8.63636363636364)(552.383838383838,-8.63636363636364)(537.454545454546,-8.63636363636364)(522.525252525253,-8.63636363636364)(507.59595959596,-8.63636363636364)(492.666666666667,-8.63636363636364)(477.737373737374,-8.63636363636364)(462.808080808081,-8.63636363636364)(447.878787878788,-8.63636363636364)(432.949494949495,-8.63636363636364)(418.020202020202,-8.63636363636364)(403.090909090909,-8.63636363636364)(388.161616161616,-8.57575757575757)(380.69696969697,-8.54545454545455)(373.232323232323,-8.51515151515152)(358.30303030303,-8.45454545454545)(343.373737373737,-8.45454545454545)(328.444444444444,-8.45454545454545)(313.515151515152,-8.45454545454545)(298.585858585859,-8.45454545454545)(283.656565656566,-8.45454545454545)(268.727272727273,-8.51515151515152)(261.262626262626,-8.54545454545455)(253.79797979798,-8.57575757575757)(238.868686868687,-8.57575757575757)(231.40404040404,-8.54545454545455)(223.939393939394,-8.51515151515152)(209.010101010101,-8.45454545454545)(194.080808080808,-8.39393939393939)(190.348484848485,-8.36363636363636)(179.151515151515,-8.22727272727273)(176.6632996633,-8.18181818181818)(179.151515151515,-8.09090909090909)(182.883838383838,-8)(194.080808080808,-7.90909090909091)(209.010101010101,-7.90909090909091)(223.939393939394,-7.90909090909091)(238.868686868687,-7.90909090909091)(253.79797979798,-7.90909090909091)(268.727272727273,-7.90909090909091)(283.656565656566,-7.90909090909091)(298.585858585859,-7.90909090909091)(313.515151515152,-7.90909090909091)(328.444444444444,-7.90909090909091)(343.373737373737,-7.90909090909091)(358.30303030303,-7.84848484848485)(365.767676767677,-7.81818181818182)(373.232323232323,-7.78787878787879)(388.161616161616,-7.66666666666667)(395.626262626263,-7.63636363636364)(403.090909090909,-7.60606060606061)(418.020202020202,-7.54545454545455)(432.949494949495,-7.48484848484848)(440.414141414141,-7.45454545454545)(447.878787878788,-7.42424242424242)(462.808080808081,-7.36363636363636)(477.737373737374,-7.36363636363636)(492.666666666667,-7.36363636363636)(507.59595959596,-7.36363636363636)(522.525252525253,-7.36363636363636)(537.454545454546,-7.36363636363636)(552.383838383838,-7.42424242424242)(559.848484848485,-7.45454545454545)};

\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (477.737373737374,-17.4545454545455)(1478,0)(0,-5.81818181818182)(1478,-10.1818181818182)(1478,-18)(418.020202020202,0)(0,-12.1818181818182)(940.545454545455,-5.45454545454545)(791.252525252525,-12.1818181818182)(1478,-4.72727272727273)(0,-18)(1478,-14.1818181818182)(462.808080808081,-8.54545454545455)(0,-1.81818181818182)(940.545454545455,-1.27272727272727)(970.40404040404,-18)(462.808080808081,-4)(1015.19191919192,-8.90909090909091)(358.30303030303,-14.3636363636364)(0,-9.09090909090909)(985.333333333333,-14.9090909090909)(1478,-7.45454545454545)(0,-15.2727272727273)(0,0)(388.161616161616,-11.2727272727273)(1194.34343434343,-12.1818181818182)(1283.91919191919,-2.54545454545455)(567.313131313131,-6.36363636363636)(1134.62626262626,0)(1283.91919191919,-16.1818181818182)(731.535353535354,0)(671.818181818182,-15.4545454545455)(776.323232323232,-3.27272727272727)(731.535353535354,-9.81818181818182)(358.30303030303,-2)(0,-3.81818181818182)(1209.27272727273,-6.54545454545455)(238.868686868687,-6.90909090909091)(1478,-12.1818181818182)(1478,-2.18181818181818)(223.939393939394,-16.5454545454545)(821.111111111111,-7.45454545454545)(716.606060606061,-18)(1149.55555555556,-4.18181818181818)(582.242424242424,-13.2727272727273)(1179.41414141414,-10.3636363636364)(1224.20202020202,-18)(209.010101010101,-10)(238.868686868687,-4.90909090909091)(268.727272727273,-18)(627.030303030303,-1.63636363636364)(1478,-16.1818181818182)(194.080808080808,-13.0909090909091)(1283.91919191919,-8.54545454545455)(1209.27272727273,-14)(0,-7.45454545454545)(910.686868686869,-16.5454545454545)(0,-13.8181818181818)(209.010101010101,0)(940.545454545455,-13.2727272727273)(955.474747474747,-10.9090909090909)(0,-10.7272727272727)(701.676767676768,-4.90909090909091)(1015.19191919192,-2.90909090909091)(194.080808080808,-3.09090909090909)(1283.91919191919,-1.09090909090909)(1478,-6)(597.171717171717,-11.0909090909091)(0,-16.7272727272727)(447.878787878788,-15.8181818181818)(223.939393939394,-8.36363636363636)(1030.12121212121,-7.27272727272727)(1478,-8.90909090909091)(925.616161616162,0)(761.393939393939,-14.1818181818182)(1478,-3.45454545454545)(462.808080808081,-9.81818181818182)(671.818181818182,-8.36363636363636)(1283.91919191919,-5.09090909090909)(179.151515151515,-1.27272727272727)(179.151515151515,-14.9090909090909)(403.090909090909,-5.63636363636364)(1104.76767676768,-16.5454545454545)(194.080808080808,-11.6363636363636)(403.090909090909,-12.7272727272727)(686.747474747475,-16.7272727272727)(552.383838383838,-2.90909090909091)(1343.63636363636,-11.0909090909091)(1313.77777777778,0)(925.616161616162,-4.18181818181818)(432.949494949495,-7.27272727272727)(1358.56565656566,-13.0909090909091)(1313.77777777778,-15.0909090909091)(776.323232323232,-6.18181818181818)(1478,-1.09090909090909)(1119.69696969697,-1.63636363636364)(567.313131313131,0)(522.525252525253,-18)(806.181818181818,-2)(1015.19191919192,-12.1818181818182)(537.454545454546,-14.5454545454545)(850.969696969697,-8.90909090909091)(447.878787878788,-1.09090909090909)(0,-4.72727272727273)(1358.56565656566,-17.2727272727273)(1104.76767676768,-5.63636363636364)(1313.77777777778,-7.27272727272727)(0,-2.72727272727273)(1313.77777777778,-3.81818181818182)(1313.77777777778,-9.63636363636364)(149.292929292929,-6)(134.363636363636,-18)(791.252525252525,-10.9090909090909)(1478,-15.2727272727273)(597.171717171717,-12.1818181818182)(1134.62626262626,-15.2727272727273)(836.040404040404,-15.4545454545455)(313.515151515152,-3.63636363636364)(0,-0.909090909090909)(1134.62626262626,-8.18181818181818)(552.383838383838,-4.90909090909091)(1478,-11.2727272727273)(1000.26262626263,-10)(627.030303030303,-7.27272727272727)(1089.83838383838,-13.4545454545455)(1478,-13.2727272727273)(761.393939393939,-0.909090909090909)(836.040404040404,-18)(1104.76767676768,-18)(313.515151515152,-9.27272727272727)(373.232323232323,-16.7272727272727)(1478,-17.0909090909091)(656.888888888889,-3.81818181818182)(89.5757575757576,-8.18181818181818)(134.363636363636,-4.18181818181818)(597.171717171717,-9.45454545454546)(298.585858585859,-15.4545454545455)(761.393939393939,-13.0909090909091)(0,-9.81818181818182)(940.545454545455,-6.54545454545455)(0,-12.9090909090909)(0,-6.72727272727273)(1119.69696969697,-11.2727272727273)(1149.55555555556,-3.09090909090909)(164.222222222222,-14)(1343.63636363636,-6)(104.505050505051,-16)(164.222222222222,-2.18181818181818)(1343.63636363636,-18)(373.232323232323,-18)(149.292929292929,-10.7272727272727)(567.313131313131,-16.3636363636364)(298.585858585859,-0.727272727272727)(1149.55555555556,-9.27272727272727)(1074.90909090909,-0.727272727272727)(418.020202020202,-2.72727272727273)(403.090909090909,-13.6363636363636)(343.373737373737,-10.5454545454545)(1478,-8.18181818181818)(119.434343434343,0)(910.686868686869,-14.1818181818182)(1328.70707070707,-12.1818181818182)(343.373737373737,-6.36363636363636)(910.686868686869,-2.54545454545455)(925.616161616162,-8)(0,-11.4545454545455)(0,-14.5454545454545)(298.585858585859,-12.1818181818182)(821.111111111111,-4.72727272727273)(821.111111111111,-17.0909090909091)(1478,-6.72727272727273)(134.363636363636,-17.0909090909091)(328.444444444444,-7.81818181818182)(1194.34343434343,-17.0909090909091)(1343.63636363636,-14.1818181818182)(1045.05050505051,-4.54545454545455)(134.363636363636,-9.27272727272727)(0,-8.36363636363636)(1358.56565656566,-1.81818181818182)(671.818181818182,-2.54545454545455)(373.232323232323,-4.72727272727273)(582.242424242424,-0.909090909090909)(1000.26262626263,-15.8181818181818)(865.89898989899,-9.81818181818182)(134.363636363636,-7.27272727272727)(656.888888888889,-5.81818181818182)(1015.19191919192,0)(895.757575757576,-11.8181818181818)(1478,-2.90909090909091)(627.030303030303,-18)(477.737373737374,-11.8181818181818)(0,-16)(1015.19191919192,-17.0909090909091)(582.242424242424,-10.3636363636364)(641.959595959596,-14.1818181818182)(1478,-5.27272727272727)(328.444444444444,0)(1478,-9.63636363636364)(119.434343434343,-12.3636363636364)(821.111111111111,0)(1463.07070707071,0)(1209.27272727273,-13.0909090909091)(1164.48484848485,-7.27272727272727)(0,-17.4545454545455)(552.383838383838,-8)(1089.83838383838,-14.3636363636364)(1239.13131313131,0)(119.434343434343,-5.27272727272727)(1478,-4.18181818181818)(477.737373737374,-15.0909090909091)(701.676767676768,-11.6363636363636)(507.59595959596,-2)(1074.90909090909,-2.18181818181818)(1328.70707070707,-10.3636363636364)(910.686868686869,-3.45454545454545)(1343.63636363636,-8)(1463.07070707071,-18)(731.535353535354,-7.09090909090909)(0,-3.27272727272727)(1059.9797979798,-6.36363636363636)(776.323232323232,-16)(1328.70707070707,-4.54545454545455)(761.393939393939,-8.54545454545455)(1343.63636363636,-3.09090909090909)(14.9292929292929,0)(283.656565656566,-13.4545454545455)(1478,-14.7272727272727)(1478,-0.909090909090909)(1373.49494949495,-16)(104.505050505051,-0.909090909090909)(627.030303030303,0)(1179.41414141414,-5.27272727272727)(283.656565656566,-17.2727272727273)(1478,-12.7272727272727)(283.656565656566,-2.54545454545455)(597.171717171717,-17.0909090909091)(477.737373737374,-6.36363636363636)(1239.13131313131,-11.2727272727273)(0,-5.27272727272727)(880.828282828283,-0.909090909090909)(268.727272727273,-5.81818181818182)(1478,-10.9090909090909)(836.040404040404,-5.63636363636364)(1179.41414141414,-15.8181818181818)(1194.34343434343,-1.09090909090909)(552.383838383838,-3.81818181818182)(1059.9797979798,-12.7272727272727)(776.323232323232,-14.9090909090909)(14.9292929292929,-18)(388.161616161616,-8.90909090909091)(507.59595959596,-12.9090909090909)(865.89898989899,-12.9090909090909)(477.737373737374,-10.7272727272727)(1074.90909090909,-10.3636363636364)(134.363636363636,-3.45454545454545)(761.393939393939,-4)(1358.56565656566,-0.727272727272727)(1059.9797979798,-3.63636363636364)(194.080808080808,-15.6363636363636)(1358.56565656566,-9.09090909090909)(298.585858585859,-1.45454545454545)(925.616161616162,-18)(253.79797979798,-11.0909090909091)(716.606060606061,-10.5454545454545)(0,-2)(1478,-16.7272727272727)(671.818181818182,-12.9090909090909)(1358.56565656566,-6.72727272727273)(522.525252525253,-5.45454545454545)(597.171717171717,-15.4545454545455)(1209.27272727273,-2.18181818181818)(0,-13.4545454545455)(0,-10.1818181818182)(582.242424242424,-8.90909090909091)(268.727272727273,-4.18181818181818)(1030.12121212121,-8.18181818181818)(477.737373737374,0)(1478,-1.81818181818182)(1224.20202020202,-14.7272727272727)(179.151515151515,-18)(283.656565656566,-14.5454545454545)(0,-6.36363636363636)(731.535353535354,-1.63636363636364)(970.40404040404,-1.81818181818182)(104.505050505051,-14.5454545454545)(432.949494949495,-18)(1030.12121212121,-11.4545454545455)(880.828282828283,-6.90909090909091)(1119.69696969697,-18)(492.666666666667,-14)(358.30303030303,-16)(1298.84848484848,-16.9090909090909)(1224.20202020202,-3.63636363636364)(104.505050505051,-11.4545454545455)(403.090909090909,-3.27272727272727)(1000.26262626263,-5.09090909090909)(1194.34343434343,-9.45454545454546)(880.828282828283,-10.5454545454545)(343.373737373737,-10)(0,-7.81818181818182)(1478,-13.8181818181818)(134.363636363636,-6.54545454545455)(119.434343434343,-13.2727272727273)(0,-4.36363636363636)(955.474747474747,-9.27272727272727)(238.868686868687,-7.63636363636364)(507.59595959596,-7.27272727272727)(477.737373737374,-16.7272727272727)(0,-1.09090909090909)(194.080808080808,-8.90909090909091)(1313.77777777778,-13.4545454545455)(1239.13131313131,-6)(1298.84848484848,-18)(1000.26262626263,-14)(746.464646464646,-17.2727272727273)(104.505050505051,-2.36363636363636)(1478,-11.8181818181818)(776.323232323232,-2.72727272727273)(1209.27272727273,-8)(1478,-15.6363636363636)(731.535353535354,-9.27272727272727)(119.434343434343,-10)(925.616161616162,-15.4545454545455)(821.111111111111,-13.8181818181818)(1358.56565656566,0)(1373.49494949495,-5.27272727272727)(1478,-7.81818181818182)(0,-16.9090909090909)(671.818181818182,-0.727272727272727)(1104.76767676768,0)(283.656565656566,-12.5454545454545)(432.949494949495,-0.909090909090909)(0,-15.0909090909091)(627.030303030303,-4.54545454545455)(238.868686868687,0)(1343.63636363636,-11.8181818181818)(746.464646464646,-7.81818181818182)(1478,-6.36363636363636)(940.545454545455,-17.0909090909091)(671.818181818182,-6.54545454545455)(0,-12)(0,-8.90909090909091)(776.323232323232,-18)(432.949494949495,-4.72727272727273)(1478,-9.09090909090909)(403.090909090909,-8)(1358.56565656566,-15.0909090909091)(627.030303030303,-3.09090909090909)(1478,-3.81818181818182)(343.373737373737,-6.90909090909091)(880.828282828283,0)(836.040404040404,-11.4545454545455)(418.020202020202,-12)(1015.19191919192,-0.909090909090909)(1149.55555555556,-12.1818181818182)(388.161616161616,-15.0909090909091)(1194.34343434343,-4.54545454545455)(134.363636363636,-4.72727272727273)(656.888888888889,-14.7272727272727)(1119.69696969697,-17.0909090909091)(582.242424242424,-11.6363636363636)(597.171717171717,-18)(731.535353535354,-5.45454545454545)(1059.9797979798,-16)(492.666666666667,-9.45454545454546)(1373.49494949495,-2.54545454545455)(134.363636363636,-16.7272727272727)(686.747474747475,-16.1818181818182)(985.333333333333,-6)(552.383838383838,-2)(880.828282828283,-8.18181818181818)(940.545454545455,-12.5454545454545)(671.818181818182,-13.6363636363636)(1478,-10.3636363636364)(1104.76767676768,-6.90909090909091)(0,-10.9090909090909)(298.585858585859,-5.27272727272727)(1478,-17.4545454545455)(1239.13131313131,-10.5454545454545)(880.828282828283,-4.54545454545455)(179.151515151515,-0.727272727272727)(1074.90909090909,-9.45454545454546)(716.606060606061,0)(313.515151515152,-18)(701.676767676768,-12.3636363636364)(880.828282828283,-14.7272727272727)(850.969696969697,-16.1818181818182)(1478,-5.45454545454545)(0,-15.8181818181818)(1283.91919191919,-1.63636363636364)(1149.55555555556,-2.72727272727273)(0,-0.181818181818182)(1089.83838383838,-14.9090909090909)(1373.49494949495,-4)(1030.12121212121,-18)(1478,-0.181818181818182)(865.89898989899,-1.81818181818182)(0,-2.90909090909091)(0,-14.1818181818182)(358.30303030303,-17.0909090909091) 
};

\end{axis}
\end{tikzpicture}%