### 2013-03-21 figurefactory/files/limno_bgape_imp_class_120.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00022117143,-4.18181818181818)(1478.00022117143,-4.36363636363636)(1478.00022117143,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478.00005529347,-8.36363636363636)(1478.00022117143,-8.54545454545455)(1478.00027646327,-8.72727272727273)(1478.00027646327,-8.90909090909091)(1478.00027646327,-9.09090909090909)(1478.00038704571,-9.27272727272727)(1478.00038704571,-9.45454545454546)(1478.00038704571,-9.63636363636364)(1478.00027646327,-9.81818181818182)(1478.00022117143,-10)(1478.00005529347,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478.00022117143,-10.7272727272727)(1478.00022117143,-10.9090909090909)(1478.00022117143,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(-0.000221171433450288,-8.72727272727273)(-0.000221171433450288,-8.54545454545455)(-0.000221171433450288,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1478,-8.63636363636364)(1478.00002764633,-8.72727272727273)(1478.00002764633,-8.90909090909091)(1478.00002764633,-9.09090909090909)(1478.00013823061,-9.27272727272727)(1478.00013823061,-9.45454545454546)(1478.00013823061,-9.63636363636364)(1478.00002764633,-9.81818181818182)(1478,-9.90909090909091)(1470.53535353535,-10)(1463.07070707071,-10.030303030303)(1448.14141414141,-10.1515151515152)(1440.67676767677,-10.1818181818182)(1433.21212121212,-10.2121212121212)(1418.28282828283,-10.3333333333333)(1410.81818181818,-10.3636363636364)(1403.35353535354,-10.3939393939394)(1388.42424242424,-10.4545454545455)(1373.49494949495,-10.5151515151515)(1366.0303030303,-10.5454545454545)(1358.56565656566,-10.5757575757576)(1343.63636363636,-10.6363636363636)(1328.70707070707,-10.6363636363636)(1313.77777777778,-10.6363636363636)(1298.84848484848,-10.6363636363636)(1283.91919191919,-10.6363636363636)(1268.9898989899,-10.6363636363636)(1254.06060606061,-10.6363636363636)(1239.13131313131,-10.6363636363636)(1224.20202020202,-10.6363636363636)(1209.27272727273,-10.5757575757576)(1194.34343434343,-10.5757575757576)(1179.41414141414,-10.5757575757576)(1164.48484848485,-10.6363636363636)(1149.55555555556,-10.6363636363636)(1134.62626262626,-10.6363636363636)(1119.69696969697,-10.6363636363636)(1104.76767676768,-10.6363636363636)(1089.83838383838,-10.6363636363636)(1074.90909090909,-10.6363636363636)(1059.9797979798,-10.6363636363636)(1045.05050505051,-10.6363636363636)(1030.12121212121,-10.6363636363636)(1015.19191919192,-10.6363636363636)(1000.26262626263,-10.6363636363636)(985.333333333333,-10.6363636363636)(970.40404040404,-10.5757575757576)(962.939393939394,-10.5454545454545)(955.474747474747,-10.5151515151515)(940.545454545455,-10.4545454545455)(925.616161616162,-10.4090909090909)(918.151515151515,-10.3636363636364)(910.686868686869,-10.3333333333333)(895.757575757576,-10.2727272727273)(880.828282828283,-10.2121212121212)(873.363636363636,-10.1818181818182)(865.89898989899,-10.1515151515152)(850.969696969697,-10.030303030303)(843.505050505051,-10)(836.040404040404,-9.96969696969697)(821.111111111111,-9.84848484848485)(817.378787878788,-9.81818181818182)(806.181818181818,-9.72727272727273)(798.717171717172,-9.63636363636364)(791.252525252525,-9.54545454545455)(780.055555555556,-9.45454545454546)(776.323232323232,-9.42424242424243)(761.393939393939,-9.3030303030303)(753.929292929293,-9.27272727272727)(746.464646464646,-9.24242424242424)(731.535353535354,-9.12121212121212)(724.070707070707,-9.09090909090909)(716.606060606061,-9.06060606060606)(701.676767676768,-9)(686.747474747475,-9)(671.818181818182,-9.06060606060606)(664.353535353535,-9.09090909090909)(656.888888888889,-9.12121212121212)(641.959595959596,-9.18181818181818)(627.030303030303,-9.18181818181818)(612.10101010101,-9.18181818181818)(597.171717171717,-9.24242424242424)(589.707070707071,-9.27272727272727)(582.242424242424,-9.3030303030303)(567.313131313131,-9.36363636363636)(552.383838383838,-9.42424242424243)(537.454545454546,-9.42424242424243)(522.525252525253,-9.42424242424243)(507.59595959596,-9.36363636363636)(492.666666666667,-9.36363636363636)(477.737373737374,-9.36363636363636)(462.808080808081,-9.36363636363636)(447.878787878788,-9.36363636363636)(432.949494949495,-9.36363636363636)(418.020202020202,-9.36363636363636)(403.090909090909,-9.36363636363636)(388.161616161616,-9.3030303030303)(380.69696969697,-9.27272727272727)(373.232323232323,-9.24242424242424)(358.30303030303,-9.18181818181818)(343.373737373737,-9.18181818181818)(328.444444444444,-9.18181818181818)(313.515151515152,-9.18181818181818)(298.585858585859,-9.18181818181818)(283.656565656566,-9.18181818181818)(268.727272727273,-9.18181818181818)(253.79797979798,-9.18181818181818)(238.868686868687,-9.24242424242424)(223.939393939394,-9.24242424242424)(209.010101010101,-9.24242424242424)(194.080808080808,-9.18181818181818)(179.151515151515,-9.18181818181818)(164.222222222222,-9.12121212121212)(160.489898989899,-9.09090909090909)(149.292929292929,-9)(138.09595959596,-8.90909090909091)(134.363636363636,-8.88636363636364)(119.434343434343,-8.84090909090909)(104.505050505051,-8.79545454545455)(97.040404040404,-8.72727272727273)(89.5757575757576,-8.63636363636364)(78.3787878787879,-8.54545454545455)(74.6464646464647,-8.51515151515152)(59.7171717171717,-8.39393939393939)(55.9848484848485,-8.36363636363636)(44.7878787878788,-8.22727272727273)(42.2996632996633,-8.18181818181818)(37.3232323232323,-8)(37.3232323232323,-7.81818181818182)(37.3232323232323,-7.63636363636364)(42.2996632996633,-7.45454545454546)(44.7878787878788,-7.40909090909091)(55.9848484848485,-7.27272727272727)(59.7171717171717,-7.24242424242424)(74.6464646464647,-7.18181818181818)(89.5757575757576,-7.18181818181818)(104.505050505051,-7.18181818181818)(119.434343434343,-7.12121212121212)(134.363636363636,-7.12121212121212)(149.292929292929,-7.12121212121212)(164.222222222222,-7.18181818181818)(179.151515151515,-7.18181818181818)(194.080808080808,-7.18181818181818)(209.010101010101,-7.18181818181818)(223.939393939394,-7.18181818181818)(238.868686868687,-7.24242424242424)(246.333333333333,-7.27272727272727)(253.79797979798,-7.3030303030303)(268.727272727273,-7.36363636363636)(283.656565656566,-7.36363636363636)(298.585858585859,-7.36363636363636)(313.515151515152,-7.36363636363636)(328.444444444444,-7.36363636363636)(343.373737373737,-7.3030303030303)(350.838383838384,-7.27272727272727)(358.30303030303,-7.24242424242424)(373.232323232323,-7.18181818181818)(384.429292929293,-7.09090909090909)(388.161616161616,-7.04545454545455)(399.358585858586,-6.90909090909091)(403.090909090909,-6.87878787878788)(418.020202020202,-6.81818181818182)(432.949494949495,-6.75757575757576)(447.878787878788,-6.75757575757576)(455.343434343434,-6.72727272727273)(462.808080808081,-6.6969696969697)(477.737373737374,-6.6969696969697)(492.666666666667,-6.63636363636364)(507.59595959596,-6.63636363636364)(522.525252525253,-6.63636363636364)(537.454545454546,-6.63636363636364)(552.383838383838,-6.6969696969697)(559.848484848485,-6.72727272727273)(567.313131313131,-6.75757575757576)(582.242424242424,-6.81818181818182)(597.171717171717,-6.87878787878788)(604.636363636364,-6.90909090909091)(612.10101010101,-6.93939393939394)(627.030303030303,-7.06060606060606)(634.494949494949,-7.09090909090909)(641.959595959596,-7.12121212121212)(656.888888888889,-7.24242424242424)(664.353535353535,-7.27272727272727)(671.818181818182,-7.3030303030303)(686.747474747475,-7.42424242424243)(694.212121212121,-7.45454545454546)(701.676767676768,-7.48484848484849)(716.606060606061,-7.60606060606061)(724.070707070707,-7.63636363636364)(731.535353535354,-7.66666666666667)(746.464646464646,-7.72727272727273)(761.393939393939,-7.72727272727273)(776.323232323232,-7.72727272727273)(791.252525252525,-7.72727272727273)(806.181818181818,-7.72727272727273)(821.111111111111,-7.66666666666667)(828.575757575758,-7.63636363636364)(836.040404040404,-7.60606060606061)(850.969696969697,-7.48484848484849)(858.434343434343,-7.45454545454546)(865.89898989899,-7.42424242424243)(878.340067340067,-7.27272727272727)(880.828282828283,-7.24242424242424)(895.757575757576,-7.18181818181818)(910.686868686869,-7.12121212121212)(925.616161616162,-7.18181818181818)(940.545454545455,-7.12121212121212)(948.010101010101,-7.09090909090909)(955.474747474747,-7.06060606060606)(970.40404040404,-6.97727272727273)(985.333333333333,-6.97727272727273)(1000.26262626263,-6.97727272727273)(1015.19191919192,-7)(1030.12121212121,-7)(1045.05050505051,-7)(1059.9797979798,-7)(1074.90909090909,-7)(1089.83838383838,-7)(1104.76767676768,-7.06060606060606)(1112.23232323232,-7.09090909090909)(1119.69696969697,-7.12121212121212)(1134.62626262626,-7.18181818181818)(1149.55555555556,-7.18181818181818)(1164.48484848485,-7.18181818181818)(1179.41414141414,-7.18181818181818)(1194.34343434343,-7.18181818181818)(1209.27272727273,-7.18181818181818)(1224.20202020202,-7.18181818181818)(1239.13131313131,-7.18181818181818)(1254.06060606061,-7.18181818181818)(1268.9898989899,-7.24242424242424)(1276.45454545455,-7.27272727272727)(1283.91919191919,-7.29545454545455)(1298.84848484848,-7.36363636363636)(1313.77777777778,-7.36363636363636)(1328.70707070707,-7.36363636363636)(1343.63636363636,-7.42424242424243)(1351.10101010101,-7.45454545454546)(1358.56565656566,-7.48484848484849)(1373.49494949495,-7.60606060606061)(1377.22727272727,-7.63636363636364)(1388.42424242424,-7.72727272727273)(1399.62121212121,-7.81818181818182)(1403.35353535354,-7.84848484848485)(1418.28282828283,-7.96969696969697)(1422.01515151515,-8)(1433.21212121212,-8.09090909090909)(1440.67676767677,-8.18181818181818)(1448.14141414141,-8.27272727272727)(1455.60606060606,-8.36363636363636)(1463.07070707071,-8.45454545454545)(1474.26767676768,-8.54545454545455)(1478,-8.63636363636364)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1403.35353535354,-8.86363636363636)(1405.84175084175,-8.90909090909091)(1410.81818181818,-9.09090909090909)(1405.84175084175,-9.27272727272727)(1403.35353535354,-9.31818181818182)(1397.75505050505,-9.45454545454546)(1388.42424242424,-9.56818181818182)(1380.9595959596,-9.63636363636364)(1373.49494949495,-9.72727272727273)(1366.0303030303,-9.81818181818182)(1358.56565656566,-9.88636363636364)(1343.63636363636,-9.94545454545455)(1328.70707070707,-9.95454545454546)(1317.5101010101,-10)(1313.77777777778,-10.0227272727273)(1298.84848484848,-10.0909090909091)(1283.91919191919,-10.1363636363636)(1268.9898989899,-10.1363636363636)(1254.06060606061,-10.1666666666667)(1239.13131313131,-10.1666666666667)(1224.20202020202,-10.1666666666667)(1209.27272727273,-10.1363636363636)(1194.34343434343,-10.1272727272727)(1179.41414141414,-10.1272727272727)(1164.48484848485,-10.1272727272727)(1149.55555555556,-10.1363636363636)(1134.62626262626,-10.1363636363636)(1119.69696969697,-10.1363636363636)(1104.76767676768,-10.1363636363636)(1089.83838383838,-10.1363636363636)(1074.90909090909,-10.1363636363636)(1059.9797979798,-10.1363636363636)(1045.05050505051,-10.1272727272727)(1030.12121212121,-10.1272727272727)(1015.19191919192,-10.1272727272727)(1000.26262626263,-10.1363636363636)(985.333333333333,-10.1060606060606)(970.40404040404,-10.0545454545455)(959.207070707071,-10)(955.474747474747,-9.98484848484848)(940.545454545455,-9.95454545454546)(925.616161616162,-9.89393939393939)(910.686868686869,-9.83636363636364)(908.198653198653,-9.81818181818182)(895.757575757576,-9.72727272727273)(880.828282828283,-9.65909090909091)(878.340067340067,-9.63636363636364)(865.89898989899,-9.52272727272727)(858.434343434343,-9.45454545454546)(850.969696969697,-9.36363636363636)(845.371212121212,-9.27272727272727)(840.519191919192,-9.09090909090909)(839.772727272727,-8.90909090909091)(839.772727272727,-8.72727272727273)(843.505050505051,-8.54545454545455)(849.103535353535,-8.36363636363636)(850.969696969697,-8.33333333333333)(863.410774410775,-8.18181818181818)(865.89898989899,-8.15151515151515)(878.340067340067,-8)(880.828282828283,-7.97727272727273)(895.757575757576,-7.83636363636364)(898.245791245791,-7.81818181818182)(910.686868686869,-7.72727272727273)(925.616161616162,-7.69090909090909)(936.813131313131,-7.63636363636364)(940.545454545455,-7.61363636363636)(955.474747474747,-7.54545454545455)(970.40404040404,-7.47272727272727)(985.333333333333,-7.47272727272727)(1000.26262626263,-7.47272727272727)(1015.19191919192,-7.5)(1030.12121212121,-7.5)(1045.05050505051,-7.5)(1059.9797979798,-7.5)(1074.90909090909,-7.50909090909091)(1089.83838383838,-7.56818181818182)(1101.03535353535,-7.63636363636364)(1104.76767676768,-7.65151515151515)(1119.69696969697,-7.68181818181818)(1134.62626262626,-7.68181818181818)(1149.55555555556,-7.68181818181818)(1164.48484848485,-7.68181818181818)(1179.41414141414,-7.68181818181818)(1194.34343434343,-7.68181818181818)(1209.27272727273,-7.68181818181818)(1224.20202020202,-7.71212121212121)(1239.13131313131,-7.76363636363636)(1254.06060606061,-7.79545454545455)(1261.52525252525,-7.81818181818182)(1268.9898989899,-7.83636363636364)(1283.91919191919,-7.83636363636364)(1298.84848484848,-7.90909090909091)(1313.77777777778,-7.97727272727273)(1316.26599326599,-8)(1328.70707070707,-8.09090909090909)(1343.63636363636,-8.15909090909091)(1346.12457912458,-8.18181818181818)(1358.56565656566,-8.29545454545455)(1366.0303030303,-8.36363636363636)(1373.49494949495,-8.43181818181818)(1382.82575757576,-8.54545454545455)(1388.42424242424,-8.61363636363636)(1397.75505050505,-8.72727272727273)(1403.35353535354,-8.86363636363636)};

\addplot [fill=green!60!black,draw=none,forget plot] coordinates{ (1373.49494949495,-9)(1377.22727272727,-9.09090909090909)(1373.49494949495,-9.18181818181818)(1371.00673400673,-9.27272727272727)(1360.43181818182,-9.45454545454546)(1358.56565656566,-9.47727272727273)(1343.63636363636,-9.60606060606061)(1339.90404040404,-9.63636363636364)(1328.70707070707,-9.68181818181818)(1313.77777777778,-9.72727272727273)(1298.84848484848,-9.79545454545455)(1295.11616161616,-9.81818181818182)(1283.91919191919,-9.86363636363636)(1268.9898989899,-9.86363636363636)(1254.06060606061,-9.87272727272727)(1239.13131313131,-9.87272727272727)(1224.20202020202,-9.87272727272727)(1209.27272727273,-9.86363636363636)(1194.34343434343,-9.83333333333333)(1179.41414141414,-9.83333333333333)(1164.48484848485,-9.83333333333333)(1149.55555555556,-9.86363636363636)(1134.62626262626,-9.83636363636364)(1119.69696969697,-9.83636363636364)(1104.76767676768,-9.83636363636364)(1089.83838383838,-9.86363636363636)(1074.90909090909,-9.86363636363636)(1059.9797979798,-9.86363636363636)(1045.05050505051,-9.83333333333333)(1030.12121212121,-9.83333333333333)(1015.19191919192,-9.83333333333333)(1000.26262626263,-9.86363636363636)(985.333333333333,-9.83333333333333)(981.60101010101,-9.81818181818182)(970.40404040404,-9.76363636363636)(955.474747474747,-9.71212121212121)(940.545454545455,-9.68181818181818)(929.348484848485,-9.63636363636364)(925.616161616162,-9.61363636363636)(910.686868686869,-9.54545454545455)(898.245791245791,-9.45454545454546)(895.757575757576,-9.43181818181818)(880.828282828283,-9.32727272727273)(875.229797979798,-9.27272727272727)(865.89898989899,-9.12121212121212)(864.654882154882,-9.09090909090909)(862.166666666667,-8.90909090909091)(862.166666666667,-8.72727272727273)(865.89898989899,-8.59090909090909)(867.765151515152,-8.54545454545455)(876.349494949495,-8.36363636363636)(880.828282828283,-8.30909090909091)(893.891414141414,-8.18181818181818)(895.757575757576,-8.16363636363636)(910.686868686869,-8.02272727272727)(918.151515151515,-8)(925.616161616162,-7.98484848484848)(940.545454545455,-7.94545454545455)(955.474747474748,-7.88636363636364)(962.939393939394,-7.81818181818182)(970.40404040404,-7.77272727272727)(985.333333333333,-7.77272727272727)(1000.26262626263,-7.77272727272727)(1015.19191919192,-7.77272727272727)(1030.12121212121,-7.77272727272727)(1045.05050505051,-7.77272727272727)(1059.9797979798,-7.77272727272727)(1074.90909090909,-7.8030303030303)(1078.64141414141,-7.81818181818182)(1089.83838383838,-7.90909090909091)(1104.76767676768,-7.94545454545455)(1119.69696969697,-7.95454545454546)(1134.62626262626,-7.95454545454546)(1149.55555555556,-7.95454545454546)(1164.48484848485,-7.95454545454546)(1179.41414141414,-7.95454545454546)(1194.34343434343,-7.95454545454546)(1209.27272727273,-7.95454545454546)(1224.20202020202,-7.98484848484848)(1227.93434343434,-8)(1239.13131313131,-8.06818181818182)(1254.06060606061,-8.12727272727273)(1268.9898989899,-8.16363636363636)(1283.91919191919,-8.13636363636364)(1295.11616161616,-8.18181818181818)(1298.84848484848,-8.20454545454546)(1313.77777777778,-8.27272727272727)(1326.21885521886,-8.36363636363636)(1328.70707070707,-8.38636363636364)(1343.63636363636,-8.49090909090909)(1349.23484848485,-8.54545454545455)(1358.56565656566,-8.6969696969697)(1360.43181818182,-8.72727272727273)(1371.00673400673,-8.90909090909091)(1373.49494949495,-9)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (492.666666666667,-7.95454545454546)(503.863636363636,-7.81818181818182)(507.59595959596,-7.78787878787879)(522.525252525253,-7.72727272727273)(537.454545454546,-7.78787878787879)(541.186868686869,-7.81818181818182)(552.383838383838,-7.90909090909091)(559.848484848485,-8)(567.313131313131,-8.13636363636364)(569.801346801347,-8.18181818181818)(567.313131313131,-8.27272727272727)(563.580808080808,-8.36363636363636)(552.383838383838,-8.45454545454545)(537.454545454546,-8.45454545454545)(522.525252525253,-8.45454545454545)(507.59595959596,-8.39393939393939)(503.863636363636,-8.36363636363636)(492.666666666667,-8.22727272727273)(488.934343434343,-8.18181818181818)(488.934343434343,-8)(492.666666666667,-7.95454545454546)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273)(447.878787878788,0)(1089.83838383838,-12.1818181818182)(313.515151515152,-6.54545454545455)(268.727272727273,-11.0909090909091)(1000.26262626263,-9.27272727272727)(477.737373737374,-8.72727272727273)(1478,-8.54545454545455)(1089.83838383838,-5.63636363636364)(0,-10.9090909090909)(0,-6.36363636363636)(358.30303030303,-4.36363636363636)(1478,-6.36363636363636)(1239.13131313131,-10.1818181818182)(761.393939393939,-8.36363636363636)(1373.49494949495,-12.1818181818182)(597.171717171717,-5.45454545454546)(223.939393939394,-9.09090909090909)(985.333333333333,-7.27272727272727)(895.757575757576,-11.2727272727273)(1298.84848484848,-5.45454545454546)(865.89898989899,-5.63636363636364)(462.808080808081,-10.3636363636364)(522.525252525253,-7.27272727272727)(179.151515151515,-7.63636363636364)(164.222222222222,-5.81818181818182)(776.323232323232,-9.63636363636364)(1239.13131313131,-8.90909090909091)(1478,-9.63636363636364)(1074.90909090909,-10.7272727272727)(119.434343434343,-9.81818181818182)(1478,-7.45454545454546)(0,-7.45454545454546)(1179.41414141414,-6.72727272727273)(447.878787878788,-6)(1283.91919191919,-11.0909090909091)(612.10101010101,-9.27272727272727)(1074.90909090909,-8.18181818181818)(343.373737373737,-8)(328.444444444444,-9.81818181818182)(836.040404040404,-7.45454545454546)(910.686868686869,-10.1818181818182)(641.959595959596,-7.81818181818182)(1358.56565656566,-9.45454545454546)(910.686868686869,-8.54545454545455)(149.292929292929,-6.72727272727273)(1343.63636363636,-7.27272727272727)(1119.69696969697,-9.63636363636364)(1015.19191919192,-6.72727272727273)(597.171717171717,-6.54545454545455)(119.434343434343,-8.54545454545455)(388.161616161616,-7.27272727272727)(1358.56565656566,-8.36363636363636)(1478,-10.1818181818182)(447.878787878788,-9.45454545454546)(1149.55555555556,-10.9090909090909)(1149.55555555556,-7.81818181818182)(1358.56565656566,-10.7272727272727)(328.444444444444,-8.72727272727273)(567.313131313131,-8.36363636363636)(0,-7.09090909090909)(850.969696969697,-9.27272727272727)(1478,-8.90909090909091)(1000.26262626263,-10.7272727272727)(701.676767676768,-7.45454545454546)(1134.62626262626,-8.90909090909091)(671.818181818182,-9.09090909090909)(253.79797979798,-7.27272727272727)(940.545454545455,-8)(462.808080808081,-6.72727272727273)(880.828282828283,-10.3636363636364)(1268.9898989899,-7.09090909090909)(104.505050505051,-7.81818181818182)(462.808080808081,-8)(1313.77777777778,-9.81818181818182)(1104.76767676768,-6.90909090909091)(895.757575757576,-7.27272727272727)(134.363636363636,-8.90909090909091)(1030.12121212121,-9.81818181818182)(1433.21212121212,-8)(1254.06060606061,-8.36363636363636)(253.79797979798,-8.18181818181818)(537.454545454546,-9.27272727272727)(298.585858585859,-9.27272727272727)(791.252525252525,-8.54545454545455)(612.10101010101,-6.90909090909091)(29.8585858585859,-8)(1015.19191919192,-8.72727272727273)(1478,-10)(119.434343434343,-7.09090909090909)(1179.41414141414,-9.81818181818182)(836.040404040404,-10)(1358.56565656566,-8.90909090909091)(1373.49494949495,-10.5454545454545)(1089.83838383838,-7.63636363636364)(656.888888888889,-8.36363636363636)(343.373737373737,-7.09090909090909)(1478,-9.09090909090909)(403.090909090909,-8.72727272727273)(1194.34343434343,-10.5454545454545)(940.545454545455,-7.09090909090909) 
};
\end{axis}
\end{tikzpicture}%