### 2013-03-21 figurefactory/files/limno_bgape_imp_class_20.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,3.30572501897308e-06)(283.656565656566,8.2640871684445e-06)(298.585858585859,1.32221788827669e-05)(313.515151515152,1.48748160261842e-05)(328.444444444444,1.48748160261842e-05)(343.373737373737,1.48748160261842e-05)(358.30303030303,1.48748160261842e-05)(373.232323232323,1.48748160261842e-05)(388.161616161616,1.48748160261842e-05)(403.090909090909,1.48748160261842e-05)(418.020202020202,1.48748160261842e-05)(432.949494949495,1.48748160261842e-05)(447.878787878788,1.48748160261842e-05)(462.808080808081,1.48748160261842e-05)(477.737373737374,1.48748160261842e-05)(492.666666666667,1.48748160261842e-05)(507.59595959596,1.48748160261842e-05)(522.525252525253,1.48748160261842e-05)(537.454545454546,1.48748160261842e-05)(552.383838383838,1.48748160261842e-05)(567.313131313131,1.48748160261842e-05)(582.242424242424,1.48748160261842e-05)(597.171717171717,1.48748160261842e-05)(612.10101010101,1.48748160261842e-05)(627.030303030303,1.48748160261842e-05)(641.959595959596,1.48748160261842e-05)(656.888888888889,9.91681445518895e-06)(671.818181818182,4.95854245421226e-06)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478.00040715143,-1.27272727272727)(1478.00081428065,-1.45454545454545)(1478.00122138767,-1.63636363636364)(1478.00122138767,-1.81818181818182)(1478.00122138767,-2)(1478.00122138767,-2.18181818181818)(1478.00122138767,-2.36363636363636)(1478.00122138767,-2.54545454545455)(1478.00122138767,-2.72727272727273)(1478.00122138767,-2.90909090909091)(1478.00122138767,-3.09090909090909)(1478.00081428065,-3.27272727272727)(1478.00040715143,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00054286364,-4.18181818181818)(1478.00054286364,-4.36363636363636)(1478.00054286364,-4.54545454545455)(1478.00040715143,-4.72727272727273)(1478.00081428065,-4.90909090909091)(1478.00122138767,-5.09090909090909)(1478.00122138767,-5.27272727272727)(1478.00122138767,-5.45454545454546)(1478.00122138767,-5.63636363636364)(1478.00122138767,-5.81818181818182)(1478.00122138767,-6)(1478.00122138767,-6.18181818181818)(1478.00122138767,-6.36363636363636)(1478.00122138767,-6.54545454545455)(1478.00122138767,-6.72727272727273)(1478.00122138767,-6.90909090909091)(1478.00122138767,-7.09090909090909)(1478.00122138767,-7.27272727272727)(1478.00122138767,-7.45454545454546)(1478.00122138767,-7.63636363636364)(1478.00122138767,-7.81818181818182)(1478.00122138767,-8)(1478.00122138767,-8.18181818181818)(1478.00122138767,-8.36363636363636)(1478.00122138767,-8.54545454545455)(1478.00122138767,-8.72727272727273)(1478.00122138767,-8.90909090909091)(1478.00122138767,-9.09090909090909)(1478.00122138767,-9.27272727272727)(1478.00122138767,-9.45454545454546)(1478.00122138767,-9.63636363636364)(1478.00122138767,-9.81818181818182)(1478.00122138767,-10)(1478.00122138767,-10.1818181818182)(1478.00122138767,-10.3636363636364)(1478.00122138767,-10.5454545454545)(1478.00149278002,-10.7272727272727)(1478.00149278002,-10.9090909090909)(1478.00149278002,-11.0909090909091)(1478.00122138767,-11.2727272727273)(1478.00122138767,-11.4545454545455)(1478.00122138767,-11.6363636363636)(1478.00122138767,-11.8181818181818)(1478.00122138767,-12)(1478.00122138767,-12.1818181818182)(1478.00122138767,-12.3636363636364)(1478.00122138767,-12.5454545454545)(1478.00122138767,-12.7272727272727)(1478.00122138767,-12.9090909090909)(1478.00122138767,-13.0909090909091)(1478.00081428065,-13.2727272727273)(1478.00040715143,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(-0.000407151430307532,-16.7272727272727)(-0.000814280653561821,-16.5454545454545)(-0.0012213876715861,-16.3636363636364)(-0.0012213876715861,-16.1818181818182)(-0.0012213876715861,-16)(-0.0012213876715861,-15.8181818181818)(-0.0012213876715861,-15.6363636363636)(-0.0012213876715861,-15.4545454545455)(-0.0012213876715861,-15.2727272727273)(-0.0012213876715861,-15.0909090909091)(-0.0012213876715861,-14.9090909090909)(-0.0012213876715861,-14.7272727272727)(-0.000814280653561821,-14.5454545454545)(-0.000407151430307532,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(-0.000407151430307532,-12.5454545454545)(-0.000814280653561821,-12.3636363636364)(-0.0012213876715861,-12.1818181818182)(-0.0012213876715861,-12)(-0.0012213876715861,-11.8181818181818)(-0.0012213876715861,-11.6363636363636)(-0.0012213876715861,-11.4545454545455)(-0.0012213876715861,-11.2727272727273)(-0.0012213876715861,-11.0909090909091)(-0.0012213876715861,-10.9090909090909)(-0.0012213876715861,-10.7272727272727)(-0.0012213876715861,-10.5454545454545)(-0.0012213876715861,-10.3636363636364)(-0.0012213876715861,-10.1818181818182)(-0.0012213876715861,-10)(-0.0012213876715861,-9.81818181818182)(-0.0012213876715861,-9.63636363636364)(-0.0012213876715861,-9.45454545454546)(-0.0012213876715861,-9.27272727272727)(-0.0012213876715861,-9.09090909090909)(-0.0012213876715861,-8.90909090909091)(-0.00149278001492805,-8.72727272727273)(-0.00149278001492805,-8.54545454545455)(-0.00149278001492805,-8.36363636363636)(-0.0012213876715861,-8.18181818181818)(-0.0012213876715861,-8)(-0.0012213876715861,-7.81818181818182)(-0.0012213876715861,-7.63636363636364)(-0.0012213876715861,-7.45454545454546)(-0.0012213876715861,-7.27272727272727)(-0.0012213876715861,-7.09090909090909)(-0.0012213876715861,-6.90909090909091)(-0.0012213876715861,-6.72727272727273)(-0.0012213876715861,-6.54545454545455)(-0.0012213876715861,-6.36363636363636)(-0.0012213876715861,-6.18181818181818)(-0.0012213876715861,-6)(-0.0012213876715861,-5.81818181818182)(-0.0012213876715861,-5.63636363636364)(-0.000814280653561821,-5.45454545454546)(-0.000407151430307532,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-5.36363636363636)(14.9292929292929,-5.36363636363636)(29.8585858585859,-5.36363636363636)(44.7878787878788,-5.3030303030303)(52.2525252525253,-5.27272727272727)(59.7171717171717,-5.24242424242424)(74.6464646464647,-5.18181818181818)(89.5757575757576,-5.18181818181818)(104.505050505051,-5.12121212121212)(111.969696969697,-5.09090909090909)(119.434343434343,-5.06060606060606)(134.363636363636,-4.93939393939394)(141.828282828283,-4.90909090909091)(149.292929292929,-4.87878787878788)(164.222222222222,-4.75757575757576)(167.954545454545,-4.72727272727273)(179.151515151515,-4.63636363636364)(190.348484848485,-4.54545454545455)(194.080808080808,-4.51515151515151)(209.010101010101,-4.39393939393939)(212.742424242424,-4.36363636363636)(223.939393939394,-4.27272727272727)(235.136363636364,-4.18181818181818)(238.868686868687,-4.15151515151515)(253.79797979798,-4.03030303030303)(261.262626262626,-4)(268.727272727273,-3.96969696969697)(283.656565656566,-3.84848484848485)(291.121212121212,-3.81818181818182)(298.585858585859,-3.78787878787879)(313.515151515152,-3.72727272727273)(328.444444444444,-3.72727272727273)(343.373737373737,-3.72727272727273)(358.30303030303,-3.72727272727273)(373.232323232323,-3.72727272727273)(388.161616161616,-3.72727272727273)(403.090909090909,-3.72727272727273)(418.020202020202,-3.72727272727273)(432.949494949495,-3.72727272727273)(447.878787878788,-3.72727272727273)(462.808080808081,-3.72727272727273)(477.737373737374,-3.72727272727273)(492.666666666667,-3.72727272727273)(507.59595959596,-3.72727272727273)(522.525252525253,-3.72727272727273)(537.454545454546,-3.72727272727273)(552.383838383838,-3.72727272727273)(567.313131313131,-3.72727272727273)(582.242424242424,-3.72727272727273)(597.171717171717,-3.72727272727273)(612.10101010101,-3.72727272727273)(627.030303030303,-3.72727272727273)(641.959595959596,-3.72727272727273)(656.888888888889,-3.72727272727273)(671.818181818182,-3.66666666666667)(679.282828282828,-3.63636363636364)(686.747474747475,-3.60606060606061)(701.676767676768,-3.54545454545455)(716.606060606061,-3.54545454545455)(731.535353535354,-3.54545454545455)(746.464646464646,-3.54545454545455)(761.393939393939,-3.60606060606061)(768.858585858586,-3.63636363636364)(776.323232323232,-3.66666666666667)(791.252525252525,-3.72727272727273)(806.181818181818,-3.72727272727273)(821.111111111111,-3.78787878787879)(828.575757575758,-3.81818181818182)(836.040404040404,-3.84848484848485)(850.969696969697,-3.96969696969697)(858.434343434343,-4)(865.89898989899,-4.03030303030303)(880.828282828283,-4.09090909090909)(895.757575757576,-4.15151515151515)(903.222222222222,-4.18181818181818)(910.686868686869,-4.21212121212121)(925.616161616162,-4.27272727272727)(940.545454545455,-4.27272727272727)(955.474747474747,-4.27272727272727)(970.40404040404,-4.27272727272727)(985.333333333333,-4.27272727272727)(1000.26262626263,-4.27272727272727)(1015.19191919192,-4.27272727272727)(1030.12121212121,-4.27272727272727)(1045.05050505051,-4.27272727272727)(1059.9797979798,-4.27272727272727)(1074.90909090909,-4.27272727272727)(1089.83838383838,-4.27272727272727)(1104.76767676768,-4.27272727272727)(1119.69696969697,-4.27272727272727)(1134.62626262626,-4.27272727272727)(1149.55555555556,-4.21212121212121)(1157.0202020202,-4.18181818181818)(1164.48484848485,-4.15151515151515)(1179.41414141414,-4.09090909090909)(1194.34343434343,-4.09090909090909)(1209.27272727273,-4.03030303030303)(1216.73737373737,-4)(1224.20202020202,-3.96969696969697)(1239.13131313131,-3.90909090909091)(1254.06060606061,-3.90909090909091)(1268.9898989899,-3.90909090909091)(1283.91919191919,-3.90909090909091)(1298.84848484848,-3.90909090909091)(1313.77777777778,-3.96969696969697)(1321.24242424242,-4)(1328.70707070707,-4.03030303030303)(1343.63636363636,-4.15151515151515)(1347.36868686869,-4.18181818181818)(1358.56565656566,-4.27272727272727)(1369.76262626263,-4.36363636363636)(1373.49494949495,-4.39393939393939)(1388.42424242424,-4.51515151515151)(1395.88888888889,-4.54545454545455)(1403.35353535354,-4.57575757575758)(1418.28282828283,-4.63636363636364)(1433.21212121212,-4.6969696969697)(1440.67676767677,-4.72727272727273)(1448.14141414141,-4.75757575757576)(1463.07070707071,-4.81818181818182)(1478,-4.81818181818182)(1478.00020357016,-4.90909090909091)(1478.00061069384,-5.09090909090909)(1478.00061069384,-5.27272727272727)(1478.00061069384,-5.45454545454546)(1478.00061069384,-5.63636363636364)(1478.00061069384,-5.81818181818182)(1478.00061069384,-6)(1478.00061069384,-6.18181818181818)(1478.00061069384,-6.36363636363636)(1478.00061069384,-6.54545454545455)(1478.00061069384,-6.72727272727273)(1478.00061069384,-6.90909090909091)(1478.00061069384,-7.09090909090909)(1478.00061069384,-7.27272727272727)(1478.00061069384,-7.45454545454546)(1478.00061069384,-7.63636363636364)(1478.00061069384,-7.81818181818182)(1478.00061069384,-8)(1478.00061069384,-8.18181818181818)(1478.00061069384,-8.36363636363636)(1478.00061069384,-8.54545454545455)(1478.00061069384,-8.72727272727273)(1478.00061069384,-8.90909090909091)(1478.00061069384,-9.09090909090909)(1478.00061069384,-9.27272727272727)(1478.00061069384,-9.45454545454546)(1478.00061069384,-9.63636363636364)(1478.00061069384,-9.81818181818182)(1478.00061069384,-10)(1478.00061069384,-10.1818181818182)(1478.00061069384,-10.3636363636364)(1478.00061069384,-10.5454545454545)(1478.00088209728,-10.7272727272727)(1478.00088209728,-10.9090909090909)(1478.00088209728,-11.0909090909091)(1478.00061069384,-11.2727272727273)(1478.00061069384,-11.4545454545455)(1478.00061069384,-11.6363636363636)(1478.00061069384,-11.8181818181818)(1478.00061069384,-12)(1478.00061069384,-12.1818181818182)(1478.00061069384,-12.3636363636364)(1478.00061069384,-12.5454545454545)(1478.00061069384,-12.7272727272727)(1478.00061069384,-12.9090909090909)(1478.00061069384,-13.0909090909091)(1478.00020357016,-13.2727272727273)(1478,-13.3636363636364)(1463.07070707071,-13.3636363636364)(1448.14141414141,-13.3636363636364)(1433.21212121212,-13.3636363636364)(1418.28282828283,-13.3636363636364)(1403.35353535354,-13.4242424242424)(1395.88888888889,-13.4545454545455)(1388.42424242424,-13.4848484848485)(1373.49494949495,-13.6060606060606)(1366.0303030303,-13.6363636363636)(1358.56565656566,-13.6666666666667)(1343.63636363636,-13.7878787878788)(1339.90404040404,-13.8181818181818)(1328.70707070707,-13.9090909090909)(1317.5101010101,-14)(1313.77777777778,-14.030303030303)(1298.84848484848,-14.1515151515152)(1295.11616161616,-14.1818181818182)(1283.91919191919,-14.3181818181818)(1281.43097643098,-14.3636363636364)(1271.47811447811,-14.5454545454545)(1268.9898989899,-14.6363636363636)(1266.50168350168,-14.7272727272727)(1256.54882154882,-14.9090909090909)(1254.06060606061,-15)(1251.57239057239,-15.0909090909091)(1246.59595959596,-15.2727272727273)(1241.61952861953,-15.4545454545455)(1239.13131313131,-15.5)(1231.66666666667,-15.6363636363636)(1224.20202020202,-15.7272727272727)(1213.00505050505,-15.8181818181818)(1209.27272727273,-15.8484848484848)(1194.34343434343,-15.9090909090909)(1179.41414141414,-15.9090909090909)(1164.48484848485,-15.9090909090909)(1149.55555555556,-15.9090909090909)(1134.62626262626,-15.8484848484848)(1127.16161616162,-15.8181818181818)(1119.69696969697,-15.7878787878788)(1104.76767676768,-15.6666666666667)(1097.30303030303,-15.6363636363636)(1089.83838383838,-15.6060606060606)(1077.39730639731,-15.4545454545455)(1074.90909090909,-15.4242424242424)(1062.46801346801,-15.2727272727273)(1059.9797979798,-15.2272727272727)(1052.51515151515,-15.0909090909091)(1045.05050505051,-15)(1037.58585858586,-14.9090909090909)(1030.12121212121,-14.8181818181818)(1022.65656565657,-14.7272727272727)(1015.19191919192,-14.6363636363636)(1003.99494949495,-14.5454545454545)(1000.26262626263,-14.5151515151515)(985.333333333333,-14.3939393939394)(981.60101010101,-14.3636363636364)(970.40404040404,-14.2727272727273)(959.207070707071,-14.1818181818182)(955.474747474747,-14.1515151515152)(940.545454545455,-14.030303030303)(933.080808080808,-14)(925.616161616162,-13.969696969697)(910.686868686869,-13.9090909090909)(895.757575757576,-13.8484848484849)(888.292929292929,-13.8181818181818)(880.828282828283,-13.7878787878788)(865.89898989899,-13.7272727272727)(850.969696969697,-13.7272727272727)(836.040404040404,-13.6666666666667)(828.575757575758,-13.6363636363636)(821.111111111111,-13.6060606060606)(806.181818181818,-13.5454545454545)(791.252525252525,-13.5454545454545)(776.323232323232,-13.5454545454545)(761.393939393939,-13.5454545454545)(746.464646464646,-13.5454545454545)(731.535353535354,-13.6060606060606)(724.070707070707,-13.6363636363636)(716.606060606061,-13.6666666666667)(701.676767676768,-13.7272727272727)(686.747474747475,-13.7272727272727)(671.818181818182,-13.7878787878788)(668.085858585859,-13.8181818181818)(656.888888888889,-13.9090909090909)(645.691919191919,-14)(641.959595959596,-14.030303030303)(627.030303030303,-14.1515151515152)(623.29797979798,-14.1818181818182)(612.10101010101,-14.2727272727273)(604.636363636364,-14.3636363636364)(597.171717171717,-14.5)(594.683501683502,-14.5454545454545)(584.73063973064,-14.7272727272727)(582.242424242424,-14.8181818181818)(579.754208754209,-14.9090909090909)(569.801346801347,-15.0909090909091)(567.313131313131,-15.1818181818182)(564.824915824916,-15.2727272727273)(559.848484848485,-15.4545454545455)(554.872053872054,-15.6363636363636)(552.383838383838,-15.7272727272727)(549.895622895623,-15.8181818181818)(539.942760942761,-16)(537.454545454546,-16.0454545454545)(529.989898989899,-16.1818181818182)(522.525252525253,-16.2727272727273)(511.328282828283,-16.3636363636364)(507.59595959596,-16.3939393939394)(492.666666666667,-16.5151515151515)(485.20202020202,-16.5454545454545)(477.737373737374,-16.5757575757576)(462.808080808081,-16.6969696969697)(459.075757575758,-16.7272727272727)(447.878787878788,-16.8181818181818)(436.681818181818,-16.9090909090909)(432.949494949495,-16.9393939393939)(418.020202020202,-17.0606060606061)(410.555555555556,-17.0909090909091)(403.090909090909,-17.1212121212121)(388.161616161616,-17.2424242424242)(384.429292929293,-17.2727272727273)(373.232323232323,-17.3636363636364)(362.035353535354,-17.4545454545455)(358.30303030303,-17.4848484848485)(343.373737373737,-17.6060606060606)(335.909090909091,-17.6363636363636)(328.444444444444,-17.6666666666667)(313.515151515152,-17.7272727272727)(298.585858585859,-17.7272727272727)(283.656565656566,-17.6666666666667)(276.191919191919,-17.6363636363636)(268.727272727273,-17.6060606060606)(253.79797979798,-17.4848484848485)(246.333333333333,-17.4545454545455)(238.868686868687,-17.4242424242424)(223.939393939394,-17.3030303030303)(216.474747474747,-17.2727272727273)(209.010101010101,-17.2424242424242)(194.080808080808,-17.1212121212121)(186.616161616162,-17.0909090909091)(179.151515151515,-17.0606060606061)(164.222222222222,-17)(149.292929292929,-16.9393939393939)(141.828282828283,-16.9090909090909)(134.363636363636,-16.8787878787879)(119.434343434343,-16.8181818181818)(104.505050505051,-16.7575757575758)(97.040404040404,-16.7272727272727)(89.5757575757576,-16.6969696969697)(74.6464646464647,-16.6363636363636)(59.7171717171717,-16.6363636363636)(44.7878787878788,-16.6363636363636)(29.8585858585859,-16.6363636363636)(14.9292929292929,-16.6363636363636)(0,-16.6363636363636)(-0.000203570163392941,-16.5454545454545)(-0.000610693835794708,-16.3636363636364)(-0.000610693835794708,-16.1818181818182)(-0.000610693835794708,-16)(-0.000610693835794708,-15.8181818181818)(-0.000610693835794708,-15.6363636363636)(-0.000610693835794708,-15.4545454545455)(-0.000610693835794708,-15.2727272727273)(-0.000610693835794708,-15.0909090909091)(-0.000610693835794708,-14.9090909090909)(-0.000610693835794708,-14.7272727272727)(-0.000203570163392941,-14.5454545454545)(0,-14.4545454545455)(14.9292929292929,-14.4545454545455)(29.8585858585859,-14.4545454545455)(44.7878787878788,-14.3939393939394)(52.2525252525253,-14.3636363636364)(59.7171717171717,-14.3333333333333)(74.6464646464647,-14.2727272727273)(89.5757575757576,-14.2121212121212)(93.3080808080808,-14.1818181818182)(104.505050505051,-14.0909090909091)(111.969696969697,-14)(119.434343434343,-13.9090909090909)(126.89898989899,-13.8181818181818)(134.363636363636,-13.6818181818182)(136.851851851852,-13.6363636363636)(141.828282828283,-13.4545454545455)(136.851851851852,-13.2727272727273)(134.363636363636,-13.2272727272727)(126.89898989899,-13.0909090909091)(119.434343434343,-13)(111.969696969697,-12.9090909090909)(104.505050505051,-12.8181818181818)(93.3080808080808,-12.7272727272727)(89.5757575757576,-12.6969696969697)(74.6464646464647,-12.6363636363636)(59.7171717171717,-12.5757575757576)(52.2525252525253,-12.5454545454545)(44.7878787878788,-12.5151515151515)(29.8585858585859,-12.4545454545455)(14.9292929292929,-12.4545454545455)(0,-12.4545454545455)(-0.000203570163392941,-12.3636363636364)(-0.000610693835794708,-12.1818181818182)(-0.000610693835794708,-12)(-0.000610693835794708,-11.8181818181818)(-0.000610693835794708,-11.6363636363636)(-0.000610693835794708,-11.4545454545455)(-0.000610693835794708,-11.2727272727273)(-0.000610693835794708,-11.0909090909091)(-0.000610693835794708,-10.9090909090909)(-0.000610693835794708,-10.7272727272727)(-0.000610693835794708,-10.5454545454545)(-0.000610693835794708,-10.3636363636364)(-0.000610693835794708,-10.1818181818182)(-0.000610693835794708,-10)(-0.000610693835794708,-9.81818181818182)(-0.000610693835794708,-9.63636363636364)(-0.000610693835794708,-9.45454545454546)(-0.000610693835794708,-9.27272727272727)(-0.000610693835794708,-9.09090909090909)(-0.000610693835794708,-8.90909090909091)(-0.000882097281547789,-8.72727272727273)(-0.000882097281547789,-8.54545454545455)(-0.000882097281547789,-8.36363636363636)(-0.000610693835794708,-8.18181818181818)(-0.000610693835794708,-8)(-0.000610693835794708,-7.81818181818182)(-0.000610693835794708,-7.63636363636364)(-0.000610693835794708,-7.45454545454546)(-0.000610693835794708,-7.27272727272727)(-0.000610693835794708,-7.09090909090909)(-0.000610693835794708,-6.90909090909091)(-0.000610693835794708,-6.72727272727273)(-0.000610693835794708,-6.54545454545455)(-0.000610693835794708,-6.36363636363636)(-0.000610693835794708,-6.18181818181818)(-0.000610693835794708,-6)(-0.000610693835794708,-5.81818181818182)(-0.000610693835794708,-5.63636363636364)(-0.000203570163392941,-5.45454545454546)(0,-5.36363636363636)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1478,-1.36363636363636)(1478.00020357016,-1.45454545454545)(1478.00061069384,-1.63636363636364)(1478.00061069384,-1.81818181818182)(1478.00061069384,-2)(1478.00061069384,-2.18181818181818)(1478.00061069384,-2.36363636363636)(1478.00061069384,-2.54545454545455)(1478.00061069384,-2.72727272727273)(1478.00061069384,-2.90909090909091)(1478.00061069384,-3.09090909090909)(1478.00020357016,-3.27272727272727)(1478,-3.36363636363636)(1463.07070707071,-3.36363636363636)(1448.14141414141,-3.3030303030303)(1440.67676767677,-3.27272727272727)(1433.21212121212,-3.24242424242424)(1418.28282828283,-3.18181818181818)(1403.35353535354,-3.18181818181818)(1388.42424242424,-3.18181818181818)(1373.49494949495,-3.18181818181818)(1358.56565656566,-3.18181818181818)(1343.63636363636,-3.18181818181818)(1328.70707070707,-3.12121212121212)(1321.24242424242,-3.09090909090909)(1313.77777777778,-3.06060606060606)(1298.84848484848,-2.93939393939394)(1295.11616161616,-2.90909090909091)(1283.91919191919,-2.77272727272727)(1281.43097643098,-2.72727272727273)(1271.47811447811,-2.54545454545455)(1268.9898989899,-2.5)(1261.52525252525,-2.36363636363636)(1254.06060606061,-2.22727272727273)(1251.57239057239,-2.18181818181818)(1241.61952861953,-2)(1239.13131313131,-1.90909090909091)(1236.6430976431,-1.81818181818182)(1236.6430976431,-1.63636363636364)(1239.13131313131,-1.59090909090909)(1250.32828282828,-1.45454545454545)(1254.06060606061,-1.42424242424242)(1268.9898989899,-1.36363636363636)(1283.91919191919,-1.36363636363636)(1298.84848484848,-1.36363636363636)(1313.77777777778,-1.36363636363636)(1328.70707070707,-1.36363636363636)(1343.63636363636,-1.36363636363636)(1358.56565656566,-1.36363636363636)(1373.49494949495,-1.36363636363636)(1388.42424242424,-1.36363636363636)(1403.35353535354,-1.36363636363636)(1418.28282828283,-1.36363636363636)(1433.21212121212,-1.36363636363636)(1448.14141414141,-1.36363636363636)(1463.07070707071,-1.36363636363636)(1478,-1.36363636363636)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (283.656565656566,8.26408717619588e-07)(298.585858585859,5.78470326217944e-06)(313.515151515152,7.4374080130921e-06)(328.444444444444,7.4374080130921e-06)(343.373737373737,7.4374080130921e-06)(358.30303030303,7.4374080130921e-06)(373.232323232323,7.4374080130921e-06)(388.161616161616,7.4374080130921e-06)(403.090909090909,7.4374080130921e-06)(418.020202020202,7.4374080130921e-06)(432.949494949495,7.4374080130921e-06)(447.878787878788,7.4374080130921e-06)(462.808080808081,7.4374080130921e-06)(477.737373737374,7.4374080130921e-06)(492.666666666667,7.4374080130921e-06)(507.59595959596,7.4374080130921e-06)(522.525252525253,7.4374080130921e-06)(537.454545454546,7.4374080130921e-06)(552.383838383838,7.4374080130921e-06)(567.313131313131,7.4374080130921e-06)(582.242424242424,7.4374080130921e-06)(597.171717171717,7.4374080130921e-06)(612.10101010101,7.4374080130921e-06)(627.030303030303,7.4374080130921e-06)(641.959595959596,7.4374080130921e-06)(656.888888888889,2.47920361444319e-06)(664.353535353535,0)(659.377104377104,-0.181818181818183)(656.888888888889,-0.227272727272728)(649.424242424242,-0.363636363636363)(641.959595959596,-0.454545454545455)(634.494949494949,-0.545454545454547)(627.030303030303,-0.636363636363637)(615.833333333333,-0.727272727272727)(612.10101010101,-0.757575757575758)(597.171717171717,-0.818181818181818)(582.242424242424,-0.818181818181818)(567.313131313131,-0.878787878787879)(559.848484848485,-0.90909090909091)(552.383838383838,-0.939393939393941)(537.454545454546,-1)(522.525252525253,-1)(507.59595959596,-1)(492.666666666667,-1)(477.737373737374,-1)(462.808080808081,-1)(447.878787878788,-1)(432.949494949495,-1)(418.020202020202,-1)(403.090909090909,-1)(388.161616161616,-1)(373.232323232323,-1)(358.30303030303,-1)(343.373737373737,-0.939393939393941)(339.641414141414,-0.90909090909091)(328.444444444444,-0.818181818181818)(320.979797979798,-0.727272727272727)(313.515151515152,-0.590909090909092)(311.026936026936,-0.545454545454547)(301.074074074074,-0.363636363636363)(298.585858585859,-0.318181818181818)(291.121212121212,-0.181818181818183)(283.656565656566,-0.0454545454545459)(281.16835016835,0)(283.656565656566,8.26408717619588e-07)};

\addplot [fill=locol,draw=none,forget plot] coordinates{ (641.959595959596,-10.6363636363636)(634.494949494949,-10.7272727272727)(629.518518518518,-10.9090909090909)(638.227272727273,-11.0909090909091)(641.959595959596,-11.1212121212121)(656.888888888889,-11.2424242424242)(664.353535353535,-11.2727272727273)(671.818181818182,-11.3030303030303)(686.747474747475,-11.3030303030303)(694.212121212121,-11.2727272727273)(701.676767676768,-11.2424242424242)(716.606060606061,-11.1212121212121)(720.338383838384,-11.0909090909091)(731.535353535354,-10.9545454545455)(734.023569023569,-10.9090909090909)(731.535353535354,-10.8181818181818)(729.047138047138,-10.7272727272727)(716.606060606061,-10.5757575757576)(709.141414141414,-10.5454545454545)(701.676767676768,-10.5151515151515)(686.747474747475,-10.4545454545455)(671.818181818182,-10.4545454545455)(656.888888888889,-10.5151515151515)(653.156565656566,-10.5454545454545)(641.959595959596,-10.6363636363636)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273) 
};
\end{axis}
\end{tikzpicture}%