### 2013-03-21 figurefactory/files/limno_bgape_imp_class_220.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00022117143,-4.18181818181818)(1478.00022117143,-4.36363636363636)(1478.00022117143,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478.00005529347,-9.09090909090909)(1478.00011058654,-9.27272727272727)(1478.00011058654,-9.45454545454546)(1478.00005529347,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478.00022117143,-10.7272727272727)(1478.00022117143,-10.9090909090909)(1478.00022117143,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(-0.000221171433450288,-8.72727272727273)(-0.000221171433450288,-8.54545454545455)(-0.000221171433450288,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1463.07070707071,-9.18181818181818)(1465.55892255892,-9.27272727272727)(1465.55892255892,-9.45454545454546)(1463.07070707071,-9.54545454545455)(1460.58249158249,-9.63636363636364)(1448.14141414141,-9.78787878787879)(1445.6531986532,-9.81818181818182)(1433.21212121212,-9.96969696969697)(1429.4797979798,-10)(1418.28282828283,-10.0909090909091)(1407.08585858586,-10.1818181818182)(1403.35353535354,-10.2121212121212)(1390.91245791246,-10.3636363636364)(1388.42424242424,-10.3939393939394)(1373.49494949495,-10.3939393939394)(1358.56565656566,-10.3939393939394)(1354.83333333333,-10.3636363636364)(1343.63636363636,-10.2954545454545)(1328.70707070707,-10.3090909090909)(1313.77777777778,-10.3181818181818)(1298.84848484848,-10.3181818181818)(1283.91919191919,-10.3181818181818)(1268.9898989899,-10.3181818181818)(1254.06060606061,-10.3484848484848)(1239.13131313131,-10.3484848484848)(1224.20202020202,-10.3484848484848)(1209.27272727273,-10.3181818181818)(1194.34343434343,-10.3181818181818)(1179.41414141414,-10.3484848484848)(1164.48484848485,-10.3484848484848)(1149.55555555556,-10.3484848484848)(1134.62626262626,-10.3181818181818)(1119.69696969697,-10.3181818181818)(1104.76767676768,-10.2878787878788)(1089.83838383838,-10.2878787878788)(1074.90909090909,-10.2878787878788)(1059.9797979798,-10.3484848484848)(1045.05050505051,-10.3484848484848)(1030.12121212121,-10.3484848484848)(1015.19191919192,-10.3181818181818)(1000.26262626263,-10.3181818181818)(985.333333333333,-10.2878787878788)(970.40404040404,-10.2363636363636)(962.939393939394,-10.1818181818182)(955.474747474747,-10.1363636363636)(940.545454545455,-10.1060606060606)(925.616161616162,-10.0909090909091)(910.686868686869,-10.0681818181818)(899.489898989899,-10)(895.757575757576,-9.98484848484848)(880.828282828283,-9.94545454545455)(865.89898989899,-9.88636363636364)(854.70202020202,-9.81818181818182)(850.969696969697,-9.8)(836.040404040404,-9.70454545454546)(828.575757575758,-9.63636363636364)(821.111111111111,-9.56818181818182)(808.670033670034,-9.45454545454546)(806.181818181818,-9.43181818181818)(793.118686868687,-9.27272727272727)(791.252525252525,-9.24242424242424)(778.811447811448,-9.09090909090909)(776.323232323232,-9)(773.835016835017,-8.90909090909091)(772.590909090909,-8.72727272727273)(772.590909090909,-8.54545454545455)(776.323232323232,-8.45454545454545)(778.811447811448,-8.36363636363636)(783.787878787879,-8.18181818181818)(791.252525252525,-8.09090909090909)(798.717171717172,-8)(806.181818181818,-7.90909090909091)(817.378787878788,-7.81818181818182)(821.111111111111,-7.78787878787879)(836.040404040404,-7.66666666666667)(843.505050505051,-7.63636363636364)(850.969696969697,-7.60606060606061)(865.89898989899,-7.47727272727273)(868.387205387205,-7.45454545454546)(880.828282828283,-7.3030303030303)(888.292929292929,-7.27272727272727)(895.757575757576,-7.24242424242424)(910.686868686869,-7.24242424242424)(918.151515151515,-7.27272727272727)(925.616161616162,-7.29090909090909)(933.080808080808,-7.27272727272727)(940.545454545455,-7.24242424242424)(955.474747474747,-7.18181818181818)(970.40404040404,-7.12121212121212)(985.333333333333,-7.12121212121212)(1000.26262626263,-7.12121212121212)(1015.19191919192,-7.22727272727273)(1030.12121212121,-7.22727272727273)(1045.05050505051,-7.22727272727273)(1059.9797979798,-7.24242424242424)(1067.44444444444,-7.27272727272727)(1074.90909090909,-7.28787878787879)(1089.83838383838,-7.31818181818182)(1101.03535353535,-7.27272727272727)(1104.76767676768,-7.24242424242424)(1119.69696969697,-7.24242424242424)(1134.62626262626,-7.24242424242424)(1138.35858585859,-7.27272727272727)(1149.55555555556,-7.36363636363636)(1164.48484848485,-7.36363636363636)(1179.41414141414,-7.3030303030303)(1194.34343434343,-7.31818181818182)(1209.27272727273,-7.31818181818182)(1220.4696969697,-7.45454545454546)(1224.20202020202,-7.46969696969697)(1239.13131313131,-7.46969696969697)(1254.06060606061,-7.46969696969697)(1268.9898989899,-7.47272727272727)(1283.91919191919,-7.47727272727273)(1298.84848484848,-7.54545454545455)(1313.77777777778,-7.54545454545455)(1328.70707070707,-7.60606060606061)(1336.17171717172,-7.63636363636364)(1343.63636363636,-7.65909090909091)(1358.56565656566,-7.78787878787879)(1373.49494949495,-7.79545454545455)(1377.22727272727,-7.81818181818182)(1388.42424242424,-7.90909090909091)(1399.62121212121,-8)(1403.35353535354,-8.03030303030303)(1412.68434343434,-8.18181818181818)(1418.28282828283,-8.31818181818182)(1420.77104377104,-8.36363636363636)(1431.34595959596,-8.54545454545455)(1433.21212121212,-8.59090909090909)(1440.67676767677,-8.72727272727273)(1448.14141414141,-8.86363636363636)(1450.62962962963,-8.90909090909091)(1460.58249158249,-9.09090909090909)(1463.07070707071,-9.18181818181818)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1403.35353535354,-8.77272727272727)(1410.81818181818,-8.90909090909091)(1410.81818181818,-9.09090909090909)(1410.81818181818,-9.27272727272727)(1410.81818181818,-9.45454545454546)(1405.84175084175,-9.63636363636364)(1403.35353535354,-9.68181818181818)(1395.88888888889,-9.81818181818182)(1388.42424242424,-9.90909090909091)(1373.49494949495,-9.96969696969697)(1358.56565656566,-9.96969696969697)(1343.63636363636,-9.96969696969697)(1336.17171717172,-10)(1328.70707070707,-10.0151515151515)(1313.77777777778,-10.0454545454545)(1298.84848484848,-10.0454545454545)(1283.91919191919,-10.0454545454545)(1268.9898989899,-10.0454545454545)(1254.06060606061,-10.0757575757576)(1239.13131313131,-10.0757575757576)(1224.20202020202,-10.0757575757576)(1209.27272727273,-10.0454545454545)(1194.34343434343,-10.0454545454545)(1179.41414141414,-10.0757575757576)(1164.48484848485,-10.0757575757576)(1149.55555555556,-10.0757575757576)(1134.62626262626,-10.0454545454545)(1119.69696969697,-10.0454545454545)(1104.76767676768,-10.0151515151515)(1089.83838383838,-10.0151515151515)(1074.90909090909,-10.0151515151515)(1059.9797979798,-10.0757575757576)(1045.05050505051,-10.0757575757576)(1030.12121212121,-10.0757575757576)(1015.19191919192,-10.0454545454545)(1000.26262626263,-10.0454545454545)(985.333333333333,-10.0151515151515)(981.60101010101,-10)(970.40404040404,-9.94545454545455)(955.474747474747,-9.86363636363636)(940.545454545455,-9.83333333333333)(933.080808080808,-9.81818181818182)(925.616161616162,-9.79545454545455)(910.686868686869,-9.75)(895.757575757576,-9.69090909090909)(880.828282828283,-9.65151515151515)(877.09595959596,-9.63636363636364)(865.89898989899,-9.56818181818182)(850.969696969697,-9.47272727272727)(848.481481481482,-9.45454545454546)(836.040404040404,-9.34090909090909)(830.441919191919,-9.27272727272727)(821.111111111111,-9.12121212121212)(818.622895622896,-9.09090909090909)(813.646464646465,-8.90909090909091)(819.244949494949,-8.72727272727273)(819.244949494949,-8.54545454545455)(821.111111111111,-8.5)(828.575757575758,-8.36363636363636)(836.040404040404,-8.22727272727273)(839.772727272727,-8.18181818181818)(850.969696969697,-8.09090909090909)(858.434343434343,-8)(865.89898989899,-7.90909090909091)(877.09595959596,-7.81818181818182)(880.828282828283,-7.79545454545455)(895.757575757576,-7.72727272727273)(910.686868686869,-7.66666666666667)(918.151515151515,-7.63636363636364)(925.616161616162,-7.61818181818182)(940.545454545455,-7.56818181818182)(955.474747474747,-7.54545454545455)(970.40404040404,-7.5)(985.333333333333,-7.5)(1000.26262626263,-7.5)(1015.19191919192,-7.60606060606061)(1030.12121212121,-7.60606060606061)(1045.05050505051,-7.60606060606061)(1059.9797979798,-7.56818181818182)(1074.90909090909,-7.58181818181818)(1089.83838383838,-7.59090909090909)(1104.76767676768,-7.61363636363636)(1112.23232323232,-7.63636363636364)(1119.69696969697,-7.66666666666667)(1134.62626262626,-7.72727272727273)(1149.55555555556,-7.72727272727273)(1164.48484848485,-7.72727272727273)(1179.41414141414,-7.68181818181818)(1194.34343434343,-7.77272727272727)(1209.27272727273,-7.77272727272727)(1224.20202020202,-7.79545454545455)(1239.13131313131,-7.76363636363636)(1254.06060606061,-7.74242424242424)(1268.9898989899,-7.77272727272727)(1283.91919191919,-7.8030303030303)(1287.65151515152,-7.81818181818182)(1298.84848484848,-7.90909090909091)(1313.77777777778,-7.90909090909091)(1328.70707070707,-7.97727272727273)(1332.43939393939,-8)(1343.63636363636,-8.09090909090909)(1354.83333333333,-8.18181818181818)(1358.56565656566,-8.21212121212121)(1371.00673400673,-8.36363636363636)(1373.49494949495,-8.39393939393939)(1388.42424242424,-8.51515151515152)(1390.91245791246,-8.54545454545455)(1400.86531986532,-8.72727272727273)(1403.35353535354,-8.77272727272727)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (74.6464646464647,-7.95454545454546)(85.8434343434344,-7.81818181818182)(89.5757575757576,-7.78787878787879)(104.505050505051,-7.66666666666667)(106.993265993266,-7.63636363636364)(116.946127946128,-7.45454545454546)(119.434343434343,-7.42424242424243)(134.363636363636,-7.36363636363636)(149.292929292929,-7.3030303030303)(164.222222222222,-7.34090909090909)(179.151515151515,-7.34090909090909)(194.080808080808,-7.34090909090909)(209.010101010101,-7.36363636363636)(223.939393939394,-7.42424242424243)(231.40404040404,-7.45454545454546)(238.868686868687,-7.48484848484849)(253.79797979798,-7.54545454545455)(268.727272727273,-7.52272727272727)(283.656565656566,-7.56818181818182)(298.585858585859,-7.61363636363636)(302.318181818182,-7.63636363636364)(313.515151515152,-7.72727272727273)(328.444444444444,-7.77272727272727)(343.373737373737,-7.77272727272727)(358.30303030303,-7.68181818181818)(362.035353535354,-7.63636363636364)(373.232323232323,-7.56818181818182)(388.161616161616,-7.47727272727273)(395.626262626263,-7.45454545454546)(403.090909090909,-7.42424242424243)(418.020202020202,-7.36363636363636)(432.949494949495,-7.31818181818182)(440.414141414141,-7.27272727272727)(447.878787878788,-7.24242424242424)(462.808080808081,-7.18181818181818)(477.737373737374,-7.18181818181818)(492.666666666667,-7.12121212121212)(496.39898989899,-7.09090909090909)(507.59595959596,-7.02272727272727)(522.525252525253,-6.97727272727273)(537.454545454546,-7.02272727272727)(548.651515151515,-7.09090909090909)(552.383838383838,-7.12121212121212)(567.313131313131,-7.18181818181818)(582.242424242424,-7.18181818181818)(597.171717171717,-7.24242424242424)(604.636363636364,-7.27272727272727)(612.10101010101,-7.3030303030303)(627.030303030303,-7.3030303030303)(641.959595959596,-7.40909090909091)(645.691919191919,-7.45454545454546)(641.959595959596,-7.54545454545455)(639.47138047138,-7.63636363636364)(641.959595959596,-7.68181818181818)(649.424242424242,-7.81818181818182)(654.400673400673,-8)(656.888888888889,-8.04545454545455)(664.353535353535,-8.18181818181818)(664.353535353535,-8.36363636363636)(664.353535353535,-8.54545454545455)(656.888888888889,-8.68181818181818)(655.022727272727,-8.72727272727273)(641.959595959596,-8.88636363636364)(634.494949494949,-8.90909090909091)(627.030303030303,-8.93939393939394)(612.10101010101,-9.06060606060606)(604.636363636364,-9.09090909090909)(597.171717171717,-9.12121212121212)(582.242424242424,-9.18181818181818)(567.313131313131,-9.18181818181818)(556.116161616162,-9.27272727272727)(552.383838383838,-9.3030303030303)(537.454545454546,-9.36363636363636)(522.525252525253,-9.36363636363636)(507.59595959596,-9.36363636363636)(492.666666666667,-9.3030303030303)(477.737373737374,-9.3030303030303)(474.005050505051,-9.27272727272727)(462.808080808081,-9.18181818181818)(447.878787878788,-9.12121212121212)(440.414141414141,-9.09090909090909)(432.949494949495,-9.06060606060606)(420.508417508418,-8.90909090909091)(418.020202020202,-8.87878787878788)(403.090909090909,-8.81818181818182)(388.161616161616,-8.75757575757576)(373.232323232323,-8.75757575757576)(365.767676767677,-8.72727272727273)(358.30303030303,-8.6969696969697)(343.373737373737,-8.63636363636364)(328.444444444444,-8.6969696969697)(324.712121212121,-8.72727272727273)(313.515151515152,-8.79545454545455)(298.585858585859,-8.84090909090909)(283.656565656566,-8.88636363636364)(276.191919191919,-8.90909090909091)(268.727272727273,-8.93939393939394)(253.79797979798,-9)(242.60101010101,-9.09090909090909)(238.868686868687,-9.12121212121212)(223.939393939394,-9.18181818181818)(209.010101010101,-9.24242424242424)(194.080808080808,-9.24242424242424)(186.616161616162,-9.27272727272727)(179.151515151515,-9.29090909090909)(171.686868686869,-9.27272727272727)(164.222222222222,-9.24242424242424)(151.781144781145,-9.09090909090909)(149.292929292929,-9.06060606060606)(136.851851851852,-8.90909090909091)(134.363636363636,-8.88636363636364)(119.434343434343,-8.88636363636364)(104.505050505051,-8.81818181818182)(97.040404040404,-8.72727272727273)(89.5757575757576,-8.63636363636364)(83.9772727272727,-8.54545454545455)(77.1346801346801,-8.36363636363636)(74.6464646464647,-8.31818181818182)(67.1818181818182,-8.18181818181818)(72.1582491582492,-8)(74.6464646464647,-7.95454545454546)};

\addplot [fill=green!60!black,draw=none,forget plot] coordinates{ (1313.77777777778,-9)(1316.26599326599,-9.09090909090909)(1313.77777777778,-9.13636363636364)(1302.58080808081,-9.27272727272727)(1298.84848484848,-9.3030303030303)(1283.91919191919,-9.36363636363636)(1268.9898989899,-9.36363636363636)(1254.06060606061,-9.42424242424243)(1246.59595959596,-9.45454545454546)(1239.13131313131,-9.48484848484848)(1224.20202020202,-9.54545454545455)(1209.27272727273,-9.5)(1201.80808080808,-9.45454545454546)(1194.34343434343,-9.42424242424243)(1179.41414141414,-9.42424242424243)(1171.94949494949,-9.45454545454546)(1164.48484848485,-9.5)(1149.55555555556,-9.54545454545455)(1134.62626262626,-9.54545454545455)(1119.69696969697,-9.54545454545455)(1104.76767676768,-9.54545454545455)(1089.83838383838,-9.48484848484848)(1074.90909090909,-9.48484848484848)(1059.9797979798,-9.48484848484848)(1045.05050505051,-9.54545454545455)(1030.12121212121,-9.54545454545455)(1015.19191919192,-9.5)(1000.26262626263,-9.5)(985.333333333333,-9.5)(970.40404040404,-9.48484848484848)(962.939393939394,-9.45454545454546)(955.474747474747,-9.42424242424243)(940.545454545455,-9.3030303030303)(933.080808080808,-9.27272727272727)(925.616161616162,-9.24242424242424)(910.686868686869,-9.12121212121212)(906.954545454546,-9.09090909090909)(895.757575757576,-8.95454545454546)(893.26936026936,-8.90909090909091)(892.025252525252,-8.72727272727273)(895.757575757576,-8.63636363636364)(898.245791245791,-8.54545454545455)(910.686868686869,-8.39393939393939)(914.419191919192,-8.36363636363636)(925.616161616162,-8.22727272727273)(929.348484848485,-8.18181818181818)(940.545454545455,-8.09090909090909)(951.742424242424,-8)(955.474747474747,-7.96969696969697)(970.40404040404,-7.96969696969697)(985.333333333333,-7.96969696969697)(992.79797979798,-8)(1000.26262626263,-8.03030303030303)(1007.72727272727,-8)(1015.19191919192,-7.96969696969697)(1030.12121212121,-7.96969696969697)(1045.05050505051,-7.90909090909091)(1056.24747474747,-8)(1059.9797979798,-8.03030303030303)(1074.90909090909,-8.09090909090909)(1089.83838383838,-8.15151515151515)(1104.76767676768,-8.09090909090909)(1119.69696969697,-8.09090909090909)(1134.62626262626,-8.15151515151515)(1142.09090909091,-8.18181818181818)(1149.55555555556,-8.22727272727273)(1164.48484848485,-8.31818181818182)(1179.41414141414,-8.31818181818182)(1194.34343434343,-8.27272727272727)(1209.27272727273,-8.33333333333333)(1224.20202020202,-8.33333333333333)(1231.66666666667,-8.36363636363636)(1239.13131313131,-8.39393939393939)(1254.06060606061,-8.39393939393939)(1268.9898989899,-8.45454545454545)(1280.18686868687,-8.54545454545455)(1283.91919191919,-8.57575757575757)(1298.84848484848,-8.6969696969697)(1301.3367003367,-8.72727272727273)(1311.28956228956,-8.90909090909091)(1313.77777777778,-9)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (447.878787878788,-8.04545454545455)(450.367003367003,-8)(462.808080808081,-7.88636363636364)(474.005050505051,-7.81818181818182)(477.737373737374,-7.78787878787879)(492.666666666667,-7.66666666666667)(507.59595959596,-7.72727272727273)(522.525252525253,-7.66666666666667)(537.454545454546,-7.66666666666667)(544.919191919192,-7.63636363636364)(552.383838383838,-7.60606060606061)(559.848484848485,-7.63636363636364)(567.313131313131,-7.66666666666667)(579.754208754209,-7.81818181818182)(582.242424242424,-7.84848484848485)(594.683501683502,-8)(597.171717171717,-8.09090909090909)(599.659932659933,-8.18181818181818)(600.90404040404,-8.36363636363636)(597.171717171717,-8.40909090909091)(585.974747474747,-8.54545454545455)(582.242424242424,-8.57575757575757)(567.313131313131,-8.6969696969697)(559.848484848485,-8.72727272727273)(552.383838383838,-8.75757575757576)(537.454545454546,-8.81818181818182)(522.525252525253,-8.81818181818182)(507.59595959596,-8.75757575757576)(503.863636363636,-8.72727272727273)(492.666666666667,-8.63636363636364)(481.469696969697,-8.54545454545455)(477.737373737374,-8.52272727272727)(462.808080808081,-8.47727272727273)(450.367003367003,-8.36363636363636)(447.878787878788,-8.31818181818182)(440.414141414141,-8.18181818181818)(447.878787878788,-8.04545454545455)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273)(447.878787878788,0)(1089.83838383838,-12.1818181818182)(313.515151515152,-6.54545454545455)(268.727272727273,-11.0909090909091)(1000.26262626263,-9.27272727272727)(477.737373737374,-8.72727272727273)(1478,-8.54545454545455)(1089.83838383838,-5.63636363636364)(0,-10.9090909090909)(0,-6.36363636363636)(358.30303030303,-4.36363636363636)(1478,-6.36363636363636)(1239.13131313131,-10.1818181818182)(761.393939393939,-8.36363636363636)(1373.49494949495,-12.1818181818182)(597.171717171717,-5.45454545454546)(223.939393939394,-9.09090909090909)(985.333333333333,-7.27272727272727)(895.757575757576,-11.2727272727273)(1298.84848484848,-5.45454545454546)(865.89898989899,-5.63636363636364)(462.808080808081,-10.3636363636364)(522.525252525253,-7.27272727272727)(179.151515151515,-7.63636363636364)(164.222222222222,-5.81818181818182)(776.323232323232,-9.63636363636364)(1239.13131313131,-8.90909090909091)(1478,-9.63636363636364)(1074.90909090909,-10.7272727272727)(119.434343434343,-9.81818181818182)(1478,-7.45454545454546)(0,-7.45454545454546)(1179.41414141414,-6.72727272727273)(447.878787878788,-6)(1283.91919191919,-11.0909090909091)(612.10101010101,-9.27272727272727)(1074.90909090909,-8.18181818181818)(343.373737373737,-8)(328.444444444444,-9.81818181818182)(836.040404040404,-7.45454545454546)(910.686868686869,-10.1818181818182)(641.959595959596,-7.81818181818182)(1358.56565656566,-9.45454545454546)(910.686868686869,-8.54545454545455)(149.292929292929,-6.72727272727273)(1343.63636363636,-7.27272727272727)(1119.69696969697,-9.63636363636364)(1015.19191919192,-6.72727272727273)(597.171717171717,-6.54545454545455)(119.434343434343,-8.54545454545455)(388.161616161616,-7.27272727272727)(1358.56565656566,-8.36363636363636)(1478,-10.1818181818182)(447.878787878788,-9.45454545454546)(1149.55555555556,-10.9090909090909)(1149.55555555556,-7.81818181818182)(1358.56565656566,-10.7272727272727)(328.444444444444,-8.72727272727273)(567.313131313131,-8.36363636363636)(0,-7.09090909090909)(850.969696969697,-9.27272727272727)(1478,-8.90909090909091)(1000.26262626263,-10.7272727272727)(701.676767676768,-7.45454545454546)(1134.62626262626,-8.90909090909091)(671.818181818182,-9.09090909090909)(253.79797979798,-7.27272727272727)(940.545454545455,-8)(462.808080808081,-6.72727272727273)(880.828282828283,-10.3636363636364)(1268.9898989899,-7.09090909090909)(104.505050505051,-7.81818181818182)(462.808080808081,-8)(1313.77777777778,-9.81818181818182)(1104.76767676768,-6.90909090909091)(895.757575757576,-7.27272727272727)(134.363636363636,-8.90909090909091)(1030.12121212121,-9.81818181818182)(1433.21212121212,-8)(1254.06060606061,-8.36363636363636)(253.79797979798,-8.18181818181818)(537.454545454546,-9.27272727272727)(298.585858585859,-9.27272727272727)(791.252525252525,-8.54545454545455)(612.10101010101,-6.90909090909091)(29.8585858585859,-8)(1015.19191919192,-8.72727272727273)(1478,-10)(119.434343434343,-7.09090909090909)(1179.41414141414,-9.81818181818182)(836.040404040404,-10)(1358.56565656566,-8.90909090909091)(1373.49494949495,-10.5454545454545)(1089.83838383838,-7.63636363636364)(656.888888888889,-8.36363636363636)(343.373737373737,-7.09090909090909)(1478,-9.09090909090909)(403.090909090909,-8.72727272727273)(1194.34343434343,-10.5454545454545)(940.545454545455,-7.09090909090909)(791.252525252525,-7.81818181818182)(522.525252525253,-7.63636363636364)(925.616161616162,-9.27272727272727)(1373.49494949495,-7.63636363636364)(761.393939393939,-9.27272727272727)(164.222222222222,-9.09090909090909)(1164.48484848485,-8.54545454545455)(1000.26262626263,-10.5454545454545)(492.666666666667,-6.72727272727273)(582.242424242424,-9.27272727272727)(1254.06060606061,-9.45454545454546)(1194.34343434343,-7.27272727272727)(209.010101010101,-8.18181818181818)(388.161616161616,-9.27272727272727)(985.333333333333,-8)(641.959595959596,-7.27272727272727)(388.161616161616,-7.81818181818182)(59.7171717171717,-8.36363636363636)(1373.49494949495,-10)(1089.83838383838,-10.1818181818182)(209.010101010101,-7.27272727272727)(880.828282828283,-8.36363636363636)(1283.91919191919,-8.18181818181818)(1074.90909090909,-9.09090909090909)(1448.14141414141,-8.36363636363636)(582.242424242424,-8.36363636363636)(89.5757575757576,-7.27272727272727)(1268.9898989899,-10.5454545454545)(1478,-9.45454545454546)(1045.05050505051,-7.09090909090909)(955.474747474747,-9.63636363636364)(821.111111111111,-9.81818181818182)(716.606060606061,-8.72727272727273)(253.79797979798,-9.09090909090909)(492.666666666667,-9.27272727272727)(865.89898989899,-7.45454545454546)(1358.56565656566,-9.09090909090909)(447.878787878788,-8.36363636363636)(418.020202020202,-7.09090909090909)(701.676767676768,-7.81818181818182)(104.505050505051,-8.72727272727273)(283.656565656566,-7.81818181818182)(1164.48484848485,-8.18181818181818)(1313.77777777778,-7.45454545454546)(1194.34343434343,-9.27272727272727)(880.828282828283,-8.90909090909091)(567.313131313131,-6.90909090909091)(1030.12121212121,-8.36363636363636)(1134.62626262626,-10.3636363636364)(1119.69696969697,-7.27272727272727)(1373.49494949495,-10.3636363636364)(298.585858585859,-8.54545454545455)(970.40404040404,-10.1818181818182)(552.383838383838,-7.63636363636364)(59.7171717171717,-7.81818181818182)(567.313131313131,-9.27272727272727)(1388.42424242424,-8)(746.464646464646,-9.09090909090909)(1478,-9.27272727272727)(806.181818181818,-7.81818181818182)(179.151515151515,-8.18181818181818)(1298.84848484848,-8.72727272727273)(1074.90909090909,-9.45454545454546)(1209.27272727273,-7.63636363636364)(910.686868686869,-7.27272727272727)(179.151515151515,-9.09090909090909)(268.727272727273,-7.45454545454546)(1254.06060606061,-9.63636363636364)(447.878787878788,-7.63636363636364)(627.030303030303,-8.54545454545455)(1015.19191919192,-8)(388.161616161616,-8.90909090909091)(1388.42424242424,-9.81818181818182)(955.474747474747,-8.90909090909091)(641.959595959596,-7.45454545454546)(1268.9898989899,-10.3636363636364)(850.969696969697,-9.81818181818182)(1119.69696969697,-8.72727272727273)(821.111111111111,-8.36363636363636)(134.363636363636,-7.45454545454546)(507.59595959596,-8.72727272727273)(1045.05050505051,-10.1818181818182)(1030.12121212121,-7.45454545454546)(1403.35353535354,-8.72727272727273)(358.30303030303,-8.18181818181818)(1343.63636363636,-7.63636363636364)(89.5757575757576,-8.54545454545455)(1224.20202020202,-8.36363636363636)(492.666666666667,-7.09090909090909)(686.747474747475,-8.18181818181818)(1164.48484848485,-10.1818181818182)(821.111111111111,-9.09090909090909)(238.868686868687,-8.54545454545455)(925.616161616162,-8.18181818181818)(507.59595959596,-9.27272727272727)(1194.34343434343,-7.45454545454546)(940.545454545455,-10)(1433.21212121212,-10)(1179.41414141414,-9.27272727272727)(582.242424242424,-7.81818181818182) 
};
\end{axis}
\end{tikzpicture}%