### 2013-03-21 figurefactory/files/limno_bgape_imp_class_320.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00022117143,-4.18181818181818)(1478.00022117143,-4.36363636363636)(1478.00022117143,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478,-8.72727272727273)(1478,-8.90909090909091)(1478,-9.09090909090909)(1478,-9.27272727272727)(1478,-9.45454545454546)(1478,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478.00022117143,-10.7272727272727)(1478.00022117143,-10.9090909090909)(1478.00022117143,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(-0.000221171433450288,-8.72727272727273)(-0.000221171433450288,-8.54545454545455)(-0.000221171433450288,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(0,-7.63636363636364)(0,-7.45454545454546)(0,-7.27272727272727)(0,-7.09090909090909)(0,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1448.14141414141,-9.18181818181818)(1450.62962962963,-9.27272727272727)(1460.58249158249,-9.45454545454546)(1455.60606060606,-9.63636363636364)(1448.14141414141,-9.77272727272727)(1446.27525252525,-9.81818181818182)(1433.21212121212,-9.97727272727273)(1429.4797979798,-10)(1418.28282828283,-10.0909090909091)(1403.35353535354,-10.1590909090909)(1400.86531986532,-10.1818181818182)(1388.42424242424,-10.2954545454545)(1373.49494949495,-10.3409090909091)(1366.0303030303,-10.3636363636364)(1358.56565656566,-10.3939393939394)(1351.10101010101,-10.3636363636364)(1343.63636363636,-10.3454545454545)(1328.70707070707,-10.3484848484848)(1313.77777777778,-10.3181818181818)(1298.84848484848,-10.2575757575758)(1283.91919191919,-10.2575757575758)(1268.9898989899,-10.2878787878788)(1254.06060606061,-10.3484848484848)(1246.59595959596,-10.3636363636364)(1239.13131313131,-10.3818181818182)(1224.20202020202,-10.3818181818182)(1216.73737373737,-10.3636363636364)(1209.27272727273,-10.3484848484848)(1194.34343434343,-10.2878787878788)(1179.41414141414,-10.3181818181818)(1164.48484848485,-10.3181818181818)(1149.55555555556,-10.3484848484848)(1134.62626262626,-10.2878787878788)(1119.69696969697,-10.2878787878788)(1104.76767676768,-10.2575757575758)(1089.83838383838,-10.2878787878788)(1074.90909090909,-10.3181818181818)(1063.71212121212,-10.3636363636364)(1059.9797979798,-10.3818181818182)(1045.05050505051,-10.3818181818182)(1037.58585858586,-10.3636363636364)(1030.12121212121,-10.3484848484848)(1015.19191919192,-10.2878787878788)(1000.26262626263,-10.2575757575758)(985.333333333333,-10.2272727272727)(970.40404040404,-10.2)(966.671717171717,-10.1818181818182)(955.474747474747,-10.1363636363636)(940.545454545455,-10.1060606060606)(925.616161616162,-10.1363636363636)(910.686868686869,-10.1272727272727)(895.757575757576,-10.0681818181818)(890.159090909091,-10)(880.828282828283,-9.90909090909091)(865.89898989899,-9.84090909090909)(863.410774410775,-9.81818181818182)(850.969696969697,-9.72727272727273)(836.040404040404,-9.65909090909091)(833.552188552189,-9.63636363636364)(821.111111111111,-9.52272727272727)(813.646464646465,-9.45454545454546)(806.181818181818,-9.4)(793.118686868687,-9.27272727272727)(791.252525252525,-9.25)(778.189393939394,-9.09090909090909)(776.323232323232,-9)(774.457070707071,-8.90909090909091)(772.590909090909,-8.72727272727273)(776.323232323232,-8.63636363636364)(778.189393939394,-8.54545454545455)(783.787878787879,-8.36363636363636)(791.252525252525,-8.27272727272727)(796.85101010101,-8.18181818181818)(806.181818181818,-8.06818181818182)(813.646464646465,-8)(821.111111111111,-7.93181818181818)(833.552188552189,-7.81818181818182)(836.040404040404,-7.79545454545455)(850.969696969697,-7.69090909090909)(858.434343434343,-7.63636363636364)(865.89898989899,-7.56818181818182)(872.119528619529,-7.45454545454546)(880.828282828283,-7.29545454545455)(888.292929292929,-7.27272727272727)(895.757575757576,-7.24242424242424)(910.686868686869,-7.20454545454546)(918.151515151515,-7.27272727272727)(925.616161616162,-7.31818181818182)(940.545454545455,-7.28787878787879)(948.010101010101,-7.27272727272727)(955.474747474747,-7.25454545454546)(970.40404040404,-7.16666666666667)(985.333333333333,-7.10606060606061)(1000.26262626263,-7.10606060606061)(1015.19191919192,-7.20454545454546)(1022.65656565657,-7.27272727272727)(1030.12121212121,-7.31818181818182)(1045.05050505051,-7.34848484848485)(1059.9797979798,-7.31818181818182)(1074.90909090909,-7.31818181818182)(1089.83838383838,-7.34848484848485)(1104.76767676768,-7.31818181818182)(1119.69696969697,-7.34848484848485)(1134.62626262626,-7.34848484848485)(1149.55555555556,-7.43636363636364)(1164.48484848485,-7.38636363636364)(1179.41414141414,-7.3030303030303)(1194.34343434343,-7.31818181818182)(1205.5404040404,-7.45454545454546)(1209.27272727273,-7.46969696969697)(1224.20202020202,-7.56060606060606)(1239.13131313131,-7.53030303030303)(1254.06060606061,-7.5)(1268.9898989899,-7.46969696969697)(1283.91919191919,-7.46969696969697)(1298.84848484848,-7.5)(1313.77777777778,-7.54545454545455)(1328.70707070707,-7.61363636363636)(1332.43939393939,-7.63636363636364)(1343.63636363636,-7.69090909090909)(1358.56565656566,-7.78787878787879)(1366.0303030303,-7.81818181818182)(1373.49494949495,-7.84090909090909)(1388.42424242424,-7.93181818181818)(1403.35353535354,-7.96969696969697)(1405.84175084175,-8)(1412.68434343434,-8.18181818181818)(1418.28282828283,-8.31818181818182)(1420.14898989899,-8.36363636363636)(1423.88131313131,-8.54545454545455)(1431.71919191919,-8.72727272727273)(1433.21212121212,-8.77272727272727)(1438.81060606061,-8.90909090909091)(1446.64848484848,-9.09090909090909)(1448.14141414141,-9.18181818181818)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1418.28282828283,-9.04545454545455)(1419.77575757576,-9.09090909090909)(1423.88131313131,-9.27272727272727)(1422.01515151515,-9.45454545454546)(1418.28282828283,-9.5)(1410.81818181818,-9.63636363636364)(1403.35353535354,-9.77272727272727)(1399.62121212121,-9.81818181818182)(1388.42424242424,-9.87272727272727)(1373.49494949495,-9.93181818181818)(1362.29797979798,-10)(1358.56565656566,-10.0227272727273)(1343.63636363636,-10.0454545454545)(1328.70707070707,-10.0757575757576)(1313.77777777778,-10.0454545454545)(1302.58080808081,-10)(1298.84848484848,-9.98181818181818)(1283.91919191919,-9.98181818181818)(1276.45454545455,-10)(1268.9898989899,-10.0151515151515)(1254.06060606061,-10.0757575757576)(1239.13131313131,-10.1060606060606)(1224.20202020202,-10.1060606060606)(1209.27272727273,-10.0757575757576)(1194.34343434343,-10.0151515151515)(1179.41414141414,-10.0454545454545)(1164.48484848485,-10.0454545454545)(1149.55555555556,-10.0757575757576)(1134.62626262626,-10.0151515151515)(1119.69696969697,-10.0151515151515)(1112.23232323232,-10)(1104.76767676768,-9.98181818181818)(1097.30303030303,-10)(1089.83838383838,-10.0151515151515)(1074.90909090909,-10.0454545454545)(1059.9797979798,-10.1060606060606)(1045.05050505051,-10.1060606060606)(1030.12121212121,-10.0757575757576)(1015.19191919192,-10.0151515151515)(1007.72727272727,-10)(1000.26262626263,-9.98181818181818)(985.333333333333,-9.95454545454546)(970.40404040404,-9.92424242424243)(955.474747474747,-9.86363636363636)(940.545454545455,-9.83333333333333)(925.616161616162,-9.86363636363636)(910.686868686869,-9.83333333333333)(906.954545454546,-9.81818181818182)(895.757575757576,-9.75)(883.316498316498,-9.63636363636364)(880.828282828283,-9.61363636363636)(865.89898989899,-9.54545454545455)(853.457912457912,-9.45454545454546)(850.969696969697,-9.43181818181818)(836.040404040404,-9.32727272727273)(830.441919191919,-9.27272727272727)(821.111111111111,-9.15909090909091)(813.646464646465,-9.09090909090909)(806.181818181818,-9)(801.70303030303,-8.90909090909091)(806.181818181818,-8.77272727272727)(808.04797979798,-8.72727272727273)(811.780303030303,-8.54545454545455)(819.618181818182,-8.36363636363636)(821.111111111111,-8.33333333333333)(830.441919191919,-8.18181818181818)(836.040404040404,-8.12727272727273)(850.969696969697,-8.02272727272727)(853.457912457912,-8)(865.89898989899,-7.88636363636364)(873.363636363636,-7.81818181818182)(880.828282828283,-7.75)(893.26936026936,-7.63636363636364)(895.757575757576,-7.61363636363636)(910.686868686869,-7.52272727272727)(925.616161616162,-7.59090909090909)(940.545454545455,-7.56060606060606)(955.474747474747,-7.53030303030303)(967.915824915825,-7.45454545454546)(970.40404040404,-7.43939393939394)(985.333333333333,-7.37878787878788)(1000.26262626263,-7.37878787878788)(1009.59343434343,-7.45454545454546)(1015.19191919192,-7.52272727272727)(1030.12121212121,-7.59090909090909)(1045.05050505051,-7.62121212121212)(1059.9797979798,-7.59090909090909)(1074.90909090909,-7.59090909090909)(1089.83838383838,-7.62121212121212)(1097.30303030303,-7.63636363636364)(1104.76767676768,-7.65454545454546)(1119.69696969697,-7.68181818181818)(1134.62626262626,-7.68181818181818)(1149.55555555556,-7.71212121212121)(1164.48484848485,-7.70454545454546)(1179.41414141414,-7.68181818181818)(1194.34343434343,-7.77272727272727)(1209.27272727273,-7.79545454545455)(1216.73737373737,-7.81818181818182)(1224.20202020202,-7.83636363636364)(1231.66666666667,-7.81818181818182)(1239.13131313131,-7.8030303030303)(1254.06060606061,-7.77272727272727)(1268.9898989899,-7.74242424242424)(1283.91919191919,-7.74242424242424)(1298.84848484848,-7.77272727272727)(1310.04545454545,-7.81818181818182)(1313.77777777778,-7.84090909090909)(1328.70707070707,-7.90909090909091)(1343.63636363636,-7.98484848484848)(1345.50252525253,-8)(1358.56565656566,-8.15909090909091)(1362.29797979798,-8.18181818181818)(1373.49494949495,-8.27272727272727)(1384.69191919192,-8.36363636363636)(1388.42424242424,-8.40909090909091)(1394.02272727273,-8.54545454545455)(1403.35353535354,-8.6969696969697)(1404.84646464646,-8.72727272727273)(1412.68434343434,-8.90909090909091)(1418.28282828283,-9.04545454545455)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (373.232323232323,-7.68181818181818)(376.964646464646,-7.63636363636364)(388.161616161616,-7.54545454545455)(403.090909090909,-7.47727272727273)(406.823232323232,-7.45454545454546)(418.020202020202,-7.36363636363636)(432.949494949495,-7.31818181818182)(440.414141414141,-7.27272727272727)(447.878787878788,-7.24242424242424)(462.808080808081,-7.18181818181818)(477.737373737374,-7.11363636363636)(492.666666666667,-7.14545454545455)(507.59595959596,-7.10606060606061)(522.525252525253,-7.10606060606061)(537.454545454546,-7.10909090909091)(552.383838383838,-7.15909090909091)(567.313131313131,-7.12121212121212)(582.242424242424,-7.18181818181818)(597.171717171717,-7.24242424242424)(599.659932659933,-7.27272727272727)(612.10101010101,-7.42424242424243)(627.030303030303,-7.40909090909091)(630.762626262626,-7.45454545454546)(629.518518518518,-7.63636363636364)(640.093434343434,-7.81818181818182)(641.959595959596,-7.90909090909091)(644.447811447811,-8)(649.424242424242,-8.18181818181818)(651.290404040404,-8.36363636363636)(651.290404040404,-8.54545454545455)(649.424242424242,-8.72727272727273)(641.959595959596,-8.87878787878788)(638.227272727273,-8.90909090909091)(627.030303030303,-9)(612.10101010101,-9.06818181818182)(597.171717171717,-9.07272727272727)(589.707070707071,-9.09090909090909)(582.242424242424,-9.12121212121212)(567.313131313131,-9.20454545454546)(556.116161616162,-9.27272727272727)(552.383838383838,-9.3030303030303)(537.454545454546,-9.3030303030303)(522.525252525253,-9.3030303030303)(507.59595959596,-9.3030303030303)(492.666666666667,-9.34090909090909)(477.737373737374,-9.3030303030303)(475.249158249158,-9.27272727272727)(462.808080808081,-9.12121212121212)(460.319865319865,-9.09090909090909)(447.878787878788,-8.97727272727273)(436.681818181818,-8.90909090909091)(432.949494949495,-8.87878787878788)(418.020202020202,-8.75757575757576)(414.287878787879,-8.72727272727273)(403.090909090909,-8.63636363636364)(395.626262626263,-8.54545454545455)(388.161616161616,-8.45454545454545)(380.69696969697,-8.36363636363636)(373.232323232323,-8.22727272727273)(370.744107744108,-8.18181818181818)(367.633838383838,-8)(365.767676767677,-7.81818181818182)(373.232323232323,-7.68181818181818)};

\addplot [fill=green!60!black,draw=none,forget plot] coordinates{ (1283.91919191919,-8.68181818181818)(1286.40740740741,-8.72727272727273)(1286.40740740741,-8.90909090909091)(1283.91919191919,-8.95454545454546)(1276.45454545455,-9.09090909090909)(1268.9898989899,-9.18181818181818)(1257.79292929293,-9.27272727272727)(1254.06060606061,-9.3030303030303)(1239.13131313131,-9.36363636363636)(1224.20202020202,-9.42424242424243)(1209.27272727273,-9.42424242424243)(1194.34343434343,-9.42424242424243)(1179.41414141414,-9.42424242424243)(1164.48484848485,-9.42424242424243)(1157.0202020202,-9.45454545454546)(1149.55555555556,-9.48484848484848)(1134.62626262626,-9.48484848484848)(1119.69696969697,-9.48484848484848)(1104.76767676768,-9.48484848484848)(1097.30303030303,-9.45454545454546)(1089.83838383838,-9.40909090909091)(1082.37373737374,-9.45454545454546)(1074.90909090909,-9.48484848484848)(1059.9797979798,-9.48484848484848)(1045.05050505051,-9.48484848484848)(1030.12121212121,-9.48484848484848)(1022.65656565657,-9.45454545454546)(1015.19191919192,-9.40909090909091)(1000.26262626263,-9.40909090909091)(985.333333333333,-9.36363636363636)(970.40404040404,-9.36363636363636)(955.474747474747,-9.3030303030303)(948.010101010101,-9.27272727272727)(940.545454545455,-9.22727272727273)(925.616161616162,-9.12121212121212)(918.151515151515,-9.09090909090909)(910.686868686869,-9.04545454545455)(903.222222222222,-8.90909090909091)(898.245791245791,-8.72727272727273)(908.198653198653,-8.54545454545455)(910.686868686869,-8.5)(921.883838383838,-8.36363636363636)(925.616161616162,-8.31818181818182)(936.813131313131,-8.18181818181818)(940.545454545455,-8.15151515151515)(955.474747474747,-8.09090909090909)(970.40404040404,-8.09090909090909)(985.333333333333,-8.09090909090909)(1000.26262626263,-8.09090909090909)(1015.19191919192,-8.13636363636364)(1030.12121212121,-8.13636363636364)(1045.05050505051,-8.13636363636364)(1059.9797979798,-8.15151515151515)(1067.44444444444,-8.18181818181818)(1074.90909090909,-8.21212121212121)(1089.83838383838,-8.31818181818182)(1104.76767676768,-8.31818181818182)(1119.69696969697,-8.31818181818182)(1134.62626262626,-8.27272727272727)(1149.55555555556,-8.31818181818182)(1157.0202020202,-8.36363636363636)(1164.48484848485,-8.39393939393939)(1179.41414141414,-8.39393939393939)(1186.87878787879,-8.36363636363636)(1194.34343434343,-8.33333333333333)(1209.27272727273,-8.33333333333333)(1216.73737373737,-8.36363636363636)(1224.20202020202,-8.40909090909091)(1239.13131313131,-8.5)(1254.06060606061,-8.5)(1268.9898989899,-8.51515151515152)(1272.72222222222,-8.54545454545455)(1283.91919191919,-8.68181818181818)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (74.6464646464647,-7.96969696969697)(87.0875420875421,-7.81818181818182)(89.5757575757576,-7.77272727272727)(100.772727272727,-7.63636363636364)(104.505050505051,-7.59090909090909)(111.969696969697,-7.45454545454546)(119.434343434343,-7.38636363636364)(134.363636363636,-7.42424242424243)(149.292929292929,-7.42424242424243)(156.757575757576,-7.45454545454546)(164.222222222222,-7.48484848484849)(179.151515151515,-7.48484848484849)(194.080808080808,-7.48484848484849)(209.010101010101,-7.54545454545455)(223.939393939394,-7.60606060606061)(231.40404040404,-7.63636363636364)(238.868686868687,-7.66666666666667)(253.79797979798,-7.66666666666667)(261.262626262626,-7.63636363636364)(268.727272727273,-7.61818181818182)(276.191919191919,-7.63636363636364)(283.656565656566,-7.66666666666667)(296.097643097643,-7.81818181818182)(298.585858585859,-7.84848484848485)(307.916666666667,-8)(313.515151515152,-8.09090909090909)(320.979797979798,-8.18181818181818)(322.84595959596,-8.36363636363636)(316.003367003367,-8.54545454545455)(313.515151515152,-8.57575757575757)(298.585858585859,-8.70454545454545)(294.853535353535,-8.72727272727273)(283.656565656566,-8.79545454545455)(271.215488215488,-8.90909090909091)(268.727272727273,-8.93939393939394)(253.79797979798,-9)(246.333333333333,-9.09090909090909)(238.868686868687,-9.15909090909091)(223.939393939394,-9.12121212121212)(209.010101010101,-9.12121212121212)(194.080808080808,-9.12121212121212)(179.151515151515,-9.24242424242424)(164.222222222222,-9.25)(149.292929292929,-9.12121212121212)(146.804713804714,-9.09090909090909)(135.856565656566,-8.90909090909091)(134.363636363636,-8.89393939393939)(119.434343434343,-8.85454545454546)(104.505050505051,-8.75757575757576)(100.772727272727,-8.72727272727273)(89.5757575757576,-8.59090909090909)(87.709595959596,-8.54545454545455)(77.1346801346801,-8.36363636363636)(74.6464646464647,-8.27272727272727)(72.1582491582492,-8.18181818181818)(72.1582491582492,-8)(74.6464646464647,-7.96969696969697)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (447.878787878788,-7.90909090909091)(450.367003367003,-7.81818181818182)(462.808080808081,-7.66666666666667)(465.296296296296,-7.63636363636364)(475.871212121212,-7.45454545454546)(477.737373737374,-7.43939393939394)(492.666666666667,-7.43939393939394)(507.59595959596,-7.43181818181818)(522.525252525253,-7.43181818181818)(529.989898989899,-7.45454545454546)(537.454545454546,-7.5)(552.383838383838,-7.48484848484849)(567.313131313131,-7.5)(582.242424242424,-7.60606060606061)(585.974747474747,-7.63636363636364)(597.171717171717,-7.72727272727273)(604.636363636364,-7.81818181818182)(612.10101010101,-7.95454545454546)(613.967171717172,-8)(619.565656565657,-8.18181818181818)(619.565656565657,-8.36363636363636)(614.589225589226,-8.54545454545455)(612.10101010101,-8.63636363636364)(597.171717171717,-8.6969696969697)(593.439393939394,-8.72727272727273)(582.242424242424,-8.79545454545455)(567.313131313131,-8.87878787878788)(559.848484848485,-8.90909090909091)(552.383838383838,-8.93181818181818)(537.454545454546,-8.97727272727273)(522.525252525253,-8.97727272727273)(507.59595959596,-8.93181818181818)(503.863636363636,-8.90909090909091)(492.666666666667,-8.77272727272727)(485.20202020202,-8.72727272727273)(477.737373737374,-8.6969696969697)(462.808080808081,-8.57575757575757)(447.878787878788,-8.56818181818182)(445.390572390572,-8.54545454545455)(445.390572390572,-8.36363636363636)(440.414141414141,-8.18181818181818)(445.390572390572,-8)(447.878787878788,-7.90909090909091)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (119.434343434343,-8.15151515151515)(131.875420875421,-8)(134.363636363636,-7.96969696969697)(149.292929292929,-7.90909090909091)(164.222222222222,-7.86363636363636)(179.151515151515,-7.86363636363636)(194.080808080808,-7.86363636363636)(209.010101010101,-7.96969696969697)(212.742424242424,-8)(223.939393939394,-8.13636363636364)(225.805555555556,-8.18181818181818)(233.270202020202,-8.36363636363636)(227.671717171717,-8.54545454545455)(223.939393939394,-8.59090909090909)(209.010101010101,-8.6969696969697)(201.545454545455,-8.72727272727273)(194.080808080808,-8.75)(179.151515151515,-8.86363636363636)(164.222222222222,-8.86363636363636)(153.025252525253,-8.72727272727273)(149.292929292929,-8.6969696969697)(134.363636363636,-8.56818181818182)(126.89898989899,-8.54545454545455)(119.434343434343,-8.5)(113.835858585859,-8.36363636363636)(116.946127946128,-8.18181818181818)(119.434343434343,-8.15151515151515)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273)(447.878787878788,0)(1089.83838383838,-12.1818181818182)(313.515151515152,-6.54545454545455)(268.727272727273,-11.0909090909091)(1000.26262626263,-9.27272727272727)(477.737373737374,-8.72727272727273)(1478,-8.54545454545455)(1089.83838383838,-5.63636363636364)(0,-10.9090909090909)(0,-6.36363636363636)(358.30303030303,-4.36363636363636)(1478,-6.36363636363636)(1239.13131313131,-10.1818181818182)(761.393939393939,-8.36363636363636)(1373.49494949495,-12.1818181818182)(597.171717171717,-5.45454545454546)(223.939393939394,-9.09090909090909)(985.333333333333,-7.27272727272727)(895.757575757576,-11.2727272727273)(1298.84848484848,-5.45454545454546)(865.89898989899,-5.63636363636364)(462.808080808081,-10.3636363636364)(522.525252525253,-7.27272727272727)(179.151515151515,-7.63636363636364)(164.222222222222,-5.81818181818182)(776.323232323232,-9.63636363636364)(1239.13131313131,-8.90909090909091)(1478,-9.63636363636364)(1074.90909090909,-10.7272727272727)(119.434343434343,-9.81818181818182)(1478,-7.45454545454546)(0,-7.45454545454546)(1179.41414141414,-6.72727272727273)(447.878787878788,-6)(1283.91919191919,-11.0909090909091)(612.10101010101,-9.27272727272727)(1074.90909090909,-8.18181818181818)(343.373737373737,-8)(328.444444444444,-9.81818181818182)(836.040404040404,-7.45454545454546)(910.686868686869,-10.1818181818182)(641.959595959596,-7.81818181818182)(1358.56565656566,-9.45454545454546)(910.686868686869,-8.54545454545455)(149.292929292929,-6.72727272727273)(1343.63636363636,-7.27272727272727)(1119.69696969697,-9.63636363636364)(1015.19191919192,-6.72727272727273)(597.171717171717,-6.54545454545455)(119.434343434343,-8.54545454545455)(388.161616161616,-7.27272727272727)(1358.56565656566,-8.36363636363636)(1478,-10.1818181818182)(447.878787878788,-9.45454545454546)(1149.55555555556,-10.9090909090909)(1149.55555555556,-7.81818181818182)(1358.56565656566,-10.7272727272727)(328.444444444444,-8.72727272727273)(567.313131313131,-8.36363636363636)(0,-7.09090909090909)(850.969696969697,-9.27272727272727)(1478,-8.90909090909091)(1000.26262626263,-10.7272727272727)(701.676767676768,-7.45454545454546)(1134.62626262626,-8.90909090909091)(671.818181818182,-9.09090909090909)(253.79797979798,-7.27272727272727)(940.545454545455,-8)(462.808080808081,-6.72727272727273)(880.828282828283,-10.3636363636364)(1268.9898989899,-7.09090909090909)(104.505050505051,-7.81818181818182)(462.808080808081,-8)(1313.77777777778,-9.81818181818182)(1104.76767676768,-6.90909090909091)(895.757575757576,-7.27272727272727)(134.363636363636,-8.90909090909091)(1030.12121212121,-9.81818181818182)(1433.21212121212,-8)(1254.06060606061,-8.36363636363636)(253.79797979798,-8.18181818181818)(537.454545454546,-9.27272727272727)(298.585858585859,-9.27272727272727)(791.252525252525,-8.54545454545455)(612.10101010101,-6.90909090909091)(29.8585858585859,-8)(1015.19191919192,-8.72727272727273)(1478,-10)(119.434343434343,-7.09090909090909)(1179.41414141414,-9.81818181818182)(836.040404040404,-10)(1358.56565656566,-8.90909090909091)(1373.49494949495,-10.5454545454545)(1089.83838383838,-7.63636363636364)(656.888888888889,-8.36363636363636)(343.373737373737,-7.09090909090909)(1478,-9.09090909090909)(403.090909090909,-8.72727272727273)(1194.34343434343,-10.5454545454545)(940.545454545455,-7.09090909090909)(791.252525252525,-7.81818181818182)(522.525252525253,-7.63636363636364)(925.616161616162,-9.27272727272727)(1373.49494949495,-7.63636363636364)(761.393939393939,-9.27272727272727)(164.222222222222,-9.09090909090909)(1164.48484848485,-8.54545454545455)(1000.26262626263,-10.5454545454545)(492.666666666667,-6.72727272727273)(582.242424242424,-9.27272727272727)(1254.06060606061,-9.45454545454546)(1194.34343434343,-7.27272727272727)(209.010101010101,-8.18181818181818)(388.161616161616,-9.27272727272727)(985.333333333333,-8)(641.959595959596,-7.27272727272727)(388.161616161616,-7.81818181818182)(59.7171717171717,-8.36363636363636)(1373.49494949495,-10)(1089.83838383838,-10.1818181818182)(209.010101010101,-7.27272727272727)(880.828282828283,-8.36363636363636)(1283.91919191919,-8.18181818181818)(1074.90909090909,-9.09090909090909)(1448.14141414141,-8.36363636363636)(582.242424242424,-8.36363636363636)(89.5757575757576,-7.27272727272727)(1268.9898989899,-10.5454545454545)(1478,-9.45454545454546)(1045.05050505051,-7.09090909090909)(955.474747474747,-9.63636363636364)(821.111111111111,-9.81818181818182)(716.606060606061,-8.72727272727273)(253.79797979798,-9.09090909090909)(492.666666666667,-9.27272727272727)(865.89898989899,-7.45454545454546)(1358.56565656566,-9.09090909090909)(447.878787878788,-8.36363636363636)(418.020202020202,-7.09090909090909)(701.676767676768,-7.81818181818182)(104.505050505051,-8.72727272727273)(283.656565656566,-7.81818181818182)(1164.48484848485,-8.18181818181818)(1313.77777777778,-7.45454545454546)(1194.34343434343,-9.27272727272727)(880.828282828283,-8.90909090909091)(567.313131313131,-6.90909090909091)(1030.12121212121,-8.36363636363636)(1134.62626262626,-10.3636363636364)(1119.69696969697,-7.27272727272727)(1373.49494949495,-10.3636363636364)(298.585858585859,-8.54545454545455)(970.40404040404,-10.1818181818182)(552.383838383838,-7.63636363636364)(59.7171717171717,-7.81818181818182)(567.313131313131,-9.27272727272727)(1388.42424242424,-8)(746.464646464646,-9.09090909090909)(1478,-9.27272727272727)(806.181818181818,-7.81818181818182)(179.151515151515,-8.18181818181818)(1298.84848484848,-8.72727272727273)(1074.90909090909,-9.45454545454546)(1209.27272727273,-7.63636363636364)(910.686868686869,-7.27272727272727)(179.151515151515,-9.09090909090909)(268.727272727273,-7.45454545454546)(1254.06060606061,-9.63636363636364)(447.878787878788,-7.63636363636364)(627.030303030303,-8.54545454545455)(1015.19191919192,-8)(388.161616161616,-8.90909090909091)(1388.42424242424,-9.81818181818182)(955.474747474747,-8.90909090909091)(641.959595959596,-7.45454545454546)(1268.9898989899,-10.3636363636364)(850.969696969697,-9.81818181818182)(1119.69696969697,-8.72727272727273)(821.111111111111,-8.36363636363636)(134.363636363636,-7.45454545454546)(507.59595959596,-8.72727272727273)(1045.05050505051,-10.1818181818182)(1030.12121212121,-7.45454545454546)(1403.35353535354,-8.72727272727273)(358.30303030303,-8.18181818181818)(1343.63636363636,-7.63636363636364)(89.5757575757576,-8.54545454545455)(1224.20202020202,-8.36363636363636)(492.666666666667,-7.09090909090909)(686.747474747475,-8.18181818181818)(1164.48484848485,-10.1818181818182)(821.111111111111,-9.09090909090909)(238.868686868687,-8.54545454545455)(925.616161616162,-8.18181818181818)(507.59595959596,-9.27272727272727)(1194.34343434343,-7.45454545454546)(940.545454545455,-10)(1433.21212121212,-10)(1179.41414141414,-9.27272727272727)(582.242424242424,-7.81818181818182)(373.232323232323,-7.63636363636364)(1313.77777777778,-9.09090909090909)(1015.19191919192,-9.09090909090909)(194.080808080808,-9.09090909090909)(627.030303030303,-8.90909090909091)(223.939393939394,-7.45454545454546)(1104.76767676768,-8.18181818181818)(940.545454545455,-7.27272727272727)(1403.35353535354,-8.18181818181818)(134.363636363636,-8)(432.949494949495,-8.36363636363636)(1059.9797979798,-7.27272727272727)(836.040404040404,-7.81818181818182)(1298.84848484848,-10.1818181818182)(791.252525252525,-9.27272727272727)(1418.28282828283,-9.27272727272727)(940.545454545455,-9.27272727272727)(1104.76767676768,-9.45454545454546)(298.585858585859,-8.72727272727273)(597.171717171717,-7.27272727272727)(447.878787878788,-7.27272727272727)(1298.84848484848,-7.81818181818182)(985.333333333333,-10.1818181818182)(74.6464646464647,-8)(1239.13131313131,-8.72727272727273)(462.808080808081,-9.09090909090909)(1119.69696969697,-10.1818181818182)(776.323232323232,-8.54545454545455)(1000.26262626263,-8.54545454545455)(1388.42424242424,-10.1818181818182)(1179.41414141414,-7.45454545454546)(298.585858585859,-7.81818181818182)(119.434343434343,-8.72727272727273)(865.89898989899,-9.81818181818182)(656.888888888889,-8.72727272727273)(149.292929292929,-7.45454545454546)(1418.28282828283,-8.36363636363636)(955.474747474747,-8)(1224.20202020202,-9.45454545454546)(447.878787878788,-7.81818181818182)(612.10101010101,-8)(1104.76767676768,-8.54545454545455)(850.969696969697,-7.63636363636364)(895.757575757576,-8.72727272727273)(1224.20202020202,-10.1818181818182)(223.939393939394,-8.36363636363636)(1045.05050505051,-7.27272727272727)(597.171717171717,-9.09090909090909)(1448.14141414141,-8.90909090909091)(1030.12121212121,-9.45454545454546)(1328.70707070707,-7.63636363636364)(373.232323232323,-8.54545454545455)(1283.91919191919,-8.54545454545455)(1448.14141414141,-9.63636363636364)(955.474747474747,-7.27272727272727)(1059.9797979798,-10.1818181818182)(791.252525252525,-8.18181818181818)(1283.91919191919,-9.27272727272727)(910.686868686869,-10)(582.242424242424,-7.27272727272727)(253.79797979798,-8.90909090909091)(104.505050505051,-7.63636363636364)(1179.41414141414,-8.36363636363636)(1134.62626262626,-7.45454545454546)(462.808080808081,-8.72727272727273)(1030.12121212121,-8)(656.888888888889,-8)(1328.70707070707,-10.1818181818182)(791.252525252525,-9.09090909090909)(462.808080808081,-7.27272727272727)(74.6464646464647,-8.18181818181818)(1149.55555555556,-9.45454545454546)(149.292929292929,-8.90909090909091)(1239.13131313131,-7.63636363636364)(268.727272727273,-7.81818181818182)(910.686868686869,-9.09090909090909)(1373.49494949495,-8)(1059.9797979798,-8.72727272727273)(1179.41414141414,-10.1818181818182)(597.171717171717,-8.72727272727273)(432.949494949495,-8.18181818181818)(836.040404040404,-9.63636363636364)(1418.28282828283,-9.45454545454546)(910.686868686869,-8.36363636363636)(1000.26262626263,-10.1818181818182)(164.222222222222,-7.45454545454546)(477.737373737374,-9.09090909090909)(895.757575757576,-7.45454545454546)(1433.21212121212,-8.72727272727273)(1418.28282828283,-10)(985.333333333333,-9.45454545454546)(1179.41414141414,-8.90909090909091)(418.020202020202,-7.45454545454546)(313.515151515152,-8.36363636363636)(776.323232323232,-8.72727272727273)(537.454545454546,-7.45454545454546)(1104.76767676768,-7.45454545454546)(1283.91919191919,-10.1818181818182)(1283.91919191919,-9.09090909090909)(1000.26262626263,-7.27272727272727) 
};
\end{axis}
\end{tikzpicture}%