### 2013-03-21 figurefactory/files/limno_bgape_imp_class_40.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00028436207,-4.18181818181818)(1478.00028436207,-4.36363636363636)(1478.00028436207,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478.00007109153,-6.54545454545455)(1478.00028436207,-6.72727272727273)(1478.00049762651,-6.90909090909091)(1478.00063979942,-7.09090909090909)(1478.00063979942,-7.27272727272727)(1478.00063979942,-7.45454545454546)(1478.00063979942,-7.63636363636364)(1478.00063979942,-7.81818181818182)(1478.00063979942,-8)(1478.00063979942,-8.18181818181818)(1478.00049762651,-8.36363636363636)(1478.00049762651,-8.54545454545455)(1478.00049762651,-8.72727272727273)(1478.00063979942,-8.90909090909091)(1478.00063979942,-9.09090909090909)(1478.00063979942,-9.27272727272727)(1478.00063979942,-9.45454545454546)(1478.00063979942,-9.63636363636364)(1478.00063979942,-9.81818181818182)(1478.00063979942,-10)(1478.00063979942,-10.1818181818182)(1478.00063979942,-10.3636363636364)(1478.00063979942,-10.5454545454545)(1478.00078196962,-10.7272727272727)(1478.0005687133,-10.9090909090909)(1478.00035545089,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(-0.000355450892342165,-8.72727272727273)(-0.000568713303462667,-8.54545454545455)(-0.000781969621696758,-8.36363636363636)(-0.00063979941985036,-8.18181818181818)(-0.00063979941985036,-8)(-0.00063979941985036,-7.81818181818182)(-0.00063979941985036,-7.63636363636364)(-0.00063979941985036,-7.45454545454546)(-0.00063979941985036,-7.27272727272727)(-0.00063979941985036,-7.09090909090909)(-0.00063979941985036,-6.90909090909091)(-0.00063979941985036,-6.72727272727273)(-0.000497626510092015,-6.54545454545455)(-0.000497626510092015,-6.36363636363636)(-0.000355450892342165,-6.18181818181818)(-0.000284362067949707,-6)(-7.1091532562972e-05,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-6.09090909090909)(7.46464646464646,-6)(14.9292929292929,-5.96969696969697)(29.8585858585859,-5.90909090909091)(44.7878787878788,-5.84848484848485)(52.2525252525253,-5.81818181818182)(59.7171717171717,-5.78787878787879)(74.6464646464647,-5.72727272727273)(89.5757575757576,-5.66666666666667)(97.040404040404,-5.63636363636364)(104.505050505051,-5.60606060606061)(119.434343434343,-5.54545454545455)(134.363636363636,-5.54545454545455)(149.292929292929,-5.54545454545455)(164.222222222222,-5.54545454545455)(179.151515151515,-5.54545454545455)(194.080808080808,-5.54545454545455)(209.010101010101,-5.54545454545455)(223.939393939394,-5.54545454545455)(238.868686868687,-5.60606060606061)(246.333333333333,-5.63636363636364)(253.79797979798,-5.66666666666667)(268.727272727273,-5.72727272727273)(283.656565656566,-5.72727272727273)(298.585858585859,-5.72727272727273)(313.515151515152,-5.72727272727273)(328.444444444444,-5.72727272727273)(343.373737373737,-5.72727272727273)(358.30303030303,-5.72727272727273)(373.232323232323,-5.72727272727273)(388.161616161616,-5.72727272727273)(403.090909090909,-5.72727272727273)(418.020202020202,-5.72727272727273)(432.949494949495,-5.72727272727273)(447.878787878788,-5.72727272727273)(462.808080808081,-5.72727272727273)(477.737373737374,-5.72727272727273)(492.666666666667,-5.72727272727273)(507.59595959596,-5.72727272727273)(522.525252525253,-5.78787878787879)(529.989898989899,-5.81818181818182)(537.454545454546,-5.84848484848485)(552.383838383838,-5.90909090909091)(567.313131313131,-5.96969696969697)(574.777777777778,-6)(582.242424242424,-6.03030303030303)(597.171717171717,-6.09090909090909)(612.10101010101,-6.15151515151515)(619.565656565657,-6.18181818181818)(627.030303030303,-6.21212121212121)(641.959595959596,-6.33333333333333)(649.424242424242,-6.36363636363636)(656.888888888889,-6.39393939393939)(671.818181818182,-6.51515151515151)(679.282828282828,-6.54545454545455)(686.747474747475,-6.57575757575758)(701.676767676768,-6.63636363636364)(712.873737373737,-6.54545454545455)(716.606060606061,-6.51515151515151)(731.535353535354,-6.51515151515151)(746.464646464646,-6.51515151515151)(753.929292929293,-6.54545454545455)(761.393939393939,-6.57575757575758)(765.126262626263,-6.54545454545455)(776.323232323232,-6.45454545454545)(783.787878787879,-6.36363636363636)(791.252525252525,-6.27272727272727)(798.717171717172,-6.18181818181818)(806.181818181818,-6.09090909090909)(813.646464646465,-6)(821.111111111111,-5.90909090909091)(832.308080808081,-5.81818181818182)(836.040404040404,-5.78787878787879)(850.969696969697,-5.66666666666667)(858.434343434343,-5.63636363636364)(865.89898989899,-5.60606060606061)(880.828282828283,-5.54545454545455)(895.757575757576,-5.54545454545455)(910.686868686869,-5.60606060606061)(918.151515151515,-5.63636363636364)(925.616161616162,-5.66666666666667)(940.545454545455,-5.72727272727273)(955.474747474747,-5.72727272727273)(970.40404040404,-5.72727272727273)(985.333333333333,-5.78787878787879)(992.79797979798,-5.81818181818182)(1000.26262626263,-5.84848484848485)(1015.19191919192,-5.90909090909091)(1030.12121212121,-5.90909090909091)(1045.05050505051,-5.90909090909091)(1059.9797979798,-5.90909090909091)(1074.90909090909,-5.90909090909091)(1089.83838383838,-5.90909090909091)(1104.76767676768,-5.90909090909091)(1119.69696969697,-5.90909090909091)(1134.62626262626,-5.84848484848485)(1142.09090909091,-5.81818181818182)(1149.55555555556,-5.78787878787879)(1164.48484848485,-5.72727272727273)(1179.41414141414,-5.66666666666667)(1186.87878787879,-5.63636363636364)(1194.34343434343,-5.60606060606061)(1209.27272727273,-5.48484848484849)(1216.73737373737,-5.45454545454546)(1224.20202020202,-5.42424242424242)(1239.13131313131,-5.36363636363636)(1254.06060606061,-5.36363636363636)(1268.9898989899,-5.36363636363636)(1283.91919191919,-5.42424242424242)(1298.84848484848,-5.42424242424242)(1313.77777777778,-5.42424242424242)(1328.70707070707,-5.42424242424242)(1336.17171717172,-5.45454545454546)(1343.63636363636,-5.48484848484849)(1358.56565656566,-5.60606060606061)(1362.29797979798,-5.63636363636364)(1373.49494949495,-5.72727272727273)(1380.9595959596,-5.81818181818182)(1388.42424242424,-5.90909090909091)(1395.88888888889,-6)(1403.35353535354,-6.09090909090909)(1410.81818181818,-6.18181818181818)(1418.28282828283,-6.27272727272727)(1425.74747474747,-6.36363636363636)(1433.21212121212,-6.45454545454545)(1444.40909090909,-6.54545454545455)(1448.14141414141,-6.57575757575758)(1463.07070707071,-6.6969696969697)(1470.53535353535,-6.72727272727273)(1478,-6.75757575757576)(1478.00017772375,-6.90909090909091)(1478.00031989971,-7.09090909090909)(1478.00031989971,-7.27272727272727)(1478.00031989971,-7.45454545454546)(1478.00031989971,-7.63636363636364)(1478.00031989971,-7.81818181818182)(1478.00031989971,-8)(1478.00031989971,-8.18181818181818)(1478.00017772375,-8.36363636363636)(1478.00017772375,-8.54545454545455)(1478.00017772375,-8.72727272727273)(1478.00031989971,-8.90909090909091)(1478.00031989971,-9.09090909090909)(1478.00031989971,-9.27272727272727)(1478.00031989971,-9.45454545454546)(1478.00031989971,-9.63636363636364)(1478.00031989971,-9.81818181818182)(1478.00031989971,-10)(1478.00031989971,-10.1818181818182)(1478.00031989971,-10.3636363636364)(1478.00031989971,-10.5454545454545)(1478.00046207296,-10.7272727272727)(1478.00024881207,-10.9090909090909)(1478.00003554509,-11.0909090909091)(1478,-11.1090909090909)(1470.53535353535,-11.0909090909091)(1463.07070707071,-11.0606060606061)(1448.14141414141,-11)(1433.21212121212,-11.0606060606061)(1425.74747474747,-11.0909090909091)(1418.28282828283,-11.1212121212121)(1403.35353535354,-11.2424242424242)(1395.88888888889,-11.2727272727273)(1388.42424242424,-11.3030303030303)(1373.49494949495,-11.3636363636364)(1358.56565656566,-11.4242424242424)(1351.10101010101,-11.4545454545455)(1343.63636363636,-11.4848484848485)(1328.70707070707,-11.5454545454545)(1313.77777777778,-11.5454545454545)(1298.84848484848,-11.5454545454545)(1283.91919191919,-11.5454545454545)(1268.9898989899,-11.5454545454545)(1254.06060606061,-11.5454545454545)(1239.13131313131,-11.5454545454545)(1224.20202020202,-11.5454545454545)(1209.27272727273,-11.5454545454545)(1194.34343434343,-11.5454545454545)(1179.41414141414,-11.5454545454545)(1164.48484848485,-11.5454545454545)(1149.55555555556,-11.5454545454545)(1134.62626262626,-11.5454545454545)(1119.69696969697,-11.5454545454545)(1104.76767676768,-11.5454545454545)(1089.83838383838,-11.5454545454545)(1074.90909090909,-11.4848484848485)(1067.44444444444,-11.4545454545455)(1059.9797979798,-11.4242424242424)(1045.05050505051,-11.3636363636364)(1030.12121212121,-11.3636363636364)(1015.19191919192,-11.3636363636364)(1000.26262626263,-11.3636363636364)(985.333333333333,-11.3030303030303)(977.868686868687,-11.2727272727273)(970.40404040404,-11.2424242424242)(955.474747474747,-11.1818181818182)(940.545454545455,-11.1212121212121)(933.080808080808,-11.0909090909091)(925.616161616162,-11.0606060606061)(910.686868686869,-11)(895.757575757576,-11)(880.828282828283,-10.9393939393939)(873.363636363636,-10.9090909090909)(865.89898989899,-10.8787878787879)(850.969696969697,-10.8181818181818)(836.040404040404,-10.7575757575758)(828.575757575758,-10.7272727272727)(821.111111111111,-10.6969696969697)(806.181818181818,-10.5757575757576)(798.717171717172,-10.5454545454545)(791.252525252525,-10.5151515151515)(776.323232323232,-10.3939393939394)(768.858585858586,-10.3636363636364)(761.393939393939,-10.3333333333333)(746.464646464646,-10.2727272727273)(731.535353535354,-10.2121212121212)(724.070707070707,-10.1818181818182)(716.606060606061,-10.1515151515152)(701.676767676768,-10.0909090909091)(686.747474747475,-10.0909090909091)(671.818181818182,-10.0909090909091)(656.888888888889,-10.030303030303)(649.424242424242,-10)(641.959595959596,-9.96969696969697)(627.030303030303,-9.96969696969697)(619.565656565657,-10)(612.10101010101,-10.030303030303)(597.171717171717,-10.0909090909091)(582.242424242424,-10.0909090909091)(567.313131313131,-10.0909090909091)(552.383838383838,-10.1515151515152)(544.919191919192,-10.1818181818182)(537.454545454546,-10.2121212121212)(522.525252525253,-10.2727272727273)(507.59595959596,-10.2727272727273)(492.666666666667,-10.3333333333333)(485.20202020202,-10.3636363636364)(477.737373737374,-10.3939393939394)(462.808080808081,-10.4545454545455)(447.878787878788,-10.4545454545455)(432.949494949495,-10.4545454545455)(418.020202020202,-10.4545454545455)(403.090909090909,-10.4545454545455)(388.161616161616,-10.3939393939394)(380.69696969697,-10.3636363636364)(373.232323232323,-10.3333333333333)(358.30303030303,-10.2727272727273)(343.373737373737,-10.2727272727273)(328.444444444444,-10.2727272727273)(313.515151515152,-10.2727272727273)(298.585858585859,-10.2727272727273)(283.656565656566,-10.2727272727273)(268.727272727273,-10.2727272727273)(253.79797979798,-10.2727272727273)(238.868686868687,-10.2727272727273)(223.939393939394,-10.2727272727273)(209.010101010101,-10.2727272727273)(194.080808080808,-10.2727272727273)(179.151515151515,-10.2727272727273)(164.222222222222,-10.2727272727273)(149.292929292929,-10.2121212121212)(141.828282828283,-10.1818181818182)(134.363636363636,-10.1515151515152)(119.434343434343,-10.0909090909091)(104.505050505051,-10.0909090909091)(89.5757575757576,-10.030303030303)(82.1111111111111,-10)(74.6464646464647,-9.96969696969697)(59.7171717171717,-9.84848484848485)(55.9848484848485,-9.81818181818182)(44.7878787878788,-9.72727272727273)(37.3232323232323,-9.63636363636364)(29.8585858585859,-9.5)(27.3703703703704,-9.45454545454546)(22.3939393939394,-9.27272727272727)(22.3939393939394,-9.09090909090909)(22.3939393939394,-8.90909090909091)(14.9292929292929,-8.81818181818182)(0,-8.74545454545454)(-3.55450892355425e-05,-8.72727272727273)(-0.00024881207026616,-8.54545454545455)(-0.000462072958277768,-8.36363636363636)(-0.000319899709928495,-8.18181818181818)(-0.000319899709928495,-8)(-0.000319899709928495,-7.81818181818182)(-0.000319899709928495,-7.63636363636364)(-0.000319899709928495,-7.45454545454546)(-0.000319899709928495,-7.27272727272727)(-0.000319899709928495,-7.09090909090909)(-0.000319899709928495,-6.90909090909091)(-0.000319899709928495,-6.72727272727273)(-0.000177723753604291,-6.54545454545455)(-0.000177723753604291,-6.36363636363636)(-3.55450892355425e-05,-6.18181818181818)(0,-6.09090909090909)};

\addplot [fill=green!60!black,draw=none,forget plot] coordinates{ (1074.90909090909,-8.93181818181818)(1087.97222222222,-9.09090909090909)(1087.97222222222,-9.27272727272727)(1074.90909090909,-9.43181818181818)(1071.17676767677,-9.45454545454546)(1059.9797979798,-9.5)(1045.05050505051,-9.5)(1033.85353535354,-9.45454545454546)(1030.12121212121,-9.43181818181818)(1015.19191919192,-9.31818181818182)(1013.32575757576,-9.27272727272727)(1013.32575757576,-9.09090909090909)(1015.19191919192,-9.06060606060606)(1030.12121212121,-8.93181818181818)(1033.85353535354,-8.90909090909091)(1045.05050505051,-8.86363636363636)(1059.9797979798,-8.86363636363636)(1071.17676767677,-8.90909090909091)(1074.90909090909,-8.93181818181818)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273)(447.878787878788,0)(1089.83838383838,-12.1818181818182)(313.515151515152,-6.54545454545455)(268.727272727273,-11.0909090909091)(1000.26262626263,-9.27272727272727)(477.737373737374,-8.72727272727273)(1478,-8.54545454545455)(1089.83838383838,-5.63636363636364)(0,-10.9090909090909)(0,-6.36363636363636)(358.30303030303,-4.36363636363636)(1478,-6.36363636363636)(1239.13131313131,-10.1818181818182)(761.393939393939,-8.36363636363636)(1373.49494949495,-12.1818181818182)(597.171717171717,-5.45454545454546)(223.939393939394,-9.09090909090909)(985.333333333333,-7.27272727272727)(895.757575757576,-11.2727272727273)(1298.84848484848,-5.45454545454546) 
};
\end{axis}
\end{tikzpicture}%