### 2013-03-21 figurefactory/files/limno_bgape_imp_class_80.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478.00022117143,-4.18181818181818)(1478.00022117143,-4.36363636363636)(1478.00022117143,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478.00005529347,-8.18181818181818)(1478.00011058654,-8.36363636363636)(1478.00027646327,-8.54545454545455)(1478.00038704571,-8.72727272727273)(1478.00049762651,-8.90909090909091)(1478.00049762651,-9.09090909090909)(1478.00049762651,-9.27272727272727)(1478.00038704571,-9.45454545454546)(1478.00038704571,-9.63636363636364)(1478.00038704571,-9.81818181818182)(1478.00038704571,-10)(1478.00022117143,-10.1818181818182)(1478.00005529347,-10.3636363636364)(1478,-10.5454545454545)(1478.00022117143,-10.7272727272727)(1478.00022117143,-10.9090909090909)(1478.00022117143,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(-0.000221171433450288,-8.72727272727273)(-0.000221171433450288,-8.54545454545455)(-0.000221171433450288,-8.36363636363636)(0,-8.18181818181818)(-5.52934727241102e-05,-8)(-0.000221171433450288,-7.81818181818182)(-0.000276463267890224,-7.63636363636364)(-0.000387045708090112,-7.45454545454546)(-0.000276463267890224,-7.27272727272727)(-0.000221171433450288,-7.09090909090909)(-5.52934727241102e-05,-6.90909090909091)(0,-6.72727272727273)(0,-6.54545454545455)(0,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-7.18181818181818)(3.73232323232323,-7.09090909090909)(14.9292929292929,-7)(26.1262626262626,-6.90909090909091)(29.8585858585859,-6.87878787878788)(44.7878787878788,-6.81818181818182)(59.7171717171717,-6.81818181818182)(74.6464646464647,-6.81818181818182)(89.5757575757576,-6.81818181818182)(104.505050505051,-6.81818181818182)(119.434343434343,-6.81818181818182)(134.363636363636,-6.81818181818182)(149.292929292929,-6.81818181818182)(164.222222222222,-6.81818181818182)(179.151515151515,-6.81818181818182)(194.080808080808,-6.81818181818182)(209.010101010101,-6.81818181818182)(223.939393939394,-6.81818181818182)(238.868686868687,-6.81818181818182)(253.79797979798,-6.81818181818182)(268.727272727273,-6.81818181818182)(283.656565656566,-6.81818181818182)(294.853535353535,-6.72727272727273)(298.585858585859,-6.6969696969697)(313.515151515152,-6.6969696969697)(328.444444444444,-6.6969696969697)(332.176767676768,-6.72727272727273)(343.373737373737,-6.81818181818182)(358.30303030303,-6.81818181818182)(373.232323232323,-6.81818181818182)(388.161616161616,-6.81818181818182)(403.090909090909,-6.81818181818182)(418.020202020202,-6.81818181818182)(432.949494949495,-6.75757575757576)(440.414141414141,-6.72727272727273)(447.878787878788,-6.6969696969697)(462.808080808081,-6.63636363636364)(477.737373737374,-6.63636363636364)(492.666666666667,-6.63636363636364)(507.59595959596,-6.63636363636364)(522.525252525253,-6.63636363636364)(537.454545454546,-6.63636363636364)(552.383838383838,-6.63636363636364)(567.313131313131,-6.6969696969697)(574.777777777778,-6.72727272727273)(582.242424242424,-6.75757575757576)(597.171717171717,-6.81818181818182)(612.10101010101,-6.87878787878788)(619.565656565657,-6.90909090909091)(627.030303030303,-6.93939393939394)(641.959595959596,-7.06060606060606)(649.424242424242,-7.09090909090909)(656.888888888889,-7.12121212121212)(671.818181818182,-7.24242424242424)(679.282828282828,-7.27272727272727)(686.747474747475,-7.3030303030303)(701.676767676768,-7.42424242424243)(709.141414141414,-7.45454545454546)(716.606060606061,-7.48484848484849)(731.535353535354,-7.60606060606061)(739,-7.63636363636364)(746.464646464646,-7.66666666666667)(761.393939393939,-7.72727272727273)(776.323232323232,-7.72727272727273)(791.252525252525,-7.72727272727273)(806.181818181818,-7.72727272727273)(821.111111111111,-7.66666666666667)(828.575757575758,-7.63636363636364)(836.040404040404,-7.60606060606061)(850.969696969697,-7.48484848484849)(858.434343434343,-7.45454545454546)(865.89898989899,-7.42424242424243)(880.828282828283,-7.3030303030303)(888.292929292929,-7.27272727272727)(895.757575757576,-7.24242424242424)(910.686868686869,-7.18181818181818)(925.616161616162,-7.12121212121212)(933.080808080808,-7.09090909090909)(940.545454545455,-7.06060606060606)(955.474747474747,-7)(970.40404040404,-6.97727272727273)(985.333333333333,-6.97727272727273)(1000.26262626263,-6.93181818181818)(1007.72727272727,-6.90909090909091)(1015.19191919192,-6.87878787878788)(1030.12121212121,-6.81818181818182)(1045.05050505051,-6.81818181818182)(1059.9797979798,-6.81818181818182)(1074.90909090909,-6.81818181818182)(1089.83838383838,-6.81818181818182)(1104.76767676768,-6.81818181818182)(1119.69696969697,-6.81818181818182)(1134.62626262626,-6.81818181818182)(1149.55555555556,-6.87878787878788)(1157.0202020202,-6.90909090909091)(1164.48484848485,-6.93939393939394)(1179.41414141414,-7)(1194.34343434343,-7)(1209.27272727273,-7)(1224.20202020202,-7)(1239.13131313131,-7)(1254.06060606061,-7)(1268.9898989899,-7.06060606060606)(1276.45454545455,-7.09090909090909)(1283.91919191919,-7.12121212121212)(1298.84848484848,-7.18181818181818)(1313.77777777778,-7.18181818181818)(1328.70707070707,-7.24242424242424)(1336.17171717172,-7.27272727272727)(1343.63636363636,-7.3030303030303)(1358.56565656566,-7.42424242424243)(1366.0303030303,-7.45454545454546)(1373.49494949495,-7.48484848484849)(1388.42424242424,-7.60606060606061)(1392.15656565657,-7.63636363636364)(1403.35353535354,-7.72727272727273)(1414.5505050505,-7.81818181818182)(1418.28282828283,-7.84848484848485)(1433.21212121212,-7.96969696969697)(1436.94444444444,-8)(1448.14141414141,-8.09090909090909)(1455.60606060606,-8.18181818181818)(1463.07070707071,-8.31818181818182)(1465.55892255892,-8.36363636363636)(1478,-8.51515151515152)(1478.00002764633,-8.54545454545455)(1478.00013823061,-8.72727272727273)(1478.00024881326,-8.90909090909091)(1478.00024881326,-9.09090909090909)(1478.00024881326,-9.27272727272727)(1478.00013823061,-9.45454545454546)(1478.00013823061,-9.63636363636364)(1478.00013823061,-9.81818181818182)(1478.00013823061,-10)(1478,-10.1515151515152)(1470.53535353535,-10.1818181818182)(1463.07070707071,-10.2121212121212)(1448.14141414141,-10.3333333333333)(1440.67676767677,-10.3636363636364)(1433.21212121212,-10.3939393939394)(1418.28282828283,-10.5151515151515)(1410.81818181818,-10.5454545454545)(1403.35353535354,-10.5757575757576)(1388.42424242424,-10.6363636363636)(1373.49494949495,-10.6363636363636)(1358.56565656566,-10.6363636363636)(1343.63636363636,-10.6363636363636)(1328.70707070707,-10.6363636363636)(1313.77777777778,-10.6363636363636)(1298.84848484848,-10.6363636363636)(1283.91919191919,-10.6363636363636)(1268.9898989899,-10.6363636363636)(1254.06060606061,-10.6363636363636)(1239.13131313131,-10.6363636363636)(1224.20202020202,-10.6363636363636)(1209.27272727273,-10.6363636363636)(1194.34343434343,-10.6363636363636)(1179.41414141414,-10.6363636363636)(1164.48484848485,-10.6363636363636)(1149.55555555556,-10.6363636363636)(1134.62626262626,-10.6363636363636)(1119.69696969697,-10.6363636363636)(1104.76767676768,-10.6363636363636)(1089.83838383838,-10.6363636363636)(1074.90909090909,-10.6363636363636)(1059.9797979798,-10.6969696969697)(1052.51515151515,-10.7272727272727)(1045.05050505051,-10.7575757575758)(1030.12121212121,-10.8181818181818)(1015.19191919192,-10.8181818181818)(1000.26262626263,-10.7575757575758)(992.79797979798,-10.7272727272727)(985.333333333333,-10.6969696969697)(970.40404040404,-10.6363636363636)(955.474747474747,-10.6363636363636)(940.545454545455,-10.6363636363636)(925.616161616162,-10.5757575757576)(918.151515151515,-10.5454545454545)(910.686868686869,-10.5)(895.757575757576,-10.4090909090909)(880.828282828283,-10.3939393939394)(877.09595959596,-10.3636363636364)(865.89898989899,-10.2727272727273)(854.70202020202,-10.1818181818182)(850.969696969697,-10.1515151515152)(836.040404040404,-10.030303030303)(832.308080808081,-10)(821.111111111111,-9.90909090909091)(813.646464646465,-9.81818181818182)(806.181818181818,-9.72727272727273)(798.717171717172,-9.63636363636364)(791.252525252525,-9.54545454545455)(780.055555555556,-9.45454545454546)(776.323232323232,-9.42424242424243)(761.393939393939,-9.3030303030303)(753.929292929293,-9.27272727272727)(746.464646464646,-9.24242424242424)(731.535353535354,-9.12121212121212)(724.070707070707,-9.09090909090909)(716.606060606061,-9.06060606060606)(701.676767676768,-9)(686.747474747475,-9.06060606060606)(679.282828282828,-9.09090909090909)(671.818181818182,-9.12121212121212)(656.888888888889,-9.18181818181818)(641.959595959596,-9.18181818181818)(627.030303030303,-9.18181818181818)(612.10101010101,-9.18181818181818)(597.171717171717,-9.24242424242424)(589.707070707071,-9.27272727272727)(582.242424242424,-9.3030303030303)(567.313131313131,-9.36363636363636)(552.383838383838,-9.36363636363636)(537.454545454546,-9.36363636363636)(522.525252525253,-9.36363636363636)(507.59595959596,-9.36363636363636)(492.666666666667,-9.36363636363636)(477.737373737374,-9.36363636363636)(462.808080808081,-9.36363636363636)(447.878787878788,-9.36363636363636)(432.949494949495,-9.36363636363636)(418.020202020202,-9.36363636363636)(403.090909090909,-9.36363636363636)(388.161616161616,-9.36363636363636)(373.232323232323,-9.36363636363636)(358.30303030303,-9.36363636363636)(343.373737373737,-9.36363636363636)(328.444444444444,-9.36363636363636)(313.515151515152,-9.36363636363636)(298.585858585859,-9.36363636363636)(283.656565656566,-9.36363636363636)(268.727272727273,-9.36363636363636)(253.79797979798,-9.36363636363636)(238.868686868687,-9.38636363636364)(223.939393939394,-9.34090909090909)(209.010101010101,-9.29545454545455)(205.277777777778,-9.27272727272727)(194.080808080808,-9.18181818181818)(179.151515151515,-9.18181818181818)(164.222222222222,-9.12121212121212)(156.757575757576,-9.09090909090909)(149.292929292929,-9.06060606060606)(134.363636363636,-8.93939393939394)(126.89898989899,-8.90909090909091)(119.434343434343,-8.88636363636364)(104.505050505051,-8.79545454545455)(97.040404040404,-8.72727272727273)(89.5757575757576,-8.63636363636364)(78.3787878787879,-8.54545454545455)(74.6464646464647,-8.51515151515152)(59.7171717171717,-8.39393939393939)(55.9848484848485,-8.36363636363636)(44.7878787878788,-8.27272727272727)(37.3232323232323,-8.18181818181818)(29.8585858585859,-8.09090909090909)(22.3939393939394,-8)(14.9292929292929,-7.90909090909091)(3.73232323232323,-7.81818181818182)(0,-7.72727272727273)(-2.76463267876964e-05,-7.63636363636364)(-0.000138230610030762,-7.45454545454546)(-2.76463267876964e-05,-7.27272727272727)(0,-7.18181818181818)};

%\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1373.49494949495,-8.95454545454546)(1377.22727272727,-9.09090909090909)(1374.98787878788,-9.27272727272727)(1373.49494949495,-9.31818181818182)(1367.89646464646,-9.45454545454546)(1358.56565656566,-9.56818181818182)(1352.96717171717,-9.63636363636364)(1343.63636363636,-9.72727272727273)(1328.70707070707,-9.8030303030303)(1324.97474747475,-9.81818181818182)(1313.77777777778,-9.86363636363636)(1298.84848484848,-9.92424242424243)(1283.91919191919,-9.98484848484848)(1280.18686868687,-10)(1268.9898989899,-10.0454545454545)(1254.06060606061,-10.0757575757576)(1239.13131313131,-10.0757575757576)(1224.20202020202,-10.0757575757576)(1209.27272727273,-10.0454545454545)(1194.34343434343,-10.0454545454545)(1179.41414141414,-10.0454545454545)(1164.48484848485,-10.0454545454545)(1149.55555555556,-10.0454545454545)(1134.62626262626,-10.0454545454545)(1119.69696969697,-10.0454545454545)(1104.76767676768,-10.0454545454545)(1089.83838383838,-10.0454545454545)(1074.90909090909,-10.0454545454545)(1059.9797979798,-10.0454545454545)(1045.05050505051,-10.0454545454545)(1033.85353535354,-10)(1030.12121212121,-9.98484848484848)(1015.19191919192,-9.92424242424243)(1000.26262626263,-9.86363636363636)(989.065656565657,-9.81818181818182)(985.333333333333,-9.8030303030303)(970.40404040404,-9.74242424242424)(957.340909090909,-9.63636363636364)(955.474747474747,-9.62121212121212)(940.545454545455,-9.52272727272727)(934.94696969697,-9.45454545454546)(925.616161616162,-9.34090909090909)(920.017676767677,-9.27272727272727)(911.930976430976,-9.09090909090909)(910.686868686869,-9.04545454545455)(906.954545454546,-8.90909090909091)(904.46632996633,-8.72727272727273)(908.820707070707,-8.54545454545455)(910.686868686869,-8.5)(916.285353535354,-8.36363636363636)(925.616161616162,-8.27272727272727)(940.545454545455,-8.1969696969697)(942.411616161616,-8.18181818181818)(955.474747474747,-8.07575757575757)(970.40404040404,-8.01515151515152)(972.270202020202,-8)(985.333333333333,-7.89393939393939)(1000.26262626263,-7.83333333333333)(1003.99494949495,-7.81818181818182)(1015.19191919192,-7.77272727272727)(1030.12121212121,-7.77272727272727)(1045.05050505051,-7.77272727272727)(1059.9797979798,-7.77272727272727)(1074.90909090909,-7.71212121212121)(1089.83838383838,-7.65151515151515)(1093.57070707071,-7.63636363636364)(1104.76767676768,-7.59090909090909)(1119.69696969697,-7.59090909090909)(1134.62626262626,-7.61818181818182)(1138.35858585859,-7.63636363636364)(1149.55555555556,-7.68181818181818)(1164.48484848485,-7.74242424242424)(1179.41414141414,-7.77272727272727)(1194.34343434343,-7.77272727272727)(1209.27272727273,-7.77272727272727)(1224.20202020202,-7.77272727272727)(1235.39898989899,-7.81818181818182)(1239.13131313131,-7.83333333333333)(1254.06060606061,-7.87272727272727)(1268.9898989899,-7.94545454545455)(1280.18686868687,-8)(1283.91919191919,-8.01515151515152)(1298.84848484848,-8.07575757575757)(1311.91161616162,-8.18181818181818)(1313.77777777778,-8.1969696969697)(1328.70707070707,-8.29545454545455)(1336.17171717172,-8.36363636363636)(1343.63636363636,-8.43181818181818)(1352.96717171717,-8.54545454545455)(1358.56565656566,-8.63636363636364)(1364.16414141414,-8.72727272727273)(1372.25084175084,-8.90909090909091)(1373.49494949495,-8.95454545454546)};

\addplot [fill=green!60!black,draw=none,forget plot] coordinates{ (1343.63636363636,-8.79545454545455)(1349.8569023569,-8.90909090909091)(1354.83333333333,-9.09090909090909)(1352.34511784512,-9.27272727272727)(1343.63636363636,-9.43181818181818)(1341.14814814815,-9.45454545454546)(1328.70707070707,-9.53030303030303)(1313.77777777778,-9.59090909090909)(1302.58080808081,-9.63636363636364)(1298.84848484848,-9.65151515151515)(1283.91919191919,-9.71212121212121)(1268.9898989899,-9.77272727272727)(1254.06060606061,-9.77272727272727)(1239.13131313131,-9.77272727272727)(1224.20202020202,-9.77272727272727)(1209.27272727273,-9.77272727272727)(1194.34343434343,-9.77272727272727)(1179.41414141414,-9.77272727272727)(1164.48484848485,-9.77272727272727)(1149.55555555556,-9.77272727272727)(1134.62626262626,-9.74242424242424)(1119.69696969697,-9.74242424242424)(1104.76767676768,-9.74242424242424)(1089.83838383838,-9.77272727272727)(1074.90909090909,-9.77272727272727)(1059.9797979798,-9.77272727272727)(1045.05050505051,-9.77272727272727)(1030.12121212121,-9.71212121212121)(1015.19191919192,-9.65151515151515)(1011.4595959596,-9.63636363636364)(1000.26262626263,-9.58181818181818)(985.333333333333,-9.50909090909091)(970.40404040404,-9.46969696969697)(968.537878787879,-9.45454545454546)(955.474747474747,-9.34848484848485)(946.143939393939,-9.27272727272727)(940.545454545455,-9.20454545454546)(934.324915824916,-9.09090909090909)(929.348484848485,-8.90909090909091)(927.109090909091,-8.72727272727273)(933.080808080808,-8.54545454545455)(940.545454545455,-8.46969696969697)(953.608585858586,-8.36363636363636)(955.474747474747,-8.34848484848485)(970.40404040404,-8.28787878787879)(983.467171717172,-8.18181818181818)(985.333333333333,-8.16666666666667)(1000.26262626263,-8.10606060606061)(1015.19191919192,-8.04545454545455)(1030.12121212121,-8.04545454545455)(1045.05050505051,-8.04545454545455)(1059.9797979798,-8.07575757575757)(1074.90909090909,-8.02272727272727)(1078.64141414141,-8)(1089.83838383838,-7.94545454545455)(1104.76767676768,-7.86363636363636)(1119.69696969697,-7.86363636363636)(1134.62626262626,-7.89393939393939)(1149.55555555556,-7.95454545454546)(1160.75252525253,-8)(1164.48484848485,-8.01818181818182)(1179.41414141414,-8.04545454545455)(1194.34343434343,-8.04545454545455)(1209.27272727273,-8.04545454545455)(1224.20202020202,-8.04545454545455)(1239.13131313131,-8.10606060606061)(1254.06060606061,-8.16666666666667)(1257.79292929293,-8.18181818181818)(1268.9898989899,-8.22727272727273)(1283.91919191919,-8.28787878787879)(1298.84848484848,-8.34848484848485)(1300.71464646465,-8.36363636363636)(1313.77777777778,-8.46969696969697)(1323.10858585859,-8.54545454545455)(1328.70707070707,-8.61363636363636)(1338.03787878788,-8.72727272727273)(1343.63636363636,-8.79545454545455)};
\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (791.252525252525,-14.9090909090909)(0,0)(1478,-4.36363636363636)(0,-8.54545454545455)(0,-18)(1478,-10.9090909090909)(1478,-18)(910.686868686869,0)(731.535353535354,-6.54545454545455)(0,-13.2727272727273)(0,-4.18181818181818)(1478,0)(686.747474747475,-10.7272727272727)(627.030303030303,-18)(537.454545454546,-2.72727272727273)(1478,-14.3636363636364)(1268.9898989899,-7.63636363636364)(1030.12121212121,-3.45454545454546)(1059.9797979798,-18)(328.444444444444,-15.2727272727273)(447.878787878788,0)(1089.83838383838,-12.1818181818182)(313.515151515152,-6.54545454545455)(268.727272727273,-11.0909090909091)(1000.26262626263,-9.27272727272727)(477.737373737374,-8.72727272727273)(1478,-8.54545454545455)(1089.83838383838,-5.63636363636364)(0,-10.9090909090909)(0,-6.36363636363636)(358.30303030303,-4.36363636363636)(1478,-6.36363636363636)(1239.13131313131,-10.1818181818182)(761.393939393939,-8.36363636363636)(1373.49494949495,-12.1818181818182)(597.171717171717,-5.45454545454546)(223.939393939394,-9.09090909090909)(985.333333333333,-7.27272727272727)(895.757575757576,-11.2727272727273)(1298.84848484848,-5.45454545454546)(865.89898989899,-5.63636363636364)(462.808080808081,-10.3636363636364)(522.525252525253,-7.27272727272727)(179.151515151515,-7.63636363636364)(164.222222222222,-5.81818181818182)(776.323232323232,-9.63636363636364)(1239.13131313131,-8.90909090909091)(1478,-9.63636363636364)(1074.90909090909,-10.7272727272727)(119.434343434343,-9.81818181818182)(1478,-7.45454545454546)(0,-7.45454545454546)(1179.41414141414,-6.72727272727273)(447.878787878788,-6)(1283.91919191919,-11.0909090909091)(612.10101010101,-9.27272727272727)(1074.90909090909,-8.18181818181818)(343.373737373737,-8)(328.444444444444,-9.81818181818182)(836.040404040404,-7.45454545454546)(910.686868686869,-10.1818181818182)(641.959595959596,-7.81818181818182)(1358.56565656566,-9.45454545454546)(910.686868686869,-8.54545454545455)(149.292929292929,-6.72727272727273)(1343.63636363636,-7.27272727272727)(1119.69696969697,-9.63636363636364)(1015.19191919192,-6.72727272727273)(597.171717171717,-6.54545454545455)(119.434343434343,-8.54545454545455)(388.161616161616,-7.27272727272727)(1358.56565656566,-8.36363636363636)(1478,-10.1818181818182)(447.878787878788,-9.45454545454546)(1149.55555555556,-10.9090909090909)(1149.55555555556,-7.81818181818182)(1358.56565656566,-10.7272727272727)(328.444444444444,-8.72727272727273)(567.313131313131,-8.36363636363636)(0,-7.09090909090909) 
};
\end{axis}
\end{tikzpicture}%