### 2013-03-22 figurefactory/files/limno_bgape_pp_180_0.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478,-8.36363636363636)(1478,-8.54545454545455)(1478.0002488174,-8.72727272727273)(1478.00049762651,-8.90909090909091)(1478.00074642733,-9.09090909090909)(1478.00074642733,-9.27272727272727)(1478.00049762651,-9.45454545454546)(1478.0002488174,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(0,-7.81818181818182)(-0.000248817401863321,-7.63636363636364)(-0.000497626510092015,-7.45454545454546)(-0.000746427325097138,-7.27272727272727)(-0.000746427325097138,-7.09090909090909)(-0.000746427325097138,-6.90909090909091)(-0.00066349464153746,-6.72727272727273)(-0.0004146910621731,-6.54545454545455)(-0.000165879189445939,-6.36363636363636)(0,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (1478,-8.81818181818182)(1478.00012440663,-8.90909090909091)(1478.00037321366,-9.09090909090909)(1478.00037321366,-9.27272727272727)(1478.00012440663,-9.45454545454546)(1478,-9.54545454545455)(1463.07070707071,-9.60606060606061)(1459.33838383838,-9.63636363636364)(1448.14141414141,-9.70454545454546)(1435.70033670034,-9.81818181818182)(1433.21212121212,-9.84090909090909)(1418.28282828283,-9.90909090909091)(1405.84175084175,-10)(1403.35353535354,-10.0227272727273)(1388.42424242424,-10.030303030303)(1373.49494949495,-10.1136363636364)(1358.56565656566,-10.1590909090909)(1354.83333333333,-10.1818181818182)(1343.63636363636,-10.2363636363636)(1328.70707070707,-10.2878787878788)(1313.77777777778,-10.3181818181818)(1298.84848484848,-10.3090909090909)(1283.91919191919,-10.2954545454545)(1268.9898989899,-10.2727272727273)(1254.06060606061,-10.2121212121212)(1246.59595959596,-10.1818181818182)(1239.13131313131,-10.1666666666667)(1224.20202020202,-10.1666666666667)(1216.73737373737,-10.1818181818182)(1209.27272727273,-10.2121212121212)(1194.34343434343,-10.2727272727273)(1179.41414141414,-10.2727272727273)(1164.48484848485,-10.2954545454545)(1149.55555555556,-10.2727272727273)(1134.62626262626,-10.2575757575758)(1119.69696969697,-10.2878787878788)(1104.76767676768,-10.3484848484848)(1089.83838383838,-10.3484848484848)(1074.90909090909,-10.2878787878788)(1059.9797979798,-10.2575757575758)(1045.05050505051,-10.2878787878788)(1030.12121212121,-10.3090909090909)(1015.19191919192,-10.2954545454545)(1000.26262626263,-10.2727272727273)(985.333333333333,-10.2121212121212)(981.60101010101,-10.1818181818182)(970.40404040404,-10.1363636363636)(955.474747474747,-10.1060606060606)(940.545454545455,-10.0545454545455)(925.616161616162,-10.0227272727273)(918.151515151515,-10)(910.686868686869,-9.98484848484848)(895.757575757576,-9.95454545454546)(880.828282828283,-9.90909090909091)(871.497474747475,-9.81818181818182)(865.89898989899,-9.77272727272727)(850.969696969697,-9.77272727272727)(836.040404040404,-9.70454545454546)(821.111111111111,-9.66666666666667)(819.244949494949,-9.63636363636364)(810.660606060606,-9.45454545454546)(806.181818181818,-9.40909090909091)(791.252525252525,-9.4)(776.323232323232,-9.29545454545455)(773.835016835017,-9.27272727272727)(763.882154882155,-9.09090909090909)(761.393939393939,-9)(758.905723905724,-8.90909090909091)(761.393939393939,-8.81818181818182)(763.882154882155,-8.72727272727273)(766.992424242424,-8.54545454545455)(773.835016835017,-8.36363636363636)(776.323232323232,-8.31818181818182)(781.921717171717,-8.18181818181818)(791.252525252525,-8.06818181818182)(798.717171717172,-8)(806.181818181818,-7.93181818181818)(818.622895622896,-7.81818181818182)(821.111111111111,-7.79545454545455)(836.040404040404,-7.72727272727273)(848.481481481482,-7.63636363636364)(850.969696969697,-7.60606060606061)(863.410774410775,-7.45454545454546)(865.89898989899,-7.43181818181818)(880.828282828283,-7.38636363636364)(895.757575757576,-7.36363636363636)(910.686868686869,-7.29090909090909)(918.151515151515,-7.27272727272727)(925.616161616162,-7.24242424242424)(940.545454545455,-7.20454545454546)(955.474747474747,-7.24242424242424)(970.40404040404,-7.24242424242424)(985.333333333333,-7.24242424242424)(1000.26262626263,-7.18181818181818)(1015.19191919192,-7.24242424242424)(1030.12121212121,-7.24242424242424)(1037.58585858586,-7.27272727272727)(1045.05050505051,-7.28787878787879)(1048.78282828283,-7.27272727272727)(1059.9797979798,-7.20454545454546)(1074.90909090909,-7.15909090909091)(1089.83838383838,-7.18181818181818)(1102.27946127946,-7.27272727272727)(1104.76767676768,-7.28787878787879)(1119.69696969697,-7.36363636363636)(1134.62626262626,-7.38636363636364)(1149.55555555556,-7.36363636363636)(1164.48484848485,-7.36363636363636)(1171.94949494949,-7.45454545454546)(1179.41414141414,-7.5)(1194.34343434343,-7.5)(1209.27272727273,-7.54545454545455)(1224.20202020202,-7.47727272727273)(1239.13131313131,-7.54545454545455)(1254.06060606061,-7.5)(1268.9898989899,-7.60606060606061)(1272.72222222222,-7.63636363636364)(1283.91919191919,-7.70454545454546)(1298.84848484848,-7.72727272727273)(1313.77777777778,-7.75)(1328.70707070707,-7.75)(1343.63636363636,-7.79545454545455)(1346.12457912458,-7.81818181818182)(1352.96717171717,-8)(1358.56565656566,-8.06818181818182)(1373.49494949495,-8.11363636363636)(1388.42424242424,-8.15909090909091)(1403.35353535354,-8.15909090909091)(1405.84175084175,-8.18181818181818)(1415.79461279461,-8.36363636363636)(1418.28282828283,-8.38181818181818)(1433.21212121212,-8.43181818181818)(1442.54292929293,-8.54545454545455)(1448.14141414141,-8.68181818181818)(1455.60606060606,-8.72727272727273)(1463.07070707071,-8.75757575757576)(1478,-8.81818181818182)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1433.21212121212,-9.20454545454546)(1440.67676767677,-9.27272727272727)(1435.07828282828,-9.45454545454546)(1433.21212121212,-9.47727272727273)(1418.28282828283,-9.61363636363636)(1415.79461279461,-9.63636363636364)(1403.35353535354,-9.72727272727273)(1388.42424242424,-9.70454545454546)(1373.49494949495,-9.78787878787879)(1366.0303030303,-9.81818181818182)(1358.56565656566,-9.83333333333333)(1343.63636363636,-9.94545454545455)(1332.43939393939,-10)(1328.70707070707,-10.0151515151515)(1313.77777777778,-10.0454545454545)(1298.84848484848,-10.0151515151515)(1291.38383838384,-10)(1283.91919191919,-9.96969696969697)(1268.9898989899,-9.90909090909091)(1254.06060606061,-9.88636363636364)(1239.13131313131,-9.87272727272727)(1224.20202020202,-9.87272727272727)(1209.27272727273,-9.88636363636364)(1194.34343434343,-9.90909090909091)(1179.41414141414,-9.90909090909091)(1164.48484848485,-9.96969696969697)(1149.55555555556,-9.97727272727273)(1134.62626262626,-9.98181818181818)(1127.16161616162,-10)(1119.69696969697,-10.0151515151515)(1104.76767676768,-10.0757575757576)(1089.83838383838,-10.0757575757576)(1074.90909090909,-10.0151515151515)(1067.44444444444,-10)(1059.9797979798,-9.98181818181818)(1052.51515151515,-10)(1045.05050505051,-10.0151515151515)(1030.12121212121,-10.0151515151515)(1022.65656565657,-10)(1015.19191919192,-9.96969696969697)(1000.26262626263,-9.90909090909091)(985.333333333333,-9.88636363636364)(970.40404040404,-9.83636363636364)(955.474747474747,-9.83333333333333)(951.742424242424,-9.81818181818182)(940.545454545455,-9.76363636363636)(925.616161616162,-9.72727272727273)(910.686868686869,-9.71212121212121)(895.757575757576,-9.68181818181818)(884.560606060606,-9.63636363636364)(880.828282828283,-9.61363636363636)(865.89898989899,-9.5)(850.969696969697,-9.5)(846.490909090909,-9.45454545454546)(836.040404040404,-9.29545454545455)(832.308080808081,-9.27272727272727)(821.111111111111,-9.18181818181818)(813.646464646465,-9.09090909090909)(806.181818181818,-9)(798.717171717172,-8.90909090909091)(806.181818181818,-8.77272727272727)(808.670033670034,-8.72727272727273)(806.181818181818,-8.63636363636364)(803.693602693603,-8.54545454545455)(806.181818181818,-8.45454545454545)(808.670033670034,-8.36363636363636)(815.512626262626,-8.18181818181818)(821.111111111111,-8.12727272727273)(836.040404040404,-8.05454545454545)(843.505050505051,-8)(850.969696969697,-7.93181818181818)(863.410774410775,-7.81818181818182)(865.89898989899,-7.8)(880.828282828283,-7.70454545454546)(895.757575757576,-7.65909090909091)(903.222222222222,-7.63636363636364)(910.686868686869,-7.61818181818182)(925.616161616162,-7.56818181818182)(940.545454545455,-7.52272727272727)(955.474747474748,-7.56818181818182)(970.40404040404,-7.56818181818182)(985.333333333333,-7.56818181818182)(1000.26262626263,-7.54545454545455)(1015.19191919192,-7.56818181818182)(1030.12121212121,-7.56818181818182)(1045.05050505051,-7.58181818181818)(1059.9797979798,-7.52272727272727)(1074.90909090909,-7.48484848484849)(1089.83838383838,-7.47727272727273)(1104.76767676768,-7.58181818181818)(1115.96464646465,-7.63636363636364)(1119.69696969697,-7.65909090909091)(1134.62626262626,-7.70454545454546)(1149.55555555556,-7.72727272727273)(1164.48484848485,-7.72727272727273)(1179.41414141414,-7.8)(1194.34343434343,-7.77272727272727)(1205.5404040404,-7.81818181818182)(1209.27272727273,-7.84090909090909)(1216.73737373737,-7.81818181818182)(1224.20202020202,-7.8030303030303)(1227.93434343434,-7.81818181818182)(1239.13131313131,-7.90909090909091)(1254.06060606061,-7.95454545454546)(1268.9898989899,-7.97727272727273)(1272.72222222222,-8)(1283.91919191919,-8.06818181818182)(1298.84848484848,-8.09090909090909)(1313.77777777778,-8.11363636363636)(1328.70707070707,-8.15909090909091)(1331.19528619529,-8.18181818181818)(1341.14814814815,-8.36363636363636)(1343.63636363636,-8.38636363636364)(1358.56565656566,-8.45454545454545)(1373.49494949495,-8.49090909090909)(1384.69191919192,-8.54545454545455)(1388.42424242424,-8.59090909090909)(1395.88888888889,-8.72727272727273)(1403.35353535354,-8.79545454545455)(1415.79461279461,-8.90909090909091)(1418.28282828283,-9)(1420.77104377104,-9.09090909090909)(1433.21212121212,-9.20454545454546)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-6.51515151515151)(7.46464646464646,-6.54545454545455)(14.9292929292929,-6.57575757575758)(29.8585858585859,-6.63636363636364)(44.7878787878788,-6.6969696969697)(52.2525252525253,-6.72727272727273)(59.7171717171717,-6.75757575757576)(74.6464646464647,-6.81818181818182)(89.5757575757576,-6.87878787878788)(97.040404040404,-6.90909090909091)(104.505050505051,-6.93939393939394)(119.434343434343,-7.06060606060606)(126.89898989899,-7.09090909090909)(134.363636363636,-7.12121212121212)(149.292929292929,-7.18181818181818)(164.222222222222,-7.24242424242424)(171.686868686869,-7.27272727272727)(179.151515151515,-7.36363636363636)(194.080808080808,-7.40909090909091)(209.010101010101,-7.36363636363636)(223.939393939394,-7.36363636363636)(238.868686868687,-7.42424242424243)(246.333333333333,-7.45454545454546)(253.79797979798,-7.48484848484849)(268.727272727273,-7.59090909090909)(283.656565656566,-7.59090909090909)(287.388888888889,-7.63636363636364)(298.585858585859,-7.72727272727273)(313.515151515152,-7.78787878787879)(328.444444444444,-7.78787878787879)(343.373737373737,-7.66666666666667)(350.838383838384,-7.63636363636364)(358.30303030303,-7.60606060606061)(373.232323232323,-7.48484848484849)(376.964646464646,-7.45454545454546)(388.161616161616,-7.36363636363636)(395.626262626263,-7.27272727272727)(403.090909090909,-7.18181818181818)(414.287878787879,-7.09090909090909)(418.020202020202,-7.06060606060606)(432.949494949495,-6.93939393939394)(440.414141414141,-6.90909090909091)(447.878787878788,-6.87878787878788)(462.808080808081,-6.81818181818182)(477.737373737374,-6.81818181818182)(492.666666666667,-6.81818181818182)(507.59595959596,-6.79545454545455)(522.525252525253,-6.79545454545455)(537.454545454546,-6.84090909090909)(544.919191919192,-6.90909090909091)(552.383838383838,-6.97727272727273)(561.714646464646,-7.09090909090909)(567.313131313131,-7.15909090909091)(579.754208754209,-7.27272727272727)(582.242424242424,-7.29090909090909)(597.171717171717,-7.34090909090909)(609.612794612795,-7.45454545454546)(612.10101010101,-7.47727272727273)(625.164141414141,-7.63636363636364)(627.030303030303,-7.68181818181818)(632.628787878788,-7.81818181818182)(641.959595959596,-7.96969696969697)(644.447811447811,-8)(655.022727272727,-8.18181818181818)(655.644781144781,-8.36363636363636)(656.888888888889,-8.45454545454545)(660.621212121212,-8.54545454545455)(660.621212121212,-8.72727272727273)(659.377104377104,-8.90909090909091)(656.888888888889,-8.95454545454546)(641.959595959596,-9.06060606060606)(634.494949494949,-9.09090909090909)(627.030303030303,-9.18181818181818)(612.10101010101,-9.22727272727273)(609.612794612795,-9.27272727272727)(597.171717171717,-9.42424242424243)(589.707070707071,-9.45454545454546)(582.242424242424,-9.48484848484848)(567.313131313131,-9.54545454545455)(552.383838383838,-9.54545454545455)(541.186868686869,-9.45454545454546)(537.454545454546,-9.42424242424243)(522.525252525253,-9.38636363636364)(507.59595959596,-9.32727272727273)(492.666666666667,-9.32727272727273)(481.469696969697,-9.27272727272727)(477.737373737374,-9.25)(462.808080808081,-9.20454545454546)(447.878787878788,-9.15909090909091)(440.414141414141,-9.09090909090909)(432.949494949495,-9)(425.484848484849,-8.90909090909091)(418.020202020202,-8.77272727272727)(410.555555555556,-8.72727272727273)(403.090909090909,-8.63636363636364)(395.626262626263,-8.54545454545455)(403.090909090909,-8.45454545454545)(406.823232323232,-8.36363636363636)(403.090909090909,-8.31818181818182)(388.161616161616,-8.31818181818182)(380.69696969697,-8.36363636363636)(373.232323232323,-8.40909090909091)(358.30303030303,-8.45454545454545)(343.373737373737,-8.45454545454545)(328.444444444444,-8.45454545454545)(313.515151515152,-8.45454545454545)(298.585858585859,-8.51515151515152)(291.121212121212,-8.54545454545455)(283.656565656566,-8.57575757575757)(268.727272727273,-8.70454545454545)(264.99494949495,-8.72727272727273)(253.79797979798,-8.79545454545455)(238.868686868687,-8.87878787878788)(231.40404040404,-8.90909090909091)(223.939393939394,-8.93939393939394)(209.010101010101,-9)(194.080808080808,-8.95454545454546)(186.616161616162,-8.90909090909091)(179.151515151515,-8.88636363636364)(164.222222222222,-8.88636363636364)(156.757575757576,-8.90909090909091)(149.292929292929,-8.95454545454546)(134.363636363636,-8.93939393939394)(126.89898989899,-8.90909090909091)(119.434343434343,-8.89090909090909)(104.505050505051,-8.79545454545455)(97.040404040404,-8.72727272727273)(89.5757575757576,-8.59090909090909)(85.8434343434344,-8.54545454545455)(74.6464646464647,-8.40909090909091)(70.9141414141414,-8.36363636363636)(59.7171717171717,-8.22727272727273)(57.2289562289562,-8.18181818181818)(47.2760942760943,-8)(44.7878787878788,-7.95454545454546)(37.3232323232323,-7.81818181818182)(29.8585858585859,-7.72727272727273)(18.6616161616162,-7.63636363636364)(14.9292929292929,-7.60606060606061)(0,-7.54545454545455)(-0.000124406627521346,-7.45454545454546)(-0.000373213662548569,-7.27272727272727)(-0.000373213662548569,-7.09090909090909)(-0.000373213662548569,-6.90909090909091)(-0.00029027890567471,-6.72727272727273)(-4.14691062186359e-05,-6.54545454545455)(0,-6.51515151515151)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (462.808080808081,-7.84848484848485)(464.674242424242,-7.81818181818182)(472.138888888889,-7.63636363636364)(477.737373737374,-7.58181818181818)(492.666666666667,-7.53030303030303)(507.59595959596,-7.47272727272727)(511.328282828283,-7.45454545454546)(522.525252525253,-7.31818181818182)(526.257575757576,-7.27272727272727)(537.454545454546,-7.20454545454546)(543.05303030303,-7.27272727272727)(552.383838383838,-7.38636363636364)(559.848484848485,-7.45454545454546)(567.313131313131,-7.52272727272727)(582.242424242424,-7.61818181818182)(585.974747474747,-7.63636363636364)(597.171717171717,-7.77272727272727)(598.664646464646,-7.81818181818182)(604.636363636364,-8)(608.368686868687,-8.18181818181818)(612.10101010101,-8.22727272727273)(617.699494949495,-8.36363636363636)(614.589225589226,-8.54545454545455)(612.10101010101,-8.57575757575757)(597.171717171717,-8.68181818181818)(593.439393939394,-8.72727272727273)(582.242424242424,-8.79545454545455)(567.313131313131,-8.87878787878788)(552.383838383838,-8.88636363636364)(548.651515151515,-8.90909090909091)(537.454545454546,-8.97727272727273)(522.525252525253,-9.02272727272727)(507.59595959596,-9)(492.666666666667,-8.96363636363636)(481.469696969697,-8.90909090909091)(477.737373737374,-8.86363636363636)(462.808080808081,-8.77272727272727)(460.319865319865,-8.72727272727273)(457.209595959596,-8.54545454545455)(450.367003367003,-8.36363636363636)(450.367003367003,-8.18181818181818)(450.367003367003,-8)(462.808080808081,-7.84848484848485)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (134.363636363636,-7.90909090909091)(149.292929292929,-7.93181818181818)(164.222222222222,-7.93181818181818)(179.151515151515,-7.98181818181818)(186.616161616162,-8)(194.080808080808,-8.03030303030303)(209.010101010101,-8.15151515151515)(211.498316498317,-8.18181818181818)(209.010101010101,-8.22727272727273)(197.813131313131,-8.36363636363636)(194.080808080808,-8.38636363636364)(179.151515151515,-8.45454545454545)(164.222222222222,-8.41818181818182)(149.292929292929,-8.47727272727273)(134.363636363636,-8.45454545454545)(123.166666666667,-8.36363636363636)(126.89898989899,-8.18181818181818)(126.89898989899,-8)(134.363636363636,-7.90909090909091)};

\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (731.535353535354,-4.36363636363636)(716.606060606061,-11.6363636363636)(627.030303030303,-18)(537.454545454546,-14.7272727272727)(447.878787878788,-7.63636363636364)(358.30303030303,-10.5454545454545)(0,-18)(0,-14.5454545454545)(0,-11.2727272727273)(0,-8)(0,-4)(0,0)(328.444444444444,-5.09090909090909)(388.161616161616,-2)(746.464646464646,0)(880.828282828283,-15.4545454545455)(925.616161616162,-8)(1045.05050505051,-18)(1104.76767676768,-11.0909090909091)(1104.76767676768,-5.45454545454546)(1119.69696969697,0)(1149.55555555556,-2.36363636363636)(1224.20202020202,-14.3636363636364)(1478,-18)(1478,-15.6363636363636)(1478,-12.9090909090909)(1478,-10.1818181818182)(1478,-7.63636363636364)(1478,-5.09090909090909)(1478,0)(1268.9898989899,-8.18181818181818)(1254.06060606061,-9.27272727272727)(1164.48484848485,-8.54545454545455)(1149.55555555556,-8.18181818181818)(1134.62626262626,-7.81818181818182)(1119.69696969697,-8.90909090909091)(1104.76767676768,-7.63636363636364)(1074.90909090909,-7.45454545454546)(1074.90909090909,-9.09090909090909)(1059.9797979798,-9.09090909090909)(1030.12121212121,-7.27272727272727)(1015.19191919192,-7.27272727272727)(970.40404040404,-9.27272727272727)(910.686868686869,-7.09090909090909)(880.828282828283,-9.09090909090909)(865.89898989899,-9.09090909090909)(865.89898989899,-7.09090909090909)(791.252525252525,-8.90909090909091)(791.252525252525,-7.27272727272727)(761.393939393939,-7.27272727272727)(746.464646464646,-8.72727272727273)(716.606060606061,-8.54545454545455)(686.747474747475,-7.63636363636364)(671.818181818182,-8.18181818181818)(656.888888888889,-8)(597.171717171717,-7.09090909090909)(537.454545454546,-8.72727272727273)(223.939393939394,-8.36363636363636)(194.080808080808,-7.09090909090909)(0,-9.45454545454546)(209.010101010101,-8.18181818181818)(268.727272727273,-8.72727272727273)(298.585858585859,-7.81818181818182)(313.515151515152,-7.81818181818182)(313.515151515152,-8.90909090909091)(373.232323232323,-9.09090909090909)(388.161616161616,-9.09090909090909)(418.020202020202,-7.63636363636364)(492.666666666667,-9.27272727272727)(522.525252525253,-9.27272727272727)(522.525252525253,-7.81818181818182)(567.313131313131,-8)(567.313131313131,-9.09090909090909)(836.040404040404,-8.36363636363636)(836.040404040404,-9.63636363636364)(850.969696969697,-9.81818181818182)(865.89898989899,-8)(940.545454545455,-10.1818181818182)(955.474747474747,-10.1818181818182)(1074.90909090909,-10.3636363636364)(1209.27272727273,-7.63636363636364)(1224.20202020202,-10.1818181818182)(1239.13131313131,-10.1818181818182)(1283.91919191919,-7.81818181818182)(1313.77777777778,-10)(1358.56565656566,-9.81818181818182)(1358.56565656566,-8.18181818181818)(1403.35353535354,-8.72727272727273)(1403.35353535354,-9.27272727272727)(1418.28282828283,-8.90909090909091)(1433.21212121212,-9.09090909090909)(1418.28282828283,-9.45454545454546)(1403.35353535354,-8.54545454545455)(1388.42424242424,-8.36363636363636)(1373.49494949495,-9.81818181818182)(1328.70707070707,-10)(1298.84848484848,-7.81818181818182)(1239.13131313131,-7.63636363636364)(1149.55555555556,-7.45454545454546)(1134.62626262626,-7.45454545454546)(1104.76767676768,-10.1818181818182)(1089.83838383838,-10.1818181818182)(1074.90909090909,-10.1818181818182)(925.616161616162,-7.45454545454546)(925.616161616162,-10)(880.828282828283,-9.81818181818182)(865.89898989899,-7.81818181818182)(821.111111111111,-9.27272727272727)(612.10101010101,-8.72727272727273)(597.171717171717,-8.90909090909091)(567.313131313131,-7.63636363636364)(477.737373737374,-7.63636363636364)(418.020202020202,-8.54545454545455)(388.161616161616,-8.36363636363636)(388.161616161616,-8.18181818181818)(268.727272727273,-8)(194.080808080808,-7.81818181818182)(164.222222222222,-8)(164.222222222222,-8.18181818181818)(149.292929292929,-8)(134.363636363636,-7.81818181818182)(119.434343434343,-8)(119.434343434343,-8.18181818181818)(134.363636363636,-8.36363636363636)(164.222222222222,-8.54545454545455)(238.868686868687,-7.81818181818182)(253.79797979798,-8.36363636363636)(403.090909090909,-8)(403.090909090909,-8.36363636363636)(447.878787878788,-8.72727272727273)(492.666666666667,-7.45454545454546)(507.59595959596,-7.45454545454546)(507.59595959596,-9.09090909090909)(552.383838383838,-7.45454545454546)(597.171717171717,-9.09090909090909)(612.10101010101,-7.81818181818182)(627.030303030303,-8.90909090909091)(641.959595959596,-8.54545454545455)(806.181818181818,-8.90909090909091)(821.111111111111,-8.18181818181818)(850.969696969697,-7.81818181818182)(955.474747474747,-10)(955.474747474747,-7.27272727272727)(1045.05050505051,-7.27272727272727)(1119.69696969697,-10.1818181818182)(1134.62626262626,-10.1818181818182)(1224.20202020202,-7.63636363636364)(1328.70707070707,-8)(1433.21212121212,-8.90909090909091)(1433.21212121212,-9.27272727272727)(1418.28282828283,-8.72727272727273)(1388.42424242424,-9.81818181818182)(1298.84848484848,-10)(1194.34343434343,-7.63636363636364)(1104.76767676768,-7.45454545454546)(910.686868686869,-7.45454545454546)(895.757575757576,-9.81818181818182)(836.040404040404,-9.45454545454546)(806.181818181818,-9.09090909090909)(806.181818181818,-8.36363636363636)(627.030303030303,-8)(582.242424242424,-9.09090909090909)(552.383838383838,-7.27272727272727)(537.454545454546,-7.27272727272727)(507.59595959596,-7.27272727272727)(492.666666666667,-7.27272727272727)(492.666666666667,-9.09090909090909)(432.949494949495,-8.72727272727273)(432.949494949495,-7.63636363636364)(253.79797979798,-7.81818181818182)(223.939393939394,-8.54545454545455)(179.151515151515,-8.72727272727273)(164.222222222222,-7.63636363636364)(149.292929292929,-7.63636363636364)(134.363636363636,-7.63636363636364)(134.363636363636,-8.54545454545455)(119.434343434343,-8.54545454545455)(104.505050505051,-7.81818181818182)(89.5757575757576,-8) 
};

\addplot [
color=lime,
draw=black,
only marks,
mark=*,
mark options={solid},
mark size=3pt,
line width=0.2pt,
forget plot
]
coordinates{
 (89.5757575757576,-8.18181818181818)
};

\end{axis}
\end{tikzpicture}%