### 2013-03-22 figurefactory/files/limno_bgape_pp_30_0.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478.00044786535,-8.18181818181818)(1478.00089570383,-8.36363636363636)(1478.00134351545,-8.54545454545455)(1478.00134351545,-8.72727272727273)(1478.00134351545,-8.90909090909091)(1478.00134351545,-9.09090909090909)(1478.00134351545,-9.27272727272727)(1478.00089570383,-9.45454545454546)(1478.00044786535,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(-0.000447865351918085,-10.5454545454545)(-0.000895703833526601,-10.3636363636364)(-0.00134351544724879,-10.1818181818182)(-0.00134351544724879,-10)(-0.00134351544724879,-9.81818181818182)(-0.00134351544724879,-9.63636363636364)(-0.00134351544724879,-9.45454545454546)(-0.00134351544724879,-9.27272727272727)(-0.00134351544724879,-9.09090909090909)(-0.00134351544724879,-8.90909090909091)(-0.00134351544724879,-8.72727272727273)(-0.00134351544724879,-8.54545454545455)(-0.00134351544724879,-8.36363636363636)(-0.000895703833526601,-8.18181818181818)(-0.000447865351918085,-8)(-0.000447865351918085,-7.81818181818182)(-0.000895703833526601,-7.63636363636364)(-0.00134351544724879,-7.45454545454546)(-0.00134351544724879,-7.27272727272727)(-0.00134351544724879,-7.09090909090909)(-0.00134351544724879,-6.90909090909091)(-0.00134351544724879,-6.72727272727273)(-0.00134351544724879,-6.54545454545455)(-0.00134351544724879,-6.36363636363636)(-0.00134351544724879,-6.18181818181818)(-0.000895703833526601,-6)(-0.000447865351918085,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-5.90909090909091)(14.9292929292929,-5.90909090909091)(29.8585858585859,-5.96969696969697)(37.3232323232323,-6)(44.7878787878788,-6.03030303030303)(59.7171717171717,-6.09090909090909)(74.6464646464647,-6.09090909090909)(89.5757575757576,-6.09090909090909)(104.505050505051,-6.09090909090909)(119.434343434343,-6.09090909090909)(134.363636363636,-6.09090909090909)(149.292929292929,-6.09090909090909)(164.222222222222,-6.09090909090909)(179.151515151515,-6.09090909090909)(194.080808080808,-6.09090909090909)(209.010101010101,-6.15151515151515)(216.474747474747,-6.18181818181818)(223.939393939394,-6.21212121212121)(238.868686868687,-6.27272727272727)(253.79797979798,-6.27272727272727)(268.727272727273,-6.27272727272727)(283.656565656566,-6.27272727272727)(298.585858585859,-6.27272727272727)(313.515151515152,-6.27272727272727)(328.444444444444,-6.27272727272727)(343.373737373737,-6.33333333333333)(350.838383838384,-6.36363636363636)(358.30303030303,-6.39393939393939)(373.232323232323,-6.39393939393939)(380.69696969697,-6.36363636363636)(388.161616161616,-6.33333333333333)(403.090909090909,-6.27272727272727)(418.020202020202,-6.27272727272727)(432.949494949495,-6.27272727272727)(447.878787878788,-6.21212121212121)(455.343434343434,-6.18181818181818)(462.808080808081,-6.15151515151515)(477.737373737374,-6.09090909090909)(492.666666666667,-6.03030303030303)(500.131313131313,-6)(507.59595959596,-5.96969696969697)(522.525252525253,-5.90909090909091)(537.454545454546,-5.90909090909091)(552.383838383838,-5.90909090909091)(567.313131313131,-5.84848484848485)(574.777777777778,-5.81818181818182)(582.242424242424,-5.78787878787879)(597.171717171717,-5.72727272727273)(612.10101010101,-5.72727272727273)(627.030303030303,-5.72727272727273)(641.959595959596,-5.72727272727273)(656.888888888889,-5.72727272727273)(671.818181818182,-5.72727272727273)(686.747474747475,-5.72727272727273)(701.676767676768,-5.72727272727273)(716.606060606061,-5.72727272727273)(731.535353535354,-5.72727272727273)(746.464646464646,-5.72727272727273)(761.393939393939,-5.72727272727273)(776.323232323232,-5.72727272727273)(791.252525252525,-5.72727272727273)(806.181818181818,-5.72727272727273)(821.111111111111,-5.72727272727273)(836.040404040404,-5.72727272727273)(850.969696969697,-5.72727272727273)(865.89898989899,-5.72727272727273)(880.828282828283,-5.72727272727273)(895.757575757576,-5.72727272727273)(910.686868686869,-5.78787878787879)(918.151515151515,-5.81818181818182)(925.616161616162,-5.84848484848485)(940.545454545455,-5.90909090909091)(955.474747474747,-5.90909090909091)(970.40404040404,-5.90909090909091)(985.333333333333,-5.96969696969697)(992.79797979798,-6)(1000.26262626263,-6.03030303030303)(1015.19191919192,-6.09090909090909)(1030.12121212121,-6.09090909090909)(1045.05050505051,-6.09090909090909)(1059.9797979798,-6.15151515151515)(1067.44444444444,-6.18181818181818)(1074.90909090909,-6.21212121212121)(1089.83838383838,-6.27272727272727)(1104.76767676768,-6.27272727272727)(1119.69696969697,-6.27272727272727)(1134.62626262626,-6.27272727272727)(1149.55555555556,-6.27272727272727)(1164.48484848485,-6.27272727272727)(1179.41414141414,-6.27272727272727)(1194.34343434343,-6.27272727272727)(1209.27272727273,-6.27272727272727)(1224.20202020202,-6.27272727272727)(1239.13131313131,-6.27272727272727)(1254.06060606061,-6.27272727272727)(1268.9898989899,-6.27272727272727)(1283.91919191919,-6.27272727272727)(1298.84848484848,-6.33333333333333)(1306.31313131313,-6.36363636363636)(1313.77777777778,-6.39393939393939)(1328.70707070707,-6.51515151515151)(1332.43939393939,-6.54545454545455)(1343.63636363636,-6.63636363636364)(1351.10101010101,-6.72727272727273)(1358.56565656566,-6.81818181818182)(1366.0303030303,-6.90909090909091)(1373.49494949495,-7.04545454545455)(1375.98316498317,-7.09090909090909)(1385.93602693603,-7.27272727272727)(1388.42424242424,-7.36363636363636)(1390.91245791246,-7.45454545454546)(1400.86531986532,-7.63636363636364)(1403.35353535354,-7.68181818181818)(1410.81818181818,-7.81818181818182)(1418.28282828283,-7.90909090909091)(1425.74747474747,-8)(1433.21212121212,-8.09090909090909)(1444.40909090909,-8.18181818181818)(1448.14141414141,-8.21212121212121)(1463.07070707071,-8.27272727272727)(1478,-8.27272727272727)(1478.00022392596,-8.36363636363636)(1478.00067175772,-8.54545454545455)(1478.00067175772,-8.72727272727273)(1478.00067175772,-8.90909090909091)(1478.00067175772,-9.09090909090909)(1478.00067175772,-9.27272727272727)(1478.00022392596,-9.45454545454546)(1478,-9.54545454545455)(1463.07070707071,-9.60606060606061)(1455.60606060606,-9.63636363636364)(1448.14141414141,-9.66666666666667)(1433.21212121212,-9.78787878787879)(1429.4797979798,-9.81818181818182)(1418.28282828283,-9.90909090909091)(1407.08585858586,-10)(1403.35353535354,-10.030303030303)(1388.42424242424,-10.1515151515152)(1384.69191919192,-10.1818181818182)(1373.49494949495,-10.2727272727273)(1362.29797979798,-10.3636363636364)(1358.56565656566,-10.3939393939394)(1343.63636363636,-10.4545454545455)(1328.70707070707,-10.5151515151515)(1321.24242424242,-10.5454545454545)(1313.77777777778,-10.5757575757576)(1298.84848484848,-10.6363636363636)(1283.91919191919,-10.6363636363636)(1268.9898989899,-10.6363636363636)(1254.06060606061,-10.6363636363636)(1239.13131313131,-10.6363636363636)(1224.20202020202,-10.6363636363636)(1209.27272727273,-10.5757575757576)(1201.80808080808,-10.5454545454545)(1194.34343434343,-10.5151515151515)(1179.41414141414,-10.4545454545455)(1164.48484848485,-10.4545454545455)(1149.55555555556,-10.4545454545455)(1134.62626262626,-10.4545454545455)(1119.69696969697,-10.4545454545455)(1104.76767676768,-10.4545454545455)(1089.83838383838,-10.4545454545455)(1074.90909090909,-10.4545454545455)(1059.9797979798,-10.4545454545455)(1045.05050505051,-10.4545454545455)(1030.12121212121,-10.4545454545455)(1015.19191919192,-10.5151515151515)(1007.72727272727,-10.5454545454545)(1000.26262626263,-10.5757575757576)(985.333333333333,-10.6363636363636)(970.40404040404,-10.6363636363636)(955.474747474747,-10.6363636363636)(940.545454545455,-10.6363636363636)(925.616161616162,-10.6363636363636)(910.686868686869,-10.6363636363636)(895.757575757576,-10.6363636363636)(880.828282828283,-10.6363636363636)(865.89898989899,-10.6363636363636)(850.969696969697,-10.6363636363636)(836.040404040404,-10.6363636363636)(821.111111111111,-10.6363636363636)(806.181818181818,-10.6363636363636)(791.252525252525,-10.6363636363636)(776.323232323232,-10.5757575757576)(768.858585858586,-10.5454545454545)(761.393939393939,-10.5151515151515)(746.464646464646,-10.4545454545455)(731.535353535354,-10.4545454545455)(716.606060606061,-10.4545454545455)(701.676767676768,-10.4545454545455)(686.747474747475,-10.4545454545455)(671.818181818182,-10.4545454545455)(656.888888888889,-10.3939393939394)(649.424242424242,-10.3636363636364)(641.959595959596,-10.3333333333333)(627.030303030303,-10.2727272727273)(612.10101010101,-10.2727272727273)(597.171717171717,-10.2727272727273)(582.242424242424,-10.2121212121212)(574.777777777778,-10.1818181818182)(567.313131313131,-10.1515151515152)(552.383838383838,-10.0909090909091)(537.454545454546,-10.030303030303)(529.989898989899,-10)(522.525252525253,-9.96969696969697)(507.59595959596,-9.90909090909091)(492.666666666667,-9.84848484848485)(485.20202020202,-9.81818181818182)(477.737373737374,-9.78787878787879)(462.808080808081,-9.72727272727273)(447.878787878788,-9.72727272727273)(432.949494949495,-9.66666666666667)(425.484848484849,-9.63636363636364)(418.020202020202,-9.60606060606061)(403.090909090909,-9.54545454545455)(388.161616161616,-9.54545454545455)(373.232323232323,-9.54545454545455)(358.30303030303,-9.54545454545455)(343.373737373737,-9.54545454545455)(328.444444444444,-9.60606060606061)(320.979797979798,-9.63636363636364)(313.515151515152,-9.66666666666667)(298.585858585859,-9.72727272727273)(283.656565656566,-9.72727272727273)(268.727272727273,-9.78787878787879)(261.262626262626,-9.81818181818182)(253.79797979798,-9.84848484848485)(238.868686868687,-9.90909090909091)(223.939393939394,-9.96969696969697)(216.474747474747,-10)(209.010101010101,-10.030303030303)(194.080808080808,-10.0909090909091)(179.151515151515,-10.1515151515152)(171.686868686869,-10.1818181818182)(164.222222222222,-10.2121212121212)(149.292929292929,-10.2727272727273)(134.363636363636,-10.2727272727273)(119.434343434343,-10.3333333333333)(111.969696969697,-10.3636363636364)(104.505050505051,-10.3939393939394)(89.5757575757576,-10.4545454545455)(74.6464646464647,-10.4545454545455)(59.7171717171717,-10.4545454545455)(44.7878787878788,-10.4545454545455)(29.8585858585859,-10.4545454545455)(14.9292929292929,-10.4545454545455)(0,-10.4545454545455)(-0.00022392595838165,-10.3636363636364)(-0.000671757723624396,-10.1818181818182)(-0.000671757723624396,-10)(-0.000671757723624396,-9.81818181818182)(-0.000671757723624396,-9.63636363636364)(-0.000671757723624396,-9.45454545454546)(-0.000671757723624396,-9.27272727272727)(-0.000671757723624396,-9.09090909090909)(-0.000671757723624396,-8.90909090909091)(-0.000671757723624396,-8.72727272727273)(-0.000671757723624396,-8.54545454545455)(-0.000671757723624396,-8.36363636363636)(-0.00022392595838165,-8.18181818181818)(0,-8.09090909090909)(11.1969696969697,-8)(11.1969696969697,-7.81818181818182)(0,-7.72727272727273)(-0.00022392595838165,-7.63636363636364)(-0.000671757723624396,-7.45454545454546)(-0.000671757723624396,-7.27272727272727)(-0.000671757723624396,-7.09090909090909)(-0.000671757723624396,-6.90909090909091)(-0.000671757723624396,-6.72727272727273)(-0.000671757723624396,-6.54545454545455)(-0.000671757723624396,-6.36363636363636)(-0.000671757723624396,-6.18181818181818)(-0.00022392595838165,-6)(0,-5.90909090909091)};

\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (731.535353535354,-4.36363636363636)(716.606060606061,-11.6363636363636)(627.030303030303,-18)(537.454545454546,-14.7272727272727)(447.878787878788,-7.63636363636364)(358.30303030303,-10.5454545454545)(0,-18)(0,-14.5454545454545)(0,-11.2727272727273)(0,-8)(0,-4)(0,0)(328.444444444444,-5.09090909090909)(388.161616161616,-2)(746.464646464646,0)(880.828282828283,-15.4545454545455)(925.616161616162,-8)(1045.05050505051,-18)(1104.76767676768,-11.0909090909091)(1104.76767676768,-5.45454545454546)(1119.69696969697,0)(1149.55555555556,-2.36363636363636)(1224.20202020202,-14.3636363636364)(1478,-18)(1478,-15.6363636363636)(1478,-12.9090909090909)(1478,-10.1818181818182)(1478,-7.63636363636364)(1478,-5.09090909090909)(1478,0) 
};

\addplot [
color=lime,
draw=black,
only marks,
mark=*,
mark options={solid},
mark size=3pt,
line width=0.2pt,
forget plot
]
coordinates{
 (1478,0)
};

\end{axis}
\end{tikzpicture}%