### 2013-03-22 figurefactory/files/limno_bgape_pp_90_0.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{locol}{rgb}{0.26, 0.45, 0.65}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\tiny},
label style={font=\tiny},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\tiny},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
tickwidth=0.1cm,
clip=false
]

\addplot [fill=locol,draw=none,forget plot] coordinates{ (0,0)(14.9292929292929,0)(29.8585858585859,0)(44.7878787878788,0)(59.7171717171717,0)(74.6464646464647,0)(89.5757575757576,0)(104.505050505051,0)(119.434343434343,0)(134.363636363636,0)(149.292929292929,0)(164.222222222222,0)(179.151515151515,0)(194.080808080808,0)(209.010101010101,0)(223.939393939394,0)(238.868686868687,0)(253.79797979798,0)(268.727272727273,0)(283.656565656566,0)(298.585858585859,0)(313.515151515152,0)(328.444444444444,0)(343.373737373737,0)(358.30303030303,0)(373.232323232323,0)(388.161616161616,0)(403.090909090909,0)(418.020202020202,0)(432.949494949495,0)(447.878787878788,0)(462.808080808081,0)(477.737373737374,0)(492.666666666667,0)(507.59595959596,0)(522.525252525253,0)(537.454545454546,0)(552.383838383838,0)(567.313131313131,0)(582.242424242424,0)(597.171717171717,0)(612.10101010101,0)(627.030303030303,0)(641.959595959596,0)(656.888888888889,0)(671.818181818182,0)(686.747474747475,0)(701.676767676768,0)(716.606060606061,0)(731.535353535354,0)(746.464646464646,0)(761.393939393939,0)(776.323232323232,0)(791.252525252525,0)(806.181818181818,0)(821.111111111111,0)(836.040404040404,0)(850.969696969697,0)(865.89898989899,0)(880.828282828283,0)(895.757575757576,0)(910.686868686869,0)(925.616161616162,0)(940.545454545455,0)(955.474747474747,0)(970.40404040404,0)(985.333333333333,0)(1000.26262626263,0)(1015.19191919192,0)(1030.12121212121,0)(1045.05050505051,0)(1059.9797979798,0)(1074.90909090909,0)(1089.83838383838,0)(1104.76767676768,0)(1119.69696969697,0)(1134.62626262626,0)(1149.55555555556,0)(1164.48484848485,0)(1179.41414141414,0)(1194.34343434343,0)(1209.27272727273,0)(1224.20202020202,0)(1239.13131313131,0)(1254.06060606061,0)(1268.9898989899,0)(1283.91919191919,0)(1298.84848484848,0)(1313.77777777778,0)(1328.70707070707,0)(1343.63636363636,0)(1358.56565656566,0)(1373.49494949495,0)(1388.42424242424,0)(1403.35353535354,0)(1418.28282828283,0)(1433.21212121212,0)(1448.14141414141,0)(1463.07070707071,0)(1478,0)(1478,-0.181818181818183)(1478,-0.363636363636363)(1478,-0.545454545454547)(1478,-0.727272727272727)(1478,-0.90909090909091)(1478,-1.09090909090909)(1478,-1.27272727272727)(1478,-1.45454545454545)(1478,-1.63636363636364)(1478,-1.81818181818182)(1478,-2)(1478,-2.18181818181818)(1478,-2.36363636363636)(1478,-2.54545454545455)(1478,-2.72727272727273)(1478,-2.90909090909091)(1478,-3.09090909090909)(1478,-3.27272727272727)(1478,-3.45454545454546)(1478,-3.63636363636364)(1478,-3.81818181818182)(1478,-4)(1478,-4.18181818181818)(1478,-4.36363636363636)(1478,-4.54545454545455)(1478,-4.72727272727273)(1478,-4.90909090909091)(1478,-5.09090909090909)(1478,-5.27272727272727)(1478,-5.45454545454546)(1478,-5.63636363636364)(1478,-5.81818181818182)(1478,-6)(1478,-6.18181818181818)(1478,-6.36363636363636)(1478,-6.54545454545455)(1478,-6.72727272727273)(1478,-6.90909090909091)(1478,-7.09090909090909)(1478,-7.27272727272727)(1478,-7.45454545454546)(1478,-7.63636363636364)(1478,-7.81818181818182)(1478,-8)(1478,-8.18181818181818)(1478.00008294006,-8.36363636363636)(1478.00033175469,-8.54545454545455)(1478.00058056104,-8.72727272727273)(1478.00074642733,-8.90909090909091)(1478.00074642733,-9.09090909090909)(1478.00074642733,-9.27272727272727)(1478.00049762651,-9.45454545454546)(1478.0002488174,-9.63636363636364)(1478,-9.81818181818182)(1478,-10)(1478,-10.1818181818182)(1478,-10.3636363636364)(1478,-10.5454545454545)(1478,-10.7272727272727)(1478,-10.9090909090909)(1478,-11.0909090909091)(1478,-11.2727272727273)(1478,-11.4545454545455)(1478,-11.6363636363636)(1478,-11.8181818181818)(1478,-12)(1478,-12.1818181818182)(1478,-12.3636363636364)(1478,-12.5454545454545)(1478,-12.7272727272727)(1478,-12.9090909090909)(1478,-13.0909090909091)(1478,-13.2727272727273)(1478,-13.4545454545455)(1478,-13.6363636363636)(1478,-13.8181818181818)(1478,-14)(1478,-14.1818181818182)(1478,-14.3636363636364)(1478,-14.5454545454545)(1478,-14.7272727272727)(1478,-14.9090909090909)(1478,-15.0909090909091)(1478,-15.2727272727273)(1478,-15.4545454545455)(1478,-15.6363636363636)(1478,-15.8181818181818)(1478,-16)(1478,-16.1818181818182)(1478,-16.3636363636364)(1478,-16.5454545454545)(1478,-16.7272727272727)(1478,-16.9090909090909)(1478,-17.0909090909091)(1478,-17.2727272727273)(1478,-17.4545454545455)(1478,-17.6363636363636)(1478,-17.8181818181818)(1478,-18)(1463.07070707071,-18)(1448.14141414141,-18)(1433.21212121212,-18)(1418.28282828283,-18)(1403.35353535354,-18)(1388.42424242424,-18)(1373.49494949495,-18)(1358.56565656566,-18)(1343.63636363636,-18)(1328.70707070707,-18)(1313.77777777778,-18)(1298.84848484848,-18)(1283.91919191919,-18)(1268.9898989899,-18)(1254.06060606061,-18)(1239.13131313131,-18)(1224.20202020202,-18)(1209.27272727273,-18)(1194.34343434343,-18)(1179.41414141414,-18)(1164.48484848485,-18)(1149.55555555556,-18)(1134.62626262626,-18)(1119.69696969697,-18)(1104.76767676768,-18)(1089.83838383838,-18)(1074.90909090909,-18)(1059.9797979798,-18)(1045.05050505051,-18)(1030.12121212121,-18)(1015.19191919192,-18)(1000.26262626263,-18)(985.333333333333,-18)(970.40404040404,-18)(955.474747474747,-18)(940.545454545455,-18)(925.616161616162,-18)(910.686868686869,-18)(895.757575757576,-18)(880.828282828283,-18)(865.89898989899,-18)(850.969696969697,-18)(836.040404040404,-18)(821.111111111111,-18)(806.181818181818,-18)(791.252525252525,-18)(776.323232323232,-18)(761.393939393939,-18)(746.464646464646,-18)(731.535353535354,-18)(716.606060606061,-18)(701.676767676768,-18)(686.747474747475,-18)(671.818181818182,-18)(656.888888888889,-18)(641.959595959596,-18)(627.030303030303,-18)(612.10101010101,-18)(597.171717171717,-18)(582.242424242424,-18)(567.313131313131,-18)(552.383838383838,-18)(537.454545454546,-18)(522.525252525253,-18)(507.59595959596,-18)(492.666666666667,-18)(477.737373737374,-18)(462.808080808081,-18)(447.878787878788,-18)(432.949494949495,-18)(418.020202020202,-18)(403.090909090909,-18)(388.161616161616,-18)(373.232323232323,-18)(358.30303030303,-18)(343.373737373737,-18)(328.444444444444,-18)(313.515151515152,-18)(298.585858585859,-18)(283.656565656566,-18)(268.727272727273,-18)(253.79797979798,-18)(238.868686868687,-18)(223.939393939394,-18)(209.010101010101,-18)(194.080808080808,-18)(179.151515151515,-18)(164.222222222222,-18)(149.292929292929,-18)(134.363636363636,-18)(119.434343434343,-18)(104.505050505051,-18)(89.5757575757576,-18)(74.6464646464647,-18)(59.7171717171717,-18)(44.7878787878788,-18)(29.8585858585859,-18)(14.9292929292929,-18)(0,-18)(0,-17.8181818181818)(0,-17.6363636363636)(0,-17.4545454545455)(0,-17.2727272727273)(0,-17.0909090909091)(0,-16.9090909090909)(0,-16.7272727272727)(0,-16.5454545454545)(0,-16.3636363636364)(0,-16.1818181818182)(0,-16)(0,-15.8181818181818)(0,-15.6363636363636)(0,-15.4545454545455)(0,-15.2727272727273)(0,-15.0909090909091)(0,-14.9090909090909)(0,-14.7272727272727)(0,-14.5454545454545)(0,-14.3636363636364)(0,-14.1818181818182)(0,-14)(0,-13.8181818181818)(0,-13.6363636363636)(0,-13.4545454545455)(0,-13.2727272727273)(0,-13.0909090909091)(0,-12.9090909090909)(0,-12.7272727272727)(0,-12.5454545454545)(0,-12.3636363636364)(0,-12.1818181818182)(0,-12)(0,-11.8181818181818)(0,-11.6363636363636)(0,-11.4545454545455)(0,-11.2727272727273)(0,-11.0909090909091)(0,-10.9090909090909)(0,-10.7272727272727)(0,-10.5454545454545)(0,-10.3636363636364)(0,-10.1818181818182)(0,-10)(0,-9.81818181818182)(0,-9.63636363636364)(0,-9.45454545454546)(0,-9.27272727272727)(0,-9.09090909090909)(0,-8.90909090909091)(0,-8.72727272727273)(0,-8.54545454545455)(0,-8.36363636363636)(0,-8.18181818181818)(0,-8)(-0.000248817401863321,-7.81818181818182)(-0.000497626510092015,-7.63636363636364)(-0.000746427325097138,-7.45454545454546)(-0.000746427325097138,-7.27272727272727)(-0.000746427325097138,-7.09090909090909)(-0.000746427325097138,-6.90909090909091)(-0.000746427325097138,-6.72727272727273)(-0.00066349464153746,-6.54545454545455)(-0.0004146910621731,-6.36363636363636)(-0.000165879189445939,-6.18181818181818)(0,-6)(0,-5.81818181818182)(0,-5.63636363636364)(0,-5.45454545454546)(0,-5.27272727272727)(0,-5.09090909090909)(0,-4.90909090909091)(0,-4.72727272727273)(0,-4.54545454545455)(0,-4.36363636363636)(0,-4.18181818181818)(0,-4)(0,-3.81818181818182)(0,-3.63636363636364)(0,-3.45454545454546)(0,-3.27272727272727)(0,-3.09090909090909)(0,-2.90909090909091)(0,-2.72727272727273)(0,-2.54545454545455)(0,-2.36363636363636)(0,-2.18181818181818)(0,-2)(0,-1.81818181818182)(0,-1.63636363636364)(0,-1.45454545454545)(0,-1.27272727272727)(0,-1.09090909090909)(0,-0.90909090909091)(0,-0.727272727272727)(0,-0.545454545454547)(0,-0.363636363636363)(0,-0.181818181818183)(0,0)};

\addplot [fill=darkgray,draw=none,forget plot] coordinates{ (0,-6.33333333333333)(7.46464646464646,-6.36363636363636)(14.9292929292929,-6.39393939393939)(29.8585858585859,-6.45454545454545)(44.7878787878788,-6.51515151515151)(52.2525252525253,-6.54545454545455)(59.7171717171717,-6.57575757575758)(74.6464646464647,-6.63636363636364)(89.5757575757576,-6.6969696969697)(93.3080808080808,-6.72727272727273)(104.505050505051,-6.81818181818182)(115.70202020202,-6.90909090909091)(119.434343434343,-6.93939393939394)(134.363636363636,-7)(149.292929292929,-7.06060606060606)(156.757575757576,-7.09090909090909)(164.222222222222,-7.12121212121212)(179.151515151515,-7.18181818181818)(194.080808080808,-7.18181818181818)(209.010101010101,-7.18181818181818)(223.939393939394,-7.18181818181818)(238.868686868687,-7.18181818181818)(253.79797979798,-7.18181818181818)(268.727272727273,-7.18181818181818)(283.656565656566,-7.18181818181818)(298.585858585859,-7.18181818181818)(313.515151515152,-7.18181818181818)(328.444444444444,-7.18181818181818)(343.373737373737,-7.18181818181818)(358.30303030303,-7.12121212121212)(365.767676767677,-7.09090909090909)(373.232323232323,-7.06060606060606)(388.161616161616,-7)(403.090909090909,-7)(418.020202020202,-6.93939393939394)(425.484848484849,-6.90909090909091)(432.949494949495,-6.87878787878788)(447.878787878788,-6.81818181818182)(462.808080808081,-6.81818181818182)(477.737373737374,-6.81818181818182)(492.666666666667,-6.81818181818182)(507.59595959596,-6.81818181818182)(522.525252525253,-6.81818181818182)(537.454545454546,-6.87878787878788)(541.186868686869,-6.90909090909091)(552.383838383838,-7)(559.848484848485,-7.09090909090909)(567.313131313131,-7.18181818181818)(578.510101010101,-7.27272727272727)(582.242424242424,-7.3030303030303)(597.171717171717,-7.42424242424243)(600.90404040404,-7.45454545454546)(612.10101010101,-7.54545454545455)(619.565656565657,-7.63636363636364)(627.030303030303,-7.72727272727273)(634.494949494949,-7.81818181818182)(641.959595959596,-7.95454545454546)(644.447811447811,-8)(654.400673400673,-8.18181818181818)(656.888888888889,-8.22727272727273)(664.353535353535,-8.36363636363636)(671.818181818182,-8.45454545454545)(679.282828282828,-8.54545454545455)(686.747474747475,-8.63636363636364)(694.212121212121,-8.72727272727273)(701.676767676768,-8.81818181818182)(712.873737373737,-8.90909090909091)(716.606060606061,-8.93939393939394)(731.535353535354,-9)(746.464646464646,-8.93939393939394)(750.19696969697,-8.90909090909091)(761.393939393939,-8.77272727272727)(765.126262626263,-8.72727272727273)(768.858585858586,-8.54545454545455)(773.835016835017,-8.36363636363636)(776.323232323232,-8.31818181818182)(783.787878787879,-8.18181818181818)(791.252525252525,-8.09090909090909)(798.717171717172,-8)(806.181818181818,-7.90909090909091)(817.378787878788,-7.81818181818182)(821.111111111111,-7.78787878787879)(836.040404040404,-7.66666666666667)(843.505050505051,-7.63636363636364)(850.969696969697,-7.60606060606061)(865.89898989899,-7.48484848484849)(873.363636363636,-7.45454545454546)(880.828282828283,-7.42424242424243)(895.757575757576,-7.3030303030303)(903.222222222222,-7.27272727272727)(910.686868686869,-7.24242424242424)(925.616161616162,-7.18181818181818)(940.545454545455,-7.12121212121212)(948.010101010101,-7.09090909090909)(955.474747474747,-7.06060606060606)(970.40404040404,-7)(985.333333333333,-7)(1000.26262626263,-6.97727272727273)(1015.19191919192,-7)(1030.12121212121,-6.93939393939394)(1037.58585858586,-6.90909090909091)(1045.05050505051,-6.87878787878788)(1059.9797979798,-6.81818181818182)(1074.90909090909,-6.81818181818182)(1089.83838383838,-6.81818181818182)(1104.76767676768,-6.81818181818182)(1119.69696969697,-6.81818181818182)(1134.62626262626,-6.81818181818182)(1149.55555555556,-6.81818181818182)(1164.48484848485,-6.81818181818182)(1179.41414141414,-6.81818181818182)(1194.34343434343,-6.81818181818182)(1209.27272727273,-6.81818181818182)(1224.20202020202,-6.81818181818182)(1239.13131313131,-6.81818181818182)(1254.06060606061,-6.87878787878788)(1261.52525252525,-6.90909090909091)(1268.9898989899,-6.93939393939394)(1283.91919191919,-7.06060606060606)(1291.38383838384,-7.09090909090909)(1298.84848484848,-7.12121212121212)(1313.77777777778,-7.24242424242424)(1317.5101010101,-7.27272727272727)(1328.70707070707,-7.36363636363636)(1339.90404040404,-7.45454545454546)(1343.63636363636,-7.48484848484849)(1358.56565656566,-7.60606060606061)(1362.29797979798,-7.63636363636364)(1373.49494949495,-7.72727272727273)(1380.9595959596,-7.81818181818182)(1388.42424242424,-7.90909090909091)(1399.62121212121,-8)(1403.35353535354,-8.03030303030303)(1418.28282828283,-8.15151515151515)(1422.01515151515,-8.18181818181818)(1433.21212121212,-8.27272727272727)(1444.40909090909,-8.36363636363636)(1448.14141414141,-8.39393939393939)(1463.07070707071,-8.51515151515152)(1470.53535353535,-8.54545454545455)(1478,-8.57575757575757)(1478.00020734323,-8.72727272727273)(1478.00037321366,-8.90909090909091)(1478.00037321366,-9.09090909090909)(1478.00037321366,-9.27272727272727)(1478.00012440663,-9.45454545454546)(1478,-9.54545454545455)(1463.07070707071,-9.60606060606061)(1455.60606060606,-9.63636363636364)(1448.14141414141,-9.66666666666667)(1433.21212121212,-9.78787878787879)(1429.4797979798,-9.81818181818182)(1418.28282828283,-9.90909090909091)(1407.08585858586,-10)(1403.35353535354,-10.030303030303)(1388.42424242424,-10.0909090909091)(1373.49494949495,-10.1590909090909)(1366.0303030303,-10.1818181818182)(1358.56565656566,-10.2121212121212)(1343.63636363636,-10.2727272727273)(1328.70707070707,-10.2954545454545)(1313.77777777778,-10.2954545454545)(1298.84848484848,-10.2954545454545)(1283.91919191919,-10.2727272727273)(1268.9898989899,-10.2727272727273)(1254.06060606061,-10.2121212121212)(1246.59595959596,-10.1818181818182)(1239.13131313131,-10.1666666666667)(1224.20202020202,-10.1666666666667)(1216.73737373737,-10.1818181818182)(1209.27272727273,-10.2121212121212)(1194.34343434343,-10.2727272727273)(1179.41414141414,-10.2727272727273)(1164.48484848485,-10.3333333333333)(1157.0202020202,-10.3636363636364)(1149.55555555556,-10.3939393939394)(1134.62626262626,-10.4545454545455)(1119.69696969697,-10.4545454545455)(1104.76767676768,-10.4545454545455)(1089.83838383838,-10.3939393939394)(1074.90909090909,-10.3939393939394)(1059.9797979798,-10.3939393939394)(1045.05050505051,-10.4545454545455)(1030.12121212121,-10.3939393939394)(1022.65656565657,-10.3636363636364)(1015.19191919192,-10.3333333333333)(1000.26262626263,-10.2727272727273)(985.333333333333,-10.2727272727273)(970.40404040404,-10.2121212121212)(962.939393939394,-10.1818181818182)(955.474747474747,-10.1666666666667)(940.545454545455,-10.1636363636364)(925.616161616162,-10.1590909090909)(910.686868686869,-10.1515151515152)(895.757575757576,-10.0909090909091)(880.828282828283,-10.0909090909091)(869.631313131313,-10)(865.89898989899,-9.96969696969697)(850.969696969697,-9.93181818181818)(836.040404040404,-9.88636363636364)(821.111111111111,-9.88636363636364)(809.914141414141,-9.81818181818182)(806.181818181818,-9.78787878787879)(791.252525252525,-9.66666666666667)(783.787878787879,-9.63636363636364)(776.323232323232,-9.60606060606061)(761.393939393939,-9.54545454545455)(746.464646464646,-9.48484848484848)(739,-9.45454545454546)(731.535353535354,-9.42424242424243)(716.606060606061,-9.42424242424243)(709.141414141414,-9.45454545454546)(701.676767676768,-9.48484848484848)(686.747474747475,-9.54545454545455)(671.818181818182,-9.60606060606061)(664.353535353535,-9.63636363636364)(656.888888888889,-9.66666666666667)(641.959595959596,-9.72727272727273)(627.030303030303,-9.72727272727273)(612.10101010101,-9.72727272727273)(597.171717171717,-9.66666666666667)(589.707070707071,-9.63636363636364)(582.242424242424,-9.60606060606061)(567.313131313131,-9.54545454545455)(552.383838383838,-9.54545454545455)(541.186868686869,-9.45454545454546)(537.454545454546,-9.42424242424243)(522.525252525253,-9.36363636363636)(511.328282828283,-9.27272727272727)(507.59595959596,-9.24242424242424)(492.666666666667,-9.24242424242424)(477.737373737374,-9.18181818181818)(462.808080808081,-9.18181818181818)(447.878787878788,-9.12121212121212)(440.414141414141,-9.09090909090909)(432.949494949495,-9.06060606060606)(418.020202020202,-9)(403.090909090909,-8.93939393939394)(395.626262626263,-8.90909090909091)(388.161616161616,-8.87878787878788)(373.232323232323,-8.81818181818182)(358.30303030303,-8.81818181818182)(343.373737373737,-8.81818181818182)(328.444444444444,-8.75757575757576)(320.979797979798,-8.72727272727273)(313.515151515152,-8.6969696969697)(298.585858585859,-8.6969696969697)(283.656565656566,-8.6969696969697)(276.191919191919,-8.72727272727273)(268.727272727273,-8.75757575757576)(253.79797979798,-8.75757575757576)(238.868686868687,-8.87878787878788)(231.40404040404,-8.90909090909091)(223.939393939394,-8.93939393939394)(209.010101010101,-9)(194.080808080808,-9)(179.151515151515,-9)(164.222222222222,-9)(149.292929292929,-9)(134.363636363636,-9)(119.434343434343,-9)(104.505050505051,-8.93939393939394)(100.772727272727,-8.90909090909091)(89.5757575757576,-8.81818181818182)(82.1111111111111,-8.72727272727273)(74.6464646464647,-8.63636363636364)(67.1818181818182,-8.54545454545455)(59.7171717171717,-8.40909090909091)(57.2289562289562,-8.36363636363636)(47.2760942760943,-8.18181818181818)(44.7878787878788,-8.09090909090909)(42.2996632996633,-8)(29.8585858585859,-7.84848484848485)(26.1262626262626,-7.81818181818182)(14.9292929292929,-7.72727272727273)(0,-7.72727272727273)(-0.000124406627521346,-7.63636363636364)(-0.000373213662548569,-7.45454545454546)(-0.000373213662548569,-7.27272727272727)(-0.000373213662548569,-7.09090909090909)(-0.000373213662548569,-6.90909090909091)(-0.000373213662548569,-6.72727272727273)(-0.00029027890567471,-6.54545454545455)(-4.14691062186359e-05,-6.36363636363636)(0,-6.33333333333333)};

\addplot [fill=red!40!yellow,draw=none,forget plot] coordinates{ (1373.49494949495,-8.68181818181818)(1377.22727272727,-8.72727272727273)(1384.69191919192,-8.90909090909091)(1384.69191919192,-9.09090909090909)(1377.22727272727,-9.27272727272727)(1373.49494949495,-9.31818181818182)(1366.0303030303,-9.45454545454546)(1358.56565656566,-9.59090909090909)(1354.83333333333,-9.63636363636364)(1343.63636363636,-9.72727272727273)(1328.70707070707,-9.77272727272727)(1313.77777777778,-9.77272727272727)(1306.31313131313,-9.81818181818182)(1298.84848484848,-9.84848484848485)(1283.91919191919,-9.84848484848485)(1268.9898989899,-9.90909090909091)(1254.06060606061,-9.88636363636364)(1239.13131313131,-9.87272727272727)(1224.20202020202,-9.87272727272727)(1209.27272727273,-9.88636363636364)(1194.34343434343,-9.90909090909091)(1179.41414141414,-9.90909090909091)(1164.48484848485,-9.90909090909091)(1149.55555555556,-9.90909090909091)(1134.62626262626,-9.90909090909091)(1119.69696969697,-9.90909090909091)(1104.76767676768,-9.90909090909091)(1089.83838383838,-9.90909090909091)(1074.90909090909,-9.90909090909091)(1059.9797979798,-9.90909090909091)(1045.05050505051,-9.90909090909091)(1030.12121212121,-9.90909090909091)(1015.19191919192,-9.90909090909091)(1000.26262626263,-9.90909090909091)(985.333333333333,-9.90909090909091)(970.40404040404,-9.88636363636364)(955.474747474747,-9.87272727272727)(940.545454545455,-9.83636363636364)(933.080808080808,-9.81818181818182)(925.616161616162,-9.78787878787879)(910.686868686869,-9.72727272727273)(895.757575757576,-9.66666666666667)(892.025252525252,-9.63636363636364)(880.828282828283,-9.54545454545455)(869.631313131313,-9.45454545454546)(865.89898989899,-9.42424242424243)(853.457912457912,-9.27272727272727)(850.969696969697,-9.22727272727273)(843.505050505051,-9.09090909090909)(843.505050505051,-8.90909090909091)(843.505050505051,-8.72727272727273)(843.505050505051,-8.54545454545455)(850.969696969697,-8.40909090909091)(854.70202020202,-8.36363636363636)(865.89898989899,-8.22727272727273)(868.387205387205,-8.18181818181818)(880.828282828283,-8.03030303030303)(888.292929292929,-8)(895.757575757576,-7.96969696969697)(910.686868686869,-7.84848484848485)(918.151515151515,-7.81818181818182)(925.616161616162,-7.78787878787879)(940.545454545455,-7.72727272727273)(955.474747474747,-7.66666666666667)(962.939393939394,-7.63636363636364)(970.40404040404,-7.60606060606061)(985.333333333333,-7.54545454545455)(1000.26262626263,-7.5)(1015.19191919192,-7.54545454545455)(1030.12121212121,-7.54545454545455)(1045.05050505051,-7.56818181818182)(1059.9797979798,-7.48484848484849)(1074.90909090909,-7.48484848484849)(1089.83838383838,-7.48484848484849)(1104.76767676768,-7.60606060606061)(1112.23232323232,-7.63636363636364)(1119.69696969697,-7.66666666666667)(1134.62626262626,-7.72727272727273)(1149.55555555556,-7.72727272727273)(1164.48484848485,-7.72727272727273)(1179.41414141414,-7.72727272727273)(1194.34343434343,-7.72727272727273)(1209.27272727273,-7.78787878787879)(1216.73737373737,-7.81818181818182)(1224.20202020202,-7.86363636363636)(1239.13131313131,-7.90909090909091)(1254.06060606061,-7.90909090909091)(1268.9898989899,-7.90909090909091)(1283.91919191919,-7.96969696969697)(1287.65151515152,-8)(1298.84848484848,-8.13636363636364)(1306.31313131313,-8.18181818181818)(1313.77777777778,-8.21212121212121)(1328.70707070707,-8.33333333333333)(1332.43939393939,-8.36363636363636)(1343.63636363636,-8.43181818181818)(1358.56565656566,-8.52272727272727)(1362.29797979798,-8.54545454545455)(1373.49494949495,-8.68181818181818)};

\addplot [
color=white,
draw=white,
only marks,
mark=x,
mark options={solid},
mark size=2.0pt,
line width=0.3pt,
forget plot
]
coordinates{
 (731.535353535354,-4.36363636363636)(716.606060606061,-11.6363636363636)(627.030303030303,-18)(537.454545454546,-14.7272727272727)(447.878787878788,-7.63636363636364)(358.30303030303,-10.5454545454545)(0,-18)(0,-14.5454545454545)(0,-11.2727272727273)(0,-8)(0,-4)(0,0)(328.444444444444,-5.09090909090909)(388.161616161616,-2)(746.464646464646,0)(880.828282828283,-15.4545454545455)(925.616161616162,-8)(1045.05050505051,-18)(1104.76767676768,-11.0909090909091)(1104.76767676768,-5.45454545454546)(1119.69696969697,0)(1149.55555555556,-2.36363636363636)(1224.20202020202,-14.3636363636364)(1478,-18)(1478,-15.6363636363636)(1478,-12.9090909090909)(1478,-10.1818181818182)(1478,-7.63636363636364)(1478,-5.09090909090909)(1478,0)(1268.9898989899,-8.18181818181818)(1254.06060606061,-9.27272727272727)(1164.48484848485,-8.54545454545455)(1149.55555555556,-8.18181818181818)(1134.62626262626,-7.81818181818182)(1119.69696969697,-8.90909090909091)(1104.76767676768,-7.63636363636364)(1074.90909090909,-7.45454545454546)(1074.90909090909,-9.09090909090909)(1059.9797979798,-9.09090909090909)(1030.12121212121,-7.27272727272727)(1015.19191919192,-7.27272727272727)(970.40404040404,-9.27272727272727)(910.686868686869,-7.09090909090909)(880.828282828283,-9.09090909090909)(865.89898989899,-9.09090909090909)(865.89898989899,-7.09090909090909)(791.252525252525,-8.90909090909091)(791.252525252525,-7.27272727272727)(761.393939393939,-7.27272727272727)(746.464646464646,-8.72727272727273)(716.606060606061,-8.54545454545455)(686.747474747475,-7.63636363636364)(671.818181818182,-8.18181818181818)(656.888888888889,-8)(597.171717171717,-7.09090909090909)(537.454545454546,-8.72727272727273)(223.939393939394,-8.36363636363636)(194.080808080808,-7.09090909090909)(0,-9.45454545454546)(209.010101010101,-8.18181818181818)(268.727272727273,-8.72727272727273)(298.585858585859,-7.81818181818182)(313.515151515152,-7.81818181818182)(313.515151515152,-8.90909090909091)(373.232323232323,-9.09090909090909)(388.161616161616,-9.09090909090909)(418.020202020202,-7.63636363636364)(492.666666666667,-9.27272727272727)(522.525252525253,-9.27272727272727)(522.525252525253,-7.81818181818182)(567.313131313131,-8)(567.313131313131,-9.09090909090909)(836.040404040404,-8.36363636363636)(836.040404040404,-9.63636363636364)(850.969696969697,-9.81818181818182)(865.89898989899,-8)(940.545454545455,-10.1818181818182)(955.474747474747,-10.1818181818182)(1074.90909090909,-10.3636363636364)(1209.27272727273,-7.63636363636364)(1224.20202020202,-10.1818181818182)(1239.13131313131,-10.1818181818182)(1283.91919191919,-7.81818181818182)(1313.77777777778,-10)(1358.56565656566,-9.81818181818182)(1358.56565656566,-8.18181818181818)(1403.35353535354,-8.72727272727273)(1403.35353535354,-9.27272727272727) 
};

\addplot [
color=lime,
draw=black,
only marks,
mark=*,
mark options={solid},
mark size=3pt,
line width=0.2pt,
forget plot
]
coordinates{
 (1418.28282828283,-8.90909090909091)
};

\end{axis}
\end{tikzpicture}%