### 2013-05-01 figurefactory/files/ev_chl_eps.tex
Most of the changes discussed with Andreas

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

\usetikzlibrary{shapes,arrows,fit,calc,positioning}

% defining custom colors
\definecolor{mycolor1}{rgb}{1,0.04,0}
\definecolor{mycolor2}{rgb}{1,0.08,0}
\definecolor{mycolor3}{rgb}{1,0.12,0}
\definecolor{mycolor4}{rgb}{1,0.16,0}
\definecolor{mycolor5}{rgb}{1,0.2,0}
\definecolor{mycolor6}{rgb}{1,0.24,0}
\definecolor{mycolor7}{rgb}{1,0.28,0}
\definecolor{mycolor8}{rgb}{1,0.32,0}
\definecolor{mycolor9}{rgb}{1,0.36,0}
\definecolor{mycolor10}{rgb}{1,0.4,0}
\definecolor{mycolor11}{rgb}{1,0.44,0}
\definecolor{mycolor12}{rgb}{1,0.48,0}
\definecolor{mycolor13}{rgb}{1,0.52,0}
\definecolor{mycolor14}{rgb}{1,0.56,0}
\definecolor{mycolor15}{rgb}{1,0.6,0}
\definecolor{mycolor16}{rgb}{1,0.64,0}
\definecolor{mycolor17}{rgb}{1,0.68,0}
\definecolor{mycolor18}{rgb}{1,0.72,0}
\definecolor{mycolor19}{rgb}{1,0.76,0}
\definecolor{mycolor20}{rgb}{1,0.8,0}
\definecolor{mycolor21}{rgb}{1,0.84,0}
\definecolor{mycolor22}{rgb}{1,0.88,0}
\definecolor{mycolor23}{rgb}{1,0.92,0}
\definecolor{mycolor24}{rgb}{1,0.96,0}
\definecolor{mycolor25}{rgb}{1,1,0}
\definecolor{mycolor26}{rgb}{0.96,1,0.04}
\definecolor{mycolor27}{rgb}{0.92,1,0.08}
\definecolor{mycolor28}{rgb}{0.88,1,0.12}
\definecolor{mycolor29}{rgb}{0.84,1,0.16}
\definecolor{mycolor30}{rgb}{0.76,1,0.24}
\definecolor{mycolor31}{rgb}{0.72,1,0.28}
\definecolor{mycolor32}{rgb}{0.68,1,0.32}
\definecolor{mycolor33}{rgb}{0.64,1,0.36}
\definecolor{mycolor34}{rgb}{0.56,1,0.44}
\definecolor{mycolor35}{rgb}{0.52,1,0.48}
\definecolor{mycolor36}{rgb}{0.48,1,0.52}
\definecolor{mycolor37}{rgb}{0.4,1,0.6}
\definecolor{mycolor38}{rgb}{0.36,1,0.64}
\definecolor{mycolor39}{rgb}{0.28,1,0.72}
\definecolor{mycolor40}{rgb}{0.24,1,0.76}
\definecolor{mycolor41}{rgb}{0.16,1,0.84}
\definecolor{mycolor42}{rgb}{0.12,1,0.88}
\definecolor{mycolor43}{rgb}{0.04,1,0.96}
\definecolor{mycolor44}{rgb}{0,0.96,1}
\definecolor{mycolor45}{rgb}{0,0.92,1}
\definecolor{mycolor46}{rgb}{0,0.84,1}
\definecolor{mycolor47}{rgb}{0,0.76,1}
\definecolor{mycolor48}{rgb}{0,0.68,1}
\definecolor{mycolor49}{rgb}{0,0.6,1}
\definecolor{mycolor50}{rgb}{0,0.52,1}
\definecolor{mycolor51}{rgb}{0,0.44,1}
\definecolor{mycolor52}{rgb}{0,0.36,1}
\definecolor{mycolor53}{rgb}{0,0.28,1}
\definecolor{mycolor54}{rgb}{0,0.16,1}
\definecolor{mycolor55}{rgb}{0,0.08,1}
\definecolor{mycolor56}{rgb}{0.44,1,0.56}
\definecolor{mycolor57}{rgb}{0.32,1,0.68}
\definecolor{mycolor58}{rgb}{0.2,1,0.8}
\definecolor{mycolor59}{rgb}{0.08,1,0.92}
\definecolor{mycolor60}{rgb}{0,1,1}
\definecolor{mycolor61}{rgb}{0,0.88,1}
\definecolor{mycolor62}{rgb}{0,0.8,1}
\definecolor{mycolor63}{rgb}{0,0.72,1}
\definecolor{mycolor64}{rgb}{0,0.64,1}
\definecolor{mycolor65}{rgb}{0,0.56,1}
\definecolor{mycolor66}{rgb}{0,0.48,1}
\definecolor{mycolor67}{rgb}{0,0.4,1}
\definecolor{mycolor68}{rgb}{0,0.32,1}
\definecolor{mycolor69}{rgb}{0,0.24,1}
\definecolor{mycolor70}{rgb}{0,0.2,1}
\definecolor{mycolor71}{rgb}{0,0.12,1}
\definecolor{mycolor72}{rgb}{0,0.04,1}
\definecolor{mycolor73}{rgb}{0,0,0.56}
\definecolor{mycolor74}{rgb}{0.8,1,0.2}
\definecolor{mycolor75}{rgb}{0.6,1,0.4}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\normalsize},
label style={font=\normalsize},
label shift={-4pt},
xlabel shift={-6pt},
legend style={font=\normalsize},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=400,
xlabel={Samples},
ymin=0.5, ymax=1,
ylabel={$F_1$-score},
name=plot1,
axis lines*=left,
colormap={mymap}{[1pt] rgb(0pt)=(0.5,0,0); rgb(8pt)=(1,0,0); rgb(24pt)=(1,1,0); rgb(40pt)=(0,1,1); rgb(56pt)=(0,0,1); rgb(63pt)=(0,0,0.5625)},
colorbar right,
colorbar style={
  tickwidth=0.03cm,
  at={(1.06, 0)},
  anchor=south east,
  yticklabel pos=right,
  yticklabel shift={-2pt}
},
tickwidth=0.1cm,
colorbar/width=0.15cm,
point meta min=0,
point meta max=1
]

\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (383,0.920627376425855) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (365,0.954618015963512) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (357,0.919687277896233) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (360,0.927679623085983) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (317,0.932016536518144) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (328,0.924838940586972) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (276,0.933055362520269) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (270,0.913882352941176) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (259,0.932475884244373) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (266,0.92346471127406) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (243,0.926806526806527) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (236,0.949767441860465) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (224,0.914419695193435) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.926267281105991) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (213,0.930856067732832) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (202,0.890752387607733) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (185,0.933861201676758) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.937427308676436) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (171,0.922470433639947) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (154,0.904059877523248) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (160,0.916763273823325) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (141,0.884670147128619) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.922135968007528) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (124,0.886009224687019) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.868557919621749) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (114,0.913023884632717) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (108,0.878591288229842) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (108,0.899885974914481) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.878853046594982) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.854586129753915) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (87,0.820249449743213) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (88,0.897721989772199) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (87,0.915261659732866) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (85,0.888432580424367) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (78,0.862354892205638) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.874582312319002) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.806558935361217) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (71,0.793864745526377) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.89149560117302) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (64,0.843316889828517) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.84857973607694) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.827240143369175) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (60,0.813210379583959) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.877560309512972) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.851585623678647) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.83169673441425) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.704498269896194) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.809187279151943) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.689636313918512) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.801229508196721) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.715773809523809) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.809985458070771) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.84080445844439) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.747091552857865) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.856391875746714) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.796696814785686) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.855006796556411) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.811505507955936) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.8166627441751) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.782567947516401) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.774969915764139) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.816287425149701) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.772799636115533) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.765501165501165) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.800239234449761) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.734396996715157) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.649639423076923) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.827707919699978) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor42,draw=mycolor42,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.599538372763993) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.533975903614458) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.769065981148243) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.768549280177187) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.802575107296137) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.672905331882481) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.696956889264581) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.58615221987315) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.621519584709769) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor51,draw=mycolor51,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.777078651685393) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor52,draw=mycolor52,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.697664543524416) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.486176836861768) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor54,draw=mycolor54,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.684233449477352) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor55,draw=mycolor55,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.753533568904594) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue,draw=blue,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.527763627101375) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.721542608335061) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!80!black,draw=blue!80!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.702539602715615) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!68!black,draw=blue!68!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.659987856709168) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!60!black,draw=blue!60!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.665330661322645) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (8,0.503211393465512) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (6,0.329348245275742) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.816740960871718) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.780425313861132) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.752901353965184) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.790649884407912) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.578715056400196) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.605374077976818) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.739073540957864) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.746866065537717) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.665319525132609) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.66875) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor57,draw=mycolor57,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.840089086859688) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.5657469550087) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.772897815051865) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.606294536817102) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.803651194406681) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.751927437641723) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.87147766323024) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.581502576620559) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.545933014354067) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor42,draw=mycolor42,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.846911800048065) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.717948717948718) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.56657567485593) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.712) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.693888588426176) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor60,draw=mycolor60,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.692395005675369) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.809546092653252) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.694828469022017) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.513928368391131) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.727987764911514) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor61,draw=mycolor61,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.669603524229075) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.753276792598304) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.702082447938801) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor62,draw=mycolor62,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.632636437686668) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.684998772403634) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.355127322780454) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor63,draw=mycolor63,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.644655847789591) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.539637599093998) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.467532467532467) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor64,draw=mycolor64,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.324404761904762) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.680245291283399) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.700901393062005) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor65,draw=mycolor65,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.658042744656918) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.718541574032903) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.768842105263158) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor66,draw=mycolor66,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.618575851393189) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor51,draw=mycolor51,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.788557806912992) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.593739916101968) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.655133295519002) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor52,draw=mycolor52,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.589428723972237) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor68,draw=mycolor68,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.668627450980392) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.543239227340267) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.622902561083309) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor69,draw=mycolor69,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.544344286129689) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor70,draw=mycolor70,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.57837528604119) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor54,draw=mycolor54,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.574266574266574) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.630492323980942) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.715287517531557) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor55,draw=mycolor55,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.680589022757697) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor72,draw=mycolor72,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.641860465116279) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue,draw=blue,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.780365296803653) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!96!black,draw=blue!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.611721611721612) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!92!black,draw=blue!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.584743057772103) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.669121514992109) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.676288659793814) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!84!black,draw=blue!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.170266836086404) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!80!black,draw=blue!80!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.41851616994293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!76!black,draw=blue!76!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.692734478203434) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!72!black,draw=blue!72!black,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.638255698711596) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!68!black,draw=blue!68!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.297386823702613) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!64!black,draw=blue!64!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.437275985663082) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!60!black,draw=blue!60!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.534194698151036) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor73,draw=mycolor73,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.54445165476963) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (8,0.317921283912877) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.592728383745799) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.427279053583855) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.494597839135654) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.586343394359228) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.809259259259259) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.763672854992474) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.6855693868142) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.611425630468348) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.779730740463725) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.743899851474645) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.819678714859438) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.840132827324478) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.729564863765758) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.693003144654088) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor57,draw=mycolor57,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.803216439580076) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.392037786774629) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.697445512069369) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.652960118900173) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.78) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.615985576923077) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.774031202818319) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.756616257088847) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.83106449453247) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor42,draw=mycolor42,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.758706467661691) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.744228157537347) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.731303288672351) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.594059405940594) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.693709327548807) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor60,draw=mycolor60,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.643767313019391) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.586734693877551) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.580510992275698) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.703553299492386) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.64476821192053) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor61,draw=mycolor61,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.658665922411387) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.647314949201741) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.718332442544094) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor62,draw=mycolor62,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.808384417638916) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.681485649971863) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.604446616173955) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor63,draw=mycolor63,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.43978349120433) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.678554886521538) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.445234708392603) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor64,draw=mycolor64,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.521162672106068) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.793095616625028) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.706996770721206) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor65,draw=mycolor65,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.802838766910623) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.205088062622309) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.810042633822833) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor66,draw=mycolor66,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.72569688451628) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor51,draw=mycolor51,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.58434026099565) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.604332327940773) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.787039106145251) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor52,draw=mycolor52,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.519722975007528) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor68,draw=mycolor68,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.381634646694888) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.613044599123485) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.628063560463237) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor69,draw=mycolor69,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.724870017331022) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor70,draw=mycolor70,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.555399719495091) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor54,draw=mycolor54,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.663329820864067) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.618096357226792) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.333671742808799) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor55,draw=mycolor55,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.713962574542463) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor72,draw=mycolor72,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.61759029500965) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue,draw=blue,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.447385003150599) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!96!black,draw=blue!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.394671107410491) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!92!black,draw=blue!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.687603305785124) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.605599104143337) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.661727133367399) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!84!black,draw=blue!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.520259820600062) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!80!black,draw=blue!80!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.315201192250373) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!76!black,draw=blue!76!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.640275387263339) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!72!black,draw=blue!72!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.652136994056043) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!68!black,draw=blue!68!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.663863337713535) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!64!black,draw=blue!64!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.378052283826487) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!60!black,draw=blue!60!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.572074618428491) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor73,draw=mycolor73,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.523722627737226) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.688920454545454) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.229361225252148) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (5,0.274867122247532) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.462440659033789) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (8,0.509664292980671) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.796888141295206) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.763816925734024) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.818261633011413) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.73870636550308) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.839582344565733) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.809427609427609) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.779917469050894) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.841784037558685) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.753956834532374) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.735879259089869) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor57,draw=mycolor57,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.822724161533196) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.602597402597402) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.793499043977055) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.83207039843559) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.723041474654378) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.810580204778157) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.702300720427609) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.76307295029051) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.790393933010322) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor42,draw=mycolor42,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.622570194384449) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.762422047877691) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.642542787286063) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.692270992366412) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.715997915581032) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor60,draw=mycolor60,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.764863828154967) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.660366552119129) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.444444444444444) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.758252427184466) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.791684711996535) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor61,draw=mycolor61,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.708605664488017) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.293010752688172) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.604067985511284) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor62,draw=mycolor62,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.108061002178649) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.79496317415063) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.41096817874069) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor63,draw=mycolor63,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.37992590097676) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.599163179916318) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.642300456866434) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor64,draw=mycolor64,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.697699304440877) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.533501896333755) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.77087114337568) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor65,draw=mycolor65,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.597621119814331) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.674875556719937) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.439812096300646) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor66,draw=mycolor66,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.705913272010512) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor51,draw=mycolor51,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.602993197278911) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.695475113122172) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.642878207018579) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor52,draw=mycolor52,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.665593129361245) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor68,draw=mycolor68,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.814687567103285) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.667370644139387) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.613364595545135) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor69,draw=mycolor69,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.67652572380754) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor70,draw=mycolor70,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.448130081300813) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor54,draw=mycolor54,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.614322691975841) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.375272727272727) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.49875173370319) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor55,draw=mycolor55,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.432941176470588) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor72,draw=mycolor72,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.568326271186441) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue,draw=blue,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.638865004299226) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!96!black,draw=blue!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.628442663995994) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!92!black,draw=blue!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.281635802469136) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.566705336426914) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.286647314949202) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!84!black,draw=blue!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.177723177723178) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!80!black,draw=blue!80!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.497709923664122) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!76!black,draw=blue!76!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.66155950752394) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!72!black,draw=blue!72!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.348736462093863) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!68!black,draw=blue!68!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.680682661390057) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!64!black,draw=blue!64!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.637261990841166) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!60!black,draw=blue!60!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.56555590158829) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor73,draw=mycolor73,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.703451882845188) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (8,0.240480961923848) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.282677391554151) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (8,0.659445843828715) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (4,0) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (6,0) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.796187683284457) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.658215010141988) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.809181544168792) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor36,draw=mycolor36,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.834630350194552) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.77797434763379) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor56,draw=mycolor56,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.804736587723538) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.730873978707601) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor37,draw=mycolor37,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.838006952491309) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.827505827505827) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor38,draw=mycolor38,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.732490272373541) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor57,draw=mycolor57,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.668514176081859) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.83093447905478) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor39,draw=mycolor39,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.682313892840208) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.752908329455561) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor40,draw=mycolor40,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.680164018452076) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (26,0.766307541625857) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor58,draw=mycolor58,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.7337957426453) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.810680779408227) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor41,draw=mycolor41,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.665826612903226) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor42,draw=mycolor42,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.77572793954212) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.688967971530249) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor59,draw=mycolor59,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.81453634085213) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (25,0.611478444855858) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor43,draw=mycolor43,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.765874939408628) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor60,draw=mycolor60,line width=0.05pt},
forget plot
]
coordinates{
 (23,0.645833333333333) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.75971515768057) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor44,draw=mycolor44,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.700170035896467) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.712065136935603) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor45,draw=mycolor45,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.699408588326048) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor61,draw=mycolor61,line width=0.05pt},
forget plot
]
coordinates{
 (24,0.770192085165471) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.551943290149108) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor46,draw=mycolor46,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.763976506417228) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor62,draw=mycolor62,line width=0.05pt},
forget plot
]
coordinates{
 (22,0.806384976525822) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.804878048780488) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor47,draw=mycolor47,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.667990293403927) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor63,draw=mycolor63,line width=0.05pt},
forget plot
]
coordinates{
 (21,0.684081632653061) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.628804347826087) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor48,draw=mycolor48,line width=0.05pt},
forget plot
]
coordinates{
 (20,0.731782945736434) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor64,draw=mycolor64,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.678260869565217) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.689445048966267) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor49,draw=mycolor49,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.724023825281271) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor65,draw=mycolor65,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.720123361603701) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (18,0.676339285714285) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor50,draw=mycolor50,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.543382997370727) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor66,draw=mycolor66,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.559206798866855) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor51,draw=mycolor51,line width=0.05pt},
forget plot
]
coordinates{
 (19,0.688484251968504) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.689798750867453) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor67,draw=mycolor67,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.681792073520965) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor52,draw=mycolor52,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.791161178509532) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor68,draw=mycolor68,line width=0.05pt},
forget plot
]
coordinates{
 (16,0.659885237350026) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.34565560821485) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor53,draw=mycolor53,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.628371089536138) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor69,draw=mycolor69,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.603129445234708) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor70,draw=mycolor70,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.423952095808383) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor54,draw=mycolor54,line width=0.05pt},
forget plot
]
coordinates{
 (17,0.708926875593542) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (14,0.761444771104578) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor71,draw=mycolor71,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.58952849291293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor55,draw=mycolor55,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.670391061452514) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor72,draw=mycolor72,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.713216957605985) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue,draw=blue,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.415430267062314) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!96!black,draw=blue!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (15,0.531341534856473) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!92!black,draw=blue!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.390929415522197) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (13,0.657091561938959) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!88!black,draw=blue!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!84!black,draw=blue!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.0748663101604278) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!80!black,draw=blue!80!black,line width=0.05pt},
forget plot
]
coordinates{
 (12,0.323261740080087) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!76!black,draw=blue!76!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.242321970975363) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!72!black,draw=blue!72!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.45557963163597) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!68!black,draw=blue!68!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.700075169130544) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!64!black,draw=blue!64!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.620746887966805) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!60!black,draw=blue!60!black,line width=0.05pt},
forget plot
]
coordinates{
 (9,0.596015015882183) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor73,draw=mycolor73,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.225817555938038) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.608452280996362) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (10,0.555043227665706) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (11,0.773273579140354) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (5,0) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=blue!52!black,draw=blue!52!black,line width=0.05pt},
forget plot
]
coordinates{
 (3,0) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (394,0.952903520804755) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (400,0.950011412919425) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (376,0.964441385638908) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (342,0.952315134761575) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (363,0.953930781572313) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (360,0.951876583007138) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (321,0.939543636791343) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (329,0.931759149940968) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (307,0.931932171005493) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (323,0.955012559945193) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (288,0.938938468764678) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (276,0.946866485013624) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (278,0.948168624740843) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (271,0.933486766398159) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (266,0.941577631279836) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (260,0.955203619909502) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (242,0.942592592592592) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (257,0.9446104343829) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (232,0.945995423340961) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (237,0.92283882378952) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (224,0.954745880714783) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (218,0.941479400749063) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (193,0.926126976634411) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (200,0.929334242217678) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (192,0.908125577100646) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (189,0.906093189964158) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (180,0.919384902143523) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (164,0.892163429336905) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (168,0.913156036714521) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (167,0.923112338858195) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (149,0.88504753673293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (149,0.921514312096029) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (156,0.918466898954704) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.911363110391113) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.920341722465943) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.917223887479825) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.915866483767718) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (108,0.899374097255657) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (125,0.921731123388582) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.878522377456784) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (114,0.89873417721519) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (101,0.894033412887828) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (109,0.905942910622368) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (105,0.912200684150513) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (104,0.880537580886013) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (100,0.902199344875994) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (99,0.92507068803016) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (99,0.883901992214335) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (87,0.907118644067797) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.930178612850847) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (85,0.880158905318914) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (85,0.881091617933723) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (82,0.898850035203004) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (77,0.792304126101066) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.880568720379147) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.857019251568246) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.880762411347518) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (70,0.917994917994918) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.887558216899534) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (69,0.889949868703748) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.896680691912108) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.85852534562212) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (64,0.872617387261739) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (62,0.836189170664839) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.737826685006877) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.819440815259995) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.805837894188168) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.863237139272271) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.875623960066556) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.873100983020554) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.832384253528101) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.834539089848308) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.76730869338886) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.814760508308895) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.852720236740269) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.825171516441921) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.81019955654102) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.900023747328426) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.854397950469684) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.845379688929552) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.872364039955605) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.724239450441609) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.847392144236961) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.735991379310345) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.736179529282977) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.853647058823529) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.842877425461429) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.827407581822463) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.812372071867182) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.813519813519813) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.742579600647598) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.760679611650485) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.769528978484202) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.678711704634721) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.79477069931521) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.770513383062747) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.636017410228509) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (396,0.952712100139082) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (392,0.947614593077643) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (392,0.936249418334109) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (397,0.945989427717766) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (376,0.948060784758449) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (373,0.952710495963091) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (349,0.937049941927991) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (368,0.939024390243902) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (354,0.953697304768486) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (315,0.956302905513612) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (303,0.899073622623111) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (327,0.933174224343675) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (295,0.940575673166202) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (295,0.93886058360352) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (300,0.922497055359246) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (293,0.896750902527076) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (261,0.921972235519387) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (269,0.93308463511308) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (247,0.930308699719364) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (252,0.939981949458484) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (230,0.867843511450382) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (252,0.848318804483188) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (225,0.886192671681631) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (218,0.92644044636757) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.893278728017345) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (192,0.923356009070295) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (201,0.938009049773756) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (189,0.914433474377057) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (186,0.930389727416085) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (183,0.91874859329282) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.918955123911587) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.913441238564391) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (162,0.908678903240426) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (151,0.85012285012285) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (140,0.916628701594533) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (151,0.890788224121557) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (149,0.878963202962277) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.869397383658589) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (138,0.901322923443938) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (134,0.874090407938258) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (127,0.884851344974044) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (125,0.881990521327014) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (115,0.847767086109055) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (116,0.919582008178101) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (113,0.877152163797115) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (109,0.905454545454545) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (107,0.904587155963303) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (96,0.788807785888078) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (90,0.870774497218656) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (93,0.911121239744758) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (86,0.830568124685772) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (90,0.859353023909986) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (87,0.882057989971659) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (85,0.908190819081908) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (88,0.901736745886654) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.864637465973769) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (77,0.893464637421665) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.817712177121771) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (73,0.878242580042066) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.831455169034787) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (69,0.865523241362168) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.842843779232928) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.874828060522696) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (62,0.88711943793911) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.84228684080828) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.848019801980198) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (58,0.837431367868226) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.885245901639344) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.824888050907377) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.862811517057827) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.830261881668283) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.84761045987376) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.84503127171647) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.824973544973545) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.84503745318352) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.875670404680644) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.787304837471206) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.823614457831325) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.794488001900689) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.854377880184332) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.689564068692206) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.877663079288381) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.838980954736582) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.879907621247113) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.883087907184293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.829443447037702) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.801331853496115) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.819965234666004) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.844422211105553) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.82573849324479) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.819451371571072) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.764828303850156) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.743595999024152) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.821473790833122) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.778629395852119) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.720295202952029) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.66336361134049) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.805212939259949) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.752428334517887) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (397,0.934160305343511) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (397,0.94524089306698) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (367,0.929148834997622) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (367,0.947294448348559) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (389,0.933301458283528) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (379,0.93422880963409) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (364,0.945616502578528) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (337,0.933944069431051) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (323,0.947319347319347) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (352,0.927371658386562) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (334,0.918444760550024) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (309,0.942599620493359) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (298,0.940340909090909) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (299,0.915570437694331) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (277,0.934536201711774) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (274,0.928488775746355) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (274,0.940456578018357) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (273,0.932895354293759) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (260,0.926910299003322) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (241,0.929078014184397) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (235,0.942850478189876) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (225,0.935612268789511) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (218,0.916131989000917) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (227,0.888836662749706) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (211,0.885599598595083) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (211,0.935409986257444) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (196,0.910611854684512) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (190,0.902900982977703) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (169,0.89730241362991) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (186,0.915159944367177) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (158,0.908758289503773) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (162,0.886347390993567) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (164,0.90384167636787) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (151,0.902107728337237) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (155,0.824947832135404) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (142,0.940032414910859) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (140,0.876129978011239) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.890436397400186) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.923581809657759) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (130,0.904063476797307) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.924432081594808) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (117,0.890865384615385) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (121,0.886433106042793) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (114,0.887542413960252) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (115,0.868705882352941) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (105,0.878974845752254) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (105,0.866553644950351) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.897697602658438) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (106,0.897867564534231) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (100,0.869216241737488) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.854842529007814) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (83,0.880474452554744) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (88,0.871479289940828) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.862068965517241) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (81,0.823916898910565) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (81,0.87471317117944) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (75,0.907271839921913) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (70,0.888687372477896) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (75,0.850431447746884) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.86952675807165) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.847666971637694) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.773658283640135) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (67,0.878610108303249) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (69,0.883764705882353) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (62,0.887003126395712) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.860380998312033) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.879873073436083) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.740056022408963) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.848177203507153) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.852720918622067) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.788466318667661) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.776524741081703) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.802967083912842) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.807327001356852) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.74825550187869) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.808443745789355) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.859515570934256) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.800415368639668) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.839123699933613) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.85170429393537) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.852880422308928) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.788147007889167) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.670103092783505) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.831607384061234) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.769647696476965) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.820658611703388) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.736505032021958) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.838047445255474) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.827401267674305) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.841885302995124) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.732270732270732) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.812539987204094) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.77541843617287) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.814403200711269) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.702660406885759) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.717720751667003) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.749895963379109) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.833855799373041) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.771508603441377) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (378,0.942979490646833) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (381,0.943245778611632) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (380,0.94035257184255) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (383,0.962669155389641) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (331,0.925541941564562) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (324,0.938212927756654) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (309,0.954211150652432) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (308,0.956339990710636) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (320,0.943022694350555) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (301,0.954808806488992) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (304,0.932391048292108) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (281,0.922196796338673) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (249,0.926408942710759) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (282,0.927337278106509) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (268,0.926519076778144) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (241,0.924154025670945) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (227,0.937823834196891) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (230,0.912776991984913) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (223,0.926325668700141) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.910007199424046) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (218,0.913730255164034) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (211,0.937013446567587) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.938219403335682) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (207,0.917582417582417) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (191,0.894787183165949) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (177,0.914111865200094) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (165,0.914821944177093) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.9206206431302) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (161,0.903480622068625) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (167,0.887135341941655) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (148,0.920459770114942) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (158,0.915129151291513) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (158,0.856050955414013) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (143,0.940582702454691) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.893748514380794) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (138,0.891955098222638) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.912106135986733) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (118,0.910209662716499) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (119,0.893617021276596) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.910199296600234) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (109,0.888345956511116) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.908580780921206) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (118,0.85981724983285) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.807237999497361) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (101,0.920968478757423) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.905616406432067) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.909007981755986) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (93,0.871028923660503) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.892279662946937) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (89,0.87518573551263) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (80,0.888736263736264) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (80,0.834558004898686) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (75,0.892543365622888) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.871602624179944) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (73,0.920446774561204) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.841117108659939) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.88423058599333) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.750309329373917) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.819299044151115) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (64,0.869031975470872) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.85938242280285) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.872703412073491) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (58,0.801672903367818) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (58,0.895564421497301) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.824200913242009) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.842509603072983) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.868368479467258) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.750543478260869) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.84572864321608) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.865376344086021) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.849712313003452) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.818655420491193) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.82167530224525) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.787552070570938) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.764059753954306) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.780626780626781) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.837742504409171) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.796061269146608) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.840107190710138) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.844711872251794) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.771172825338535) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.826396872844332) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.782767266925005) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.793881644934804) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.758937198067633) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.804933299773471) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.816696914700544) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.858898398068905) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.750996015936255) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.766917293233083) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.865286236297198) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.814657008947593) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.683255572835666) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.889537494387068) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.70027397260274) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.710260499734184) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (359,0.938007894125842) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (373,0.929148396030464) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (373,0.930340199028003) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (365,0.943217665615142) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (351,0.934422509643749) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (326,0.938672768878719) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (330,0.943181818181818) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (314,0.917971488665576) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (298,0.927476549988561) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (280,0.927379341452413) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (275,0.939483900433889) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (277,0.919571656413762) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (282,0.903471231573942) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (279,0.936288088642659) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (259,0.926994141505182) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (238,0.910505836575875) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (251,0.93389395001168) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (222,0.89699472356045) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (196,0.922896097790315) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.929571106094808) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (206,0.941471191072649) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (209,0.915285880980163) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (187,0.937081659973226) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (169,0.891990846681922) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (173,0.928407526388251) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (179,0.935153583617747) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (175,0.893796992481203) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (159,0.896208530805687) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (164,0.891215293582157) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (144,0.894282380396733) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (160,0.906091928744227) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (147,0.896969696969697) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (152,0.902559286217422) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (135,0.926115241635688) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (128,0.910437398750289) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.884047786366831) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (123,0.919790956600773) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (116,0.889611797874648) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (118,0.894700300856283) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (107,0.862725940400586) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (107,0.89745323416723) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.875083500334001) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.917658035922556) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (99,0.877098723975823) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.900353356890459) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.861282660332542) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (88,0.864287338031371) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (90,0.864826941066417) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (84,0.90782686701054) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.827603513174404) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.876328283970156) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (78,0.816810877626699) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (71,0.852161100196464) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.876665022200296) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.892138063279003) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.852275379229872) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (69,0.80523890371089) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (69,0.89031971580817) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (64,0.838799720865318) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.879193056532958) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (58,0.869746923612723) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.859432531947152) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.855797263960529) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.861531190926276) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.877043673012318) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.830406757057124) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.823343135037322) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.815963511972634) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.80306905370844) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.886635944700461) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.821601661666282) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.864347420032687) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.852624643092466) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.806205250596659) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.832639059054153) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.864088909469785) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.838546458141674) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.841983317886932) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.810797782598216) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.859231313369488) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.729853259562184) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.824464561713402) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.766083312792704) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.815245478036176) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.61463133640553) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.796529284164859) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.747038037829973) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.837807854448635) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.774766962930847) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.862701252236136) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.838320575272558) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.796355577443979) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.837251769002511) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.723449903454194) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (390,0.943111720356409) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (394,0.941344098788886) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (363,0.956684617185671) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (388,0.913378248315688) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (356,0.943857176415316) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (338,0.944365865157944) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (341,0.942908926144087) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (339,0.923076923076923) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (282,0.941947142534448) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (288,0.934869503879614) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (300,0.948609146628948) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (289,0.948664343786295) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (292,0.927440321437012) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (263,0.919101666264187) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (260,0.939753431030472) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (261,0.931168201648085) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (239,0.930678819860878) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (238,0.925429310750411) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (226,0.950766372503483) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (223,0.939647168059424) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (238,0.919540229885057) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (216,0.908223534756929) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (198,0.930068283494231) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (190,0.927849264705882) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (186,0.926075588305205) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (190,0.932804479701353) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.901950975487744) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (165,0.94096025729382) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (168,0.936338289962825) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (165,0.925866421849897) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (153,0.918179674605046) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (151,0.889051450865642) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (147,0.93466172381835) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (146,0.912822878228782) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (127,0.92406264831514) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.87672514619883) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (132,0.890973036342321) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (125,0.905815279361459) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (110,0.889400921658986) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (112,0.865304226660474) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (118,0.910045335242186) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (116,0.897238569488456) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (110,0.854803756320732) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (107,0.90625) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (97,0.903271692745377) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (101,0.913876702839991) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.88991888760139) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (95,0.865896771152486) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (93,0.901109275430729) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (90,0.901492537313433) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (86,0.852832055404403) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (89,0.858653183176281) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (82,0.858048780487805) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (80,0.869932852524248) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.889090496711272) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (78,0.896317829457364) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (77,0.892119009766069) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.856290174471993) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (71,0.890079180251514) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.726404494382022) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (67,0.905402325051288) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.826590962857819) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.88936678614098) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.838297872340425) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.797135032652201) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.848924341310128) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.885818181818182) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.880097087378641) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.832529867952211) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.850841830685321) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.820338983050847) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.67741935483871) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.855822143957737) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.835058993498676) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.815170008718396) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.865622809067539) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.833570412517781) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.807608957380207) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.785352927005541) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.833264033264033) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.862557077625571) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.778743481499876) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.818950177935943) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.807204803202135) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.835551545228965) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.778224852071006) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.744557016598405) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.73802669458257) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.75384214639229) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.847930283224401) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.842250287026406) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.838967136150235) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.822881761608425) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.7851966873706) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.720473192138905) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.758868074930251) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (383,0.951429868361325) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (382,0.935750344194585) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (371,0.949725776965265) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (344,0.944087992667278) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (392,0.955411054342778) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (361,0.949798296727925) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (354,0.954296160877514) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (326,0.907050500121981) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (321,0.947368421052631) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (293,0.935015472506546) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (308,0.94629963898917) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (328,0.937470835277648) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (288,0.93393536121673) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (285,0.949372967951695) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (258,0.926022216969983) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (252,0.913256955810147) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (250,0.943482224247949) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (251,0.932195353247985) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (227,0.941825707059094) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (230,0.940886699507389) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (219,0.920431093785829) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (203,0.929538030797947) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (200,0.941863180014009) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (208,0.934316991908615) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (204,0.903396771862202) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (193,0.892315104794025) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (177,0.875905118601748) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (163,0.898557158712541) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (178,0.936907788305986) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (170,0.912552891396333) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (159,0.91599045346062) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (144,0.922380629101607) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (145,0.910709987966306) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (151,0.910662824207493) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (134,0.91804788213628) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (136,0.903957528957529) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (130,0.904576436222006) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.9032410693163) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (126,0.910507569011576) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (121,0.871679197994987) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.899589228662711) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (105,0.90691908947487) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (108,0.882266731802638) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.909937165464277) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (104,0.898093359631821) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (103,0.898517145505097) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.90556453474948) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (89,0.873345935727788) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (90,0.860771184483953) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (92,0.87972972972973) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (89,0.912517449976733) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (82,0.867244205104191) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (77,0.860787824529991) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.8750598372427) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (73,0.881859573448358) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.850269784172662) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.862982273201251) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (73,0.854074247023114) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (72,0.809160305343511) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.898933704218822) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.827188940092166) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.857617465590887) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (62,0.86075378018506) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (60,0.837376460017969) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (60,0.799814643188137) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (58,0.876528410453129) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.822458270106221) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.752043596730245) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.879429250891795) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.792210993344836) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.900909302867801) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.822479928635147) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.87355801854784) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (47,0.844191343963554) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.799615107048352) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.857925993604385) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.836230110159119) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.840919133893062) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.823117338003503) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.832502708559047) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.800888230940044) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.843180808529542) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.776632302405498) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.866138265081462) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.796213515645543) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.770543331836551) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.718188530642382) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.834423117485762) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.753682072117826) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.803141875306824) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.638873550524572) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.759153175591532) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.759052652182423) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.830610134436401) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.742883170120701) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.800578034682081) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.714539795464651) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (387,0.933061039255729) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (367,0.930166270783848) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (393,0.950263580105432) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (386,0.941776849918812) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (371,0.94829584975655) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (327,0.90861494526416) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (333,0.934071222329163) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (312,0.940333024976873) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (319,0.952755905511811) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (291,0.928819846974264) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (320,0.924105011933174) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (286,0.937973484848485) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (306,0.946532804109269) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (278,0.946003301108229) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (267,0.92492351141445) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (256,0.944150769937945) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (261,0.943794448267951) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (243,0.917827967323402) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (229,0.936320754716981) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (233,0.917488573490498) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (237,0.907488986784141) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (223,0.943184475099369) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (212,0.927693761814745) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (203,0.945964589560819) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (211,0.930276087973795) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (189,0.900637243332546) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (182,0.923324150596878) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (189,0.910302165120152) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (174,0.916034046468829) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (171,0.907547602661161) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (156,0.933886201336098) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (157,0.9039780521262) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (150,0.889532293986637) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (158,0.888785046728972) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (141,0.932767478222024) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (139,0.891423140888572) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (137,0.877201332698715) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (136,0.886715955163367) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.882861819032032) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (115,0.899142128448875) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (121,0.905923344947735) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (115,0.85835493754419) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (111,0.892103391036282) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (107,0.859504132231405) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (100,0.819444444444444) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (106,0.848375451263538) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.841389728096677) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (96,0.858062918008422) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.866728366496992) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (94,0.904174573055028) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.883874000940291) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (84,0.854593682179889) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (83,0.906145995118704) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (87,0.89874593590339) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (80,0.867531264474293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (71,0.871425545435289) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (74,0.886390255329117) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.787721573713791) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (71,0.845849802371542) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.826116221673771) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.891566265060241) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.833047315518509) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.883359963478658) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.786316776007498) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.839675914559293) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.829219645973573) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.841117216117216) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.824148296593186) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.837520165936852) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.877200902934537) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.863420158550396) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.736785904965296) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.851279788172992) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.741027627162406) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.697504906083544) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.754951085659747) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.815701292484442) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (44,0.691448216587881) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.840065374737333) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.829336437718277) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (42,0.724161533196441) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.858744394618834) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.795233463035019) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.766457023060796) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.816364054240404) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.788235294117647) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.763184584178499) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.680154972018941) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.794871794871795) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.747634875990795) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.821968815452641) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.814132104454685) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (29,0.680098012523822) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.651342850229857) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.63776918829376) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.836888888888889) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.75130507412821) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (392,0.939098093909809) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (379,0.957686882933709) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (371,0.924756154203437) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (383,0.942587041733917) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (397,0.947464180122811) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (356,0.929844871498032) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (352,0.957913998170174) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (360,0.937442502299908) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (335,0.945193171608266) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (352,0.939683952815491) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (312,0.946127167630058) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (320,0.93780573025856) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (275,0.948735169017237) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (298,0.936179981634527) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (272,0.925908881730327) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (281,0.928959276018099) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (266,0.942338617045716) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (252,0.916311420237828) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (262,0.943153049482163) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (240,0.933608058608059) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (222,0.939130434782609) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (220,0.925796696906257) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (206,0.908452353078904) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (211,0.931767833407178) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (222,0.901987979657882) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (200,0.928455661032646) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (196,0.928901200369344) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (181,0.935962877030162) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (162,0.912796208530806) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (167,0.936327853669831) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (172,0.914111865200094) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (167,0.932306995698438) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (160,0.931144550843593) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (154,0.917685794920038) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (156,0.91789667896679) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (143,0.93822573153739) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (135,0.930113768284188) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (120,0.901171788635861) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (135,0.910498747437941) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (127,0.875512995896033) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (120,0.927276795703737) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.880533079485959) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (119,0.942141230068337) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (110,0.908587257617728) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (112,0.897986577181208) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (104,0.893064876957494) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (104,0.853658536585366) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (98,0.881488286632981) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (102,0.921433657726086) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (96,0.911570811548079) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (82,0.877422164162856) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (80,0.893617021276596) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (83,0.923837339923613) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (81,0.844239631336405) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (81,0.886069889899473) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.85734496580995) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (78,0.866697631212262) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (79,0.874169251218431) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.852757793764988) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.872311212814645) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.845542427497315) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (61,0.830812854442344) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (60,0.818329090488313) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (66,0.821681864235056) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (64,0.898277082841633) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.852543464262717) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.843390467245832) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.864185826025124) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.740112994350282) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.880052436093511) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (54,0.853464419475655) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.828899194532585) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (52,0.838430050565856) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.855004677268475) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (48,0.867371456030754) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.836273179556761) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.861166116611661) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.83525656764056) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.712680577849117) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.773505333664103) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.811254980079681) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.803274494755692) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.784564272936366) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.821192052980132) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.850057937427578) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.797594263243118) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.808585194918966) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.825700615174299) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.703475474488538) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (34,0.834313085129545) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.808871168701719) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.83205930044012) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.510006455777921) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.83009814612868) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.766931340495096) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (31,0.854600796066495) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (28,0.694024108745832) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.805599458116956) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (395,0.949461862423959) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!84!black,draw=red!84!black,line width=0.05pt},
forget plot
]
coordinates{
 (400,0.940305635148042) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (399,0.951202749140893) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (376,0.94974406700791) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (374,0.957481038841646) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (388,0.946094106898127) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!88!black,draw=red!88!black,line width=0.05pt},
forget plot
]
coordinates{
 (355,0.954766875434934) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (348,0.945183111733147) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (325,0.934083978418954) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (337,0.948700253048079) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (315,0.935748792270531) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (303,0.93956934475573) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!92!black,draw=red!92!black,line width=0.05pt},
forget plot
]
coordinates{
 (309,0.926012098650535) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (290,0.946436888681882) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (284,0.94877096255456) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (285,0.932323710364411) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red!96!black,draw=red!96!black,line width=0.05pt},
forget plot
]
coordinates{
 (283,0.881557085212438) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (267,0.926106092193653) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (240,0.91988821611551) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (242,0.936298355339356) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (242,0.944292237442922) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=red,draw=red,line width=0.05pt},
forget plot
]
coordinates{
 (220,0.948069852941176) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (231,0.894094680331869) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (215,0.910953891192974) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (207,0.929101019462465) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor1,draw=mycolor1,line width=0.05pt},
forget plot
]
coordinates{
 (185,0.932905885046943) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (191,0.919963201471941) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (176,0.931339225016995) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (183,0.833289577316881) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor2,draw=mycolor2,line width=0.05pt},
forget plot
]
coordinates{
 (173,0.870991432068543) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (177,0.909341923607915) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (153,0.891376846117198) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (148,0.924608819345661) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor3,draw=mycolor3,line width=0.05pt},
forget plot
]
coordinates{
 (146,0.904174573055028) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (157,0.922968859751814) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (143,0.905770635500365) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor4,draw=mycolor4,line width=0.05pt},
forget plot
]
coordinates{
 (141,0.900096061479347) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (126,0.899784740492705) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (134,0.893432764012509) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor5,draw=mycolor5,line width=0.05pt},
forget plot
]
coordinates{
 (129,0.908745247148289) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (133,0.893189964157706) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (125,0.900207421064761) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor6,draw=mycolor6,line width=0.05pt},
forget plot
]
coordinates{
 (123,0.886419204267615) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (108,0.929667812142039) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (110,0.920957542908762) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor7,draw=mycolor7,line width=0.05pt},
forget plot
]
coordinates{
 (96,0.905677655677656) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (98,0.88590899098703) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (98,0.861262112975656) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor8,draw=mycolor8,line width=0.05pt},
forget plot
]
coordinates{
 (92,0.853032845840326) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (91,0.913465987968533) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (89,0.91643059490085) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor9,draw=mycolor9,line width=0.05pt},
forget plot
]
coordinates{
 (92,0.837742917349567) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (86,0.878060153882024) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor10,draw=mycolor10,line width=0.05pt},
forget plot
]
coordinates{
 (82,0.901427913403961) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (78,0.838621940163191) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (73,0.877507670521595) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor11,draw=mycolor11,line width=0.05pt},
forget plot
]
coordinates{
 (76,0.874259102455546) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (81,0.865921787709497) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor12,draw=mycolor12,line width=0.05pt},
forget plot
]
coordinates{
 (70,0.828997221520586) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (65,0.831875147789075) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor13,draw=mycolor13,line width=0.05pt},
forget plot
]
coordinates{
 (68,0.882807327300816) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.850534584333406) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor14,draw=mycolor14,line width=0.05pt},
forget plot
]
coordinates{
 (63,0.86932804363997) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (56,0.824389225307997) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor15,draw=mycolor15,line width=0.05pt},
forget plot
]
coordinates{
 (60,0.864516129032258) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (59,0.878966455122393) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor16,draw=mycolor16,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.79732165851146) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (55,0.808812729498164) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor17,draw=mycolor17,line width=0.05pt},
forget plot
]
coordinates{
 (57,0.802688430148824) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (51,0.868092691622103) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor18,draw=mycolor18,line width=0.05pt},
forget plot
]
coordinates{
 (49,0.835748792270531) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (53,0.806613226452906) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor19,draw=mycolor19,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.88034188034188) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.779572694111516) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor20,draw=mycolor20,line width=0.05pt},
forget plot
]
coordinates{
 (50,0.84643179765131) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor21,draw=mycolor21,line width=0.05pt},
forget plot
]
coordinates{
 (45,0.825348432055749) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.795289499509323) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor22,draw=mycolor22,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.856514084507042) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (40,0.805523051720103) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor23,draw=mycolor23,line width=0.05pt},
forget plot
]
coordinates{
 (46,0.726723427629772) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor24,draw=mycolor24,line width=0.05pt},
forget plot
]
coordinates{
 (43,0.80864040660737) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (41,0.807138249838744) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor25,draw=mycolor25,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.873322720036388) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor26,draw=mycolor26,line width=0.05pt},
forget plot
]
coordinates{
 (37,0.824190064794816) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor27,draw=mycolor27,line width=0.05pt},
forget plot
]
coordinates{
 (38,0.831571529245948) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (39,0.828000775644754) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor28,draw=mycolor28,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.834297520661157) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor29,draw=mycolor29,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.774317201724964) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (36,0.826616682286785) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor74,draw=mycolor74,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.708015768725361) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor30,draw=mycolor30,line width=0.05pt},
forget plot
]
coordinates{
 (32,0.739744155271283) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor31,draw=mycolor31,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.73972602739726) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor32,draw=mycolor32,line width=0.05pt},
forget plot
]
coordinates{
 (35,0.720064291454594) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.735064280312579) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor33,draw=mycolor33,line width=0.05pt},
forget plot
]
coordinates{
 (33,0.800338051975491) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor75,draw=mycolor75,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.777827248441674) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor34,draw=mycolor34,line width=0.05pt},
forget plot
]
coordinates{
 (30,0.638319672131147) 
};
\addplot [
%
mark size=1.9pt,
only marks,
mark=*,
mark options={fill=mycolor35,draw=mycolor35,line width=0.05pt},
forget plot
]
coordinates{
 (27,0.785007808433108)
};

\node at (axis cs:230, 0.57) [shape=circle,draw=black,line width=1.2pt,inner sep=0.2pt,anchor=west,minimum size=20pt] {\LARGE$\epsilon$};
\node at (axis cs:90, 0.6) [anchor=west,minimum size=16pt] {\large \emph{large}};
\node at (axis cs:390, 0.6) [anchor=east,minimum size=16pt] {\large \emph{small}};

\draw[<-,line width=1.2pt] (axis cs:90, 0.57) -- (axis cs:230, 0.57);
\draw[->,line width=1.2pt] (axis cs:258, 0.57) -- (axis cs:390, 0.57);

\end{axis}
\end{tikzpicture}%