### 2013-07-31 figurefactory/files/limno_bgape_sc.tex
Many small fixes

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
%

% defining custom colors
\definecolor{mycolor1}{rgb}{0,0,0.5625}
\definecolor{mycolor2}{rgb}{0,0,0.625}
\definecolor{mycolor3}{rgb}{0,0,0.6875}
\definecolor{mycolor4}{rgb}{0,0,0.8125}
\definecolor{mycolor5}{rgb}{0,0,0.875}
\definecolor{mycolor6}{rgb}{0,0,0.9375}
\definecolor{mycolor7}{rgb}{0,0.0625,1}
\definecolor{mycolor8}{rgb}{0,0.125,1}
\definecolor{mycolor9}{rgb}{0,0.1875,1}
\definecolor{mycolor10}{rgb}{0,0.25,1}
\definecolor{mycolor11}{rgb}{0,0.3125,1}
\definecolor{mycolor12}{rgb}{0,0.375,1}
\definecolor{mycolor13}{rgb}{0,0.4375,1}
\definecolor{mycolor14}{rgb}{0,0.5,1}
\definecolor{mycolor15}{rgb}{0,0.5625,1}
\definecolor{mycolor16}{rgb}{0,0.625,1}
\definecolor{mycolor17}{rgb}{0,0.6875,1}
\definecolor{mycolor18}{rgb}{0,0.75,1}
\definecolor{mycolor19}{rgb}{0,0.8125,1}
\definecolor{mycolor20}{rgb}{0,0.875,1}
\definecolor{mycolor21}{rgb}{0,0.9375,1}
\definecolor{mycolor22}{rgb}{0,1,1}
\definecolor{mycolor23}{rgb}{0.0625,1,0.9375}
\definecolor{mycolor24}{rgb}{0.125,1,0.875}
\definecolor{mycolor25}{rgb}{0.1875,1,0.8125}
\definecolor{mycolor26}{rgb}{0.25,1,0.75}
\definecolor{mycolor27}{rgb}{0.3125,1,0.6875}
\definecolor{mycolor28}{rgb}{0.375,1,0.625}
\definecolor{mycolor29}{rgb}{0.4375,1,0.5625}
\definecolor{mycolor30}{rgb}{0.5,1,0.5}
\definecolor{mycolor31}{rgb}{0.5625,1,0.4375}
\definecolor{mycolor32}{rgb}{0.625,1,0.375}
\definecolor{mycolor33}{rgb}{0.6875,1,0.3125}
\definecolor{mycolor34}{rgb}{0.75,1,0.25}
\definecolor{mycolor35}{rgb}{0.8125,1,0.1875}
\definecolor{mycolor36}{rgb}{0.875,1,0.125}
\definecolor{mycolor37}{rgb}{0.9375,1,0.0625}
\definecolor{mycolor38}{rgb}{1,1,0}
\definecolor{mycolor39}{rgb}{1,0.9375,0}
\definecolor{mycolor40}{rgb}{1,0.875,0}
\definecolor{mycolor41}{rgb}{1,0.8125,0}
\definecolor{mycolor42}{rgb}{1,0.75,0}
\definecolor{mycolor43}{rgb}{1,0.6875,0}
\definecolor{mycolor44}{rgb}{1,0.625,0}
\definecolor{mycolor45}{rgb}{1,0.5625,0}
\definecolor{mycolor46}{rgb}{1,0.4375,0}
\definecolor{mycolor47}{rgb}{1,0.375,0}
\definecolor{mycolor48}{rgb}{1,0.3125,0}
\definecolor{mycolor49}{rgb}{1,0.25,0}
\definecolor{mycolor50}{rgb}{1,0.1875,0}
\definecolor{mycolor51}{rgb}{1,0.125,0}
\definecolor{mycolor52}{rgb}{1,0.0625,0}
\definecolor{mycolor53}{rgb}{0.9375,0,0}
\definecolor{mycolor54}{rgb}{0.875,0,0}
\definecolor{mycolor55}{rgb}{0.8125,0,0}
\definecolor{mycolor56}{rgb}{0.6875,0,0}
\definecolor{mycolor57}{rgb}{0.625,0,0}

\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\Large},
label style={font=\Large},
xlabel shift={-10pt},
ylabel shift={-17pt},
legend style={font=\Large},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=1478,
xtick={0, 400, 1000, 1400},
xlabel={Length (m)},
ymin=-18, ymax=0,
ytick={0, -4, -14, -18},
ylabel={Depth (m)},
name=plot1,
axis lines*=box,
unbounded coords=jump,
colormap/jet,
colorbar right,
colorbar style={
  tickwidth=0.03cm,
  at={(1.04, 0)},
  anchor=south east,
  yticklabel pos=right,
  yticklabel shift={-2pt}
},
colorbar/width=0.15cm,
tickwidth=0.1cm,
point meta min=0.9,
point meta max=9.3
]

\addplot[scatter,only marks,mark size=1pt,scatter src=explicit,scatter/use mapped color={mark=*,draw=mapped color,fill=mapped color}] plot coordinates{ (0,-0.831) [9.000000e-01] (0.396123095452957,-0.832) [9.000000e-01] (16.8850657170843,-2.228) [9.000000e-01] (19.386309378027,-2.485) [1] (20.7170231970803,-2.774) [1] (21.6989675078734,-3.031) [1.200000e+00] (22.8071472777201,-3.289) [1.400000e+00] (24.5610830029506,-3.559) [1.500000e+00] (25.9984308826544,-3.773) [1.700000e+00] (27.8988551242371,-4.005) [1.800000e+00] (29.1617886372451,-4.262) [2.200000e+00] (62.5565899293934,-9.656) [4.900000e+00] (62.5565899293934,-9.727) [4.800000e+00] (70.9331967712539,-9.716) [4.700000e+00] (70.5118107778974,-9.674) [4.600000e+00] (71.7410292630623,-9.816) [4.500000e+00] (72.8642052039513,-9.938) [4.400000e+00] (75.0082362859188,-9.97) [4.300000e+00] (76.7726194267189,-10.033) [4.200000e+00] (78.6602249298857,-10.233) [4.100000e+00] (80.0953847487801,-10.526) [4] (81.4237947228505,-10.853) [3.800000e+00] (82.4094012394071,-11.137) [3.800000e+00] (90.4064215094236,-12.599) [3.200000e+00] (91.7791118891167,-12.761) [3.200000e+00] (93.2690607870255,-12.927) [3.300000e+00] (94.9258643776602,-13.131) [3.200000e+00] (96.4233119768856,-13.315) [3.100000e+00] (97.7794091531292,-13.539) [3] (99.2566050421831,-13.721) [2.900000e+00] (100.834528136853,-13.918) [2.800000e+00] (102.038059375172,-14.076) [2.700000e+00] (103.564851467609,-14.21) [2.700000e+00] (105.10292010624,-14.316) [2.500000e+00] (106.352328039151,-14.45) [2.500000e+00] (107.570853984965,-14.606) [2.400000e+00] (109.162235182079,-14.748) [2.400000e+00] (110.512369431851,-14.892) [2.300000e+00] (111.678350262332,-15.037) [2.300000e+00] (113.235254720706,-15.144) [2.200000e+00] (114.500486265143,-15.249) [2.200000e+00] (115.691203492305,-15.382) [2.100000e+00] (115.815129572001,-15.528) [2.100000e+00] (117.676456857818,-15.642) [2] (119.477680213669,-15.649) [2] (121.68110463208,-15.667) [1.900000e+00] (123.336430499725,-15.756) [1.900000e+00] (124.837654253971,-15.889) [1.900000e+00] (125.874974432477,-16.079) [1.900000e+00] (127.464940685869,-16.178) [1.800000e+00] (129.19548862963,-16.275) [1.800000e+00] (130.857541771519,-16.416) [1.800000e+00] (132.21363502171,-16.627) [1.800000e+00] (133.550247559459,-16.799) [1.900000e+00] (134.959009363721,-16.962) [2] (136.520336217567,-17.12) [1.800000e+00] (137.872711335487,-17.243) [1.800000e+00] (139.076945489566,-17.415) [1.700000e+00] (140.399268842806,-17.544) [1.800000e+00] (141.977956830779,-17.703) [1.800000e+00] (143.325077182371,-17.802) [1.700000e+00] (144.660151796424,-17.868) [1.800000e+00] (146.571133738,-17.888) [1.800000e+00] (148.269322831534,-17.76) [1.800000e+00] (149.626182665483,-17.551) [1.800000e+00] (151.662688380933,-17.391) [1.800000e+00] (152.718137136586,-17.26) [1.900000e+00] (153.837406694506,-17.099) [1.800000e+00] (155.125065343297,-16.946) [1.800000e+00] (157.211165026066,-16.724) [1.700000e+00] (158.777742720471,-16.503) [1.700000e+00] (160.189513695575,-16.363) [1.900000e+00] (161.502800912108,-16.229) [1.900000e+00] (162.89502879595,-16.13) [1.800000e+00] (164.148212512045,-15.928) [1.800000e+00] (165.084159331749,-15.722) [1.900000e+00] (166.140376448098,-15.548) [1.900000e+00] (168.073974836101,-15.419) [1.900000e+00] (171.903368566451,-15.398) [1.700000e+00] (173.569902526982,-17.024) [1.300000e+00] (175.53259191065,-16.566) [1.400000e+00] (177.935861325938,-16.19) [1.500000e+00] (179.811407222947,-15.874) [1.600000e+00] (181.880781436237,-15.569) [1.600000e+00] (183.237574077208,-15.215) [1.700000e+00] (184.995797300345,-14.884) [1.800000e+00] (186.783495133997,-14.578) [1.900000e+00] (187.952482126714,-14.274) [2.100000e+00] (189.258334913618,-13.899) [2.200000e+00] (190.738599459577,-13.554) [2.200000e+00] (192.463118747332,-13.211) [2.200000e+00] (194.003421319868,-12.947) [2.300000e+00] (195.57383981248,-12.748) [2.300000e+00] (196.983300606632,-12.597) [2.300000e+00] (198.481506135811,-12.432) [2.300000e+00] (200.168477251637,-12.264) [2.300000e+00] (201.114162264182,-12.13) [2.400000e+00] (202.394383846677,-12.051) [2.400000e+00] (203.75431396949,-11.953) [2.400000e+00] (205.330754700682,-11.83) [2.400000e+00] (206.892138374923,-11.74) [2.500000e+00] (207.741581757429,-11.685) [2.500000e+00] (209.102985706053,-11.61) [2.600000e+00] (211.135704464084,-11.467) [2.700000e+00] (212.557210594941,-11.291) [2.700000e+00] (214.108788883075,-11.101) [2.800000e+00] (215.157375520155,-11.056) [2.900000e+00] (216.15643461017,-10.923) [3.100000e+00] (217.950858027716,-10.784) [3.300000e+00] (219.060511686867,-10.537) [3.300000e+00] (220.557241700015,-10.4) [3.500000e+00] (221.391560315826,-10.222) [3.800000e+00] (222.696512075325,-10.14) [4.100000e+00] (224.639715159097,-9.899) [4.500000e+00] (226.372497123793,-9.749) [4.800000e+00] (227.744407109268,-9.586) [5.500000e+00] (228.880265172045,-9.464) [5.900000e+00] (230.235578800823,-9.304) [6.500000e+00] (231.748069617228,-9.108) [6.800000e+00] (233.558252978639,-8.875) [7] (235.063246470297,-8.73) [7.100000e+00] (236.020206310127,-8.554) [7.100000e+00] (237.432741377094,-8.429) [7.200000e+00] (238.858024551416,-8.262) [7.100000e+00] (240.511864989205,-8.107) [6.800000e+00] (242.258165228102,-7.934) [6.600000e+00] (243.650385776785,-7.836) [6.400000e+00] (245.109501714279,-7.66) [6.200000e+00] (246.41003125226,-7.468) [6] (247.91047415366,-7.229) [5.800000e+00] (249.408674496533,-7.087) [5.500000e+00] (250.905335539608,-6.901) [5.300000e+00] (252.429101641981,-6.732) [5.100000e+00] (254.261837265445,-6.508) [4.900000e+00] (256.077338317457,-6.347) [4.700000e+00] (257.562018696169,-6.134) [4.500000e+00] (258.874594918112,-6.023) [4.300000e+00] (260.562968311294,-5.811) [4.100000e+00] (262.146131130528,-5.613) [3.900000e+00] (263.724810180359,-5.365) [3.800000e+00] (265.261324877772,-5.226) [3.500000e+00] (266.695641668371,-4.999) [3.400000e+00] (268.071397382789,-4.805) [3.200000e+00] (269.50725304322,-4.552) [3] (270.879226899651,-4.345) [2.800000e+00] (272.529283359421,-4.089) [2.700000e+00] (274.421414868838,-3.899) [2.500000e+00] (276.528515938972,-3.657) [2.400000e+00] (278.224450469698,-3.467) [2.300000e+00] (279.852654000009,-3.213) [2.200000e+00] (281.380900404519,-2.967) [2.100000e+00] (282.910687914953,-2.692) [2.100000e+00] (284.340519646842,-2.457) [2] (285.643290082154,-2.181) [1.900000e+00] (286.883581618075,-1.945) [1.800000e+00] (288.129196146879,-1.684) [1.800000e+00] (289.668018103901,-1.46) [1.700000e+00] (291.718798937234,-1.22) [1.700000e+00] (293.480855222687,-0.988) [1.600000e+00] (295.087271814318,-0.83) [1.600000e+00] (296.946972721786,-0.798) [1.500000e+00] (298.652642321077,-0.83) [1.500000e+00] (300.10874040534,-0.909) [1.400000e+00] (301.305531104619,-0.978) [1.400000e+00] (302.585043928458,-1.136) [1.400000e+00] (303.913378375254,-1.404) [1.500000e+00] (305.470971595757,-1.672) [1.500000e+00] (306.92854127026,-1.939) [1.500000e+00] (308.210996786529,-2.209) [1.500000e+00] (309.714513617752,-2.477) [1.500000e+00] (310.96089607739,-2.74) [1.500000e+00] (312.301209440349,-3.024) [1.500000e+00] (313.842201832416,-3.269) [1.600000e+00] (314.94653232983,-3.551) [1.800000e+00] (316.189901233967,-3.763) [1.900000e+00] (317.843665738257,-4.031) [2] (319.262080173156,-4.255) [2.100000e+00] (320.718176108592,-4.52) [2.200000e+00] (321.911954198816,-4.732) [2.500000e+00] (323.378556571735,-5.026) [2.800000e+00] (324.913593203712,-5.256) [3.100000e+00] (326.006645836546,-5.507) [3.500000e+00] (327.553729582038,-5.694) [4] (328.869245116202,-5.916) [4.500000e+00] (330.669681375372,-6.094) [4.900000e+00] (331.861986586285,-6.351) [5.200000e+00] (333.832351781839,-6.545) [5.700000e+00] (335.219313382338,-6.757) [6.100000e+00] (336.676178566863,-6.917) [6.500000e+00] (337.793956904831,-7.134) [6.800000e+00] (339.0560970724,-7.311) [7] (340.45279368349,-7.528) [7.100000e+00] (341.845707518994,-7.672) [7.100000e+00] (343.47089012962,-7.851) [7.200000e+00] (345.114845293166,-7.963) [7.200000e+00] (346.776171422735,-8.152) [7.100000e+00] (348.35036028571,-8.301) [7.100000e+00] (349.459871648953,-8.528) [7.100000e+00] (350.616033398694,-8.703) [7] (351.497758437856,-8.86) [6.900000e+00] (353.034334026721,-8.989) [6.700000e+00] (354.586669071716,-9.074) [6.600000e+00] (356.179630516457,-9.155) [6.400000e+00] (357.438757186451,-9.271) [6.200000e+00] (358.529425133756,-9.39) [6.100000e+00] (359.478813004749,-9.532) [5.800000e+00] (361.113870203384,-9.675) [5.700000e+00] (362.460972905081,-9.782) [5.500000e+00] (363.784751087263,-9.879) [5.300000e+00] (365.233348925488,-9.969) [5.200000e+00] (366.592499170642,-10.101) [5] (368.268111476218,-10.237) [4.900000e+00] (369.818974559425,-10.419) [4.700000e+00] (371.642733276156,-10.685) [4.600000e+00] (372.877833114694,-11.001) [4.400000e+00] (373.847532867175,-11.19) [4.300000e+00] (375.693074138188,-11.412) [4.100000e+00] (377.630304671522,-11.463) [4] (379.09697275044,-11.528) [3.800000e+00] (380.372556803692,-11.64) [3.700000e+00] (381.856454982116,-11.776) [3.500000e+00] (383.005119923821,-11.919) [3.300000e+00] (384.857455777906,-12.088) [3.200000e+00] (386.078266002381,-12.239) [3.200000e+00] (387.260475696,-12.362) [3.200000e+00] (389.108956134072,-12.409) [3.100000e+00] (390.953653796453,-12.432) [2.900000e+00] (392.525527147113,-12.5) [2.800000e+00] (393.960604190631,-12.689) [2.800000e+00] (395.779878336545,-12.859) [2.800000e+00] (397.904264438766,-13.094) [2.700000e+00] (399.214520508063,-13.314) [2.600000e+00] (400.393232341897,-13.571) [2.500000e+00] (401.647874011237,-13.774) [2.500000e+00] (402.939357986557,-13.984) [2.400000e+00] (404.591714086049,-14.176) [2.400000e+00] (406.2718054043,-14.32) [2.400000e+00] (407.533939067413,-14.519) [2.400000e+00] (408.688554066189,-14.637) [2.400000e+00] (410.549851784699,-14.788) [2.300000e+00] (412.090132781494,-14.83) [2.200000e+00] (413.592872714638,-15.015) [2.200000e+00] (414.827198547541,-15.209) [2.200000e+00] (416.789364884208,-15.394) [2.200000e+00] (418.043378717052,-15.55) [2.100000e+00] (419.387464912905,-15.93) [2] (420.472873767022,-16.192) [2] (421.724427922257,-16.478) [2] (423.503552062762,-16.562) [1.900000e+00] (424.828169183932,-16.685) [1.900000e+00] (426.354087009002,-16.739) [1.900000e+00] (427.785377809461,-16.807) [1.700000e+00] (429.073076401396,-16.926) [1.700000e+00] (430.106652771831,-17.068) [1.600000e+00] (431.401219934598,-17.158) [1.600000e+00] (433.205431349417,-17.248) [1.600000e+00] (434.578866109566,-17.353) [1.500000e+00] (436.204110128149,-17.507) [1.500000e+00] (437.46701398703,-17.658) [1.400000e+00] (438.906569154545,-17.852) [1.400000e+00] (440.101808404137,-17.969) [1.500000e+00] (441.330806400502,-18.097) [1.500000e+00] (443.360487222009,-18.182) [1.400000e+00] (444.585774223355,-18.25) [1.400000e+00] (445.447251768488,-18.17) [1.400000e+00] (447.12362865369,-17.912) [1.300000e+00] (448.757908036852,-17.697) [1.400000e+00] (450.639589132766,-17.439) [1.400000e+00] (451.816205698007,-17.237) [1.300000e+00] (452.9941430965,-17.065) [1.400000e+00] (454.324782973168,-16.989) [1.400000e+00] (455.645539375936,-16.878) [1.400000e+00] (456.434723634116,-16.729) [1.500000e+00] (456.720912797513,-16.525) [1.400000e+00] (457.811644272833,-16.327) [1.600000e+00] (459.498525417922,-16.033) [1.600000e+00] (461.469577023641,-15.747) [1.600000e+00] (463.191832955192,-15.465) [1.600000e+00] (464.677192693975,-15.222) [1.600000e+00] (466.37086931058,-14.943) [1.600000e+00] (468.175076857479,-14.739) [1.600000e+00] (470.070274909105,-14.536) [1.500000e+00] (471.191898749391,-14.413) [1.600000e+00] (472.485619237177,-14.316) [1.800000e+00] (473.894352188464,-14.297) [1.800000e+00] (475.2843152821,-14.247) [1.800000e+00] (476.772760414979,-14.174) [1.900000e+00] (477.794282650237,-14.062) [1.900000e+00] (479.272147329054,-13.973) [2] (481.272548904457,-13.839) [2] (483.080465268393,-13.733) [2.100000e+00] (484.381601955964,-13.665) [2.200000e+00] (485.775969040986,-13.563) [2.300000e+00] (487.077181092446,-13.401) [2.300000e+00] (489.279028768061,-13.204) [2.400000e+00] (490.763687600189,-13.02) [2.500000e+00] (491.869402232016,-12.837) [2.500000e+00] (493.070662947826,-12.64) [2.500000e+00] (494.53802035846,-12.458) [2.700000e+00] (496.492539449225,-12.275) [2.800000e+00] (497.893005483946,-12.122) [2.800000e+00] (499.123616082126,-11.968) [2.800000e+00] (500.490249812515,-11.835) [2.900000e+00] (501.848619204718,-11.682) [2.900000e+00] (503.312963977375,-11.566) [3] (504.493136941834,-11.409) [3.100000e+00] (505.878540989002,-11.274) [3.300000e+00] (507.491032689709,-11.118) [3.400000e+00] (508.862918091176,-10.935) [3.500000e+00] (511.147488495327,-10.796) [3.800000e+00] (512.948604172493,-10.614) [4] (514.328755259145,-10.438) [4.200000e+00] (516.083989741004,-10.3) [4.400000e+00] (517.537827675343,-10.172) [4.800000e+00] (519.392317458851,-10.061) [5.200000e+00] (520.815259503113,-9.944) [5.600000e+00] (522.379553032858,-9.792) [6] (523.810063323146,-9.61) [6.400000e+00] (525.637516703518,-9.414) [6.800000e+00] (527.023691376214,-9.321) [7.300000e+00] (528.15651341941,-9.149) [7.400000e+00] (529.172082217439,-8.984) [7.700000e+00] (530.85293204054,-8.836) [7.800000e+00] (532.357130095883,-8.71) [8] (533.626051093809,-8.573) [8.200000e+00] (534.898679503406,-8.452) [8.300000e+00] (536.107429634038,-8.264) [8.200000e+00] (537.799631396382,-8.075) [8.100000e+00] (540.384899305627,-7.876) [8] (541.339910154951,-7.715) [7.700000e+00] (542.511119680717,-7.661) [7.300000e+00] (543.880839780806,-7.633) [7.100000e+00] (545.200893674116,-7.409) [6.800000e+00] (546.475838257982,-7.245) [6.600000e+00] (547.608736611513,-7.032) [6.300000e+00] (549.143057840604,-6.86) [6] (550.521736593632,-6.726) [5.800000e+00] (552.116151548903,-6.522) [5.500000e+00] (553.811287547934,-6.305) [5.400000e+00] (555.298181574116,-6.105) [5.300000e+00] (556.74073906526,-5.926) [5.100000e+00] (557.826209757301,-5.717) [4.800000e+00] (559.27401937813,-5.489) [4.600000e+00] (559.306150016763,-5.28) [4.500000e+00] (554.03698251721,-5.057) [4.200000e+00] (550.895504015747,-5.033) [4.100000e+00] (557.717371908828,-5.156) [3.900000e+00] (561.272322234517,-5.281) [3.800000e+00] (563.747903483793,-5.203) [3.600000e+00] (564.998287509892,-4.978) [3.500000e+00] (566.837092085808,-4.748) [3.300000e+00] (568.695439782222,-4.514) [3.200000e+00] (569.834361750772,-4.302) [3.100000e+00] (571.248417125722,-4.023) [2.800000e+00] (572.80529153421,-3.801) [2.600000e+00] (574.549320164143,-3.571) [2.500000e+00] (576.188068158497,-3.327) [2.400000e+00] (578.026021500922,-3.09) [2.200000e+00] (579.600891674788,-2.832) [2.200000e+00] (581.740317445334,-2.59) [2.100000e+00] (583.303910184611,-2.314) [1.900000e+00] (584.752411942598,-2.064) [1.800000e+00] (586.321176517195,-1.813) [1.800000e+00] (588.424530119685,-1.551) [1.700000e+00] (590.44747590734,-1.304) [1.700000e+00] (592.019332554757,-1.041) [1.700000e+00] (593.690365127009,-0.8) [1.600000e+00] (595.153075403151,-0.696) [1.600000e+00] (596.676036840976,-0.739) [1.500000e+00] (598.709327188053,-0.795) [1.400000e+00] (601.300749078113,-2.544) [1.300000e+00] (602.105991355403,-2.71) [1.400000e+00] (602.931707421628,-2.856) [1.400000e+00] (604.097821108785,-3.021) [1.400000e+00] (605.605105678937,-3.159) [1.500000e+00] (606.909319136346,-3.331) [1.600000e+00] (607.979110811731,-3.493) [1.600000e+00] (609.188014567845,-3.646) [1.800000e+00] (610.690046607554,-3.809) [1.900000e+00] (612.411437982475,-4.003) [2] (614.375596230447,-4.17) [2.100000e+00] (616.187897873787,-4.376) [2.400000e+00] (617.94682586952,-4.576) [2.400000e+00] (618.749668182145,-4.779) [2.600000e+00] (620.273480998677,-4.971) [2.900000e+00] (621.701663395038,-5.196) [3.300000e+00] (623.128379160792,-5.442) [3.700000e+00] (624.704017811359,-5.668) [4.100000e+00] (625.890205708112,-5.859) [4.500000e+00] (627.184768119437,-6.025) [4.900000e+00] (628.376288411573,-6.173) [5.400000e+00] (629.95416652508,-6.417) [5.900000e+00] (631.361339658358,-6.624) [6.100000e+00] (632.467039445835,-6.799) [6.500000e+00] (633.877225704956,-6.956) [6.700000e+00] (634.377821438691,-7.18) [6.900000e+00] (635.886569630999,-7.328) [7] (637.87413280914,-7.457) [7.100000e+00] (639.081327361336,-7.556) [7.200000e+00] (640.297640667434,-7.738) [7.300000e+00] (641.772475802155,-7.909) [7.300000e+00] (643.080473498388,-8.053) [7.500000e+00] (644.590686114394,-8.199) [7.400000e+00] (646.167016013546,-8.389) [7.300000e+00] (647.243442658802,-8.559) [7.300000e+00] (648.343970668986,-8.723) [7.100000e+00] (649.684333638446,-8.831) [7] (651.167431792837,-8.938) [6.700000e+00] (652.901641882458,-9.081) [6.600000e+00] (654.338248404371,-9.243) [6.300000e+00] (655.474846404656,-9.424) [6.100000e+00] (657.13915376985,-9.627) [6] (658.702740091544,-9.773) [5.900000e+00] (660.580459086924,-9.973) [5.700000e+00] (662.026020678454,-10.183) [5.400000e+00] (663.454278583737,-10.426) [5.200000e+00] (664.918608256291,-10.574) [5.100000e+00] (666.458173966181,-10.646) [4.900000e+00] (668.330721643186,-10.788) [4.600000e+00] (669.558311123957,-10.94) [4.400000e+00] (671.25188905276,-11.118) [4.200000e+00] (672.389951047011,-11.246) [4] (674.707480115744,-11.353) [3.900000e+00] (675.728374949706,-11.489) [3.700000e+00] (677.108510448648,-11.599) [3.600000e+00] (679.2440557551,-11.705) [3.500000e+00] (681.03990082214,-11.873) [3.300000e+00] (682.434326451909,-12.131) [3.200000e+00] (683.467879725312,-12.398) [3] (684.649502814835,-12.614) [2.900000e+00] (685.759066706,-12.651) [2.900000e+00] (687.378953344612,-12.667) [2.900000e+00] (689.062024894014,-12.646) [2.700000e+00] (690.203181255074,-12.758) [2.600000e+00] (691.420265703543,-12.874) [2.600000e+00] (692.68013752651,-13.063) [2.400000e+00] (694.521863875757,-13.218) [2.300000e+00] (695.88632611914,-13.466) [2.400000e+00] (697.552177888773,-13.663) [2.300000e+00] (698.993949654389,-13.954) [2.200000e+00] (699.936593147191,-14.207) [2.100000e+00] (701.709024055777,-14.439) [2.100000e+00] (703.513131741756,-14.529) [2.200000e+00] (704.763963356725,-14.689) [2.100000e+00] (706.036575719774,-14.831) [2.100000e+00] (707.20080998498,-15.004) [2.100000e+00] (708.929071683957,-15.065) [2] (710.481543569578,-15.128) [1.900000e+00] (712.215914938905,-15.152) [1.900000e+00] (713.715535694763,-15.381) [1.900000e+00] (714.842317245742,-15.592) [2] (716.293816454677,-15.756) [1.900000e+00] (717.707006268394,-15.856) [1.900000e+00] (718.838180573656,-15.98) [2] (719.965567172751,-16.117) [1.900000e+00] (721.823820049284,-16.151) [1.900000e+00] (723.697136289684,-16.202) [1.900000e+00] (725.150266789064,-16.221) [1.900000e+00] (726.538141129001,-16.297) [1.900000e+00] (727.790604063571,-16.461) [1.800000e+00] (729.143013490269,-16.715) [1.800000e+00] (730.396856512014,-16.988) [1.800000e+00] (731.709323504927,-17.227) [1.700000e+00] (733.194736747184,-17.449) [1.600000e+00] (734.552313078727,-17.626) [1.500000e+00] (736.01284813964,-17.818) [1.400000e+00] (737.570231800915,-18) [1.400000e+00] (739.057883158724,-18.103) [1.300000e+00] (740.58160404782,-18.17) [1.300000e+00] (741.891057055759,-18.245) [1.400000e+00] (743.346254819787,-18.143) [1.400000e+00] (743.715041281529,-17.892) [1.500000e+00] (743.372501862502,-17.724) [1.400000e+00] (745.983919066963,-17.728) [1.400000e+00] (747.118878011834,-17.766) [1.400000e+00] (748.204335323046,-17.714) [1.500000e+00] (749.921836001063,-17.627) [1.600000e+00] (751.85808119249,-17.399) [1.600000e+00] (753.443353552948,-17.167) [1.500000e+00] (755.020276379409,-16.882) [1.500000e+00] (756.722107327212,-16.622) [1.600000e+00] (758.217849965371,-16.404) [1.700000e+00] (759.600472320502,-16.237) [1.700000e+00] (760.890382266074,-16.048) [1.800000e+00] (762.251915034731,-15.834) [1.900000e+00] (763.556890150241,-15.676) [1.900000e+00] (764.824330880797,-15.518) [1.800000e+00] (766.085660622746,-15.363) [1.700000e+00] (767.303515412042,-15.243) [1.800000e+00] (768.707746775803,-15.12) [1.800000e+00] (770.035275901486,-14.985) [1.900000e+00] (771.45749898875,-14.849) [1.800000e+00] (772.960126365527,-14.688) [1.900000e+00] (774.436410707631,-14.525) [2] (775.964604761534,-14.353) [2] (777.462755076765,-14.189) [1.900000e+00] (778.832378667063,-14.032) [1.900000e+00] (780.093793662851,-13.878) [1.800000e+00] (781.446113566901,-13.752) [1.900000e+00] (782.834503461074,-13.586) [1.800000e+00] (784.288319505995,-13.497) [1.800000e+00] (785.599318236394,-13.405) [1.900000e+00] (786.924607986195,-13.283) [2.100000e+00] (788.248433341269,-13.18) [2.200000e+00] (789.524138038227,-13.07) [2.400000e+00] (790.91709024823,-12.952) [2.500000e+00] (792.288176484898,-12.877) [2.500000e+00] (793.366829955702,-12.776) [2.600000e+00] (794.642446761998,-12.629) [2.800000e+00] (796.212734421098,-12.47) [3] (797.200481206599,-12.267) [3.100000e+00] (798.647581243616,-12.004) [3.200000e+00] (800.583991269068,-11.717) [3.500000e+00] (802.90744059022,-11.483) [3.700000e+00] (804.338700017224,-11.325) [3.800000e+00] (806.056278527151,-11.18) [4.200000e+00] (807.985111391972,-11.039) [4.300000e+00] (809.40285440908,-10.905) [4.600000e+00] (810.612357085689,-10.72) [4.800000e+00] (811.859479372142,-10.588) [5] (813.389994187876,-10.469) [5.300000e+00] (814.731034966986,-10.353) [5.900000e+00] (816.271277168218,-10.164) [6.400000e+00] (817.730428419094,-10.007) [6.800000e+00] (819.284273172684,-9.883) [7] (820.654668793693,-9.801) [7.300000e+00] (822.52108780309,-9.643) [7.500000e+00] (824.253041726866,-9.508) [7.800000e+00] (824.840833412025,-9.375) [8.100000e+00] (825.621798949205,-9.168) [8.300000e+00] (826.630547947332,-8.977) [8.400000e+00] (831.387618680305,-8.688) [8.100000e+00] (832.009593889613,-8.736) [8] (833.13129049664,-8.678) [7.700000e+00] (834.55738448232,-8.491) [7.300000e+00] (836.29381328966,-8.28) [7.100000e+00] (837.783693679666,-8.09) [6.800000e+00] (838.642402030511,-7.945) [6.500000e+00] (839.946772788831,-7.77) [6.200000e+00] (841.338348668701,-7.611) [6] (842.878677529188,-7.483) [5.800000e+00] (844.115207810356,-7.353) [5.600000e+00] (845.405796651774,-7.164) [5.400000e+00] (847.348138962983,-6.937) [5.300000e+00] (849.439057126407,-6.71) [5.100000e+00] (851.269312551894,-6.508) [4.800000e+00] (852.633501755337,-6.361) [4.500000e+00] (853.946733414982,-6.169) [4.400000e+00] (855.284935676456,-5.981) [4.200000e+00] (856.693903191344,-5.759) [4] (857.946273251773,-5.559) [3.800000e+00] (859.17385121571,-5.328) [3.600000e+00] (860.447224319571,-5.155) [3.400000e+00] (861.974037716558,-4.943) [3.300000e+00] (863.154179017472,-4.784) [3.100000e+00] (864.559686758026,-4.588) [2.900000e+00] (865.969848287846,-4.383) [2.900000e+00] (867.286867711739,-4.149) [2.800000e+00] (868.260319609999,-3.964) [2.600000e+00] (869.90360330309,-3.753) [2.500000e+00] (871.737734871017,-3.54) [2.400000e+00] (873.290023959297,-3.36) [2.400000e+00] (875.221085368602,-3.144) [2.400000e+00] (876.822958650073,-2.923) [2.300000e+00] (878.168558335542,-2.678) [2.200000e+00] (879.80133966195,-2.437) [2.100000e+00] (881.004900626457,-2.218) [2] (882.346711287478,-1.994) [1.900000e+00] (883.697561755581,-1.767) [1.700000e+00] (885.452750792101,-1.535) [1.600000e+00] (887.555980861665,-1.306) [1.500000e+00] (889.205200278853,-1.061) [1.600000e+00] (890.644715820389,-0.827) [1.600000e+00] (892.012869115298,-0.748) [1.600000e+00] (893.553879938004,-0.801) [1.500000e+00] (894.906967633035,-0.868) [1.500000e+00] (896.429897494422,-0.934) [1.500000e+00] (898.206359610865,-0.995) [1.300000e+00] (899.742210724292,-1.063) [1.300000e+00] (901.288470982272,-1.254) [1.200000e+00] (903.073700720065,-1.51) [1.100000e+00] (904.056099232407,-1.782) [1.200000e+00] (905.661577359442,-2.038) [1.300000e+00] (907.309240966211,-2.306) [1.400000e+00] (909.364346881288,-2.578) [1.500000e+00] (910.446869560287,-2.838) [1.600000e+00] (911.88862146923,-3.104) [1.700000e+00] (913.17326682145,-3.367) [1.900000e+00] (914.652547754957,-3.618) [2.100000e+00] (916.223188387228,-3.881) [2.100000e+00] (916.929453947779,-4.175) [2.300000e+00] (949.594894040151,-9.763) [7.400000e+00] (949.808442506302,-9.812) [7.200000e+00] (954.119267223631,-10.245) [6.100000e+00] (955.077367636151,-10.285) [6] (955.870193070796,-10.436) [5.900000e+00] (956.973622504513,-10.602) [5.700000e+00] (957.974781187876,-10.738) [5.400000e+00] (959.174455579098,-10.807) [5.200000e+00] (960.59579086236,-10.891) [5.100000e+00] (962.045024189117,-10.985) [4.900000e+00] (963.425911060813,-11.064) [4.800000e+00] (964.813321619232,-11.136) [4.700000e+00] (966.231642458392,-11.247) [4.600000e+00] (966.694102289736,-11.395) [4.400000e+00] (967.312982504803,-11.498) [4.200000e+00] (968.686475163626,-11.494) [4] (970.423758967149,-11.394) [3.800000e+00] (971.967186212811,-11.307) [3.700000e+00] (973.608225517697,-11.336) [3.600000e+00] (975.241000484524,-11.459) [3.600000e+00] (976.294257303585,-11.632) [3.500000e+00] (977.599308908957,-11.904) [3.500000e+00] (978.903584586514,-12.164) [3.400000e+00] (980.006235914641,-12.393) [3.200000e+00] (981.364566598335,-12.584) [3.100000e+00] (982.397411548232,-12.729) [2.900000e+00] (983.378530622847,-12.912) [2.900000e+00] (989.79094801466,-13.235) [2.600000e+00] (995.224455677119,-13.41) [2.500000e+00] (1000.14019726013,-13.706) [2.400000e+00] (1002.25207355658,-14.247) [2.200000e+00] (1000.65341649522,-14.931) [2.200000e+00] (997.295887116171,-15.812) [2.200000e+00] (994.919876553112,-16.289) [2.100000e+00] (994.332094927043,-16.622) [2.100000e+00] (995.485103280614,-16.682) [2.100000e+00] (996.923150762594,-16.607) [2] (998.196112979616,-16.548) [1.900000e+00] (1000.67604527668,-17.357) [1.500000e+00] (1001.93259553678,-17.288) [1.500000e+00] (1002.22561334395,-17.297) [1.600000e+00] (1003.20868579264,-17.338) [1.600000e+00] (1004.78410622502,-17.339) [1.600000e+00] (1005.48465970795,-17.355) [1.600000e+00] (1005.85593613862,-17.341) [1.500000e+00] (1007.06785016757,-17.295) [1.600000e+00] (1008.06297600792,-17.305) [1.500000e+00] (1008.80348316173,-17.345) [1.600000e+00] (1009.85654903618,-17.359) [1.600000e+00] (1011.20282240245,-17.344) [1.600000e+00] (1012.60324586046,-17.296) [1.600000e+00] (1014.21473272895,-17.231) [1.600000e+00] (1015.62342006652,-17.223) [1.600000e+00] (1017.13758234084,-17.182) [1.700000e+00] (1018.17431059987,-17.194) [1.700000e+00] (1019.64239789447,-17.222) [1.700000e+00] (1021.02026602436,-17.262) [1.600000e+00] (1022.49030262215,-17.315) [1.700000e+00] (1023.69162054579,-17.364) [1.600000e+00] (1024.8659118696,-17.41) [1.500000e+00] (1026.33623415529,-17.439) [1.500000e+00] (1027.80879272144,-17.5) [1.500000e+00] (1029.3551368915,-17.576) [1.500000e+00] (1030.66543718858,-17.673) [1.500000e+00] (1031.78538741735,-17.819) [1.400000e+00] (1033.27379069361,-17.983) [1.300000e+00] (1034.9056886297,-18.147) [1.400000e+00] (1036.38719108639,-18.326) [1.500000e+00] (1038.29470453214,-18.468) [1.500000e+00] (1039.41241759212,-18.662) [1.500000e+00] (1039.85018497153,-18.872) [1.500000e+00] (1041.09272505679,-18.793) [1.500000e+00] (1041.98129417478,-18.667) [1.500000e+00] (1043.34030054281,-18.513) [1.500000e+00] (1044.48960879633,-18.295) [1.500000e+00] (1045.85843405255,-18.043) [1.600000e+00] (1047.17096979954,-17.765) [1.600000e+00] (1048.3632461971,-17.507) [1.600000e+00] (1049.54725960384,-17.334) [1.600000e+00] (1050.59594536694,-17.217) [1.600000e+00] (1051.87085792077,-17.119) [1.500000e+00] (1052.99090194634,-17.03) [1.500000e+00] (1054.12971004,-16.957) [1.500000e+00] (1055.32266730831,-16.84) [1.500000e+00] (1056.72979576895,-16.701) [1.500000e+00] (1058.11738334515,-16.545) [1.500000e+00] (1059.3733385923,-16.392) [1.500000e+00] (1060.89401664355,-16.143) [1.600000e+00] (1062.18623248286,-15.915) [1.600000e+00] (1063.40329834032,-15.696) [1.600000e+00] (1064.57895149249,-15.546) [1.500000e+00] (1065.70579996237,-15.429) [1.600000e+00] (1066.97701680064,-15.293) [1.600000e+00] (1068.1782366588,-15.225) [1.600000e+00] (1069.49067382348,-15.171) [1.700000e+00] (1070.63327145846,-15.094) [1.700000e+00] (1071.61050304224,-14.96) [1.800000e+00] (1072.80938929815,-14.838) [1.800000e+00] (1073.96248623326,-14.688) [1.800000e+00] (1075.20648142696,-14.503) [1.700000e+00] (1076.64150979507,-14.343) [1.800000e+00] (1078.19601714235,-14.142) [1.900000e+00] (1079.76851039381,-13.994) [1.900000e+00] (1081.18623283987,-13.781) [2] (1082.59792771605,-13.623) [2.100000e+00] (1084.10431214011,-13.514) [2.100000e+00] (1085.90234959274,-13.381) [2.100000e+00] (1087.05398663402,-13.231) [2.200000e+00] (1088.57087081702,-13.053) [2.100000e+00] (1090.09008725511,-12.9) [2.100000e+00] (1091.4769918407,-12.772) [2.100000e+00] (1092.84513281712,-12.649) [2.200000e+00] (1094.24341218875,-12.522) [2.300000e+00] (1095.51394781334,-12.438) [2.400000e+00] (1096.1619996381,-12.383) [2.600000e+00] (1097.01206795742,-12.425) [2.700000e+00] (1099.014838651,-12.447) [2.900000e+00] (1100.39396114136,-12.309) [3.400000e+00] (1102.24594919868,-11.997) [3.800000e+00] (1104.00344501749,-11.681) [4.100000e+00] (1105.51128452111,-11.427) [4.400000e+00] (1107.12577098928,-11.181) [4.800000e+00] (1107.86715291421,-10.985) [5.200000e+00] (1109.350979832,-10.768) [5.500000e+00] (1110.98977308767,-10.523) [5.700000e+00] (1112.74367271982,-10.308) [6] (1114.19308776837,-10.146) [6.400000e+00] (1115.51086868174,-10.013) [6.800000e+00] (1116.88056518392,-9.896) [7.200000e+00] (1118.40123673035,-9.858) [7.700000e+00] (1119.77161229719,-9.812) [8.100000e+00] (1120.82456904281,-9.688) [8.500000e+00] (1121.82123851495,-9.578) [8.800000e+00] (1122.88575493397,-9.357) [9] (1124.42451016549,-9.221) [9.300000e+00] (1126.24355474329,-9.142) [9] (1127.67945653535,-8.966) [9] (1129.14763301534,-8.868) [9.100000e+00] (1130.56612978604,-8.765) [9] (1131.53665172696,-8.636) [8.900000e+00] (1132.91976023767,-8.484) [8.800000e+00] (1134.07684413984,-8.301) [8.500000e+00] (1135.01780587406,-8.116) [8.200000e+00] (1136.86260253337,-7.922) [7.700000e+00] (1143.78281335461,-7.406) [5.900000e+00] (1144.95183244868,-7.148) [5.700000e+00] (1147.24158909297,-6.876) [5.400000e+00] (1147.44895053885,-6.646) [5.200000e+00] (1146.53928710166,-6.497) [5] (1145.94752157509,-6.361) [4.900000e+00] (1147.76499124173,-6.235) [4.700000e+00] (1149.71565228853,-6.108) [4.600000e+00] (1151.26402471399,-5.943) [4.400000e+00] (1152.86976213877,-5.753) [4.200000e+00] (1154.37011162885,-5.579) [4] (1155.88854477508,-5.444) [3.800000e+00] (1157.56709413638,-5.27) [3.500000e+00] (1158.84511433255,-5.088) [3.400000e+00] (1160.10796309328,-4.93) [3.100000e+00] (1161.46492281255,-4.745) [3] (1162.9720785239,-4.565) [2.800000e+00] (1164.44841634584,-4.378) [2.600000e+00] (1165.9428369353,-4.157) [2.400000e+00] (1167.50035013104,-3.934) [2.300000e+00] (1169.04600948072,-3.706) [2.200000e+00] (1170.1113188371,-3.482) [2.100000e+00] (1171.51531979886,-3.269) [2] (1172.97279402392,-3.059) [1.900000e+00] (1174.45204455426,-2.837) [1.800000e+00] (1176.02296795244,-2.635) [1.700000e+00] (1177.56940506485,-2.405) [1.700000e+00] (1179.12691785991,-2.181) [1.400000e+00] (1179.88384728466,-1.953) [1.400000e+00] (1180.87157080448,-1.718) [1.400000e+00] (1182.22230031307,-1.514) [1.300000e+00] (1183.39191592362,-1.309) [1.200000e+00] (1183.84659478021,-1.093) [1.200000e+00] (1186.07762491598,-0.919) [1.200000e+00] (1188.46440227703,-0.789) [1.200000e+00] (1190.661499174,-0.73) [1.200000e+00] (1192.57219419704,-0.788) [1.300000e+00] (1195.23748059964,-0.842) [1.300000e+00] (1196.96939848037,-0.906) [1.300000e+00] (1198.78337303955,-0.975) [1.300000e+00] (1200.24250369573,-1.105) [1.300000e+00] (1201.64145577211,-1.351) [1.300000e+00] (1203.01707432754,-1.622) [1.300000e+00] (1204.39648551966,-1.878) [1.400000e+00] (1205.68266482191,-2.171) [1.500000e+00] (1206.99508854765,-2.433) [1.500000e+00] (1208.22555873589,-2.689) [1.600000e+00] (1209.34335883212,-2.959) [1.700000e+00] (1210.60464559154,-3.221) [1.600000e+00] (1211.8946179894,-3.473) [1.600000e+00] (1213.3189327837,-3.719) [1.700000e+00] (1214.61784457113,-3.975) [1.800000e+00] (1216.00619472316,-4.227) [1.700000e+00] (1217.34652371687,-4.479) [1.900000e+00] (1218.87097976344,-4.721) [2] (1220.7248737587,-4.977) [2.100000e+00] (1221.98626234077,-5.247) [2.200000e+00] (1222.78938611222,-5.557) [2.400000e+00] (1223.98009727317,-5.796) [2.600000e+00] (1224.93486893376,-6.018) [3.100000e+00] (1225.9325136813,-6.22) [3.500000e+00] (1227.33447808197,-6.436) [3.800000e+00] (1228.50195886306,-6.668) [4.200000e+00] (1229.63238698508,-6.932) [4.700000e+00] (1230.80433677276,-7.124) [5.100000e+00] (1231.89511164539,-7.285) [5.600000e+00] (1233.07667630422,-7.435) [6.200000e+00] (1234.66354595031,-7.585) [6.500000e+00] (1236.33964681948,-7.713) [7] (1238.30573416926,-7.911) [7.400000e+00] (1239.19284199324,-8.064) [7.800000e+00] (1240.2670895661,-8.168) [8.200000e+00] (1241.95088108564,-8.293) [8.400000e+00] (1243.16793852173,-8.474) [8.600000e+00] (1244.56756398881,-8.701) [8.800000e+00] (1246.10485348372,-8.901) [8.900000e+00] (1247.40745218822,-9.047) [9] (1248.77703927679,-9.187) [8.900000e+00] (1250.789999579,-9.341) [8.900000e+00] (1251.89563609244,-9.476) [8.600000e+00] (1253.28486677653,-9.659) [8.300000e+00] (1254.62955908615,-9.837) [8] (1256.26075601036,-10.015) [7.600000e+00] (1257.4101500688,-10.185) [7.300000e+00] (1258.91438674125,-10.325) [7] (1260.31099690458,-10.502) [6.700000e+00] (1261.60688989225,-10.664) [6.500000e+00] (1263.21350636205,-10.809) [6.300000e+00] (1264.71685952279,-10.954) [6] (1265.98338961353,-11.19) [5.700000e+00] (1266.73226112732,-11.393) [5.500000e+00] (1267.7071425454,-11.554) [5.300000e+00] (1269.72903828236,-11.626) [5] (1270.27325560719,-11.698) [4.800000e+00] (1271.18437252059,-11.796) [4.700000e+00] (1272.74691494272,-11.9) [4.500000e+00] (1274.05309607591,-11.959) [4.400000e+00] (1276.06158390439,-11.954) [4.200000e+00] (1278.22795215536,-11.968) [3.800000e+00] (1280.0078803631,-12.01) [3.700000e+00] (1281.5564452019,-12.113) [3.500000e+00] (1283.19143353155,-12.287) [3.300000e+00] (1285.03008670274,-12.514) [3.200000e+00] (1286.80476550434,-12.713) [3.100000e+00] (1288.14353123638,-12.931) [2.900000e+00] (1289.43484858004,-13.167) [2.900000e+00] (1290.8908552334,-13.364) [2.900000e+00] (1292.49112588285,-13.555) [2.800000e+00] (1294.02986696077,-13.741) [2.700000e+00] (1294.75384418361,-13.958) [2.600000e+00] (1295.804755227,-14.182) [2.500000e+00] (1297.13000945056,-14.372) [2.400000e+00] (1298.48217956594,-14.497) [2.300000e+00] (1299.9620885554,-14.599) [2.200000e+00] (1301.49033216845,-14.715) [2.200000e+00] (1302.84349389597,-14.882) [2.200000e+00] (1304.19977636202,-15.022) [2.200000e+00] (1305.4980057917,-15.231) [2] (1307.01788060733,-15.459) [1.900000e+00] (1308.40622256463,-15.637) [1.900000e+00] (1309.7847402734,-15.805) [1.800000e+00] (1311.17542256977,-15.963) [1.800000e+00] (1312.40364871271,-16.098) [1.700000e+00] (1313.88298765571,-16.257) [1.600000e+00] (1315.07436739274,-16.415) [1.600000e+00] (1316.41458462095,-16.57) [1.400000e+00] (1317.71001305514,-16.709) [1.400000e+00] (1318.98558132077,-16.852) [1.400000e+00] (1319.83116048681,-17.064) [1.400000e+00] (1320.98959498518,-17.268) [1.400000e+00] (1322.46882723301,-17.461) [1.400000e+00] (1323.97362799299,-17.56) [1.300000e+00] (1325.32221361779,-17.597) [1.300000e+00] (1326.85627103847,-17.614) [1.300000e+00] (1328.30177718925,-17.633) [1.300000e+00] (1329.71043785175,-17.698) [1.300000e+00] (1331.10827995892,-17.768) [1.300000e+00] (1332.6523732995,-17.822) [1.400000e+00] (1334.47783129625,-17.829) [1.400000e+00] (1335.85120592223,-17.945) [1.400000e+00] (1337.10688025979,-18.177) [1.500000e+00] (1338.26063147237,-18.392) [1.500000e+00] (1339.68323188759,-18.67) [1.300000e+00] (1339.9581389525,-18.799) [1.300000e+00] (1340.67732573557,-18.784) [1.300000e+00] (1341.21753081328,-18.63) [1.300000e+00] (1342.94957195777,-18.381) [1.300000e+00] (1344.40323176751,-18.093) [1.400000e+00] (1346.17052779674,-17.84) [1.400000e+00] (1347.9066835279,-17.536) [1.400000e+00] (1349.55060342417,-17.276) [1.400000e+00] (1351.14047768772,-17.017) [1.400000e+00] (1352.44473305526,-16.732) [1.500000e+00] (1353.96839755711,-16.419) [1.500000e+00] (1355.50635264854,-16.138) [1.500000e+00] (1357.11129881878,-15.86) [1.500000e+00] (1358.53513639307,-15.575) [1.500000e+00] (1359.70261216194,-15.361) [1.500000e+00] (1361.2488302731,-15.218) [1.600000e+00] (1362.58088957364,-15.111) [1.600000e+00] (1363.61236525458,-15.043) [1.600000e+00] (1364.85041088957,-15.045) [1.600000e+00] (1366.07036971485,-15.082) [1.500000e+00] (1367.13400802831,-15.091) [1.500000e+00] (1368.40588516614,-15.031) [1.500000e+00] (1369.5347367272,-14.91) [1.600000e+00] (1370.90576622876,-14.792) [1.600000e+00] (1372.24140296951,-14.659) [1.600000e+00] (1373.52700687035,-14.42) [1.600000e+00] (1375.30111058419,-14.254) [1.600000e+00] (1376.63384049369,-14.122) [1.500000e+00] (1377.83404143197,-14.026) [1.500000e+00] (1378.87590570485,-13.925) [1.600000e+00] (1379.9274857251,-13.773) [1.500000e+00] (1381.43852547062,-13.578) [1.600000e+00] (1382.91104905107,-13.409) [1.500000e+00] (1384.29569773182,-13.288) [1.500000e+00] (1385.56679272203,-13.196) [1.700000e+00] (1386.99130139299,-13.068) [1.600000e+00] (1388.34524157429,-12.941) [1.700000e+00] (1389.21124576909,-12.901) [1.800000e+00] (1389.98334522926,-12.9) [1.800000e+00] (1391.0488726273,-12.801) [2] (1393.74950605546,-12.776) [2.400000e+00] (1394.97002432265,-12.698) [2.700000e+00] (1396.22504644985,-12.473) [2.900000e+00] (1398.05555037586,-12.144) [3.200000e+00] (1399.1181873029,-11.824) [3.600000e+00] (1401.64920663085,-11.48) [4.200000e+00] (1403.44642385463,-11.258) [4.600000e+00] (1404.68926236817,-11.005) [4.800000e+00] (1405.90709396026,-10.804) [5.100000e+00] (1407.43600136379,-10.614) [5.400000e+00] (1409.00622151415,-10.435) [5.600000e+00] (1410.64722942181,-10.298) [6] (1411.94634451724,-10.143) [6.400000e+00] (1413.21297101755,-10.041) [6.700000e+00] (1414.46385239472,-9.968) [7] (1415.92431418593,-9.906) [7.200000e+00] (1417.36155049859,-9.807) [7.200000e+00] (1418.68221953631,-9.675) [7.200000e+00] (1420.20967378253,-9.48) [7.300000e+00] (1421.67091600235,-9.317) [7.200000e+00] (1422.98923979945,-9.171) [7.200000e+00] (1424.22582926871,-8.984) [7.200000e+00] (1425.78477066576,-8.755) [6.900000e+00] (1427.58712355013,-8.535) [6.600000e+00] (1429.05573673518,-8.346) [6.300000e+00] (1430.79142762825,-8.097) [6.200000e+00] (1432.37147033703,-7.846) [5.900000e+00] (1433.98701834985,-7.676) [5.600000e+00] (1435.64912808587,-7.556) [5.500000e+00] (1437.12320908543,-7.416) [5.300000e+00] (1438.75282710198,-7.268) [5.200000e+00] (1440.11402407316,-7.118) [5] (1441.70590362958,-6.951) [4.800000e+00] (1443.1476036757,-6.699) [4.700000e+00] (1444.73378655567,-6.53) [4.600000e+00] (1446.3830544746,-6.366) [4.500000e+00] (1447.90001042533,-6.182) [4.400000e+00] (1449.34405574091,-5.969) [4.400000e+00] (1450.64607334099,-5.769) [4.200000e+00] (1452.09670606906,-5.551) [4.100000e+00] (1453.60740936325,-5.32) [4] (1454.97072700506,-5.149) [3.900000e+00] (1456.47640506205,-4.964) [3.700000e+00] (1458.17245669575,-4.764) [3.500000e+00] (1459.46085205005,-4.569) [3.300000e+00] (1460.82818876259,-4.343) [3] (1462.50827288093,-4.125) [2.800000e+00] (1463.84244686445,-3.909) [2.600000e+00] (1465.40149558962,-3.698) [2.400000e+00] (1466.52676194585,-3.463) [2.300000e+00] (1467.92871193727,-3.244) [2.100000e+00] (1469.68617260176,-3.026) [2.100000e+00] (1471.11201506838,-2.817) [2.100000e+00] (1472.59748306645,-2.601) [2] (1474.07546982089,-2.369) [1.800000e+00] (1475.78156797342,-2.13) [1.700000e+00] (1477.61472914717,-1.899) [1.600000e+00] (1467.60446488409,-18.751) [1.100000e+00] (1466.58437849054,-18.578) [1.100000e+00] (1465.49115594547,-18.426) [1.200000e+00] (1464.4061968499,-18.23) [1.100000e+00] (1463.37929844506,-18.173) [1.100000e+00] (1462.43658902727,-18.086) [1.200000e+00] (1461.23204683625,-17.887) [1.200000e+00] (1459.85287508218,-17.681) [1.200000e+00] (1458.36015006729,-17.466) [1.200000e+00] (1456.86731162039,-17.203) [1.300000e+00] (1455.5192925405,-16.868) [1.300000e+00] (1454.31419145351,-16.542) [1.300000e+00] (1450.64071654329,-15.98) [1.500000e+00] (1448.138509899,-15.593) [1.600000e+00] (1446.51134828992,-15.316) [1.600000e+00] (1444.69694014989,-14.952) [1.700000e+00] (1442.94628759967,-14.61) [1.700000e+00] (1441.41503928239,-14.241) [1.700000e+00] (1440.04692088087,-13.935) [1.700000e+00] (1438.41953602032,-13.647) [1.900000e+00] (1436.88002474145,-13.396) [2.100000e+00] (1435.2195896949,-13.153) [2.100000e+00] (1433.90182745959,-12.964) [2.300000e+00] (1432.13029292342,-12.785) [2.500000e+00] (1430.76217231493,-12.641) [2.600000e+00] (1429.47600842649,-12.489) [2.700000e+00] (1428.02749550952,-12.368) [2.700000e+00] (1426.41976101012,-12.238) [2.700000e+00] (1424.70595225822,-12.071) [2.700000e+00] (1423.23633762869,-11.893) [2.700000e+00] (1421.55479823343,-11.719) [2.800000e+00] (1420.07479730604,-11.573) [2.900000e+00] (1417.8192306451,-11.462) [3.100000e+00] (1415.249793885,-11.356) [3.100000e+00] (1413.73606909636,-11.402) [3.400000e+00] (1413.25438261309,-11.716) [3.400000e+00] (1413.83120554272,-12.001) [3.500000e+00] (1413.97837015015,-12.097) [3.700000e+00] (1413.27080885722,-12.012) [3.900000e+00] (1412.51813951515,-11.777) [4.200000e+00] (1411.39130711559,-11.52) [4.600000e+00] (1410.17649085623,-11.307) [4.900000e+00] (1408.79720226103,-11.078) [5.300000e+00] (1407.37213280066,-10.896) [5.700000e+00] (1406.17462295122,-10.718) [6] (1405.24084633161,-10.546) [6.600000e+00] (1403.56767805647,-10.304) [7] (1402.06299671626,-10.116) [7.300000e+00] (1400.53029312182,-9.965) [7.600000e+00] (1399.1975649316,-9.821) [7.900000e+00] (1397.75128348039,-9.642) [7.900000e+00] (1396.53501406379,-9.446) [7.900000e+00] (1395.05724325484,-9.277) [8] (1393.55691886572,-9.113) [7.800000e+00] (1392.22865747505,-8.993) [7.600000e+00] (1390.74641900698,-8.828) [7.500000e+00] (1389.36177164464,-8.623) [7.200000e+00] (1387.89673019638,-8.443) [7] (1386.15589676663,-8.302) [6.600000e+00] (1384.65177529787,-8.161) [6.400000e+00] (1383.41808922181,-7.995) [6.100000e+00] (1382.23933724953,-7.785) [5.900000e+00] (1380.6360629875,-7.596) [5.800000e+00] (1378.8869657072,-7.399) [5.500000e+00] (1377.46781319561,-7.191) [5.200000e+00] (1376.0102530205,-6.949) [5.100000e+00] (1374.66088951669,-6.727) [4.800000e+00] (1373.08619735126,-6.59) [4.700000e+00] (1371.94585264192,-6.422) [4.500000e+00] (1370.66940066178,-6.242) [4.400000e+00] (1369.69576279155,-6.048) [4.300000e+00] (1368.43214999165,-5.859) [4.100000e+00] (1367.11750557345,-5.68) [4] (1365.82687018668,-5.514) [3.800000e+00] (1358.0491112121,-4.428) [3] (1356.91412223239,-4.282) [2.800000e+00] (1355.4671650152,-4.083) [2.700000e+00] (1354.08631040554,-3.901) [2.600000e+00] (1352.36802622931,-3.685) [2.500000e+00] (1350.49576365711,-3.505) [2.400000e+00] (1349.08263794274,-3.322) [2.300000e+00] (1347.21473544621,-3.093) [2.200000e+00] (1345.5444662084,-2.864) [2.200000e+00] (1343.97647319727,-2.6) [2] (1342.33177247173,-2.367) [1.900000e+00] (1340.28900331513,-2.124) [1.800000e+00] (1338.82317714085,-1.871) [1.800000e+00] (1336.97346597111,-1.607) [1.600000e+00] (1335.13113023837,-1.339) [1.700000e+00] (1333.24893362696,-1.07) [1.600000e+00] (1331.79807220526,-0.794) [1.600000e+00] (1330.28412264903,-0.73) [1.600000e+00] (1328.05924709374,-0.783) [1.600000e+00] (1326.7399172119,-0.846) [1.500000e+00] (1325.14567940565,-0.905) [1.500000e+00] (1323.7947525533,-0.995) [1.500000e+00] (1321.92305171687,-1.226) [1.500000e+00] (1320.43321613918,-1.499) [1.500000e+00] (1319.30113144111,-1.77) [1.600000e+00] (1318.25993743423,-2.048) [1.600000e+00] (1316.62193764001,-2.31) [1.700000e+00] (1315.36133409805,-2.583) [1.700000e+00] (1314.59515654598,-2.852) [1.800000e+00] (1313.05049450631,-3.124) [1.800000e+00] (1311.33911983747,-3.365) [1.900000e+00] (1310.13344862224,-3.644) [2] (1309.00895235099,-3.905) [2.200000e+00] (1307.89864040397,-4.17) [2.300000e+00] (1306.45245882977,-4.393) [2.400000e+00] (1305.17454847706,-4.631) [2.400000e+00] (1304.16115840331,-4.913) [2.500000e+00] (1303.13269599313,-5.173) [2.600000e+00] (1301.98028958134,-5.379) [2.700000e+00] (1300.21096728202,-5.571) [2.900000e+00] (1299.56582512303,-5.78) [3.200000e+00] (1298.72684237099,-6.01) [3.500000e+00] (1296.73062616818,-6.193) [4] (1295.25965387372,-6.341) [4.500000e+00] (1294.18696787251,-6.551) [5.100000e+00] (1292.99111826921,-6.768) [5.600000e+00] (1291.57429744316,-6.969) [6.100000e+00] (1289.70883195093,-7.141) [6.600000e+00] (1288.40010416754,-7.292) [6.900000e+00] (1286.97860315129,-7.521) [7.300000e+00] (1285.72090744522,-7.685) [7.600000e+00] (1284.36483533216,-7.817) [7.800000e+00] (1282.95539129053,-7.971) [7.900000e+00] (1281.72705808084,-8.094) [7.900000e+00] (1280.01848277533,-8.247) [8.100000e+00] (1278.56704875759,-8.42) [8.300000e+00] (1276.54536417516,-8.63) [8.300000e+00] (1275.55083824411,-8.945) [8.100000e+00] (1274.74490616559,-9.244) [8.100000e+00] (1274.4520582803,-9.45) [7.800000e+00] (1272.55857732872,-9.553) [7.600000e+00] (1270.99951456589,-9.636) [7.400000e+00] (1269.97562573783,-9.723) [7.200000e+00] (1268.23030811423,-9.793) [7] (1267.44012253526,-9.92) [6.800000e+00] (1265.98556826663,-10.057) [6.600000e+00] (1264.77687956989,-10.214) [6.400000e+00] (1263.81093573556,-10.345) [6.200000e+00] (1262.33684147126,-10.419) [6] (1260.453389761,-10.418) [5.800000e+00] (1259.69246261962,-10.516) [5.600000e+00] (1258.29517832209,-10.658) [5.300000e+00] (1256.87377918654,-10.789) [5.200000e+00] (1255.54628833979,-10.886) [5] (1253.79270600637,-10.99) [4.800000e+00] (1252.19554818548,-11.13) [4.700000e+00] (1250.33755918074,-11.308) [4.500000e+00] (1248.78384727515,-11.516) [4.400000e+00] (1247.65300567139,-11.736) [4.100000e+00] (1246.14731314699,-11.923) [3.900000e+00] (1244.15690734271,-12.099) [3.700000e+00] (1242.85254201573,-12.238) [3.700000e+00] (1241.5257249022,-12.453) [3.600000e+00] (1239.70827050957,-12.646) [3.400000e+00] (1238.47401072197,-12.812) [3.300000e+00] (1237.18102511146,-12.951) [3.200000e+00] (1235.77838608894,-13.092) [3] (1234.92724201865,-13.271) [2.800000e+00] (1233.25862419952,-13.427) [2.800000e+00] (1231.69373668181,-13.516) [2.700000e+00] (1230.16413608929,-13.684) [2.600000e+00] (1228.80067933978,-13.848) [2.500000e+00] (1227.52879323466,-13.989) [2.300000e+00] (1225.76169411969,-14.134) [2.300000e+00] (1224.39064988976,-14.318) [2.200000e+00] (1223.29374430628,-14.448) [2.100000e+00] (1221.75655623334,-14.598) [2] (1220.28178216083,-14.719) [1.900000e+00] (1218.84754398989,-14.884) [1.800000e+00] (1217.48320418753,-15.031) [1.600000e+00] (1216.02864650973,-15.162) [1.700000e+00] (1214.52305365948,-15.266) [1.700000e+00] (1213.09629942697,-15.383) [1.600000e+00] (1211.73866451105,-15.477) [1.600000e+00] (1210.3088953757,-15.589) [1.600000e+00] (1208.98509239242,-15.678) [1.600000e+00] (1207.56514484015,-15.779) [1.600000e+00] (1206.20984579138,-15.881) [1.500000e+00] (1204.76063891286,-15.962) [1.300000e+00] (1203.44821429867,-16.075) [1.400000e+00] (1202.04255779459,-16.209) [1.500000e+00] (1200.58129397006,-16.332) [1.500000e+00] (1199.21326203556,-16.458) [1.500000e+00] (1197.77980076775,-16.585) [1.600000e+00] (1196.51695501576,-16.73) [1.600000e+00] (1195.36901388762,-16.876) [1.500000e+00] (1193.99942338997,-16.995) [1.600000e+00] (1192.64110946744,-17.121) [1.600000e+00] (1191.25197854405,-17.236) [1.600000e+00] (1189.8470998028,-17.356) [1.600000e+00] (1188.55489161248,-17.476) [1.700000e+00] (1187.30555785631,-17.587) [1.600000e+00] (1185.71956521724,-17.698) [1.700000e+00] (1184.48899151613,-17.768) [1.700000e+00] (1183.20504677524,-17.891) [1.700000e+00] (1181.87006597298,-17.857) [1.600000e+00] (1180.51408748793,-17.71) [1.700000e+00] (1179.21527491556,-17.532) [1.700000e+00] (1177.55267730216,-17.397) [1.700000e+00] (1175.8991208432,-17.283) [1.800000e+00] (1174.75798642428,-17.231) [1.900000e+00] (1174.10827590382,-17.186) [1.800000e+00] (1173.41559139117,-17.132) [1.700000e+00] (1171.92049305529,-16.974) [1.700000e+00] (1170.73736461812,-16.889) [1.700000e+00] (1169.62044235932,-16.756) [1.700000e+00] (1168.32901196076,-16.614) [1.800000e+00] (1167.16833665748,-16.422) [1.800000e+00] (1166.17759826385,-16.263) [1.800000e+00] (1164.65100297407,-16.072) [1.900000e+00] (1163.09047622275,-15.905) [1.800000e+00] (1161.57194348978,-15.672) [1.800000e+00] (1160.0347489412,-15.497) [1.800000e+00] (1158.69966508499,-15.299) [1.900000e+00] (1156.7492062204,-15.122) [2] (1155.41577959662,-14.99) [2.100000e+00] (1154.00944150736,-14.904) [2.200000e+00] (1152.70283927249,-14.772) [2.200000e+00] (1151.45485883103,-14.676) [2.200000e+00] (1149.95305257152,-14.522) [2.200000e+00] (1148.63615195832,-14.381) [2.300000e+00] (1147.34102719149,-14.204) [2.400000e+00] (1145.82103639964,-14.04) [2.400000e+00] (1144.74892179738,-13.848) [2.600000e+00] (1143.56792615695,-13.716) [2.600000e+00] (1142.35018458587,-13.602) [2.600000e+00] (1141.65187477676,-13.491) [2.600000e+00] (1140.30425583196,-13.382) [2.600000e+00] (1138.82791586104,-13.233) [2.600000e+00] (1137.18251807842,-13.105) [2.700000e+00] (1135.59875521519,-12.947) [2.700000e+00] (1134.26075561296,-12.762) [2.800000e+00] (1133.03698558365,-12.587) [2.800000e+00] (1131.26015143414,-12.428) [2.800000e+00] (1130.22575918218,-12.33) [2.900000e+00] (1128.87269085595,-12.251) [3] (1127.56463102108,-12.13) [3.100000e+00] (1126.03793230557,-12.02) [3.100000e+00] (1124.6570592156,-11.9) [3.300000e+00] (1123.48141170852,-11.745) [3.400000e+00] (1122.05251433599,-11.596) [3.500000e+00] (1120.54166164881,-11.451) [3.700000e+00] (1119.30359974042,-11.308) [3.900000e+00] (1117.97084917283,-11.201) [4.200000e+00] (1116.76506195247,-11.128) [4.700000e+00] (1115.52388738955,-10.998) [5.100000e+00] (1114.06708615421,-10.884) [5.500000e+00] (1112.90203572806,-10.767) [5.900000e+00] (1111.48742851512,-10.605) [6.400000e+00] (1109.30596956348,-10.398) [6.800000e+00] (1107.88902697108,-10.193) [7.100000e+00] (1106.47286300133,-10.049) [7.600000e+00] (1105.07322590136,-9.923) [8] (1103.8073583699,-9.776) [8.300000e+00] (1102.3896380727,-9.614) [8.500000e+00] (1101.1365061219,-9.466) [8.700000e+00] (1099.64742802074,-9.296) [8.700000e+00] (1098.08096586433,-9.195) [8.800000e+00] (1097.40520745745,-9.079) [8.800000e+00] (1097.40520745745,-8.921) [8.900000e+00] (1094.21064647718,-8.686) [8.600000e+00] (1091.40446483745,-8.448) [8.200000e+00] (1089.8424730639,-8.213) [8] (1087.99475856794,-7.98) [7.600000e+00] (1086.07160169445,-7.758) [7.300000e+00] (1084.60935340063,-7.561) [7] (1083.08799843377,-7.451) [6.500000e+00] (1081.81619933602,-7.328) [6.300000e+00] (1079.1385388049,-6.985) [5.700000e+00] (1077.74570397125,-6.806) [5.500000e+00] (1075.8613386601,-6.601) [5.300000e+00] (1074.6226901946,-6.406) [5.200000e+00] (1073.3772368842,-6.223) [5.100000e+00] (1071.81679794179,-6.058) [4.900000e+00] (1070.29388578812,-5.885) [4.700000e+00] (1069.06184758176,-5.705) [4.400000e+00] (1067.97097000306,-5.499) [4.100000e+00] (1066.61410523025,-5.298) [3.900000e+00] (1065.01001506618,-5.076) [3.700000e+00] (1063.42099365522,-4.917) [3.500000e+00] (1061.67934088511,-4.744) [3.300000e+00] (1060.29311502131,-4.554) [3.200000e+00] (1058.8272692515,-4.356) [3.100000e+00] (1057.35101926743,-4.137) [2.900000e+00] (1055.91375421798,-3.9) [2.800000e+00] (1054.31345448368,-3.696) [2.600000e+00] (1052.81309462103,-3.482) [2.500000e+00] (1051.1007989871,-3.245) [2.500000e+00] (1049.3847134108,-3.021) [2.200000e+00] (1048.03309632509,-2.8) [2.100000e+00] (1046.54547161327,-2.547) [2] (1044.95256033064,-2.301) [1.800000e+00] (1043.36723208579,-2.06) [1.800000e+00] (1041.79240268213,-1.822) [1.600000e+00] (1040.44904904586,-1.568) [1.600000e+00] (1039.08693180037,-1.321) [1.600000e+00] (1037.69623386343,-1.079) [1.600000e+00] (1036.08465451642,-0.868) [1.500000e+00] (1034.44371585572,-0.783) [1.400000e+00] (1033.01169895211,-0.809) [1.300000e+00] (1031.81047651529,-0.884) [1.300000e+00] (1030.48452302764,-0.952) [1.300000e+00] (1029.18034673384,-1.076) [1.200000e+00] (1027.77389653641,-1.321) [1.100000e+00] (1026.29161594517,-1.603) [1.200000e+00] (1024.95963488762,-1.873) [1.100000e+00] (1023.55473969911,-2.15) [1.100000e+00] (1022.15509437549,-2.42) [1.200000e+00] (1020.91547149935,-2.691) [1.300000e+00] (1019.53760293993,-2.956) [1.400000e+00] (1018.0906119683,-3.226) [1.500000e+00] (1016.85176547873,-3.472) [1.700000e+00] (1015.44609304413,-3.729) [1.700000e+00] (1013.98549327128,-3.968) [1.900000e+00] (1012.72564643332,-4.216) [2] (1011.50040877402,-4.458) [2.100000e+00] (1010.18271655804,-4.693) [2.200000e+00] (1008.81690184902,-4.906) [2.400000e+00] (1007.49541834364,-5.12) [2.600000e+00] (1006.07924486067,-5.318) [2.700000e+00] (1004.80656649648,-5.51) [2.900000e+00] (1003.43919637738,-5.701) [3.300000e+00] (1001.96439935303,-5.881) [3.800000e+00] (1000.42952018238,-6.067) [4.100000e+00] (999.025399661174,-6.246) [4.500000e+00] (997.639265044581,-6.463) [5.100000e+00] (996.359876921423,-6.666) [5.600000e+00] (995.352506686828,-6.863) [6.400000e+00] (994.088185969619,-7.043) [6.900000e+00] (992.909420795723,-7.216) [7.500000e+00] (991.676506447192,-7.366) [7.800000e+00] (990.383509522842,-7.546) [8.300000e+00] (989.014679080286,-7.697) [8.600000e+00] (987.664612705648,-7.809) [8.900000e+00] (986.142468285034,-7.963) [9] (984.908188211044,-8.161) [9.200000e+00] (983.809139806535,-8.371) [9.300000e+00] (982.538014273264,-8.517) [9.200000e+00] (981.584017314541,-8.678) [9.100000e+00] (980.236963636116,-8.833) [9] (978.820009840593,-8.966) [8.900000e+00] (977.579702196686,-9.142) [8.700000e+00] (976.333366983251,-9.291) [8.400000e+00] (974.995354029198,-9.436) [8.100000e+00] (973.748335279327,-9.584) [8] (972.383977025846,-9.742) [7.700000e+00] (970.971589587974,-9.878) [7.400000e+00] (969.771046389772,-9.985) [7] (968.438965977565,-10.063) [7] (967.252713368873,-10.184) [6.700000e+00] (965.947755072429,-10.319) [6.500000e+00] (964.704432125032,-10.502) [6.200000e+00] (963.615200860472,-10.611) [5.900000e+00] (962.416210435395,-10.753) [5.700000e+00] (961.219550058383,-10.836) [5.400000e+00] (960.044667305046,-10.976) [5.300000e+00] (958.835175372754,-11.115) [5.100000e+00] (957.424246998191,-11.266) [4.900000e+00] (956.046466991233,-11.354) [4.700000e+00] (954.647592008332,-11.481) [4.500000e+00] (953.57187362845,-11.668) [4.300000e+00] (952.314259690758,-11.837) [4.300000e+00] (950.79357218468,-11.974) [4.100000e+00] (949.145137406695,-12.058) [3.900000e+00] (947.571076682795,-12.207) [3.700000e+00] (946.170647637567,-12.346) [3.400000e+00] (944.661235818084,-12.487) [3.400000e+00] (943.210447999136,-12.585) [3.300000e+00] (941.662730797541,-12.775) [3.200000e+00] (940.293027279379,-12.973) [3.200000e+00] (939.004594592698,-13.165) [3.100000e+00] (937.636536209829,-13.33) [3.100000e+00] (936.11507059816,-13.467) [3] (934.62150079421,-13.592) [3] (933.23759910671,-13.692) [2.900000e+00] (931.893555574444,-13.848) [2.800000e+00] (930.551748781447,-13.986) [2.700000e+00] (929.248247399485,-14.097) [2.600000e+00] (927.747098170069,-14.231) [2.600000e+00] (926.301559385583,-14.371) [2.500000e+00] (924.867388650044,-14.492) [2.400000e+00] (923.572242650441,-14.625) [2.300000e+00] (922.27097892514,-14.741) [2.200000e+00] (920.924697335324,-14.918) [2.200000e+00] (919.688082514908,-15.092) [2.100000e+00] (918.430465539635,-15.243) [2] (917.186364114915,-15.374) [2.100000e+00] (915.82200202236,-15.502) [2.100000e+00] (914.424583065523,-15.625) [2.100000e+00] (913.015111634495,-15.772) [2] (911.719873731302,-15.879) [1.900000e+00] (910.212788033911,-15.976) [1.900000e+00] (908.793591023742,-16.075) [1.900000e+00] (907.31577047193,-16.206) [1.800000e+00] (906.010031061392,-16.345) [1.800000e+00] (904.674249259621,-16.479) [1.700000e+00] (903.277606718308,-16.586) [1.700000e+00] (902.046332405872,-16.711) [1.500000e+00] (900.691784759011,-16.787) [1.500000e+00] (899.168855327436,-16.949) [1.500000e+00] (897.633871185916,-17.075) [1.500000e+00] (896.286811060699,-17.175) [1.500000e+00] (894.941212205776,-17.279) [1.500000e+00] (893.451337431782,-17.408) [1.500000e+00] (892.046429692659,-17.514) [1.700000e+00] (890.675262842406,-17.627) [1.600000e+00] (889.465676485758,-17.746) [1.500000e+00] (888.108799766306,-17.857) [1.500000e+00] (886.656545061023,-17.962) [1.500000e+00] (885.383859907284,-17.912) [1.500000e+00] (884.01889959068,-17.724) [1.600000e+00] (882.901082477351,-17.523) [1.600000e+00] (881.802806248483,-17.323) [1.600000e+00] (880.498616544693,-17.138) [1.700000e+00] (879.026820095559,-16.931) [1.600000e+00] (877.809744039519,-16.644) [1.700000e+00] (876.429627107761,-16.397) [1.700000e+00] (875.025493577463,-16.178) [1.800000e+00] (873.410104774697,-16) [1.700000e+00] (871.791017962395,-15.804) [1.800000e+00] (870.537186602619,-15.615) [1.800000e+00] (868.694701458534,-15.415) [1.700000e+00] (867.08757745238,-15.223) [1.800000e+00] (865.614227399732,-15.017) [1.900000e+00] (864.385873937612,-14.841) [2] (862.610365940251,-14.677) [2] (861.020455064246,-14.538) [2] (859.453187436334,-14.406) [2] (857.99033937838,-14.318) [2] (856.571822954067,-14.22) [2] (854.978212291562,-14.087) [1.900000e+00] (853.520614127641,-13.95) [2] (852.117851340282,-13.827) [2.100000e+00] (850.861692255118,-13.705) [2.100000e+00] (849.461255932683,-13.577) [2.100000e+00] (848.147248345326,-13.504) [2.100000e+00] (846.819725236414,-13.433) [2] (845.412486304952,-13.343) [2] (844.093228422723,-13.258) [2] (843.008464706853,-13.176) [2] (841.739565598268,-13.075) [2.100000e+00] (840.369171696223,-12.955) [2.100000e+00] (838.998090674602,-12.825) [2.100000e+00] (837.579663116557,-12.692) [2.100000e+00] (836.165709954415,-12.531) [2.100000e+00] (834.610492651815,-12.388) [2.200000e+00] (833.273929335062,-12.189) [2.100000e+00] (831.820893974603,-12.039) [2.300000e+00] (830.445248000617,-11.842) [2.300000e+00] (829.067364399022,-11.769) [2.400000e+00] (827.567756652624,-11.585) [2.500000e+00] (826.543678783466,-11.478) [2.700000e+00] (825.166569332433,-11.338) [2.800000e+00] (824.108060359763,-11.327) [3] (822.891246831437,-11.222) [3.100000e+00] (821.550981951035,-11.135) [3.200000e+00] (820.181273908338,-10.933) [3.300000e+00] (818.591447513916,-10.779) [3.300000e+00] (816.733293301337,-10.624) [3.500000e+00] (815.407406012638,-10.487) [3.600000e+00] (814.196441607155,-10.307) [3.800000e+00] (812.971186182143,-10.191) [4.200000e+00] (811.725702753407,-10.067) [4.500000e+00] (810.427532796457,-9.957) [4.900000e+00] (809.264602505868,-9.838) [5.200000e+00] (807.970133253125,-9.748) [5.600000e+00] (806.181878268759,-9.589) [5.900000e+00] (804.594287782957,-9.559) [6.300000e+00] (802.105555486717,-9.486) [6.700000e+00] (800.381863293485,-9.376) [7] (798.783770412051,-9.249) [7.300000e+00] (797.011179855664,-9.063) [7.800000e+00] (795.607037184189,-8.843) [7.700000e+00] (794.659749559802,-8.668) [7.700000e+00] (793.302178844221,-8.469) [7.500000e+00] (791.987392378577,-8.3) [7.300000e+00] (790.411766075439,-8.112) [7.200000e+00] (788.89407616534,-7.907) [6.900000e+00] (787.509387404899,-7.692) [6.600000e+00] (786.169032943847,-7.578) [6.300000e+00] (784.654355729727,-7.359) [6.100000e+00] (783.2185318382,-7.2) [5.900000e+00] (781.634727798284,-7.013) [5.600000e+00] (780.26510374907,-6.825) [5.400000e+00] (778.882739217905,-6.62) [5.200000e+00] (777.507088946264,-6.467) [5] (776.020216065958,-6.23) [4.900000e+00] (774.687436157821,-6.063) [4.600000e+00] (773.289231904957,-5.865) [4.500000e+00] (771.727205517642,-5.652) [4.200000e+00] (770.211063508091,-5.418) [4] (768.843677144825,-5.289) [3.600000e+00] (767.339499718543,-5.098) [3.400000e+00] (765.648947064977,-4.894) [3.300000e+00] (763.967431982819,-4.717) [3.100000e+00] (762.349793431481,-4.537) [3] (760.780966270365,-4.291) [2.800000e+00] (759.322585789947,-4.052) [2.700000e+00] (757.678604662749,-3.841) [2.500000e+00] (756.224011888574,-3.615) [2.500000e+00] (754.822708024827,-3.376) [2.400000e+00] (753.404959821386,-3.137) [2.400000e+00] (752.036797588258,-2.926) [2.300000e+00] (750.655119833938,-2.693) [2.300000e+00] (749.306499373713,-2.454) [2.300000e+00] (747.881175786812,-2.226) [2.200000e+00] (746.286089956341,-2.016) [2.100000e+00] (744.90819901951,-1.767) [2] (743.457393550743,-1.543) [1.900000e+00] (742.086218221031,-1.313) [1.900000e+00] (740.592541529239,-1.089) [1.900000e+00] (739.105665640431,-0.847) [1.900000e+00] (737.7164968392,-0.676) [1.700000e+00] (736.467131767602,-0.628) [1.700000e+00] (735.135123890202,-0.685) [1.600000e+00] (733.718148548226,-0.741) [1.600000e+00] (732.301173664802,-0.81) [1.700000e+00] (730.732427802551,-0.883) [1.600000e+00] (729.376400202102,-0.948) [1.600000e+00] (728.017359475809,-1.048) [1.700000e+00] (726.529017903125,-1.291) [1.700000e+00] (725.048855443399,-1.575) [1.700000e+00] (724.063259916447,-1.853) [1.600000e+00] (722.862015102309,-2.127) [1.600000e+00] (721.779400983767,-2.401) [1.700000e+00] (721.059813058517,-2.663) [1.700000e+00] (719.74355749989,-2.924) [1.700000e+00] (718.640712252961,-3.171) [1.700000e+00] (717.080312399596,-3.439) [1.800000e+00] (715.637162724711,-3.711) [1.900000e+00] (714.717848681874,-3.989) [2] (713.795522402028,-4.25) [2.200000e+00] (712.4732401908,-4.505) [2.300000e+00] (711.563571891685,-4.745) [2.400000e+00] (710.429215845819,-5.008) [2.700000e+00] (708.613226077895,-5.21) [2.800000e+00] (707.463805349647,-5.419) [2.900000e+00] (706.149788365231,-5.616) [3.200000e+00] (704.48264035703,-5.833) [3.600000e+00] (703.082964775302,-5.999) [4.100000e+00] (701.378971766709,-6.211) [4.700000e+00] (699.766577063555,-6.414) [5] (698.412010820508,-6.62) [5.100000e+00] (697.28135820226,-6.828) [5.300000e+00] (695.723027419745,-7.009) [5.500000e+00] (693.925028085888,-7.176) [5.500000e+00] (692.01046930519,-7.311) [5.500000e+00] (690.546833372907,-7.521) [5.500000e+00] (688.969042947106,-7.744) [5.300000e+00] (687.477600932657,-7.97) [5.200000e+00] (686.001222740964,-8.176) [5.100000e+00] (684.535347403717,-8.413) [5.100000e+00] (683.18448415858,-8.566) [5] (681.766730826862,-8.751) [4.800000e+00] (680.154334404616,-8.881) [4.900000e+00] (678.643350216711,-9.073) [4.800000e+00] (677.427647231597,-9.246) [4.700000e+00] (675.909172397437,-9.432) [4.500000e+00] (674.521463292103,-9.613) [4.400000e+00] (673.066173827466,-9.773) [4.200000e+00] (671.757405935361,-9.883) [4.100000e+00] (670.499079748215,-10.018) [3.900000e+00] (669.162587845816,-10.214) [3.800000e+00] (667.686982056748,-10.411) [3.700000e+00] (666.335425013387,-10.614) [3.700000e+00] (664.983012665348,-10.763) [3.700000e+00] (663.78245649205,-10.844) [3.500000e+00] (662.517331307941,-10.916) [3.400000e+00] (661.261162053384,-11.024) [3.300000e+00] (659.833676951407,-11.13) [3.100000e+00] (658.893880722283,-11.263) [3] (657.489723293332,-11.381) [3] (656.038218468331,-11.511) [2.800000e+00] (654.64077845257,-11.601) [2.700000e+00] (653.198310310169,-11.77) [2.600000e+00] (652.005325995222,-11.966) [2.600000e+00] (650.819832756161,-12.174) [2.500000e+00] (649.680913250323,-12.358) [2.400000e+00] (648.514961354301,-12.52) [2.400000e+00] (647.255779032033,-12.684) [2.300000e+00] (645.969645043784,-12.852) [2.200000e+00] (644.835285588577,-12.973) [2.200000e+00] (643.809057558051,-13.065) [2.100000e+00] (642.577681346561,-13.175) [2] (641.42532707258,-13.286) [2.100000e+00] (640.168464937323,-13.368) [2] (639.223335911926,-13.427) [2] (638.222755741307,-13.542) [2] (637.205727647829,-13.651) [2] (636.106746591512,-13.781) [2] (635.017575881011,-13.872) [1.900000e+00] (633.795317993951,-13.939) [1.800000e+00] (632.0468972664,-13.954) [1.800000e+00] (630.951008283034,-13.974) [1.800000e+00] (629.405737615541,-13.993) [1.800000e+00] (627.225769032791,-14.015) [1.800000e+00] (625.121887779129,-14.154) [1.700000e+00] (621.551514734018,-14.31) [1.800000e+00] (619.900719910773,-14.515) [1.800000e+00] (618.054506527998,-14.815) [1.900000e+00] (616.722406206452,-15.044) [1.900000e+00] (615.389453246023,-15.222) [1.900000e+00] (613.606504179905,-15.332) [1.900000e+00] (611.7717275025,-15.455) [1.800000e+00] (610.45383747623,-15.615) [1.700000e+00] (609.309504508075,-15.77) [1.700000e+00] (607.701578271135,-15.912) [1.700000e+00] (606.007140826818,-16.065) [1.600000e+00] (604.578876846683,-16.26) [1.600000e+00] (604.579944728651,-16.448) [1.600000e+00] (602.816058880144,-17.675) [1.500000e+00] (601.360921292768,-17.85) [1.400000e+00] (600.091391460551,-17.933) [1.500000e+00] (598.824794459762,-17.917) [1.500000e+00] (597.396609455813,-17.918) [1.400000e+00] (596.325431902945,-17.884) [1.400000e+00] (594.77543626897,-17.841) [1.400000e+00] (593.20219325654,-17.849) [1.400000e+00] (591.218648551032,-17.893) [1.400000e+00] (589.956529988152,-18.024) [1.400000e+00] (588.412638664266,-18.182) [1.400000e+00] (582.336623377122,-17.886) [1.400000e+00] (581.114974725035,-17.673) [1.500000e+00] (579.406245030146,-17.449) [1.500000e+00] (577.859181396896,-17.216) [1.500000e+00] (576.218036074547,-16.96) [1.500000e+00] (574.871644878632,-16.696) [1.600000e+00] (574.101622676206,-16.423) [1.500000e+00] (573.088755141335,-16.269) [1.600000e+00] (572.01765346925,-16.235) [1.500000e+00] (571.115710711763,-16.258) [1.500000e+00] (570.11366093119,-16.202) [1.500000e+00] (568.835089143089,-16.04) [1.500000e+00] (567.203989259495,-15.781) [1.500000e+00] (565.714856658791,-15.552) [1.500000e+00] (564.110633535895,-15.333) [1.600000e+00] (562.323118373242,-15.117) [1.700000e+00] (560.845956633901,-14.925) [1.700000e+00] (558.693089904352,-14.678) [1.700000e+00] (556.77395379491,-14.459) [1.700000e+00] (555.322281132898,-14.267) [1.800000e+00] (553.985698563551,-14.124) [1.800000e+00] (552.830168170264,-13.968) [1.800000e+00] (551.241779437419,-13.825) [1.900000e+00] (550.045775579096,-13.684) [2] (548.78365311927,-13.566) [2] (547.64094573849,-13.475) [2.100000e+00] (546.247281709027,-13.393) [2.200000e+00] (544.864894225709,-13.29) [2.200000e+00] (543.675610016303,-13.124) [2.200000e+00] (542.262402895629,-12.959) [2.100000e+00] (540.754421576806,-12.819) [2.100000e+00] (539.812463344637,-12.683) [2.200000e+00] (538.417254196222,-12.56) [2.200000e+00] (537.355726574113,-12.45) [2.300000e+00] (536.175479506121,-12.368) [2.300000e+00] (535.025355553696,-12.339) [2.400000e+00] (533.976650514154,-12.292) [2.300000e+00] (532.90840327187,-12.193) [2.400000e+00] (531.945358606919,-12.092) [2.400000e+00] (530.807284462883,-12.001) [2.400000e+00] (527.235027128636,-11.747) [2.900000e+00] (526.042883274045,-11.672) [3] (524.60040034535,-11.509) [3.300000e+00] (522.839914622353,-11.305) [3.400000e+00] (521.397431263821,-11.086) [3.700000e+00] (519.694103741061,-10.819) [3.800000e+00] (518.208827218271,-10.547) [3.900000e+00] (516.93241250787,-10.353) [4] (515.450071510865,-10.218) [4.300000e+00] (514.217993765289,-10.058) [4.700000e+00] (512.707073453311,-9.95) [5.100000e+00] (511.298496015165,-9.809) [5.400000e+00] (509.637339639398,-9.679) [5.900000e+00] (508.222356493428,-9.598) [6.400000e+00] (506.457387488947,-9.468) [6.500000e+00] (503.337214264038,-9.275) [6.900000e+00] (504.013696442417,-9.182) [7.100000e+00] (500.802379777837,-8.83) [7.400000e+00] (496.377132778696,-8.687) [7.400000e+00] (494.824658580862,-8.493) [7.700000e+00] (493.371515019616,-8.271) [8] (491.46364903889,-8.071) [8] (490.176029649865,-7.884) [7.900000e+00] (488.813799911123,-7.688) [7.700000e+00] (487.44716532124,-7.51) [7.400000e+00] (485.841467802819,-7.344) [7] (484.422154797968,-7.157) [6.800000e+00] (482.920192767225,-6.919) [6.600000e+00] (481.396448211589,-6.719) [6.400000e+00] (479.886992869078,-6.51) [6.200000e+00] (478.461654988851,-6.314) [5.800000e+00] (477.141597285648,-6.113) [5.400000e+00] (476.059754372661,-5.864) [5.100000e+00] (474.662298856012,-5.656) [4.800000e+00] (473.372288015914,-5.452) [4.500000e+00] (472.003336909058,-5.306) [4.300000e+00] (470.420267446521,-5.114) [4] (469.318957264921,-4.93) [3.800000e+00] (468.328178443739,-4.74) [3.700000e+00] (466.802192276234,-4.557) [3.400000e+00] (466.072110445456,-4.38) [3.300000e+00] (464.397353067545,-4.175) [3.200000e+00] (463.323849885504,-3.982) [3] (462.121816453522,-3.843) [2.900000e+00] (460.885179290108,-3.638) [2.800000e+00] (459.689169873697,-3.435) [2.700000e+00] (458.499109538703,-3.256) [2.600000e+00] (457.153558373643,-3.062) [2.400000e+00] (455.846319727767,-2.881) [2.200000e+00] (454.849515601345,-2.71) [2.200000e+00] (453.717386270204,-2.514) [2.100000e+00] (452.289108938242,-2.32) [2.100000e+00] (450.942784998619,-2.134) [2] (449.543858896415,-1.925) [1.900000e+00] (448.331392642486,-1.733) [1.800000e+00] (447.124876387937,-1.526) [1.700000e+00] (445.53037884952,-1.316) [1.500000e+00] (444.043474938491,-1.118) [1.600000e+00] (442.586692337034,-0.911) [1.500000e+00] (441.311814492722,-0.735) [1.600000e+00] (440.0640461795,-0.656) [1.400000e+00] (438.629046132091,-0.692) [1.400000e+00] (437.292602570498,-0.756) [1.300000e+00] (435.862927751959,-0.824) [1.300000e+00] (434.345202885215,-0.889) [1.300000e+00] (432.884562631887,-0.966) [1.300000e+00] (431.354788444461,-1.04) [1.200000e+00] (429.827255038153,-1.249) [1.100000e+00] (428.154660527987,-1.524) [1.100000e+00] (426.650452558797,-1.795) [1.200000e+00] (425.147016934149,-2.061) [1.100000e+00] (423.91794452047,-2.341) [1.100000e+00] (422.603906506131,-2.611) [1.200000e+00] (421.437942262887,-2.879) [1.200000e+00] (420.039710992459,-3.129) [1.300000e+00] (418.866951765254,-3.384) [1.400000e+00] (417.506262475924,-3.65) [1.400000e+00] (416.27417696535,-3.909) [1.500000e+00] (414.786499019738,-4.183) [1.600000e+00] (413.449907053332,-4.418) [1.700000e+00] (411.87208275713,-4.638) [1.900000e+00] (410.711371532222,-4.879) [2.300000e+00] (409.558925075052,-5.088) [2.600000e+00] (408.143464651823,-5.339) [3.100000e+00] (406.213105668329,-5.576) [3.500000e+00] (404.934368453107,-5.82) [4] (403.509798706878,-6.026) [4.500000e+00] (402.231832488814,-6.248) [4.700000e+00] (400.939648408512,-6.508) [5.200000e+00] (399.8000218685,-6.782) [5.600000e+00] (398.428897091865,-7.009) [6.100000e+00] (397.144207670668,-7.19) [6.400000e+00] (395.809926976744,-7.373) [6.500000e+00] (394.264390812682,-7.487) [6.500000e+00] (392.887242049321,-7.659) [6.400000e+00] (391.595057539338,-7.824) [6.400000e+00] (389.945084922732,-7.994) [6.200000e+00] (388.619768242425,-8.137) [6] (387.257607478938,-8.319) [5.800000e+00] (386.070628670456,-8.481) [5.600000e+00] (384.63709276194,-8.642) [5.500000e+00] (383.298256547178,-8.788) [5.400000e+00] (381.759442296891,-8.932) [5.200000e+00] (380.392028114153,-9.128) [5.100000e+00] (378.620245872335,-9.316) [5] (377.146853672631,-9.45) [4.800000e+00] (375.758425755201,-9.586) [4.600000e+00] (374.431636654369,-9.812) [4.500000e+00] (373.026749634537,-9.992) [4.200000e+00] (371.633767944389,-10.181) [4] (370.147556344525,-10.372) [3.900000e+00] (368.772645754694,-10.537) [3.800000e+00] (367.419520384683,-10.72) [3.800000e+00] (365.718340112779,-10.891) [3.700000e+00] (364.53731402696,-11.024) [3.500000e+00] (363.142860804626,-11.194) [3.400000e+00] (361.837854716421,-11.343) [3.200000e+00] (360.308845968063,-11.465) [3.100000e+00] (358.891067337667,-11.541) [3.100000e+00] (357.468805826649,-11.64) [3] (356.081846833763,-11.805) [3.100000e+00] (354.687393640509,-12.03) [3.100000e+00] (353.56436406112,-12.177) [2.900000e+00] (352.092439955415,-12.367) [2.800000e+00] (350.815941430591,-12.459) [2.700000e+00] (349.210231907574,-12.532) [2.600000e+00] (347.835320487121,-12.681) [2.500000e+00] (346.313734675275,-12.842) [2.400000e+00] (344.722313476962,-12.962) [2.400000e+00] (343.578129296234,-13.139) [2.400000e+00] (342.343797417454,-13.309) [2.400000e+00] (341.160528831294,-13.438) [2.300000e+00] (339.802917084292,-13.576) [2.200000e+00] (338.470872415071,-13.772) [2.100000e+00] (337.123067430278,-13.946) [2] (335.963122504192,-14.113) [1.900000e+00] (334.516833731827,-14.28) [2] (333.135967010073,-14.407) [1.900000e+00] (331.800138705769,-14.518) [1.900000e+00] (330.585278625369,-14.621) [1.900000e+00] (328.944333951,-14.743) [1.900000e+00] (327.462672256065,-14.817) [2] (325.863053656869,-14.956) [1.900000e+00] (324.267147858687,-15.048) [1.800000e+00] (322.789267162244,-15.187) [1.800000e+00] (321.547368901744,-15.314) [1.900000e+00] (320.434774066272,-15.475) [1.900000e+00] (318.941063968179,-15.593) [1.900000e+00] (317.374505816666,-15.746) [1.800000e+00] (315.678644240174,-15.888) [1.800000e+00] (314.236836199551,-16.018) [1.700000e+00] (313.025055694277,-16.127) [1.700000e+00] (311.47194713154,-16.285) [1.600000e+00] (309.951197964876,-16.378) [1.600000e+00] (308.346255112257,-16.52) [1.600000e+00] (307.066041265706,-16.728) [1.600000e+00] (305.757950015311,-16.904) [1.500000e+00] (304.228166000962,-17.04) [1.400000e+00] (302.837420393518,-17.21) [1.500000e+00] (301.38888694296,-17.305) [1.500000e+00] (300.125972740781,-17.421) [1.400000e+00] (298.885545669475,-17.498) [1.400000e+00] (297.497812299326,-17.64) [1.400000e+00] (296.186777955744,-17.75) [1.400000e+00] (295.077193179508,-17.95) [1.400000e+00] (293.818061323182,-18.021) [1.300000e+00] (292.831688676552,-18.15) [1.400000e+00] (291.631116119339,-18.135) [1.500000e+00] (290.767321429905,-18.024) [1.600000e+00] (289.851691562145,-17.851) [1.500000e+00] (288.879675867614,-17.737) [1.500000e+00] (287.788162191773,-17.6) [1.500000e+00] (286.795130956376,-17.523) [1.500000e+00] (285.551827183263,-17.406) [1.500000e+00] (284.290452797649,-17.276) [1.500000e+00] (282.823845639159,-17.113) [1.600000e+00] (281.499290901627,-16.954) [1.500000e+00] (280.021405878581,-16.781) [1.500000e+00] (278.72767166716,-16.576) [1.500000e+00] (277.142268008781,-16.423) [1.500000e+00] (275.607996729204,-16.27) [1.500000e+00] (274.272934235597,-16.157) [1.500000e+00] (272.625255381976,-15.987) [1.500000e+00] (271.167749090406,-15.859) [1.600000e+00] (269.557615771968,-15.682) [1.600000e+00] (268.169177453186,-15.495) [1.500000e+00] (267.007054756922,-15.455) [1.500000e+00] (266.115383180306,-15.444) [1.500000e+00] (265.284515238342,-15.483) [1.400000e+00] (264.289942954844,-15.486) [1.400000e+00] (263.362968860198,-15.512) [1.500000e+00] (262.368396576777,-15.477) [1.500000e+00] (261.624099037979,-15.475) [1.600000e+00] (260.974571138378,-15.504) [1.600000e+00] (260.234825136624,-15.507) [1.700000e+00] (259.516929398633,-15.546) [1.700000e+00] (258.667489942075,-15.551) [1.700000e+00] (257.718797047891,-15.541) [1.700000e+00] (256.513805422619,-15.453) [1.700000e+00] (255.238842044868,-15.289) [1.700000e+00] (253.892500397529,-15.069) [1.700000e+00] (252.540905301887,-14.791) [1.800000e+00] (251.135934381069,-14.547) [1.800000e+00] (249.629469270326,-14.262) [1.700000e+00] (248.307154512368,-14.009) [1.800000e+00] (247.062306625014,-13.859) [1.800000e+00] (245.870064483537,-13.79) [1.800000e+00] (244.640275918067,-13.678) [1.900000e+00] (243.466872720712,-13.62) [1.800000e+00] (242.302504399783,-13.493) [1.800000e+00] (241.107250050036,-13.362) [1.900000e+00] (240.115622294194,-13.256) [1.900000e+00] (239.07145656035,-13.162) [1.900000e+00] (237.810011427514,-13.081) [2] (236.76731771346,-12.946) [2] (235.572062933929,-12.859) [2] (234.058166030276,-12.752) [2.100000e+00] (232.803513215808,-12.602) [2.200000e+00] (231.524832652155,-12.382) [2.400000e+00] (230.312276909916,-12.242) [2.500000e+00] (228.961382465307,-12.025) [2.500000e+00] (227.440625992459,-11.781) [2.600000e+00] (225.910834642978,-11.517) [2.600000e+00] (224.328437147144,-11.306) [2.700000e+00] (222.966265726449,-11.074) [2.900000e+00] (221.624405728609,-10.955) [2.900000e+00] (220.277291393918,-10.845) [3] (219.176034396724,-10.817) [3.200000e+00] (217.876271471216,-10.759) [3.300000e+00] (216.625397907096,-10.632) [3.400000e+00] (215.474481010066,-10.524) [3.600000e+00] (214.282940866645,-10.368) [3.900000e+00] (212.944860519937,-10.21) [4.100000e+00] (211.566222000067,-10.085) [4.500000e+00] (210.420559010051,-9.937) [4.800000e+00] (209.215497984919,-9.768) [5.200000e+00] (208.040487836099,-9.618) [5.900000e+00] (206.379213886203,-9.469) [6.300000e+00] (205.008775090025,-9.312) [6.700000e+00] (203.713495991791,-9.097) [7] (202.159674285664,-8.827) [7.200000e+00] (201.018496441257,-8.636) [7.300000e+00] (199.746540973915,-8.446) [7.500000e+00] (198.303952481099,-8.305) [7.900000e+00] (196.941075223403,-8.194) [7.800000e+00] (195.637530038682,-8.056) [7.700000e+00] (194.636163375407,-7.879) [7.600000e+00] (193.422195413169,-7.748) [7.300000e+00] (192.200538519048,-7.601) [7] (190.756347918267,-7.416) [6.700000e+00] (189.429479590574,-7.281) [6.500000e+00] (188.240244012092,-7.165) [6.200000e+00] (187.114956199569,-7.004) [6] (185.752076734914,-6.85) [5.700000e+00] (184.34408844842,-6.649) [5.400000e+00] (182.866899406683,-6.41) [5] (181.335629962636,-6.237) [4.800000e+00] (179.841203225041,-6.022) [4.400000e+00] (178.464803796198,-5.851) [4.200000e+00] (177.109421255634,-5.659) [4] (175.454870195167,-5.414) [3.800000e+00] (173.674795848668,-5.182) [3.600000e+00] (171.965332357637,-4.993) [3.400000e+00] (170.288226915292,-4.724) [3.200000e+00] (168.632138361122,-4.528) [3] (166.890380117385,-4.347) [2.900000e+00] (165.223781571413,-4.123) [2.700000e+00] (163.664702090405,-3.934) [2.600000e+00] (162.044750347582,-3.72) [2.400000e+00] (160.566853662593,-3.461) [2.200000e+00] (159.021292856571,-3.226) [2.100000e+00] (157.400571463901,-2.964) [2] (155.943754504227,-2.722) [2] (154.368910264961,-2.511) [2] (152.584284927235,-2.268) [2] (150.923708189979,-2.007) [1.900000e+00] (149.460804894204,-1.76) [1.700000e+00] (147.84309307329,-1.507) [1.700000e+00] (146.244924802141,-1.238) [1.500000e+00] (144.900815770371,-1.004) [1.600000e+00] (143.470269634734,-0.819) [1.600000e+00] (141.974301936405,-0.758) [1.500000e+00] (140.533951733203,-0.813) [1.500000e+00] (139.065024938698,-0.907) [1.400000e+00] (137.659980837027,-0.95) [1.400000e+00] (136.259484689078,-1.016) [1.400000e+00] (134.810100097212,-1.159) [1.300000e+00] (133.352449848955,-1.397) [1.300000e+00] (132.065432597066,-1.666) [1.200000e+00] (130.707802712561,-1.954) [1.200000e+00] (129.47492781155,-2.229) [1.100000e+00] (128.01958207023,-2.481) [1.200000e+00] (126.909988785689,-2.77) [1.200000e+00] (125.711714491731,-3.053) [1.200000e+00] (124.473584824369,-3.305) [1.300000e+00] (123.188103776746,-3.56) [1.400000e+00] (121.783120764148,-3.843) [1.500000e+00] (120.439010013312,-4.074) [1.600000e+00] (119.167817290445,-4.314) [1.800000e+00] (118.014589468897,-4.555) [2] (116.73359445556,-4.806) [2.200000e+00] (115.336876669515,-5.006) [2.500000e+00] (114.089006230463,-5.247) [2.900000e+00] (112.845621366624,-5.501) [3.200000e+00] (111.442112638754,-5.695) [3.700000e+00] (110.116776246778,-5.903) [4.300000e+00] (108.766641997032,-6.131) [4.700000e+00] (107.363900253314,-6.301) [5] (105.871709197718,-6.442) [5.300000e+00] (104.399829536001,-6.645) [5.900000e+00] (103.164710357685,-6.891) [6.200000e+00] (101.764211631149,-7.088) [6.500000e+00] (100.452455403654,-7.248) [6.900000e+00] (99.1256434665944,-7.419) [6.800000e+00] (97.8138262082112,-7.572) [6.900000e+00] (96.4456837531614,-7.747) [6.800000e+00] (95.0910621799263,-7.904) [6.700000e+00] (93.8258894888612,-8.108) [6.600000e+00] (92.4712674857886,-8.265) [6.500000e+00] (91.1993042826323,-8.413) [6.300000e+00] (89.8108500040459,-8.586) [6.200000e+00] (88.4675045472981,-8.778) [6] (87.2233505504801,-8.882) [5.900000e+00] (85.9964957392214,-9.078) [5.600000e+00] (84.5538977914222,-9.213) [5.500000e+00] (82.9722548164121,-9.334) [5.300000e+00] (81.6845268547643,-9.468) [5.200000e+00] (80.2584611374513,-9.613) [5] (78.8421963632209,-9.803) [4.900000e+00] (77.4747594605468,-9.955) [4.700000e+00] (76.0201768357193,-10.138) [4.600000e+00] (74.4649253703916,-10.29) [4.400000e+00] (73.0658395257547,-10.487) [4.300000e+00] (72.0216639070358,-10.694) [4.200000e+00] (70.9030332254584,-10.903) [4.100000e+00] (69.6167808958216,-11.103) [4] (68.3109841567949,-11.223) [3.900000e+00] (66.8631301533118,-11.33) [3.800000e+00] (65.4543627619866,-11.421) [3.700000e+00] (64.0959585290907,-11.544) [3.500000e+00] (62.7428099801227,-11.67) [3.400000e+00] (61.499422137068,-11.78) [3.300000e+00] (60.0959086527029,-11.931) [3.100000e+00] (58.7314822392054,-12.087) [3] (57.4038976438461,-12.225) [3] (56.0259503486373,-12.377) [2.900000e+00] (54.5811059167043,-12.505) [2.800000e+00] (53.3647589774474,-12.611) [2.700000e+00] (51.9417028894792,-12.698) [2.600000e+00] (50.5945747730972,-12.829) [2.600000e+00] (49.2136152988279,-12.976) [2.500000e+00] (47.7905574337383,-13.137) [2.400000e+00] (46.4434302349562,-13.28) [2.200000e+00] (45.2263154517417,-13.464) [2.100000e+00] (43.9813901439844,-13.587) [2] (42.7237127157393,-13.721) [2] (41.2721370546646,-13.854) [2] (39.9024541265132,-13.976) [1.900000e+00] (38.4576088640612,-14.069) [1.800000e+00] (37.0571054390084,-14.173) [1.800000e+00] (35.595727516336,-14.264) [1.800000e+00] (34.1794588741853,-14.401) [1.700000e+00] (32.8030437995526,-14.528) [1.800000e+00] (31.4867349197922,-14.672) [1.700000e+00] (30.2493666784941,-14.846) [1.800000e+00] (28.8631506897011,-14.954) [1.700000e+00] (27.5378083096434,-15.088) [1.600000e+00] (26.0328551041201,-15.203) [1.700000e+00] (24.6022977364807,-15.333) [1.600000e+00] (23.2303696133712,-15.447) [1.600000e+00] (21.8215983541257,-15.571) [1.500000e+00] (20.4962555443606,-15.687) [1.600000e+00] (19.0717190389646,-15.846) [1.500000e+00] (17.706521773341,-15.959) [1.500000e+00] (16.3338258065844,-16.087) [1.400000e+00] (14.8694353038863,-16.235) [1.500000e+00] (13.476428402087,-16.436) [1.400000e+00] (11.8399253450121,-16.599) [1.300000e+00] (10.4807502888891,-16.697) [1.300000e+00] (9.08927689692154,-16.787) [1.200000e+00] (7.78347631897693,-16.917) [1.200000e+00] (5.9876150635571,-17.024) [1.200000e+00] (4.93215682376435,-17.209) [1.200000e+00] (3.75188262730727,-17.434) [1.200000e+00] };

\end{axis}
\end{tikzpicture}%