### 2013-07-31 figurefactory/files/voned_0_3_poster.tex

% This file was created by matlab2tikz v0.2.3.
% Copyright (c) 2008--2012, Nico Schlömer <nico.schloemer@gmail.com>
% All rights reserved.
% 
% 
% 
\begin{tikzpicture}

\begin{axis}[%
tick label style={font=\Huge},
label style={font=\Huge},
label shift={-4pt},
xlabel shift={-6pt},
legend style={font=\Huge},
view={0}{90},
width=\figurewidth,
height=\figureheight,
scale only axis,
xmin=0, xmax=2,
xtick={0, 0.5, 1, 1.5, 2},
ymin=-2, ymax=4,
tickwidth=0cm,
]

\addplot [fill=olive!20!lightgray,draw=none,forget plot] coordinates{ (0,3)(0.01,3)(0.02,3)(0.03,3)(0.04,3)(0.05,3)(0.06,3)(0.07,3)(0.08,3)(0.09,3)(0.1,3)(0.11,3)(0.12,3)(0.13,3)(0.14,3)(0.15,3)(0.16,3)(0.17,3)(0.18,3)(0.19,3)(0.2,3)(0.21,3)(0.22,3)(0.23,3)(0.24,3)(0.25,3)(0.26,3)(0.27,3)(0.28,3)(0.29,3)(0.3,3)(0.31,3)(0.32,3)(0.33,3)(0.34,3)(0.35,3)(0.36,3)(0.37,3)(0.38,3)(0.39,3)(0.4,3)(0.41,3)(0.42,3)(0.43,3)(0.44,3)(0.45,3)(0.46,3)(0.47,3)(0.48,3)(0.49,3)(0.5,3)(0.51,3)(0.52,3)(0.53,3)(0.54,3)(0.55,3)(0.56,3)(0.57,3)(0.58,3)(0.59,3)(0.6,3)(0.61,3)(0.62,3)(0.63,3)(0.64,3)(0.65,3)(0.66,3)(0.67,3)(0.68,3)(0.69,3)(0.7,3)(0.71,3)(0.72,3)(0.73,3)(0.74,3)(0.75,3)(0.76,3)(0.77,3)(0.78,3)(0.79,3)(0.8,3)(0.81,3)(0.82,3)(0.83,3)(0.84,3)(0.85,3)(0.86,3)(0.87,3)(0.88,3)(0.89,3)(0.9,3)(0.91,3)(0.92,3)(0.93,3)(0.94,3)(0.95,3)(0.96,3)(0.97,3)(0.98,3)(0.99,3)(1,3)(1.01,3)(1.02,3)(1.03,3)(1.04,3)(1.05,3)(1.06,3)(1.07,3)(1.08,3)(1.09,3)(1.1,3)(1.11,3)(1.12,3)(1.13,3)(1.14,3)(1.15,3)(1.16,3)(1.17,3)(1.18,3)(1.19,3)(1.2,3)(1.21,3)(1.22,3)(1.23,3)(1.24,3)(1.25,3)(1.26,3)(1.27,3)(1.28,3)(1.29,3)(1.3,3)(1.31,3)(1.32,3)(1.33,3)(1.34,3)(1.35,3)(1.36,3)(1.37,3)(1.38,3)(1.39,3)(1.4,3)(1.41,3)(1.42,3)(1.43,3)(1.44,3)(1.45,3)(1.46,3)(1.47,3)(1.48,3)(1.49,3)(1.5,3)(1.51,3)(1.52,3)(1.53,3)(1.54,3)(1.55,3)(1.56,3)(1.57,3)(1.58,3)(1.59,3)(1.6,3)(1.61,3)(1.62,3)(1.63,3)(1.64,3)(1.65,3)(1.66,3)(1.67,3)(1.68,3)(1.69,3)(1.7,3)(1.71,3)(1.72,3)(1.73,3)(1.74,3)(1.75,3)(1.76,3)(1.77,3)(1.78,3)(1.79,3)(1.8,3)(1.81,3)(1.82,3)(1.83,3)(1.84,3)(1.85,3)(1.86,3)(1.87,3)(1.88,3)(1.89,3)(1.9,3)(1.91,3)(1.92,3)(1.93,3)(1.94,3)(1.95,3)(1.96,3)(1.97,3)(1.98,3)(1.99,3)(2,3)(2,-1)(1.99,-1)(1.98,-1)(1.97,-1)(1.96,-1)(1.95,-1)(1.94,-1)(1.93,-1)(1.92,-1)(1.91,-1)(1.9,-1)(1.89,-1)(1.88,-1)(1.87,-1)(1.86,-1)(1.85,-1)(1.84,-1)(1.83,-1)(1.82,-1)(1.81,-1)(1.8,-1)(1.79,-1)(1.78,-1)(1.77,-1)(1.76,-1)(1.75,-1)(1.74,-1)(1.73,-1)(1.72,-1)(1.71,-1)(1.7,-1)(1.69,-1)(1.68,-1)(1.67,-1)(1.66,-1)(1.65,-1)(1.64,-1)(1.63,-1)(1.62,-1)(1.61,-1)(1.6,-1)(1.59,-1)(1.58,-1)(1.57,-1)(1.56,-1)(1.55,-1)(1.54,-1)(1.53,-1)(1.52,-1)(1.51,-1)(1.5,-1)(1.49,-1)(1.48,-1)(1.47,-1)(1.46,-1)(1.45,-1)(1.44,-1)(1.43,-1)(1.42,-1)(1.41,-1)(1.4,-1)(1.39,-1)(1.38,-1)(1.37,-1)(1.36,-1)(1.35,-1)(1.34,-1)(1.33,-1)(1.32,-1)(1.31,-1)(1.3,-1)(1.29,-1)(1.28,-1)(1.27,-1)(1.26,-1)(1.25,-1)(1.24,-1)(1.23,-1)(1.22,-1)(1.21,-1)(1.2,-1)(1.19,-1)(1.18,-1)(1.17,-1)(1.16,-1)(1.15,-1)(1.14,-1)(1.13,-1)(1.12,-1)(1.11,-1)(1.1,-1)(1.09,-1)(1.08,-1)(1.07,-1)(1.06,-1)(1.05,-1)(1.04,-1)(1.03,-1)(1.02,-1)(1.01,-1)(1,-1)(0.99,-1)(0.98,-1)(0.97,-1)(0.96,-1)(0.95,-1)(0.94,-1)(0.93,-1)(0.92,-1)(0.91,-1)(0.9,-1)(0.89,-1)(0.88,-1)(0.87,-1)(0.86,-1)(0.85,-1)(0.84,-1)(0.83,-1)(0.82,-1)(0.81,-1)(0.8,-1)(0.79,-1)(0.78,-1)(0.77,-1)(0.76,-1)(0.75,-1)(0.74,-1)(0.73,-1)(0.72,-1)(0.71,-1)(0.7,-1)(0.69,-1)(0.68,-1)(0.67,-1)(0.66,-1)(0.65,-1)(0.64,-1)(0.63,-1)(0.62,-1)(0.61,-1)(0.6,-1)(0.59,-1)(0.58,-1)(0.57,-1)(0.56,-1)(0.55,-1)(0.54,-1)(0.53,-1)(0.52,-1)(0.51,-1)(0.5,-1)(0.49,-1)(0.48,-1)(0.47,-1)(0.46,-1)(0.45,-1)(0.44,-1)(0.43,-1)(0.42,-1)(0.41,-1)(0.4,-1)(0.39,-1)(0.38,-1)(0.37,-1)(0.36,-1)(0.35,-1)(0.34,-1)(0.33,-1)(0.32,-1)(0.31,-1)(0.3,-1)(0.29,-1)(0.28,-1)(0.27,-1)(0.26,-1)(0.25,-1)(0.24,-1)(0.23,-1)(0.22,-1)(0.21,-1)(0.2,-1)(0.19,-1)(0.18,-1)(0.17,-1)(0.16,-1)(0.15,-1)(0.14,-1)(0.13,-1)(0.12,-1)(0.11,-1)(0.1,-1)(0.09,-1)(0.08,-1)(0.07,-1)(0.06,-1)(0.05,-1)(0.04,-1)(0.03,-1)(0.02,-1)(0.01,-1)(0,-1)};

\addplot [
color=olive!70!black,
solid,
line width=3.0pt,
forget plot
]
coordinates{
(0,1)(2,1)
};

\addplot [
color=black,
solid,
line width=2.5pt,
mark size=4pt,
mark=-,
forget plot
]
coordinates{
 (1.25,-1)(1.25,3)
};

\node at (axis cs:0.1, 1.5) [anchor=west,minimum size=20pt] {\Huge$\mu(\*x)$};
\draw[->,line width=1.5pt] (axis cs:0.43, 1.5) to[in=90,out=0] (axis cs:0.58, 1.1);

\node at (axis cs:1.6, 1.5) [anchor=east,minimum size=20pt] {\Huge$C(\*x)$};

\node at (axis cs:0.1, 3.5) [anchor=west,minimum size=20pt] {\Huge$\mu(\*x)+\beta\sigma(\*x)$};
\draw[->,line width=1.5pt] (axis cs:0.95, 3.5) to[in=90,out=0] (axis cs:1.1, 3.1);
\node at (axis cs:0.1, -1.5) [anchor=west,minimum size=20pt] {\Huge$\mu(\*x)-\beta\sigma(\*x)$};
\draw[->,line width=1.5pt] (axis cs:0.95, -1.5) to[in=270,out=0] (axis cs:1.1, -1.1);

\end{axis}
\end{tikzpicture}%